---
title: "Quality Checks"
author: "Rachel Voyt"
date: "`r Sys.Date()`"
output: html_document
---

# Fastp

For quality checks, I originally used the program *fastp*, which - unlike fastqc - allows for processing of paired-end reads.

Set up environment: - Go to directory:

```{bash}
cd /home/rachelvoyt/Documents/UT-Grad/Development/repos/tamarinGenetics_primatesPeru/seqAnalysis/01_run2_fecalHairBlood/
```

-   Activate fastp environment:

```{bash eval = F}
conda activate fastp-env
```

-   Use a bash loop to run fastp on all files:

```{bash eval = F}
for i in 01_run2Seqs/*R1*; do name=$(basename $i _L001_R1_001.fastq.gz); mkdir 00_qualityChecks/${name}; cd 00_qualityChecks/${name}; fastp -i ../../01_run2Seqs/${name}_L001_R1_001.fastq.gz -I ../../01_run2Seqs/${name}_L001_R2_001.fastq.gz; mv fastp.html ../${name}_fastp.html; mv fastp.json ../${name}_fastp.json; cd ../; rm -r ${name}; cd ../; done
```

-   Run multiqc to unify all reports:

```{bash eval = F}
conda activate multiqc-env
cd /home/rachelvoyt/Documents/UT-Grad/Development/repos/tamarinGenetics_primatesPeru/seqAnalysis/01_run2_fecalHairBlood/00_qualityChecks
multiqc .
```

```{r}
library(TidyMultiqc)
MultiQCpath <- file.path("/home/rachelvoyt/Documents/UT-Grad/Development/repos/tamarinGenetics_primatesPeru/seqAnalysis/01_run2_fecalHairBlood/00_qualityChecks/multiqc_data_1", "multiqc_data.json")
MultiQC <- TidyMultiqc::load_multiqc(MultiQCpath)
MultiQC$sampleNo <- substr(MultiQC$metadata.sample_id, start = 9, stop = 11)
View(MultiQC)

md <- read.csv('/home/rachelvoyt/Documents/UT-Grad/Development/repos/tamarinGenetics_primatesPeru/seqAnalysis/01_run2_fecalHairBlood/03_run2GTscore/metadata_tamRun2.csv') %>%
  mutate(sampleNo = substr(sampleID, start = 9, stop = 11))
head(md)

MultiQC_md <- merge(MultiQC, md, by = "sampleNo")
```

# FastQC

Note: use "\--interactive" to force plots to tell you which line belongs to which sample, otherwise it will remove this option since there are so many samples in this run.

```{bash eval = F}
conda activate fastqc-env
for i in ../01_run2Seqs/*; do fastqc $i; done
mv ../01_run2Seqs/*fastqc* .
conda activate multiqc-env
multiqc --interactive .
```

```{r}
MultiQCfastQCpath <- file.path("/home/rachelvoyt/Documents/UT-Grad/Development/repos/tamarinGenetics_primatesPeru/seqAnalysis/01_run2_fecalHairBlood/00_qualityChecksFastQC/multiqc_data", "multiqc_data.json")
MultiQCfastQC <- TidyMultiqc::load_multiqc(MultiQCfastQCpath, sections = c("general", "raw"))
```

```{r}
MultiQCfastQC %>% dplyr::select(dplyr::contains('quality'))
```

```{r}
TidyMultiqc::list_plots(MultiQCfastQCpath)
```

asdf
