---
title: "paper1_nanoporeValidation"
author: "Rachel Voyt"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1 Overview

This set of scripts is for tamGenetics_paper1, which focuses on validating the tamGenetics SNP panel on the Nanopore sequencing platform.

# 2 Packages

```{r, echo=FALSE}
library(readxl)
library(tidyverse)
```

# 3 Data

## 3.1 Metadata

```{r, echo=FALSE}
# Metadata w/both Illumina and Nanopore sampleIDs/sampleFiles
md_illONT_samplesAll <- read.csv("./04_tamRun4/illONT_tamRun4_metadata.csv")

# Metadata subsets by species
md_illONT_samplesLWED <- md_illONT_samplesAll %>%
  filter(species == "LWED")

md_illONT_samplesSIMP <- md_illONT_samplesAll %>%
  filter(species == "SIMP")

md_illONT_samplesNegs <- md_illONT_samplesAll %>%
  filter(str_detect(sampleType, "Neg"))

# Metadata subsets by sample type
md_illONT_samplesBlood <- md_illONT_samplesAll %>%
  filter(sampleType == "blood")

md_illONT_samplesFecal <- md_illONT_samplesAll %>%
  filter(sampleType == "fecal")

md_illONT_samplesHair <- md_illONT_samplesAll %>%
  filter(sampleType == "hair")
```

## 3.2 Primer list

We have 140 general loci (excluding those that are species specific) - these are the ones we'll look at to get sample/loci combos with 100% genotype success.

```{r, echo=FALSE}
primerList.v3 <- read.csv("./primers/03_lociChoices/tamGenetics_primerList_v3.csv")

primerList.v3_lociDual <- primerList.v3 %>%
  select(locus) %>%
  filter(!str_detect(locus, "LWED|SIMP"))

nrow(primerList.v3_lociDual)
```

## 3.3 Genotypes (dual loci only)

Import genotypes and subset to dual loci only 

### gtscore.10x

```{r, echo=FALSE}
# All samples
genos_ill_gtscore_samplesAll_lociDual <- read.table("./04_tamRun4/00_illumina/03_run4GTscore_ill/ill_fullSet_polyGenResults_singleSNP_10x.txt", header = T) %>%
  rownames_to_column("locus") %>%
  mutate(locus = sub('[_][^_]+$', '', locus)) %>%
  filter(locus %in% primerList.v3_lociDual$locus) %>%
  column_to_rownames("locus") %>%
  # recode 0 as NA
  mutate(across(everything(), ~ as.character(.))) %>%
  mutate(across(everything(), ~ 
                  case_when(. == "0" ~ gsub("0", NA, .),
                            .default = .)
                ))

genos_ont_gtscore_samplesAll_lociDual <- read.table("./04_tamRun4/01_nanopore/03_run4GTscore_ont/ont_fullSet_polyGenResults_singleSNP_10x.txt", header = T) %>%
  rownames_to_column("locus") %>%
  mutate(locus = sub('[_][^_]+$', '', locus)) %>%
  filter(locus %in% primerList.v3_lociDual$locus) %>%
  column_to_rownames("locus") %>%
  # recode 0 as NA
  mutate(across(everything(), ~ as.character(.))) %>%
  mutate(across(everything(), ~ 
                  case_when(. == "0" ~ gsub("0", NA, .),
                            .default = .)
                ))

# Blood samples
genos_ill_gtscore_samplesBlood_lociDual <- genos_ill_gtscore_samplesAll_lociDual %>%
  select(c(md_illONT_samplesBlood$sampleID_ill))

genos_ont_gtscore_samplesBlood_lociDual <- genos_ont_gtscore_samplesAll_lociDual %>%
  select(c(md_illONT_samplesBlood$sampleID_ont))

# Fecal samples
genos_ill_gtscore_samplesFecal_lociDual <- genos_ill_gtscore_samplesAll_lociDual %>%
  select(c(md_illONT_samplesFecal$sampleID_ill))

genos_ont_gtscore_samplesFecal_lociDual <- genos_ont_gtscore_samplesAll_lociDual %>%
  select(c(md_illONT_samplesFecal$sampleID_ont))
```

### gtseq

```{r}
# All samples
genos_ill_gtseq_samplesAll_lociDual <- 
  # need read_csv vs. read.csv here
  read_csv("./04_tamRun4/00_illumina/04_run4GTseq_ill/genos/ill_tamRun4_compiledGenos_gtseq.csv") %>%
  as.data.frame() %>%
  mutate(Sample = sub("-", "\\.", Sample)) %>%
  dplyr::rename_all(funs(make.names(.))) %>%
  # read_csv adds a comma to last column - remove this
  mutate(
    SPECIESID_9 = gsub(",", "", SPECIESID_9)
  ) %>%
  column_to_rownames("Sample") %>%
  select(c(primerList.v3_lociDual$locus)) %>%
  t() %>%
  as.data.frame() %>%
  mutate(across(everything(), ~ str_c(str_sub(., 1, 1), ",", str_sub(., 2, 2)))) %>%
  # recode 0,0 as NA
  mutate(across(everything(), ~ 
                  case_when(. == "0,0" ~ gsub("0,0", NA, .),
                            .default = .)
                ))

genos_ont_gtseq_samplesAll_lociDual <- 
  # need read_csv vs. read.csv here
  read_csv("./04_tamRun4/01_nanopore/04_run4GTseq_ont/genos/ont_tamRun4_compiledGenos_gtseq.csv") %>%
  as.data.frame() %>%
  mutate(Sample = str_sub(Sample, 1, 9)) %>%
  dplyr::rename_all(funs(make.names(.))) %>%
  # read_csv adds a comma to last column - remove this
  mutate(
    SPECIESID_9 = gsub(",", "", SPECIESID_9)
  ) %>%
  column_to_rownames("Sample") %>%
  select(c(primerList.v3_lociDual$locus)) %>%
  t() %>%
  as.data.frame() %>%
  mutate(across(everything(), ~ str_c(str_sub(., 1, 1), ",", str_sub(., 2, 2)))) %>%
  # recode 0,0 as NA
  mutate(across(everything(), ~ 
                  case_when(. == "0,0" ~ gsub("0,0", NA, .),
                            .default = .)
                ))

# Blood samples
genos_ill_gtseq_samplesBlood_lociDual <- genos_ill_gtseq_samplesAll_lociDual %>%
  select(c(md_illONT_samplesBlood$sampleID_ill))

genos_ont_gtseq_samplesBlood_lociDual <- genos_ont_gtseq_samplesAll_lociDual %>%
  select(c(md_illONT_samplesBlood$sampleID_ont))

# Fecal samples
genos_ill_gtseq_samplesFecal_lociDual <- genos_ill_gtseq_samplesAll_lociDual %>%
  select(c(md_illONT_samplesFecal$sampleID_ill))

genos_ont_gtseq_samplesFecal_lociDual <- genos_ont_gtseq_samplesAll_lociDual %>%
  select(c(md_illONT_samplesFecal$sampleID_ont))
```

# 4 Geno success

**NEEDS UPDATING**

## Blood samples

### gtscore.10x

```{r}
# Illumina
genoSuccess_bySample_ill_gtscore_samplesBlood_lociDual <- genos_ill_gtscore_samplesBlood_lociDual %>%
  t() %>%
  as.data.frame() %>%
  mutate(
    totalGenos = rowSums(!is.na(.))
  )

# Nanopore
genoSuccess_bySample_ont_gtscore_samplesBlood_lociDual <- genos_ont_gtscore_samplesBlood_lociDual %>%
  t() %>%
  as.data.frame() %>%
  mutate(
    totalGenos = rowSums(!is.na(.))
  )
```

Plots

```{r}
# Illumina
ggplot(genoSuccess_bySample_ill_gtscore_samplesBlood_lociDual, aes(y = totalGenos)) +
  geom_bar() +
  scale_y_continuous(breaks = seq(0, 140, 5)) + 
  ggtitle(label = "[ILLUMINA] 10x gtscore: Number of blood samples per total genotypes assigned") +
  theme_bw()

# Nanopore
ggplot(genoSuccess_bySample_ont_gtscore_samplesBlood_lociDual, aes(y = totalGenos)) +
  geom_bar() +
  scale_y_continuous(breaks = seq(0, 140, 5)) + 
  ggtitle(label = "[NANOPORE] 10x gtscore: Number of blood samples per total genotypes assigned") +
  theme_bw()
```

### gtseq

```{r}
# Illumina
genoSuccess_bySample_ill_gtseq_samplesBlood_lociDual <- genos_ill_gtseq_samplesBlood_lociDual %>%
  mutate(across(everything(), ~ as.character(.))) %>%
  t() %>%
  as.data.frame() %>%
  mutate(
    totalGenos = rowSums(!is.na(.))
  )

# Nanopore
genoSuccess_bySample_ont_gtseq_samplesBlood_lociDual <- genos_ont_gtseq_samplesBlood_lociDual %>%
  mutate(across(everything(), ~ as.character(.))) %>%
  t() %>%
  as.data.frame() %>%
  mutate(
    totalGenos = rowSums(!is.na(.))
  )
```

Plots

```{r}
# Illumina
ggplot(genoSuccess_bySample_ill_gtseq_samplesBlood_lociDual, aes(y = totalGenos)) +
  geom_bar() +
  scale_y_continuous(breaks = seq(0, 140, 5)) + 
  ggtitle(label = "[ILLUMINA] 10x gtseq: Number of blood samples per total genotypes assigned") +
  theme_bw()

# Nanopore
ggplot(genoSuccess_bySample_ont_gtseq_samplesBlood_lociDual, aes(y = totalGenos)) +
  geom_bar() +
  scale_y_continuous(breaks = seq(0, 140, 5)) + 
  ggtitle(label = "[NANOPORE] 10x gtseq: Number of blood samples per total genotypes assigned") +
  theme_bw()
```

### Summary

```{r}
table_countSamples_v_countGenoSuccess <- data.frame(
  lociGenotyped = as.numeric(seq_len(140))
) %>%
  arrange(desc(lociGenotyped)) %>%
  mutate(
    propLoci = round(lociGenotyped/140, 2)
         ) %>%
  rowwise() %>%
  mutate(
    # gtscore
    samples_ill.gtscore10x = sum(genoSuccess_bySample_ill_gtscore_samplesBlood_lociDual$totalGenos >= lociGenotyped),
    samples_ont.gtscore10x = sum(genoSuccess_bySample_ont_gtscore_samplesBlood_lociDual$totalGenos >= lociGenotyped),
    # gtseq
    samples_ill.gtseq = sum(genoSuccess_bySample_ill_gtseq_samplesBlood_lociDual$lociGenotyped >= lociGenotyped),
    samples_ont.gtseq = sum(genoSuccess_bySample_ont_gtscore_samplesBlood_lociDual$lociGenotyped >= lociGenotyped)
    )

table_countSamples_v_countGenoSuccess %>%
  kable(caption = "Number of loci vs. number of samples 100% genotyped") %>%
  kable_material(full_width = F) %>%
  scroll_box(height = "500px", fixed_thead = list(enabled = T, background = "#363636"))
```

Export

```{r}
write.csv(table_countSamples_v_countGenoSuccess, "./paper1_nanoporeValidation/countSamples_vs_genLociGenos.csv", row.names = F)
```

## By locus

### gtscore.10x

```{r}
# Illumina
genoSuccess_byLocus_ill_gtscore_samplesBlood_lociDual <- genos_ill_gtscore_samplesBlood_lociDual %>%
  mutate(
    samplesGenotyped = rowSums(!is.na(.))
  )

# Nanopore
genoSuccess_byLocus_ont_gtscore_samplesBlood_lociDual <- genos_ont_gtscore_samplesBlood_lociDual %>%
  mutate(
    samplesGenotyped = rowSums(!is.na(.))
  )
```

Plots

```{r}
# Illumina
ggplot(genoSuccess_byLocus_ill_gtscore_samplesBlood_lociDual, aes(y = samplesGenotyped)) +
  geom_bar() +
  scale_y_continuous(breaks = seq(0, 30, 5)) + 
  ggtitle(label = "[ILLUMINA] 10x gtscore: Number of blood samples genotyped per locus") +
  labs(y = "Samples genotyped", x = "# of loci") +
  theme_bw()

# Nanopore
ggplot(genoSuccess_byLocus_ont_gtscore_samplesBlood_lociDual, aes(y = samplesGenotyped)) +
  geom_bar() +
  scale_y_continuous(breaks = seq(0, 30, 5)) + 
  ggtitle(label = "[NANOPORE] 10x gtscore: Number of blood samples genotyped per locus") +
   labs(y = "Samples genotyped", x = "# of loci") +
  theme_bw()
```

### gtseq

```{r}
# Illumina
genoSuccess_byLocus_ill_gtseq_samplesBlood_lociDual <- genos_ill_gtseq_samplesBlood_lociDual %>%
  mutate(
    samplesGenotyped = rowSums(!is.na(.))
  )

# Nanopore
genoSuccess_byLocus_ont_gtseq_samplesBlood_lociDual <- genos_ont_gtseq_samplesBlood_lociDual %>%
  mutate(
    samplesGenotyped = rowSums(!is.na(.))
  )
```

Plots

```{r}
# Illumina
ggplot(genoSuccess_byLocus_ill_gtseq_samplesBlood_lociDual, aes(y = samplesGenotyped)) +
  geom_bar() +
  scale_y_continuous(breaks = seq(0, 30, 5)) + 
  ggtitle(label = "[ILLUMINA] gtseq: Number of blood samples genotyped per locus") +
  labs(y = "Samples genotyped", x = "# of loci") +
  theme_bw()

# Nanopore
ggplot(genoSuccess_byLocus_ont_gtseq_samplesBlood_lociDual, aes(y = samplesGenotyped)) +
  geom_bar() +
  scale_y_continuous(breaks = seq(0, 30, 5)) + 
  ggtitle(label = "[NANOPORE] gtseq: Number of blood samples genotyped per locus") +
  labs(y = "Samples genotyped", x = "# of loci") +
  theme_bw()
```

### Summary

Based on the table below, I would recommend reporting results based on the 20 blood samples that have ~100 loci genotyped... BUT are these the same loci??

```{r}
table_countLoci_v_countGenoSuccess <- data.frame(
  samplesGenotyped = as.numeric(seq_len(30))
) %>%
  arrange(desc(samplesGenotyped)) %>%
  mutate(
    propSamples = round(samplesGenotyped/30, 2)
         ) %>%
  rowwise() %>%
  mutate(
    # gtscore
    loci_ill.gtscore10x = sum(genoSuccess_byLocus_ill_gtscore_samplesBlood_lociDual$samplesGenotyped >= samplesGenotyped),
    loci_ont.gtscore10x = sum(genoSuccess_byLocus_ont_gtscore_samplesBlood_lociDual$samplesGenotyped >= samplesGenotyped),
    # gtseq
    loci_ill.gtseq = sum(genoSuccess_byLocus_ill_gtseq_samplesBlood_lociDual$samplesGenotyped >= samplesGenotyped),
    loci_ont.gtseq = sum(genoSuccess_byLocus_ont_gtscore_samplesBlood_lociDual$samplesGenotyped >= samplesGenotyped)
    )

write.csv(table_countSamples_v_countGenoSuccess, "./paper1_nanoporeValidation/countSamples_vs_genLociGenos.csv", row.names = F)
```

# 5 Identify optimal sample/loci combos

## 5.1 Functions

Function 1: For a given dataframe of genotypes, this function returns the number of loci genotyped at the desired proportion of n samples ("propSamples") for subsets of 2 to 30 samples

```{r}
# number of loci genotyped at propSamples (0 to 1.0) of samples from 10 to 30 samples
optimize_samplesLoci <- function(df, propSamples) {
  # Initialize an empty list to store results
  result_list <- list()

  for (n in 2:30) {
    # create matrix where NAs are 0 and non-NAs are 1
    is_na_vector <- ifelse(is.na(df), 0, 1)

    crossprod_vector <- crossprod(is_na_vector)
    col_sums <- colSums(crossprod_vector)

    x_df <- as.data.frame(df)  # to get meaningful colnames
    # use colSums vector to select n columns; will rank all columns and give the n first; gives n columns w/the max number of non-NA rows
    res <- x_df[, rank(-col_sums, ties.method = "first") <= n]

    # Store the result in the list
    result_list[[as.character(n)]] <- c(n, nrow(res[which(rowMeans(!is.na(res)) >= propSamples), ]))
  }

  # Convert the list to a dataframe
  result_df <- do.call(rbind, result_list)

  # Rename columns
  colnames(result_df) <- c("samples", "loci")

  return(result_df)
}
```

Function 2: For a given dataframe of genotypes, this function returns an optimized set of genotypes for n samples, where the number of samples (n) and the proportion of those n samples successfully genotyped per locus (propSamples) is chosen based on the results of function 1 above

```{r}
optimized_genos <- function(df, n, propSamples) {
  is_na_vector <- ifelse(is.na(df), 0, 1)
  crossprod_vector <- crossprod(is_na_vector)
  col_sums <- colSums(crossprod_vector)

  x_df <- as.data.frame(df)
  res <- x_df[, rank(-col_sums, ties.method = "first") <= n]

  result <- res[which(rowMeans(!is.na(res)) >= propSamples), ]
  return(result)
}
```

## 5.2 Blood samples

### Identify optimal sample/loci combos

#### gtscore

```{r}
# loci genotyped at 100% of samples among subsets of 10 to 30 samples
opt100_samplesLoci_ill_gtscore_blood <- optimize_samplesLoci(genos_ill_gtscore_samplesBlood_lociDual, 1) %>%
  as.data.frame() %>%
  dplyr::rename("loci100_ill.gtscore" = "loci")

opt100_samplesLoci_ont_gtscore_blood <- optimize_samplesLoci(genos_ont_gtscore_samplesBlood_lociDual, 1) %>%
  as.data.frame() %>%
  dplyr::rename("loci100_ont.gtscore" = "loci")

# loci genotyped at 90% of samples among subsets of 10 to 30 samples
opt90_samplesLoci_ill_gtscore_blood <- optimize_samplesLoci(genos_ill_gtscore_samplesBlood_lociDual, 0.9) %>%
  as.data.frame() %>%
  dplyr::rename("loci90_ill.gtscore" = "loci")

opt90_samplesLoci_ont_gtscore_blood <- optimize_samplesLoci(genos_ont_gtscore_samplesBlood_lociDual, 0.9) %>%
  as.data.frame() %>%
  dplyr::rename("loci90_ont.gtscore" = "loci")
```

#### gtseq

```{r}
# loci genotyped at 90% of samples among subsets of 10 to 30 samples
opt100_samplesLoci_ill_gtseq_blood <- optimize_samplesLoci(genos_ill_gtseq_samplesBlood_lociDual, 1) %>%
  as.data.frame() %>%
  dplyr::rename("loci100_ill.gtseq" = "loci")

opt100_samplesLoci_ont_gtseq_blood <- optimize_samplesLoci(genos_ont_gtseq_samplesBlood_lociDual, 1) %>%
  as.data.frame() %>%
  dplyr::rename("loci100_ont.gtseq" = "loci")

# loci genotyped at 90% of samples among subsets of 10 to 30 samples
opt90_samplesLoci_ill_gtseq_blood <- optimize_samplesLoci(genos_ill_gtseq_samplesBlood_lociDual, 0.9) %>%
  as.data.frame() %>%
  dplyr::rename("loci90_ill.gtseq" = "loci")

opt90_samplesLoci_ont_gtseq_blood <- optimize_samplesLoci(genos_ont_gtseq_samplesBlood_lociDual, 0.9) %>%
  as.data.frame() %>%
  dplyr::rename("loci90_ont.gtseq" = "loci")
```

#### Summaries

```{r}
# loci genotyped at 100% of samples among subsets of 10 to 30 samples
opt100_samplesLoci_blood_summary <- merge(opt100_samplesLoci_ill_gtscore_blood, opt100_samplesLoci_ont_gtscore_blood, by = "samples") %>%
  merge(., opt100_samplesLoci_ill_gtseq_blood, by = "samples") %>%
  merge(., opt100_samplesLoci_ont_gtseq_blood, by = "samples")

# loci genotyped at 90% of samples among subsets of 10 to 30 samples
opt90_samplesLoci_blood_summary <- merge(opt90_samplesLoci_ill_gtscore_blood, opt90_samplesLoci_ont_gtscore_blood, by = "samples") %>%
  merge(., opt90_samplesLoci_ill_gtseq_blood, by = "samples") %>%
  merge(., opt90_samplesLoci_ont_gtseq_blood, by = "samples")
```

### Dataframes

#### 100% geno success

##### gtscore

```{r}
optGenos100_ill_gtscore_blood <- optimized_genos(genos_ill_gtscore_samplesBlood_lociDual, 20, 1.0)

optGenos100_ont_gtscore_blood <- optimized_genos(genos_ont_gtscore_samplesBlood_lociDual, 20, 1.0)
```

##### gtseq

```{r}
optGenos100_ill_gtseq_blood <- optimized_genos(genos_ill_gtseq_samplesBlood_lociDual, 10, 1.0)

optGenos100_ont_gtseq_blood <- optimized_genos(genos_ont_gtseq_samplesBlood_lociDual, 10, 1.0)
```

#### 90% geno success

##### gtscore

```{r}
optGenos90_ill_gtscore_blood <- optimized_genos(genos_ill_gtscore_samplesBlood_lociDual, 20, 0.9)

optGenos90_ont_gtscore_blood <- optimized_genos(genos_ont_gtscore_samplesBlood_lociDual, 20, 0.9)
```

##### gtseq

```{r}
optGenos90_ill_gtseq_blood <- optimized_genos(genos_ill_gtseq_samplesBlood_lociDual, 10, 0.9)

optGenos90_ont_gtseq_blood <- optimized_genos(genos_ont_gtseq_samplesBlood_lociDual, 10, 0.9)
```

## 5.3 Fecal samples

### Identify optimal sample/loci combos

#### gtscore

```{r}
# loci genotyped at 100% of samples among subsets of 10 to 30 samples
opt100_samplesLoci_ill_gtscore_fecal <- optimize_samplesLoci(genos_ill_gtscore_samplesFecal_lociDual, 1) %>%
  as.data.frame() %>%
  dplyr::rename("loci100_ill.gtscore" = "loci")

opt100_samplesLoci_ont_gtscore_fecal <- optimize_samplesLoci(genos_ont_gtscore_samplesFecal_lociDual, 1) %>%
  as.data.frame() %>%
  dplyr::rename("loci100_ont.gtscore" = "loci")

# loci genotyped at 90% of samples among subsets of 10 to 30 samples
opt90_samplesLoci_ill_gtscore_fecal <- optimize_samplesLoci(genos_ill_gtscore_samplesFecal_lociDual, 0.9) %>%
  as.data.frame() %>%
  dplyr::rename("loci90_ill.gtscore" = "loci")

opt90_samplesLoci_ont_gtscore_fecal <- optimize_samplesLoci(genos_ont_gtscore_samplesFecal_lociDual, 0.9) %>%
  as.data.frame() %>%
  dplyr::rename("loci90_ont.gtscore" = "loci")
```

#### gtseq

```{r}
# loci genotyped at 90% of samples among subsets of 10 to 30 samples
opt100_samplesLoci_ill_gtseq_fecal <- optimize_samplesLoci(genos_ill_gtseq_samplesFecal_lociDual, 1) %>%
  as.data.frame() %>%
  dplyr::rename("loci100_ill.gtseq" = "loci")

opt100_samplesLoci_ont_gtseq_fecal <- optimize_samplesLoci(genos_ont_gtseq_samplesFecal_lociDual, 1) %>%
  as.data.frame() %>%
  dplyr::rename("loci100_ont.gtseq" = "loci")

# loci genotyped at 90% of samples among subsets of 10 to 30 samples
opt90_samplesLoci_ill_gtseq_fecal <- optimize_samplesLoci(genos_ill_gtseq_samplesFecal_lociDual, 0.9) %>%
  as.data.frame() %>%
  dplyr::rename("loci90_ill.gtseq" = "loci")

opt90_samplesLoci_ont_gtseq_fecal <- optimize_samplesLoci(genos_ont_gtseq_samplesFecal_lociDual, 0.9) %>%
  as.data.frame() %>%
  dplyr::rename("loci90_ont.gtseq" = "loci")
```

#### Summaries

```{r}
# loci genotyped at 100% of samples among subsets of 10 to 30 samples
opt100_samplesLoci_fecal_summary <- merge(opt100_samplesLoci_ill_gtscore_fecal, opt100_samplesLoci_ont_gtscore_fecal, by = "samples") %>%
  merge(., opt100_samplesLoci_ill_gtseq_fecal, by = "samples") %>%
  merge(., opt100_samplesLoci_ont_gtseq_fecal, by = "samples")

# loci genotyped at 90% of samples among subsets of 10 to 30 samples
opt90_samplesLoci_fecal_summary <- merge(opt90_samplesLoci_ill_gtscore_fecal, opt90_samplesLoci_ont_gtscore_fecal, by = "samples") %>%
  merge(., opt90_samplesLoci_ill_gtseq_fecal, by = "samples") %>%
  merge(., opt90_samplesLoci_ont_gtseq_fecal, by = "samples")
```

### Dataframes

#### 100% geno success

##### gtscore

```{r}
optGenos100_ill_gtscore_fecal <- optimized_genos(genos_ill_gtscore_samplesFecal_lociDual, 20, 1.0)

optGenos100_ont_gtscore_fecal <- optimized_genos(genos_ont_gtscore_samplesFecal_lociDual, 20, 1.0)
```

##### gtseq

```{r}
optGenos100_ill_gtseq_fecal <- optimized_genos(genos_ill_gtseq_samplesFecal_lociDual, 10, 1.0)

optGenos100_ont_gtseq_fecal <- optimized_genos(genos_ont_gtseq_samplesFecal_lociDual, 10, 1.0)
```

#### 90% geno success

##### gtscore

```{r}
optGenos90_ill_gtscore_fecal <- optimized_genos(genos_ill_gtscore_samplesFecal_lociDual, 20, 0.9)

optGenos90_ont_gtscore_fecal <- optimized_genos(genos_ont_gtscore_samplesFecal_lociDual, 20, 0.9)
```

##### gtseq

```{r}
optGenos90_ill_gtseq_fecal <- optimized_genos(genos_ill_gtseq_samplesFecal_lociDual, 10, 0.9)

optGenos90_ont_gtseq_fecal <- optimized_genos(genos_ont_gtseq_samplesFecal_lociDual, 10, 0.9)
```

# 6 Geno consistency

Using the desired optimized set of genotypes, we can compare the consistency between Illumina and Nanopore genotype results

## 6.1 Blood samples

### 100% geno success

#### gtscore

We'll start by looking at the 20 blood samples + 122 loci w/100% genotype success using Illumina sequences + gtscore pipeline

```{r}
# Convert genotypes to long-format
optGenos100_ill_gtscore_blood_l <- optGenos100_ill_gtscore_blood %>%
  t() %>%
  as.data.frame() %>%
  rownames_to_column("sampleID") %>%
  pivot_longer(!sampleID, names_to = "locus", values_to = "ill_geno") %>%
  merge(., md_illONT_samplesAll[, c("sampleID_unique", "sampleID_ill")], by.x = "sampleID", by.y = "sampleID_ill") %>%
  relocate(sampleID_unique) %>%
  select(-sampleID)

genos_ont_gtscore_samplesBlood_lociDual_l <- genos_ont_gtscore_samplesBlood_lociDual %>%
  t() %>%
  as.data.frame() %>%
  rownames_to_column("sampleID") %>%
  pivot_longer(!sampleID, names_to = "locus", values_to = "ont_geno") %>%
  merge(., md_illONT_samplesAll[, c("sampleID_unique", "sampleID_ont")], by.x = "sampleID", by.y = "sampleID_ont") %>%
  relocate(sampleID_unique) %>%
  select(-sampleID)

# compare ill vs. ont genos
genoCompare100_illONT_optGenos_gtscore_blood <- merge(optGenos100_ill_gtscore_blood_l, genos_ont_gtscore_samplesBlood_lociDual_l, by = c("sampleID_unique", "locus")) %>%
  mutate(
    match = case_when(
      is.na(ont_geno) ~ "NA",
      ill_geno == ont_geno ~ "yes",
      ill_geno != ont_geno ~ "no"
      )
    )
```

#### gtseq

For gtseq, we'll use the Illumina blood genotypes w/10 samples + 56 loci w/100% genotypes

```{r}
# Convert genotypes to long-format
optGenos100_ill_gtseq_blood_l <- optGenos100_ill_gtseq_blood %>%
  t() %>%
  as.data.frame() %>%
  rownames_to_column("sampleID") %>%
  pivot_longer(!sampleID, names_to = "locus", values_to = "ill_geno") %>%
  merge(., md_illONT_samplesAll[, c("sampleID_unique", "sampleID_ill")], by.x = "sampleID", by.y = "sampleID_ill") %>%
  relocate(sampleID_unique) %>%
  select(-sampleID)

genos_ont_gtseq_samplesBlood_lociDual_l <- genos_ont_gtseq_samplesBlood_lociDual %>%
  t() %>%
  as.data.frame() %>%
  rownames_to_column("sampleID") %>%
  pivot_longer(!sampleID, names_to = "locus", values_to = "ont_geno") %>%
  merge(., md_illONT_samplesAll[, c("sampleID_unique", "sampleID_ont")], by.x = "sampleID", by.y = "sampleID_ont") %>%
  relocate(sampleID_unique) %>%
  select(-sampleID)

# compare ill vs. ont genos
genoCompare100_illONT_optGenos_gtseq_blood <- merge(optGenos100_ill_gtseq_blood_l, genos_ont_gtseq_samplesBlood_lociDual_l, by = c("sampleID_unique", "locus")) %>%
  mutate(
    match = case_when(
      is.na(ont_geno) ~ "NA",
      ill_geno == ont_geno ~ "yes",
      ill_geno != ont_geno ~ "no"
      )
    )
```

### 90% geno success

#### gtscore

Using 20 samples + 131 loci w/genotypes for at least 90% of samples in Illumina results as a base

```{r}
# Convert genotypes to long-format
optGenos90_ill_gtscore_blood_l <- optGenos90_ill_gtscore_blood %>%
  t() %>%
  as.data.frame() %>%
  rownames_to_column("sampleID") %>%
  pivot_longer(!sampleID, names_to = "locus", values_to = "ill_geno") %>%
  merge(., md_illONT_samplesAll[, c("sampleID_unique", "sampleID_ill")], by.x = "sampleID", by.y = "sampleID_ill") %>%
  relocate(sampleID_unique) %>%
  select(-sampleID)

genos_ont_gtscore_samplesBlood_lociDual_l <- genos_ont_gtscore_samplesBlood_lociDual %>%
  t() %>%
  as.data.frame() %>%
  rownames_to_column("sampleID") %>%
  pivot_longer(!sampleID, names_to = "locus", values_to = "ont_geno") %>%
  merge(., md_illONT_samplesAll[, c("sampleID_unique", "sampleID_ont")], by.x = "sampleID", by.y = "sampleID_ont") %>%
  relocate(sampleID_unique) %>%
  select(-sampleID)

# compare ill vs. ont genos
genoCompare90_illONT_optGenos_gtscore_blood <- merge(optGenos90_ill_gtscore_blood_l, genos_ont_gtscore_samplesBlood_lociDual_l, by = c("sampleID_unique", "locus")) %>%
  mutate(
    match = case_when(
      is.na(ill_geno) ~ "NA",
      is.na(ont_geno) ~ "NA",
      ill_geno == ont_geno ~ "yes",
      ill_geno != ont_geno ~ "no"
      )
    )
```

#### gtseq

Using 10 samples + 94 loci w/genotypes for at least 90% of samples in Illumina results as a base

```{r}
# Convert genotypes to long-format
optGenos90_ill_gtseq_blood_l <- optGenos90_ill_gtseq_blood %>%
  t() %>%
  as.data.frame() %>%
  rownames_to_column("sampleID") %>%
  pivot_longer(!sampleID, names_to = "locus", values_to = "ill_geno") %>%
  merge(., md_illONT_samplesAll[, c("sampleID_unique", "sampleID_ill")], by.x = "sampleID", by.y = "sampleID_ill") %>%
  relocate(sampleID_unique) %>%
  select(-sampleID)

genos_ont_gtseq_samplesBlood_lociDual_l <- genos_ont_gtseq_samplesBlood_lociDual %>%
  t() %>%
  as.data.frame() %>%
  rownames_to_column("sampleID") %>%
  pivot_longer(!sampleID, names_to = "locus", values_to = "ont_geno") %>%
  merge(., md_illONT_samplesAll[, c("sampleID_unique", "sampleID_ont")], by.x = "sampleID", by.y = "sampleID_ont") %>%
  relocate(sampleID_unique) %>%
  select(-sampleID)

# compare ill vs. ont genos
genoCompare90_illONT_optGenos_gtseq_blood <- merge(optGenos90_ill_gtseq_blood_l, genos_ont_gtseq_samplesBlood_lociDual_l, by = c("sampleID_unique", "locus")) %>%
  mutate(
    match = case_when(
      is.na(ont_geno) ~ "NA",
      ill_geno == ont_geno ~ "yes",
      ill_geno != ont_geno ~ "no"
      )
    )
```

### Summary

```{r}
genoCompare_illONT_blood_summary <- data.frame(
  match = c("illONT_NA", "illONT_mismatch", "illONT_match"),
  gtscore100 = table(genoCompare100_illONT_optGenos_gtscore_blood$match),
  gtscore90 = table(genoCompare90_illONT_optGenos_gtscore_blood$match),
  gtseq100 = table(genoCompare100_illONT_optGenos_gtseq_blood$match),
  gtseq90 = table(genoCompare90_illONT_optGenos_gtseq_blood$match)
) %>%
  select(-ends_with("Var1")) %>%
  column_to_rownames("match") %>%
  t() %>%
  as.data.frame() %>%
  mutate(
    samples = c("20", "20", "10", "10"),
    loci = c("122", "131", "56", "94")
  ) %>%
  relocate(c(samples, loci)) %>%
  mutate(
    propNA = round(illONT_NA/(illONT_NA + illONT_match + illONT_mismatch), 2),
    propMatch = round(illONT_match/(illONT_match + illONT_mismatch), 2),
    propMismatch = round(illONT_mismatch/(illONT_match + illONT_mismatch), 2)
  ) %>%
  rowwise() %>%
  mutate(
    genosTotal = illONT_NA + illONT_match + illONT_mismatch,
    genosCommon = illONT_match + illONT_mismatch
  ) %>%
  select(c(samples, loci, genosTotal, genosCommon, illONT_NA, propNA, illONT_mismatch, propMismatch, illONT_match, propMatch))
```

## 6.2 Fecal: samples vs. loci

### 100% geno success

#### gtscore

We'll start by looking at the 20 blood samples + 122 loci w/100% genotype success using Illumina sequences + gtscore pipeline

```{r}
optGenos100_ill_gtscore_fecal_l <- optGenos100_ill_gtscore_fecal %>%
  t() %>%
  as.data.frame() %>%
  rownames_to_column("sampleID") %>%
  pivot_longer(!sampleID, names_to = "locus", values_to = "ill_geno") %>%
  merge(., md_illONT_samplesAll[, c("sampleID_unique", "sampleID_ill")], by.x = "sampleID", by.y = "sampleID_ill") %>%
  relocate(sampleID_unique) %>%
  select(-sampleID)

genos_ont_gtscore_samplesFecal_lociDual_l <- genos_ont_gtscore_samplesFecal_lociDual %>%
  t() %>%
  as.data.frame() %>%
  rownames_to_column("sampleID") %>%
  pivot_longer(!sampleID, names_to = "locus", values_to = "ont_geno") %>%
  merge(., md_illONT_samplesAll[, c("sampleID_unique", "sampleID_ont")], by.x = "sampleID", by.y = "sampleID_ont") %>%
  relocate(sampleID_unique) %>%
  select(-sampleID)

# compare ill vs. ont genos
genoCompare100_illONT_optGenos_gtscore_fecal <- merge(optGenos100_ill_gtscore_fecal_l, genos_ont_gtscore_samplesFecal_lociDual_l, by = c("sampleID_unique", "locus")) %>%
  mutate(
    match = case_when(
      is.na(ont_geno) ~ "NA",
      ill_geno == ont_geno ~ "yes",
      ill_geno != ont_geno ~ "no"
      )
    )

table(genoCompare100_illONT_optGenos_gtscore_fecal$match)
```

#### gtseq

For gtseq, we'll use the Illumina blood genotypes w/10 samples + 56 loci w/100% genotypes

```{r}
optGenos100_ill_gtseq_fecal_l <- optGenos100_ill_gtseq_fecal %>%
  t() %>%
  as.data.frame() %>%
  rownames_to_column("sampleID") %>%
  pivot_longer(!sampleID, names_to = "locus", values_to = "ill_geno") %>%
  merge(., md_illONT_samplesAll[, c("sampleID_unique", "sampleID_ill")], by.x = "sampleID", by.y = "sampleID_ill") %>%
  relocate(sampleID_unique) %>%
  select(-sampleID)

genos_ont_gtseq_samplesFecal_lociDual_l <- genos_ont_gtseq_samplesFecal_lociDual %>%
  t() %>%
  as.data.frame() %>%
  rownames_to_column("sampleID") %>%
  pivot_longer(!sampleID, names_to = "locus", values_to = "ont_geno") %>%
  merge(., md_illONT_samplesAll[, c("sampleID_unique", "sampleID_ont")], by.x = "sampleID", by.y = "sampleID_ont") %>%
  relocate(sampleID_unique) %>%
  select(-sampleID)

# compare ill vs. ont genos
genoCompare100_illONT_optGenos_gtseq_fecal <- merge(optGenos100_ill_gtseq_fecal_l, genos_ont_gtseq_samplesFecal_lociDual_l, by = c("sampleID_unique", "locus")) %>%
  mutate(
    match = case_when(
      is.na(ont_geno) ~ "NA",
      ill_geno == ont_geno ~ "yes",
      ill_geno != ont_geno ~ "no"
      )
    )

table(genoCompare100_illONT_optGenos_gtseq_fecal$match)
```

### 90% geno success

#### gtscore

Using 20 samples + 131 loci w/genotypes for at least 90% of samples in Illumina results as a base

```{r}
optGenos90_ill_gtscore_fecal_l <- optGenos90_ill_gtscore_fecal %>%
  t() %>%
  as.data.frame() %>%
  rownames_to_column("sampleID") %>%
  pivot_longer(!sampleID, names_to = "locus", values_to = "ill_geno") %>%
  merge(., md_illONT_samplesAll[, c("sampleID_unique", "sampleID_ill")], by.x = "sampleID", by.y = "sampleID_ill") %>%
  relocate(sampleID_unique) %>%
  select(-sampleID)

genos_ont_gtscore_samplesFecal_lociDual_l <- genos_ont_gtscore_samplesFecal_lociDual %>%
  t() %>%
  as.data.frame() %>%
  rownames_to_column("sampleID") %>%
  pivot_longer(!sampleID, names_to = "locus", values_to = "ont_geno") %>%
  merge(., md_illONT_samplesAll[, c("sampleID_unique", "sampleID_ont")], by.x = "sampleID", by.y = "sampleID_ont") %>%
  relocate(sampleID_unique) %>%
  select(-sampleID)

# compare ill vs. ont genos
genoCompare90_illONT_optGenos_gtscore_fecal <- merge(optGenos90_ill_gtscore_fecal_l, genos_ont_gtscore_samplesFecal_lociDual_l, by = c("sampleID_unique", "locus")) %>%
  mutate(
    match = case_when(
      is.na(ill_geno) ~ "NA",
      is.na(ont_geno) ~ "NA",
      ill_geno == ont_geno ~ "yes",
      ill_geno != ont_geno ~ "no"
      )
    )

table(genoCompare90_illONT_optGenos_gtscore_fecal$match)
```

#### gtseq

Using 10 samples + 94 loci w/genotypes for at least 90% of samples in Illumina results as a base

```{r}
optGenos90_ill_gtseq_fecal_l <- optGenos90_ill_gtseq_fecal %>%
  t() %>%
  as.data.frame() %>%
  rownames_to_column("sampleID") %>%
  pivot_longer(!sampleID, names_to = "locus", values_to = "ill_geno") %>%
  merge(., md_illONT_samplesAll[, c("sampleID_unique", "sampleID_ill")], by.x = "sampleID", by.y = "sampleID_ill") %>%
  relocate(sampleID_unique) %>%
  select(-sampleID)

genos_ont_gtseq_samplesFecal_lociDual_l <- genos_ont_gtseq_samplesFecal_lociDual %>%
  t() %>%
  as.data.frame() %>%
  rownames_to_column("sampleID") %>%
  pivot_longer(!sampleID, names_to = "locus", values_to = "ont_geno") %>%
  merge(., md_illONT_samplesAll[, c("sampleID_unique", "sampleID_ont")], by.x = "sampleID", by.y = "sampleID_ont") %>%
  relocate(sampleID_unique) %>%
  select(-sampleID)

# compare ill vs. ont genos
genoCompare90_illONT_optGenos_gtseq_fecal <- merge(optGenos90_ill_gtseq_fecal_l, genos_ont_gtseq_samplesFecal_lociDual_l, by = c("sampleID_unique", "locus")) %>%
  mutate(
    match = case_when(
      is.na(ont_geno) ~ "NA",
      ill_geno == ont_geno ~ "yes",
      ill_geno != ont_geno ~ "no"
      )
    )

table(genoCompare90_illONT_optGenos_gtseq_fecal$match)
```

### Summary

```{r}
genoCompare_illONT_fecal_summary <- data.frame(
  match = c("NA", "illONT_mismatch", "illONT_match"),
  gtscore100 = table(genoCompare100_illONT_optGenos_gtscore_fecal$match),
  gtscore90 = table(genoCompare90_illONT_optGenos_gtscore_fecal$match),
  gtseq100 = table(genoCompare100_illONT_optGenos_gtseq_fecal$match),
  gtseq90 = table(genoCompare90_illONT_optGenos_gtseq_fecal$match)
) %>%
  select(-ends_with("Var1")) %>%
  column_to_rownames("match") %>%
  t() %>%
  as.data.frame() %>%
  mutate(
    samples = c("20", "10", "20", "10"),
    loci = c("122", "56", "131", "94")
  ) %>%
  relocate(c(samples, loci)) %>%
  mutate(
    propMismatch = round(illONT_mismatch/(illONT_match + illONT_mismatch), 2)
  )
```

### Dataframes

#### gtscore

```{r}
optGenos100_ill_gtscore_fecal <- optimized_genos(genos_ill_gtscore_samplesFecal_lociDual, 20)
```

#### gtseq

```{r}

```

genoRate100 metadata, including whether samples have whole genome sequenced

```{r}
wholeGenomeSeqs <- read_excel("./04_tamRun4/DNASubmissionForm_MEW.xlsx",
                              range = "A17:R44") %>%
  as.data.frame() %>%
  dplyr::rename("animalID" = "Animal ID") %>%
  select(animalID)

genoRate100_blood_md_10x <- genoRate100_blood_10x %>%
  t() %>%
  as.data.frame() %>%
  rownames_to_column("sampleID") %>%
  select(sampleID) %>%
  merge(., md_tamRun4, by = "sampleID") %>%
  mutate(
    wholeGenome = 
      case_when(
        animalID %in% wholeGenomeSeqs$animalID ~ "yes",
        .default = "no"
      )
  ) %>%
  relocate(wholeGenome, .after = sampleID)
```

# RESOURCES

for optimizing samples/loci to include: 
https://stackoverflow.com/questions/24016612/subset-of-data-frame-columns-to-maximize-complete-observations

alternative approach:
https://stackoverflow.com/questions/42858959/best-possible-subset-of-data-frame-without-na

# SCRAPS

works well; updated above to return df instead of print
```{r}
for (n in 20:30) {
  is_na_vector <- ifelse(is.na(df), 0, 1)

  crossprod_vector <- crossprod(is_na_vector)
  col_sums <- colSums(crossprod_vector)

  x_df <- as.data.frame(df)  # to get meaningful colnames
  res <- x_df[, rank(-col_sums, ties.method = "first") <= n]

  # Print the value of n and the number of rows in the na.omit result for each iteration
  cat("For sample subset =", n, "loci w/complete genos =", nrow(na.omit(res)), "\n")
}
```


## Blood samples

### 10x coverage

Filtering to 10x coverage gives 27 blood samples + 71 loci with 100% genotype success + 3 blood samples with whole-genome seqs

```{r}
passSamples_blood_10x <- sampleSum_blood_10x %>%
  filter(GenotypeRate >= 0.5)

genoRate100_blood_10x <- genos_10x %>%
  replace(., .==0, NA) %>%
  t() %>%
  as.data.frame() %>%
  rownames_to_column("sampleID") %>%
  # filter to blood samples w/ >50% geno success
  filter(sampleID %in% passSamples_blood_10x$sampleID) %>%
  column_to_rownames("sampleID") %>%
  t() %>%
  as.data.frame() %>%
  # remove species-specific loci
  rownames_to_column("locus") %>%
  filter(!str_detect(locus, "LWED|SIMP")) %>%
  column_to_rownames("locus") %>%
  # filter to only samples/loci w/100% genotype success
  na.omit()
```

### 30x coverage

If we want 100% genotype success for each locus, filtering to 30x coverage gives 27 blood samples + 71 loci with 100% genotype success + 3 blood samples with whole-genome seqs (same as 10x)

```{r}
passSamples_blood_30x <- sampleSum_blood_30x %>%
  filter(GenotypeRate >= 0.5)

genoRate100_blood_30x <- genos_30x %>%
  replace(., .==0, NA) %>%
  t() %>%
  as.data.frame() %>%
  rownames_to_column("sampleID") %>%
  # filter to blood samples w/ >50% geno success
  filter(sampleID %in% passSamples_blood_30x$sampleID) %>%
  column_to_rownames("sampleID") %>%
  t() %>%
  as.data.frame() %>%
  # remove species-specific loci
  rownames_to_column("locus") %>%
  filter(!str_detect(locus, "LWED|SIMP")) %>%
  column_to_rownames("locus") %>%
  # filter to SAMPLES with no missing genos
  select_if(~mean(is.na(.)) <= 0)

genoRate90_blood_30x <- genos_30x %>%
  replace(., .==0, NA) %>%
  t() %>%
  as.data.frame() %>%
  rownames_to_column("sampleID") %>%
  # filter to blood samples w/ >50% geno success
  filter(sampleID %in% passSamples_blood_30x$sampleID) %>%
  column_to_rownames("sampleID") %>%
  # filter to LOCI with <= 10% missing genos
  select_if(~mean(is.na(.)) <= 0.1) %>%
  t() %>%
  as.data.frame() %>%
  # remove species-specific loci
  rownames_to_column("locus") %>%
  filter(!str_detect(locus, "LWED|SIMP")) %>%
  column_to_rownames("locus") %>%
  # filter to SAMPLES with <= 10% missing genos
  select_if(~mean(is.na(.)) <= 0.1)
```

genoRate100 metadata, including whether samples have whole genome sequenced

```{r}
wholeGenomeSeqs <- read_excel("./04_tamRun4/DNASubmissionForm_MEW.xlsx",
                              range = "A17:R44") %>%
  as.data.frame() %>%
  dplyr::rename("animalID" = "Animal ID") %>%
  select(animalID)

genoRate100_blood_md_30x <- genoRate100_blood_30x %>%
  t() %>%
  as.data.frame() %>%
  rownames_to_column("sampleID") %>%
  select(sampleID) %>%
  merge(., md_tamRun4, by = "sampleID") %>%
  mutate(
    wholeGenome = 
      case_when(
        animalID %in% wholeGenomeSeqs$animalID ~ "yes",
        .default = "no"
      )
  ) %>%
  relocate(wholeGenome, .after = sampleID)

genoRate90_blood_md_30x <- genoRate90_blood_30x %>%
  t() %>%
  as.data.frame() %>%
  rownames_to_column("sampleID") %>%
  select(sampleID) %>%
  merge(., md_tamRun4, by = "sampleID") %>%
  mutate(
    wholeGenome = 
      case_when(
        animalID %in% wholeGenomeSeqs$animalID ~ "yes",
        .default = "no"
      )
  ) %>%
  relocate(wholeGenome, .after = sampleID)
```

## Fecal samples

### 10x coverage

Filtering to 10x coverage gives 26 samples + 53 loci with 100% genotype success + 3 fecal samples with whole-genome seqs

```{r}
passSamples_fecal_10x <- sampleSum_fecal_10x %>%
  filter(GenotypeRate >= 0.5)

genoRate100_fecal_10x <- genos_10x %>%
  replace(., .==0, NA) %>%
  t() %>%
  as.data.frame() %>%
  rownames_to_column("sampleID") %>%
  # filter to fecal samples w/ >50% geno success
  filter(sampleID %in% passSamples_fecal_10x$sampleID) %>%
  column_to_rownames("sampleID") %>%
  t() %>%
  as.data.frame() %>%
  # remove species-specific loci
  rownames_to_column("locus") %>%
  filter(!str_detect(locus, "LWED|SIMP")) %>%
  column_to_rownames("locus") %>%
  # filter to only samples/loci w/100% genotype success
  na.omit()
```

genoRate100 metadata, including whether samples have whole genome sequenced

```{r}
wholeGenomeSeqs <- read_excel("./04_tamRun4/DNASubmissionForm_MEW.xlsx",
                              range = "A17:R44") %>%
  as.data.frame() %>%
  dplyr::rename("animalID" = "Animal ID") %>%
  select(animalID)

genoRate100_fecal_md_10x <- genoRate100_fecal_10x %>%
  t() %>%
  as.data.frame() %>%
  rownames_to_column("sampleID") %>%
  select(sampleID) %>%
  merge(., md_tamRun4, by = "sampleID") %>%
  mutate(
    wholeGenome = 
      case_when(
        animalID %in% wholeGenomeSeqs$animalID ~ "yes",
        .default = "no"
      )
  ) %>%
  relocate(wholeGenome, .after = sampleID)
```

### 30x coverage

Filtering to 30x coverage gives 26 samples + 53 loci with 100% genotype success + 3 fecal samples with whole-genome seqs (same as 10x)

```{r}
passSamples_fecal_30x <- sampleSum_fecal_30x %>%
  filter(GenotypeRate >= 0.5)

genoRate100_fecal_30x <- genos_30x %>%
  replace(., .==0, NA) %>%
  t() %>%
  as.data.frame() %>%
  rownames_to_column("sampleID") %>%
  # filter to fecal samples w/ >50% geno success
  filter(sampleID %in% passSamples_fecal_30x$sampleID) %>%
  column_to_rownames("sampleID") %>%
  t() %>%
  as.data.frame() %>%
  # remove species-specific loci
  rownames_to_column("locus") %>%
  filter(!str_detect(locus, "LWED|SIMP")) %>%
  column_to_rownames("locus") %>%
  # filter to only samples/loci w/100% genotype success
  na.omit()

genoRate90_fecal_30x <- genos_30x %>%
  replace(., .==0, NA) %>%
  t() %>%
  as.data.frame() %>%
  rownames_to_column("sampleID") %>%
  # filter to blood samples w/ >50% geno success
  filter(sampleID %in% passSamples_fecal_30x$sampleID) %>%
  column_to_rownames("sampleID") %>%
  t() %>%
  as.data.frame() %>%
  # remove species-specific loci
  rownames_to_column("locus") %>%
  filter(!str_detect(locus, "LWED|SIMP")) %>%
  column_to_rownames("locus") %>%
  # filter to samples with < 10% missing genos
  select_if(~mean(is.na(.)) <= 0.1)
```

genoRate100 metadata, including whether samples have whole genome sequenced

```{r}
wholeGenomeSeqs <- read_excel("./04_tamRun4/DNASubmissionForm_MEW.xlsx",
                              range = "A17:R44") %>%
  as.data.frame() %>%
  dplyr::rename("animalID" = "Animal ID") %>%
  select(animalID)

genoRate100_fecal_md_30x <- genoRate100_fecal_30x %>%
  t() %>%
  as.data.frame() %>%
  rownames_to_column("sampleID") %>%
  select(sampleID) %>%
  merge(., md_tamRun4, by = "sampleID") %>%
  mutate(
    wholeGenome = 
      case_when(
        animalID %in% wholeGenomeSeqs$animalID ~ "yes",
        .default = "no"
      )
  ) %>%
  relocate(wholeGenome, .after = sampleID)

genoRate90_fecal_md_30x <- genoRate90_fecal_30x %>%
  t() %>%
  as.data.frame() %>%
  rownames_to_column("sampleID") %>%
  select(sampleID) %>%
  merge(., md_tamRun4, by = "sampleID") %>%
  mutate(
    wholeGenome = 
      case_when(
        animalID %in% wholeGenomeSeqs$animalID ~ "yes",
        .default = "no"
      )
  ) %>%
  relocate(wholeGenome, .after = sampleID)
```

# Blood: loci vs. samples

## Define function

```{r}
# number of samples genotyped at propLoci (0 to 1.0) of 2 to 140 loci
optimize_lociSamples <- function(df, propLoci) {
  # reformat genos
  df <- df %>%
    t() %>%
    as.data.frame()
  
  # Initialize an empty list to store results
  result_list <- list()

  for (n in 2:140) {
    # create matrix where NAs are 0 and non-NAs are 1
    is_na_vector <- ifelse(is.na(df), 0, 1)

    crossprod_vector <- crossprod(is_na_vector)
    col_sums <- colSums(crossprod_vector)

    x_df <- as.data.frame(df)  # to get meaningful colnames
    res <- x_df[, rank(-col_sums, ties.method = "first") <= n]

    # Store the result in the list
    result_list[[as.character(n)]] <- c(n, nrow(res[which(rowMeans(!is.na(res)) >= propLoci), ]))
  }

  # Convert the list to a dataframe
  result_df <- do.call(rbind, result_list)

  # Rename columns
  colnames(result_df) <- c("loci", "samples")

  return(result_df)
}
```

## Identify optimal sample/loci combos

### gtscore

```{r}
# samples genotyped at 100% of loci among subsets of 1 to 140 loci
opt100_lociSamples_ill_gtscore <-
  optimize_lociSamples(genos_ill_gtscore_samplesBlood_lociDual, 1) %>%
  as.data.frame() %>%
  dplyr::rename("samples100_ill.gtscore" = "samples")

opt100_samplesLoci_ont_gtscore <- optimize_samplesLoci(genos_ont_gtscore_samplesBlood_lociDual, 1) %>%
  as.data.frame() %>%
  dplyr::rename("loci100_ont.gtscore" = "loci")

# loci genotyped at 90% of samples among subsets of 10 to 30 samples
opt90_samplesLoci_ill_gtscore <- optimize_samplesLoci(genos_ill_gtscore_samplesBlood_lociDual, 0.9) %>%
  as.data.frame() %>%
  dplyr::rename("loci90_ill.gtscore" = "loci")

opt90_samplesLoci_ont_gtscore <- optimize_samplesLoci(genos_ont_gtscore_samplesBlood_lociDual, 0.9) %>%
  as.data.frame() %>%
  dplyr::rename("loci90_ont.gtscore" = "loci")
```

### gtseq

```{r}
# loci genotyped at 90% of samples among subsets of 10 to 30 samples
opt100_samplesLoci_ill_gtseq <- optimize_samplesLoci(genos_ill_gtseq_samplesBlood_lociDual, 1) %>%
  as.data.frame() %>%
  dplyr::rename("loci100_ill.gtseq" = "loci")

opt100_samplesLoci_ont_gtseq <- optimize_samplesLoci(genos_ont_gtseq_samplesBlood_lociDual, 1) %>%
  as.data.frame() %>%
  dplyr::rename("loci100_ont.gtseq" = "loci")

# loci genotyped at 90% of samples among subsets of 10 to 30 samples
opt90_samplesLoci_ill_gtseq <- optimize_samplesLoci(genos_ill_gtseq_samplesBlood_lociDual, 0.9) %>%
  as.data.frame() %>%
  dplyr::rename("loci90_ill.gtseq" = "loci")

opt90_samplesLoci_ont_gtseq <- optimize_samplesLoci(genos_ont_gtseq_samplesBlood_lociDual, 0.9) %>%
  as.data.frame() %>%
  dplyr::rename("loci90_ont.gtseq" = "loci")
```

### Summaries

```{r}
# loci genotyped at 100% of samples among subsets of 10 to 30 samples
opt100_samplesLoci_summary <- merge(opt100_samplesLoci_ill_gtscore, opt100_samplesLoci_ont_gtscore, by = "samples") %>%
  merge(., opt100_samplesLoci_ill_gtseq, by = "samples") %>%
  merge(., opt100_samplesLoci_ont_gtseq, by = "samples")

# loci genotyped at 90% of samples among subsets of 10 to 30 samples
opt90_samplesLoci_summary <- merge(opt90_samplesLoci_ill_gtscore, opt90_samplesLoci_ont_gtscore, by = "samples") %>%
  merge(., opt90_samplesLoci_ill_gtseq, by = "samples") %>%
  merge(., opt90_samplesLoci_ont_gtseq, by = "samples")
```
