---
title: "ch2_popGen"
author: "Rachel Voyt"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1 Overview

Pop gen stats

Resources: - Tom Jenkins tutorial
(<https://tomjenkins.netlify.app/tutorials/r-popgen-getting-started/>)

# 2 Packages

```{r}
library(adegenet)
library(conflicted)
library(ggpubr)
library(gt)
library(hierfstat)

#install_github("nikostourvas/PopGenUtils")
library("PopGenUtils")

library(poppr)
library(reshape2)
library(RColorBrewer)
library(scales)
library(tidyverse)
conflicts_prefer(dplyr::filter)
conflicts_prefer(dplyr::lag)
conflicts_prefer(dplyr::select)
conflicted::conflicts_prefer(dplyr::slice)
conflicts_prefer(base::unique)
conflicts_prefer(base::duplicated)

source("./project_scripts/ravo_ch2Scripts.R")
source("./project_scripts/ravo_gtScripts.R")
#if (!require("BiocManager", quietly = TRUE))
#    install.packages("BiocManager")
#BiocManager::install("msa")
source("./project_scripts/GTscore/GTscore_modified.R")
```

# 3 Data

## metadata

```{r}
md <- read.csv("./metadataReconciliation/tamRun5_metadata_v5.csv") %>%
  mutate(captureYear = str_sub(captureDate, 1, 4)) %>%
  relocate(captureYear, .after = captureDate)

sampleRef <- read.csv("./project_data/master_sampleInfo.csv")

lociRef <- read.csv("./project_data/master_lociInfo.csv")
```

## capData

```{r}
# 2009 to 2023
capData_2009to2023 <- read.csv("./DISSERTATION/ch2_demography/00_data/01_demoData_clean/captureData_byIndividual_v6.csv") %>%
  # ditch UNK animalIDs
  filter(animalID != "UNK") %>%
  # add captureYear
  mutate(
    captureYear = str_sub(captureDate, 1, 4),
    captureYear = as.numeric(captureYear)
  ) %>%
  relocate(captureYear, .after = captureDate)

capData_2009to2023_firstEntry <- capData_2009to2023[match(unique(capData_2009to2023$animalID), capData_2009to2023$animalID),] %>%
  select(rowID, captureDate, animalID, ageClass, animalName1, animalName2, groupName, species, sex, notes_MD, notes_RV)

# 2009 to 2019 only
capData_2009to2019 <- capData_2009to2023 %>%
  filter(rowID <= 613)

capData_2009to2019_firstEntry <- capData_2009to2019[match(unique(capData_2009to2019$animalID), capData_2009to2019$animalID),] %>%
  select(rowID, captureDate, animalID, ageClass, animalName1, animalName2, groupName, species, sex, notes_MD, notes_RV)
```

## capHist

Filtered to 2009-2019 only

```{r}
capHist_2009to2019 <- read.csv("./DISSERTATION/ch2_demography/00_data/01_demoData_clean/captureHistories_byIndiv_2009to2023.csv", colClasses = "character") %>%
  # filter to 2009-2019 only
  mutate(ch = str_sub(ch, 1, 11)) %>%
  filter(ch != "00000000000")
```

## birth/death year

```{r}
birthAssignments <- read.csv("./DISSERTATION/ch2_demography/00_data/01_demoData_clean/birthAssignments_capData_tamRun5_v3.csv") %>%
  select(animalID, birthYear_est) %>%
  na.omit() %>%
  distinct() %>%
  arrange(animalID) %>%
  merge(., capData_2009to2023[, c("animalID", "captureYear", "groupName")], by.x = c("animalID", "birthYear_est"), by.y = c("animalID", "captureYear"), all.x = T) %>%
  distinct()

deathAssignments <- read.csv("./DISSERTATION/ch2_demography/00_data/01_demoData_clean/deathAssignments_capData_tamRun5_v1.csv") %>%
  select(animalID, deathYear_est) %>%
  na.omit() %>%
  distinct() %>%
  arrange(animalID)
```

## parityAssignments

Using "parity2" -- meaning either left or right (vs avg) meets threshold

```{r}
parityAssignments <- get_parityStatus(capData_file = capData_2009to2019) %>%
  select(animalID, captureDate, parity2) %>%
  dplyr::rename("parity" = "parity2") %>%
  mutate(
    captureYear = str_sub(captureDate, 1, 4)
  ) %>%
  #select(-captureDate) %>%
  distinct() %>%
  merge(., birthAssignments, by = "animalID", all.x = T) %>%
  dplyr::rename("natalGroup" = "groupName") %>%
  merge(., capData_2009to2019[, c("animalID", "captureDate", "groupName")], by = c("animalID", "captureDate")) %>%
  dplyr::rename("capGroup" = "groupName")

parousOnly <- parityAssignments %>%
  filter(parity == "parous") %>%
  arrange(animalID, captureYear)

parityAssigments_firstParous <- parousOnly[match(unique(parousOnly$animalID), parousOnly$animalID),] %>%
  filter(parity == "parous") %>%
  dplyr::rename("parityYear" = "captureYear") %>%
  mutate(natalCap_match = natalGroup == capGroup)
```

## likelyTwins

created in ch2_demography.Rmd

```{r}
likelyTwins <- read.csv("./DISSERTATION/ch2_demography/00_data/01_demoData_clean/twinList_tamRun5_6June2024.csv")

twinList <- likelyTwins %>%
  select(twin1, twin2, twin3) %>%
  arrange(twin1) %>%
  mutate(twinSet = str_c("twinSet", row_number())) %>%
  pivot_longer(-twinSet,
               names_to = "twinID",
               values_to = "animalID") %>%
  na.omit() %>%
  merge(., sampleRef[, c("animalID", "sampleID_franz", "seqRun", "sampleType")], by = "animalID", all.x = T) %>%
  filter(seqRun == "run5") %>%
  filter(sampleType == "hair") %>%
  relocate(twinSet) %>%
  arrange(twinSet)
```

## genos

```{r}
genos10x <- read.table("./05_tamRun5/03_run5GTscore/fullSet_polyGenResults_singleSNP_10x.txt", header = T, na.strings = "0") %>%
  `rownames<-`(sub('[_][^_]+$', '', rownames(.)))
```

quick overview --

species sex count 1 LWED F 64 2 LWED M 70 3 SIMP F 47 4 SIMP M 52

samples from indivs born during study period: species count 1 LWED 62 2
SIMP 52

```{r}
hairGenoData_overview <- md %>%
  filter(species %in% c("LWED", "SIMP")) %>%
  filter(sampleType == "hair") %>%
  select(animalID, species, sex) %>%
  distinct()

# number of F & M by species
hairGenoData_overview %>%
  group_by(species, sex) %>%
  summarise(count = n()) %>%
  as.data.frame()

# number born during study period (by species)
hairGenoData_overview %>%
  filter(animalID %in% birthAssignments$animalID) %>%
  group_by(species) %>%
  summarise(count = n()) %>%
  as.data.frame()
```

# Mark-recapture

```{r}
library(marked)

capHist_2009to2019
```

## Capture overview

### Tables

**by group**

Summary of captures for each year (a la Thorley et al. 2023)

```{r}
lwed_capSummary_byGroup <- get_capData_summary(capData_2009to2019, "LWED")

simp_capSummary_byGroup <- get_capData_summary(capData_2009to2019, "SIMP")

lwedSIMP_capSummary_byGroup <- rbind(lwed_capSummary_byGroup, simp_capSummary_byGroup) %>%
  mutate(
    species = case_when(
      species == "LWED" ~ "Saddleback tamarins",
      species == "SIMP" ~ "Emperor tamarins"
    )
  )
```

```{r}
capSummary_table <- lwedSIMP_capSummary_byGroup %>%
  gt(groupname_col = "species",
     rowname_col = "captureYear") %>%
  cols_label(
    captureYear = md("**captureYear**"),
    species = md("**species**"),
    capPeriod_start = md("**periodStart**"),
    capPeriod_end = md("**periodEnd**"),
    capPeriod_days = md("**periodDays**"),
    nGroups = md("**nGroups**"),
    nF = md("**nFemales**"),
    nM = md("**nMales**"),
    nTotal = md("**nTotal**"),
    nRecap = md("**nRecap**"),
    nNew = md("**nNew**"),
    propRecap = md("**propRecap**"),
    propNew = md("**propNew**")
  ) %>%
  fmt(
    columns = everything(),
    rows = 12:13,
    fns = function(x) ifelse(x == "0", "—", NA)
  ) %>%
  sub_missing(columns = everything(),
              rows = 12:13)

capSummary_table %>%
  gtsave("./DISSERTATION/ch2_demography/02_results/00_tablesFigures/table_capSummary.png", vwidth = 1500, vheight = 1000)
```

**by indiv**

```{r}

```

### Plots

Cleveland plot showing capture history for LWED and SIMP social groups
from 2009 to 2019

```{r}
# keep all cap records
capOverview_continuous <- capData_2009to2019 %>%
  select(captureDate, captureYear, species, groupName, animalID) %>%
  distinct()

# only keep one per year
capOverview_discrete <- capData_2009to2019 %>%
  select(captureYear, species, groupName, animalID) %>%
  distinct() %>%
  # add first year captured
  arrange(species, groupName, captureYear) %>%
  group_by(groupName) %>%
  mutate(firstCapture = min(captureYear)) %>%
  ungroup() %>%
  # add number of indv captured each event
  group_by(species, captureYear, groupName, firstCapture) %>%
  summarise(nCaptured = n()) %>%
  # add capID
  group_by(species, groupName) %>%
  mutate(capID = row_number()) %>%
  ungroup() %>%
  arrange(species, groupName) %>%
  # add nextCapture
  group_by(groupName) %>%
  mutate(nextCapture = lead(captureYear)) %>%
  #filter(!is.na(nextCapture)) %>%
  ungroup() %>%
  relocate(species, groupName, firstCapture, captureYear, nextCapture, capID, nCaptured) %>%
  # remove UNK group name & Loners
  filter(!groupName %in% c("UNK", "Loners"))
```

```{r}
ggplot(capOverview_discrete) +
  geom_segment(aes(x = captureYear,
                   xend = nextCapture,
                   y = groupName,
                   yend = groupName,
                   color = as.factor(capID))) +
  geom_point(aes(x = captureYear, y = groupName, color = as.factor(capID), size = nCaptured)) +
  scale_x_continuous(breaks = 2009:2019) +
  facet_grid(rows = vars(species), scales = "free_y") +
  theme_bw()
```

# Basic demography stuff

## Time in study

The dataframe below includes the animalIDs of all individuals in
capData_byIndiv_v5 from 2009 to 2023. It includes the first and last
years and groups in which the individual was captured as well as the
difference between those two years.

```{r}
# first anti-sort for last entries in capture records
capData_2009to2023_antiSort <- capData_2009to2023 %>%
  arrange(desc(rowID))

# then create df
capData_2009to2023_first.lastEntry <- capData_2009to2023_antiSort[match(unique(capData_2009to2023_antiSort$animalID), capData_2009to2023_antiSort$animalID),] %>%
  
  arrange(as.numeric(animalID)) %>%
  select(animalID, captureYear, groupName, species, sex) %>%
  dplyr::rename("captureYear_last" = "captureYear",
                "groupName_last" = "groupName") %>%
  merge(., capData_2009to2023_firstEntry[, c("animalID", "captureDate", "groupName")], by = "animalID") %>%
  mutate(
    captureYear_first = str_sub(captureDate, 1, 4)
  ) %>%
  select(-captureDate) %>%
  dplyr::rename("groupName_first" = "groupName") %>%
  mutate(
    captureYear_first = as.numeric(captureYear_first),
    captureYear_last = as.numeric(captureYear_last),
    diff_lastFirst = captureYear_last - captureYear_first,
    groupMatch = case_when(
      groupName_first == groupName_last ~ "yes",
      .default = "no"
    )
  ) %>%
  relocate(animalID, groupName_first, groupName_last, groupMatch, captureYear_first, captureYear_last, diff_lastFirst) %>%
  
  # see who started in study as a juvenile
  merge(., birthAssignments[, c("animalID", "birthYear_est")], by = "animalID", all.x = T) %>%
  
  # then calculate age at last capture
  mutate(
    age_lastCap = case_when(
      !is.na(birthYear_est) ~ diff_lastFirst,
      .default = NA
    )
  )
```

Maybe good to get a list of who was present in the population each
year??

```{r}
library(dplyr)
library(purrr)
conflicts_prefer(purrr::reduce)

presenceMatrix <- capData_2009to2023_first.lastEntry %>%
  select(animalID, species, sex, captureYear_first, captureYear_last) %>%
  reduce(2009:2023, .init = ., ~ .x %>%
    mutate("{.y}" := ifelse(.y >= captureYear_first & .y <= captureYear_last, animalID, NA))) %>%
  merge(., birthAssignments, by = "animalID", all.x = T) %>%
  relocate(birthYear_est, .before = captureYear_first)

colSums(!is.na(presenceMatrix)) %>%
  as.data.frame()
```

## Group numbers

```{r}
# 16 lwed groups, 10 simp groups
capData_2009to2019 %>%
  filter(groupName != "Loners") %>%
  group_by(species) %>%
  summarise(nGroups = length(unique(groupName)))

# really quick and dirty avg group size estimates...
lwed_grpSizes <- capData_2009to2019 %>%
  filter(species == "LWED") %>%
  group_by(groupName, captureYear) %>%
  summarise(count = n()) %>%
  filter(groupName != "Loners")
round(mean(lwed_grpSizes$count)) # 5

simp_grpSizes <- capData_2009to2019 %>%
  filter(species == "SIMP") %>%
  group_by(groupName, captureYear) %>%
  summarise(count = n()) %>%
  filter(groupName != "Loners")
round(mean(simp_grpSizes$count)) # 5
```

## Group membership

```{r}
groupMembership <- capData_2009to2019 %>%
  select(animalID, captureDate, captureYear, groupName) %>%
  # first need to get cases w/multiple groups in one year into single col
  distinct() %>%
  mutate(temp = str_c(animalID, "_", captureYear)) %>%
  arrange(temp, captureDate) %>%
  mutate(dup = if_else(duplicated(temp), "entry2", "entry1")) %>%
  select(temp, groupName, dup) %>%
  distinct() %>%
  pivot_wider(names_from = dup,
              values_from = groupName) %>%
  mutate(
    groupName = case_when(
      is.na(entry2) ~ entry1,
      entry1 == entry2 ~ entry1,
      entry1 != entry2 ~ str_c(entry1, entry2, sep = "_")
    )
  ) %>%
  separate("temp", into = c("animalID", "captureYear")) %>%
  select(animalID, captureYear, groupName) %>%
  mutate(captureYear = str_c("grp", captureYear)) %>%
  # now can make big df
  pivot_wider(names_from = captureYear,
              values_from = groupName) %>%
  arrange(as.numeric(animalID)) %>%
  relocate(c(animalID, grp2009, grp2010, grp2011, grp2012, grp2013, grp2014))
```

# 4 Filter genos

Prior to using genotyping results in additional analyses, I'm using the
following filters for quality control:

1.  Species mismatches - remove any that conflict w/metadata
2.  Sex mismatches - remove any where majority of sex loci conflict
    w/metadata
3.  Duplicates - for any duplicate pairs identified, remove the sample
    with lower geno success
4.  Separate genos by species & filter loci to retain only those for individual identification, then (following Arpin et al. 2024):
    1.  Remove loci with \>50% missing data
    2.  Remove samples with \>50% missing data
    3.  Remove loci with minor allele freq \<0.03

## 4.1 Species/sex checks

### species checks

4 species mismatches, though each was genotyped at only one locus - will
probably get filtered out when ditching poorly performing loci/samples

```{r}
genos_forChecks <- genos10x %>%
  # subset to hair samples
  select(md[md$species %in% c("LWED", "SIMP") & md$sampleType == "hair", "sampleID"])

speciesChecks <- assignSpecies(genos_forChecks, md, "sampleID")

toRemove_speciesChecks <- speciesChecks %>%
  filter(mdMatch == F) %>%
  select(sampleID) %>%
  pull()
```

### sex checks

15 sex mismatches, though 12 of these are MIX. Vast majority clear up if
go by the majority of calls -- not terribly worried about it though, bc
when I tried before to first ditch poorly performing samples and then
run assignSex, only got 3 mismatches and all were MIX (one 50/50, two
with majority correct).

--just checked the sampleSummary for the samples where mdMajMatch were
either F or NA -- all of these had shit genoSuccess so they'll be
ditched regardless

```{r}
# run sexChecks (can use same geno file as speciesChecks)
sexChecks <- assignSex(genos_forChecks, md, "sampleID", exclude_nonTargetSp = "yes") %>%
  filter(mdMatch == FALSE) %>%
  mutate(
    majSex = case_when(
      propF > propM ~ "F",
      propM > propF ~ "M",
      .default = NA
    ),
    mdMajMatch = mdSex == majSex
  ) %>%
  relocate(c(majSex, mdMajMatch), .after = "mdSex")

toRemove_sexChecks <- sexChecks %>%
  filter(mdMajMatch == F | is.na(mdMajMatch)) %>%
  select(sampleID) %>%
  pull()
```

## 4.2 Dup checks

Here I'm using my version of the GTscore "IDduplicateSamples" function;
I modified it only slightly so that I could use a dataframe as input vs.
the polygen output.

Results show tamRun5.211/tamRun5.264 as the only likely duplicate pair -
this is the set that I already caught in metadataReconciliation.Rmd for
animalID 200

**tamRun5.211 has the higher genotype success -- ditch tamRun5.264**

```{r}
run5_dupTest_10x <- get_dupSamples(genos_forChecks)

# view likely dups - only one pair (tamRun5.211/tamRun5.264)
run5_likelyDups <- run5_dupTest_10x %>%
  filter(proportionCommon >= 0.5) %>%
  filter(proportionMatch > 0.9)
run5_likelyDups

View(md %>% filter(sampleID %in% c(run5_likelyDups$Sample1, run5_likelyDups$Sample2)))

# tamRun5.211 has the higher genotype success -- ditch tamRun5.264
run5_sampleSum <- read.csv("./05_tamRun5/03_run5GTscore/summaryFiles/master_sampleSummary_10x.csv")

run5_sampleSum %>% filter(Sample %in% c(run5_likelyDups$Sample1, run5_likelyDups$Sample2)) %>% select(Sample, GenotypeRate)

toRemove_dupChecks <- "tamRun5.264"
```

## 4.3 Poor performance

I found that some of the loci that were supposed to be sex-specific
sometimes showed some variation in the other species -- as such, I'm
opting to ignore the species designation for INDID SNPs and instead add
a blanket filter to remove any loci whose minAF is less than 0.03
(following Arpin et al. 2024, who used this cutoff)

First though, apply filters based on species/sex checks, dupChecks, and
genotype success--

```{r}
lwed_genos_temp <- genos10x %>%
  # subset to hair + INDID only
  select(md[md$species == "LWED" & md$sampleType == "hair", "sampleID"]) %>%
  filter(!str_detect(rownames(.), "SEXID|SPECIESID")) %>%
  t() %>%
  as.data.frame() %>%
  # 1. remove any species/sex mismatches
  filter(!rownames(.) %in% toRemove_speciesChecks) %>%
  filter(!rownames(.) %in% toRemove_sexChecks) %>%
  # 2. remove any dups
  filter(!rownames(.) %in% toRemove_dupChecks) %>%
  # 3. remove SNPs with >50% missing data
  select(where(~sum(!is.na(.x))/length(.x) >= 0.5)) %>%
  # 4. remove samples with >50% missing data
  filter((rowSums(!is.na(.)) / ncol(.)) >= 0.5)

simp_genos_temp <- genos10x %>%
  # subset to hair + INDID only
  select(md[md$species == "SIMP" & md$sampleType == "hair", "sampleID"]) %>%
  filter(!str_detect(rownames(.), "SEXID|SPECIESID")) %>%
  t() %>%
  as.data.frame() %>%
  # 1. remove any species/sex mismatches
  filter(!rownames(.) %in% toRemove_speciesChecks) %>%
  filter(!rownames(.) %in% toRemove_sexChecks) %>%
  # 2. remove any dups
  filter(!rownames(.) %in% toRemove_dupChecks) %>%
  # 3. remove SNPs with >50% missing data
  select(where(~sum(!is.na(.x))/length(.x) >= 0.5)) %>%
  # 4. remove samples with >50% missing data
  filter((rowSums(!is.na(.)) / ncol(.)) >= 0.5)
```

Then use these genos to check and filter based on minAF

**LWED**

```{r}
locusTable <- read.table("./05_tamRun5/03_run5GTscore/fullSet_LocusTable_singleSNPs.txt", header = T) %>%
  mutate(
    Locus_ID = sub('[_][^_]+$', '', Locus_ID)
  )

lwed_locusTable <- locusTable %>%
  filter(Locus_ID %in% colnames(lwed_genos_temp))

lwed_readCounts_forFreqChecks <- read.table("./05_tamRun5/03_run5GTscore/fullSet_AlleleReads_singleSNPs_10x.txt") %>%
  select(md[md$species == "LWED" & md$sampleType == "hair", "sampleID"]) %>%
  `rownames<-`(sub('[_][^_]+$', '', rownames(.))) %>%
  filter(rownames(.) %in% colnames(lwed_genos_temp)) %>%
  select(rownames(lwed_genos_temp))

lwed_genos_forFreqChecks <- read.table("./05_tamRun5/03_run5GTscore/fullSet_polyGenResults_singleSNP_10x.txt", header = T) %>%
  select(md[md$species == "LWED" & md$sampleType == "hair", "sampleID"]) %>%
  `rownames<-`(sub('[_][^_]+$', '', rownames(.))) %>%
  filter(rownames(.) %in% colnames(lwed_genos_temp)) %>%
  select(rownames(lwed_genos_temp))

lwed_alleleFreqs <- summarizeGTscore(lwed_readCounts_forFreqChecks, lwed_locusTable, lwed_genos_forFreqChecks)

lwed_toRemove_afChecks <- lwed_alleleFreqs %>%
  filter(minAF < 0.03) %>%
  select(Locus_ID) %>%
  pull()
```

**SIMP**

```{r}
simp_locusTable <- locusTable %>%
  filter(Locus_ID %in% colnames(simp_genos_temp))

simp_readCounts_forFreqChecks <- read.table("./05_tamRun5/03_run5GTscore/fullSet_AlleleReads_singleSNPs_10x.txt") %>%
  select(md[md$species == "SIMP" & md$sampleType == "hair", "sampleID"]) %>%
  `rownames<-`(sub('[_][^_]+$', '', rownames(.))) %>%
  filter(rownames(.) %in% colnames(simp_genos_temp)) %>%
  select(rownames(simp_genos_temp))

simp_genos_forFreqChecks <- read.table("./05_tamRun5/03_run5GTscore/fullSet_polyGenResults_singleSNP_10x.txt", header = T) %>%
  select(md[md$species == "SIMP" & md$sampleType == "hair", "sampleID"]) %>%
  `rownames<-`(sub('[_][^_]+$', '', rownames(.))) %>%
  filter(rownames(.) %in% colnames(simp_genos_temp)) %>%
  select(rownames(simp_genos_temp))

simp_alleleFreqs <- summarizeGTscore(simp_readCounts_forFreqChecks, simp_locusTable, simp_genos_forFreqChecks)

simp_toRemove_afChecks <- simp_alleleFreqs %>%
  filter(minAF < 0.03) %>%
  select(Locus_ID) %>%
  pull()
```

## 4.4 Final genos

### genos

```{r}
lwed_genos <- lwed_genos_temp %>%
  select(!all_of(lwed_toRemove_afChecks)) %>%
  rownames_to_column("sampleID") %>%
  mutate(pop = "lwed") %>%
  merge(., md[, c("animalID", "sampleID", "sex", "captureYear")], by = "sampleID", all.x = T) %>%
  merge(., birthAssignments, by = "animalID", all.x = T) %>%
  relocate(sampleID, animalID, sex, birthYear_est, pop, captureYear)

simp_genos <- simp_genos_temp %>%
  select(!all_of(simp_toRemove_afChecks)) %>%
  rownames_to_column("sampleID") %>%
  mutate(pop = "simp") %>%
  merge(., md[, c("animalID", "sampleID", "sex", "captureYear")], by = "sampleID", all.x = T) %>%
  merge(., birthAssignments, by = "animalID", all.x = T) %>%
  relocate(sampleID, animalID, sex, birthYear_est, pop, captureYear)
```

### genind

```{r}
lwed_genind <- adegenet::df2genind(X = lwed_genos[,c(7:59)],
                                  sep = ",",
                                  ind.names = lwed_genos$sampleID,
                                  pop = lwed_genos$pop,
                                  NA.char = "NA",
                                  ploidy = 2,
                                  type = "codom")
lwed_genind@other$sex <- lwed_genos$sex
lwed_genind@other$birthYear <- lwed_genos$birthYear_est
lwed_genind@other$captureYear <- lwed_genos$captureYear

simp_genind <- adegenet::df2genind(X = simp_genos[,c(7:54)],
                                  sep = ",",
                                  ind.names = simp_genos$sampleID,
                                  pop = simp_genos$pop,
                                  NA.char = "NA",
                                  ploidy = 2,
                                  type = "codom")
simp_genind@other$sex <- simp_genos$sex
simp_genind@other$birthYear <- simp_genos$birthYear_est
simp_genind@other$captureYear <- simp_genos$captureYear
```

## 4.5 Quick overview of filtered geno data:

Retained LWED data include... - 72 indivs + 53 loci - avg 94% geno
success among samples - 43 non-juvs, 29 juvs

Retained SIMP data include... - 55 indivs + 48 loci - avg 93% geno
success among samples - 27 non-juvs, 28 juvs

```{r}
# LWED
lwed_genind
mean(propTyped(lwed_genind, by = "ind")) # 0.9360587
median(propTyped(lwed_genind, by = "ind")) # 1

sum(is.na(lwed_genos$birthYear_est)) # 43 non-juvs remaining
sum(!is.na(lwed_genos$birthYear_est)) # 29 juvs remaining

# SIMP
simp_genind
mean(propTyped(simp_genind, by = "ind")) # 0.932197
median(propTyped(simp_genind, by = "ind")) # 1

sum(is.na(simp_genos$birthYear_est)) # 27 non-juvs remaining
sum(!is.na(simp_genos$birthYear_est)) # 28 juvs remaining
```

# 5 Basic popGen

## 5.1 adegenet summary

we can use the summary() function from 'adegenet' to provide the
following: - Number of alleles per locus - Observed heterozygosity -
Expected heterozygosity

```{r}
lwed_genind_sum <- summary(lwed_genind)
simp_genind_sum <- summary(simp_genind)

lwed_genind_sum
simp_genind_sum
```

## Bartlett test

We can then test whether there's a significant difference exists between
mean observed and expected heterozygosity values using the **Bartlett
test of homogeneity of variances**.

There appears to be a significant different between observed and
expected heterozygosity for both LWED and SIMP

```{r}
# LWED
## p = 0.001065
bartlett.test(list(lwed_genind_sum$Hexp, lwed_genind_sum$Hobs)) 

t.test(lwed_genind_sum$Hexp, lwed_genind_sum$Hobs, paired = T, var.equal = T, alternative = "greater")

# SIMP
## p = 0.001576
bartlett.test(list(simp_genind_sum$Hexp, simp_genind_sum$Hobs)) 

t.test(simp_genind_sum$Hexp, simp_genind_sum$Hobs, paired = T, var.equal = T, alternative = "greater")
```

## HWE

The Hardy-Weinberg test function in 'adegenet' has been removed and
replaced by 'hw.test' in the package 'pegas'. This function uses both
the classical chi-squared test based on expected genotype frequencies
calculated from the allelic frequencies as well as an exact test based
on Monte Carlo permutations of alleles.

**NOTE** that B = 0 is used for the parametric version of the test;
larger numbers will indicate the number of permutations to use for the
Monte-Carlo version.

```{r}
library(pegas)
lwed_hwt <- hw.test(lwed_genind, B = 0)
simp_hwt <- hw.test(simp_genind, B = 0)
```

## Population structure

As of version 2.0.0, adegenet relies on hierfstat and pegas for most F
statistics.

```{r}
library(hierfstat)
wc(lwed_genind) # FST 0, FIS -0.005851178
wc(simp_genind) # FST 0, FIS 0.01158646
```

Use pegas to provide estimates by locus:

```{r}
library(pegas)
lwed_ftab <- Fst(as.loci(lwed_genind))
simp_ftab <- Fst(as.loci(simp_genind))

lwed_ftab
simp_ftab
```

```{r}
lwed_basicStats <- basic.stats(lwed_genind, diploid = T)
simp_basicStats <- basic.stats(simp_genind, diploid = T)

lwed_basicStats$overall
simp_basicStats$overall
```

lwed_basicStats\$overall Ho Hs Ht Dst Htp Dstp Fst Fstp Fis Dest 0.4599
0.4573 0.4573 0.0000 NaN NaN 0.0000 NaN -0.0059 NaN

simp_basicStats\$overall Ho Hs Ht Dst Htp Dstp Fst Fstp Fis Dest 0.4339
0.4390 0.4390 0.0000 NaN NaN 0.0000 NaN 0.0116 NaN

Fis

```{r}
# Fis per locus
lwed_basicStats$Fis
simp_basicStats$Fis

# mean Fis
apply(lwed_basicStats$Fis, MARGIN = 2, FUN = mean, na.rm = T) # -0.0013
apply(simp_basicStats$Fis, MARGIN = 2, FUN = mean, na.rm = T) # 0.020525
```

Pairwise Fst

```{r}
lwed_fst <- genet.dist(lwed_genind, method = "WC84")
```

Is mean Hobs significantly lower than mean Hexp?

```{r}
lwed_genSum <- summary(lwed_genind)
names(lwed_genSum)

bartlett.test(list(lwed_genSum$Hexp, lwed_genSum$Hobs))
t.test(lwed_genSum$Hexp, lwed_genSum$Hobs, paired = T, var.equal = T, alternative = "greater")
```

# 6 PID, PIDsibs

## 6.1 Calculations

```{r}
lwed_pidPerm <- PopGenUtils::pid_permute(obj = lwed_genind, nrep = 1000)

simp_pidPerm <- PopGenUtils::pid_permute(obj = simp_genind, nrep = 1000)
```

## 5.2 Analysis

```{r}
# LWED
lwed_pidPerm_data <- lwed_pidPerm$median_values

## PID meets 0.0001 threshold at 11 loci
lwed_pid.thresh <- as.numeric(lwed_pidPerm_data[which(lwed_pidPerm_data$PID<0.0001, arr.ind=TRUE)[1],][[1]])

## PIDsibs meets 0.0001 threshold at 20 loci
lwed_pidSibs.thresh <- as.numeric(lwed_pidPerm_data[which(lwed_pidPerm_data$PIDsibs<0.0001, arr.ind=TRUE)[1],][[1]])

# SIMP
simp_pidPerm_data <- simp_pidPerm$median_values

## PID meets 0.0001 threshold at 11 loci
simp_pid.thresh <- as.numeric(simp_pidPerm_data[which(simp_pidPerm_data$PID<0.0001, arr.ind=TRUE)[1],][[1]])

## PIDsibs meets 0.0001 threshold at 21 loci
simp_pidSibs.thresh <- as.numeric(simp_pidPerm_data[which(simp_pidPerm_data$PIDsibs<0.0001, arr.ind=TRUE)[1],][[1]])
```

## 5.3 Plots

```{r}
pidPerm_data_forFig <- lwed_pidPerm$results %>%
  mutate(species = "LWED") %>%
  rbind(., simp_pidPerm$results %>% mutate(species = "SIMP"))

pidThreshLines_forFig <- data.frame(species = c("LWED", "SIMP"),
                                    vline = c(lwed_pid.thresh,
                                              simp_pid.thresh))
pidSibThreshLines_forFig <- data.frame(species = c("LWED", "SIMP"),
                                    vline = c(lwed_pidSibs.thresh,
                                              simp_pidSibs.thresh))

pidPerm_fig <- ggplot(pidPerm_data_forFig,
                      aes(y = value,
                          x = loci,
                          color = statistic)) +
  geom_boxplot() +
  scale_colour_Publication() +
  theme_Publication() +
  
  scale_x_discrete(limits = as.character(1:25)) +
  geom_vline(data = pidThreshLines_forFig,
             aes(xintercept = vline),
                 color = "#386cb0") +
  geom_vline(data = pidSibThreshLines_forFig,
             aes(xintercept = vline),
             color = "#fdb462") +
  
  labs(x = "Number of loci",
       y = "Probability of identity") +
  facet_grid(rows = vars(species))

pidPerm_fig

  xlab("Number of loci")
```

# 6 CKMRsim

## 6.1 Background

Resources: - CKMRsim-example-1
<https://eriqande.github.io/CKMRsim/articles/CKMRsim-example-1.html> -
Shedd et al. GitHub
<https://github.com/krshedd/Relative-fitness-of-Pink-Salmon/blob/main/analysis/8_Hogan_Stockdale_13_16_MS_CKMRSIM.Rmd>

## 6.2 Packages

```{r}
remotes::install_github("eriqande/CKMRsim")
library(CKMRsim)
library(tidyverse)
# the next step tests to see if mendel was already installed.
# if not, it installs it
if(system.file("bin", package = "CKMRsim") == "") {
  install_mendel(Dir = system.file(package = "CKMRsim"))
}
```

## 6.3 Format genos

I'm using the LWED and SIMP genos that I filtered earlier since these
are the data I'm using for FRANz.

------------------------------------------------------------------------

Long-format genos should have four columns: Indiv, Locus, gene_copy, and
Allele.

-   Alleles are named with characters (even if they are numbers, they
    must be coerced to characters)
-   The gene_copy column contains either a 1 or a 2 in every row,
    telling us which copy of the gene (in a diploid) is which allele
-   Missing data in the Allele column is given by NA

**LWED**

```{r}
lwed_genos_forCKMR_temp <- lwed_genos %>%
  column_to_rownames("sampleID")

lwed_genos_forCKMR <- names(lwed_genos_forCKMR_temp) %>%
  # split loci columns
  map_dfc(~ lwed_genos_forCKMR_temp %>%
            select(all_of(.x)) %>%
            separate(.x,
                     into = paste0(.x, c(".1", ".2")),
                     sep = ",")
          ) %>%
  mutate(across(everything(), ~ gsub("0", NA, .))) %>%
  rownames_to_column("Indiv") %>%
  # make long-format
  pivot_longer(
    cols = -Indiv, 
    names_to = c("Locus", "gene_copy"), 
    names_sep = "\\.", 
    values_to = "Allele"
  )
```

**SIMP**

```{r}
simp_genos_forCKMR_temp <- simp_genos %>%
  column_to_rownames("sampleID")

simp_genos_forCKMR <- names(simp_genos_forCKMR_temp) %>%
  map_dfc(~ simp_genos_forCKMR_temp %>%
            select(all_of(.x)) %>%
            separate(.x,
                     into = paste0(.x, c(".1", ".2")),
                     sep = ",")
          ) %>%
  mutate(across(everything(), ~ gsub("0", NA, .))) %>%
  rownames_to_column("Indiv") %>%
  pivot_longer(
    cols = -Indiv, 
    names_to = c("Locus", "gene_copy"), 
    names_sep = "\\.", 
    values_to = "Allele"
  )
```

## 6.4 Allele frequency

Allele frequency input format:

-   **Chrom** should be a character or integer denoting which chromosome
    the marker is on. (For example 1 or X or Omy12)
-   **Locus** gives a character vector with the names of the
    markers/loci. Please don't use spaces in the marker names!
-   **Pos** gives the genome coordinates of the marker. This needs to be
    a number (double or integer)
-   **Allele** gives the name of each allele at each locus. This must be
    a character vector. So, in our example, we will coerce the numbers
    we have for each locus into a character vector. Please do not use
    spaces in allele names.
-   **Freq** is the frequency of each allele in the population. These
    should sum to 1.0 over each locus.
-   The remaining columns, **LocIdx** and **AlleIdx** are integer
    indices that get assigned to each locus and to each allele within
    each index. These columns get filled by using the reindex_markers()
    function.

**LWED**

```{r}
lwed_locusNames <- unique(lwed_genos_forCKMR$Locus)

lwed_afreqs_ready <- lwed_genos_forCKMR %>%
  count(Locus, Allele) %>%  
  group_by(Locus) %>%
  mutate(
    Freq = n / sum(n),
    #Chrom = "Unk",
    #Pos = as.integer(factor(Locus, levels = lwed_locusNames))
  ) %>%
  
  merge(., lociRef[, c("locus", "chr_num", "snpPos_inChr")], by.x = "Locus", by.y = "locus") %>%
  dplyr::rename("Chrom" = "chr_num",
                "Pos" = "snpPos_inChr") %>%
  
  ungroup() %>%
  select(Chrom, Pos, Locus, Allele, Freq) %>%
  arrange(Pos, desc(Freq)) %>%
  mutate(AlleIdx = NA,
         LocIdx = NA) %>%
  # NOTE - is very impt to remove NAs from allele frequencies
  filter(!is.na(Allele)) %>%
  # pass to reindex_markers to give allele-freq df for CKRMsim
  reindex_markers()
```

**SIMP**

```{r}
simp_locusNames <- unique(simp_genos_forCKMR$Locus)

simp_afreqs_ready <- simp_genos_forCKMR %>%
  count(Locus, Allele) %>%  
  group_by(Locus) %>%
  mutate(
    Freq = n / sum(n),
    #Chrom = "Unk",
    #Pos = as.integer(factor(Locus, levels = lwed_locusNames))
  ) %>%
  
  merge(., lociRef[, c("locus", "chr_num", "snpPos_inChr")], by.x = "Locus", by.y = "locus") %>%
  dplyr::rename("Chrom" = "chr_num",
                "Pos" = "snpPos_inChr") %>%
  
  ungroup() %>%
  select(Chrom, Pos, Locus, Allele, Freq) %>%
  arrange(Pos, desc(Freq)) %>%
  mutate(AlleIdx = NA,
         LocIdx = NA) %>%
  # NOTE - is very impt to remove NAs from allele frequencies
  filter(!is.na(Allele)) %>%
  # pass to reindex_markers to give allele-freq df for CKRMsim
  reindex_markers()
```

## 6.5 Create a CKMR object (create_ckmr)

-   **D** : the tibble of allele frequencies that has been run through
    reindex_markers().
-   **kappa_matrix** : A matrix that describes the pairwise
    relationships that you will be wanting to jointly simulate genotypes
    and likelihoods for. Each row is named by what you want to call the
    relationship and there are three columns which give, respectively,
    the probability that a pair with such a relationship share 0, 1, or
    2 genes identical-by-descent. The CKMRsim package comes with a
    matrix called kappas that has this information for 12 relationships:
    -   **MZ** : monozgotic twins (or "self"). This can be used to
        figure out how much power you have for identifying the same
        individual, sampled twice.
    -   **PO** : parent-offspring.
    -   **FS** : full siblings.
    -   **HS** : half siblings.
    -   **GP** : grandparent - grandoffspring.
    -   **AN** : aunt-neice (same as uncle-nephew or any such avuncular
        relationship)
    -   **DFC** : double first cousins.
    -   **FC** : first cousins.
    -   **HC** : half cousins.
    -   **U** : unrelated
-   **ge_mod_assumed**

**LWED**

```{r}
lwed_ckmr <- create_ckmr(
  D = lwed_afreqs_ready,
  kappa_matrix = kappas[c("PO", "FS", "HS", "U"), ],
  ge_mod_assumed = ge_model_TGIE,
  ge_mod_true = ge_model_TGIE,
  ge_mod_assumed_pars_list = list(epsilon = 0.01),
  ge_mod_true_pars_list = list(epsilon = 0.01)
)
lwed_ckmr
```

**SIMP**

```{r}
simp_ckmr <- create_ckmr(
  D = simp_afreqs_ready,
  kappa_matrix = kappas[c("PO", "FS", "HS", "U"), ],
  ge_mod_assumed = ge_model_TGIE,
  ge_mod_true = ge_model_TGIE,
  ge_mod_assumed_pars_list = list(epsilon = 0.01),
  ge_mod_true_pars_list = list(epsilon = 0.01)
)
simp_ckmr
```

## 6.6 Simulations & log-probabilities

simulate_Qij() -- This function simulates genotypes from different
relationships, and then for each simulated genotype pair it also
calculates the log probability of the pair of genotypes conditional on
the pair being of one or several relationships. This function has four
main inputs that we discuss here (the others allow the addition of
missing data and physical linkage, and are discussed in other
vignettes).

-   **C** : the ckmr object to use for simulation and probabilty
    calculation.
-   **sim_relats** : the set of true relationships you want to simulate
    genotypes from.
-   **calc_relats** : the set of assumed relationships you wish to
    compute genotype probabilities for, from the simulated data.
-   **reps** : for each relationship in sim_relats, the number of
    genotype pairs to simulate. This is, by default 10,000.

PO = parent-offspring FS = full sib HS = half sib HAN = half aunt-niece
(or uncle/niece) U = unrelated

### Run simulations

**LWED**

```{r}
# This simulates a large number of pairwise genotype probabilites
# at each locus
lwed_Qs <- simulate_Qij(lwed_ckmr,
                        calc_relats = c("PO", "FS", "HS", "U"),
                        sim_relats = c("PO", "FS", "HS", "U")
                        )
lwed_Qs
```

**SIMP**

```{r}
simp_Qs <- simulate_Qij(simp_ckmr,
                        calc_relats = c("PO", "FS", "HS", "U"),
                        sim_relats = c("PO", "FS", "HS", "U")
                        )
simp_Qs
```

### Extract logls

extract_logls() function lets us get simulated log-likelihood ratios out
of the Qij object we created above and plot a histogram or a density
plot of the distributions. This is helpful for developing intuition
about things and understanding what is going on.

In the following, we extract those genotype probabilities in different
ways to calculate a variety of different log-likelihood ratios for
different relationships.

**LWED**

```{r}
# log likelihood ratios for the hypothesis of PO vs Unrelated
lwed_PO_U_logls <- extract_logls(lwed_Qs,
                                 numer = c(PO = 1),
                                 denom = c(U = 1)
                                 ) %>%
  mutate(true_relat_pub = case_when(true_relat == "FS" ~ "Full Sibling",
                                    true_relat == "HS" ~ "Half Sibling",
                                    true_relat == "PO" ~ "Parent-Offspring",
                                    true_relat == "U" ~ "Unrelated")) %>% 
  mutate(true_relat_pub = factor(x = true_relat_pub, levels = c("Parent-Offspring", "Full Sibling", "Half Sibling", "Unrelated")))

# FS vs. U
lwed_FS_U_logls <- extract_logls(lwed_Qs,
                            numer = c(FS = 1),
                            denom = c(U = 1))
```

**SIMP**

```{r}
# log likelihood ratios for the hypothesis of PO vs Unrelated
simp_PO_U_logls <- extract_logls(simp_Qs,
                                 numer = c(PO = 1),
                                 denom = c(U = 1)
                                 ) %>%
  mutate(true_relat_pub = case_when(true_relat == "FS" ~ "Full Sibling",
                                    true_relat == "HS" ~ "Half Sibling",
                                    true_relat == "PO" ~ "Parent-Offspring",
                                    true_relat == "U" ~ "Unrelated")) %>% 
  mutate(true_relat_pub = factor(x = true_relat_pub, levels = c("Parent-Offspring", "Full Sibling", "Half Sibling", "Unrelated")))

# FS vs. U
simp_FS_U_logls <- extract_logls(simp_Qs,
                            numer = c(FS = 1),
                            denom = c(U = 1))
```

### Plots

Plot the distribution of those logl ratios for each of the different
true relationships --

**LWED**

```{r}
# all comparisons
ckmrSim_data_forFig <- lwed_PO_U_logls %>%
  mutate(species = "LWED") %>%
  rbind(., simp_PO_U_logls %>% mutate(species = "SIMP"))

ckmrSim_fig <- ggplot(ckmrSim_data_forFig,
                      aes(x = logl_ratio,
                          fill = true_relat_pub)) +
  geom_density(alpha = 0.25) +
  theme_bw(base_size = 14) +
  #scale_x_continuous(limits = c(-60, 30), breaks = seq(-60, 30, by = 10)) +
  #scale_y_continuous(limits = c(0, 0.06), breaks = seq(0, 0.6, by = 0.2)) +
  theme(legend.position = c(0.15, 0.7),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 10),
        legend.key.size = unit(1, "line"),
        panel.grid = element_blank()) + 
  labs(x = "Log likelihood ratio", y = "Density", fill = "True Relationship") +
  facet_grid(rows = vars(species))

ckmrSim_fig


# just to PO vs. U pairs
ggplot(lwed_PO_U_logls %>% filter(true_relat %in% c("PO", "U")),
            aes(x = logl_ratio, fill = true_relat_pub)) +
  geom_density(alpha = 0.25) +
  theme_bw(base_size = 14) +
  #scale_x_continuous(limits = c(-60, 30), breaks = seq(-60, 30, by = 10)) +
  #scale_y_continuous(limits = c(0, 0.06), breaks = seq(0, 0.6, by = 0.2)) +
  theme(legend.position = c(0.15, 0.7),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 10),
        legend.key.size = unit(1, "line"),
        panel.grid = element_blank()) + 
  labs(title = "LWED", x = "Log likelihood ratio", y = "Density", fill = "True Relationship")
```

**SIMP**

```{r}
# all comparisons
ggplot(simp_PO_U_logls, aes(x = logl_ratio, fill = true_relat_pub)) +
  geom_density(alpha = 0.25) +
  theme_bw(base_size = 14) +
  #scale_x_continuous(limits = c(-60, 30), breaks = seq(-60, 30, by = 10)) +
  #scale_y_continuous(limits = c(0, 0.06), breaks = seq(0, 0.6, by = 0.2)) +
  theme(legend.position = c(0.15, 0.7),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 10),
        legend.key.size = unit(1, "line"),
        panel.grid = element_blank()) + 
  labs(title = "SIMP", x = "Log likelihood ratio", y = "Density", fill = "True Relationship")

# just to PO vs. U pairs
ggplot(simp_PO_U_logls %>% filter(true_relat %in% c("PO", "U")),
            aes(x = logl_ratio, fill = true_relat)) +
  geom_density(alpha = 0.25) +
  theme_bw(base_size = 14) +
  #scale_x_continuous(limits = c(-60, 30), breaks = seq(-60, 30, by = 10)) +
  #scale_y_continuous(limits = c(0, 0.06), breaks = seq(0, 0.6, by = 0.2)) +
  theme(legend.position = c(0.15, 0.7),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 10),
        legend.key.size = unit(1, "line"),
        panel.grid = element_blank()) + 
  labs(title = "SIMP", x = "Log likelihood ratio", y = "Density", fill = "True Relationship")
```

## False neg/pos rates

### Run mc_sample_simple()

mc_sample_simple() lets you estimate false positive rates and false
negative rates from the simulated values in an object of class Qij. The
function is configured so that it will automatically compute both
regular or "vanilla" Monte Carlo estimates of these probabilities, and
also importance sampling (IS) Monte Carlo estimates, which are
particularly good for estimating very small probabilities.

**estimate false-pos rates when true relationship is U but we're looking
for PO pairs (uses importance sampling by default)**

```{r}
# LWED
lwed_PO_is <- mc_sample_simple(lwed_Qs,
                               nu = "PO",
                               de = "U")
lwed_PO_is

# SIMP
simp_PO_is <- mc_sample_simple(simp_Qs,
                               nu = "PO",
                               de = "U")
simp_PO_is
```

lwed_PO_is shows false positive rates (FPR) at \~0.006 and smaller when
the false negative rate is 0.01 and greater

simp_PO_is shows false positive rates (FPR) at \~0.01 and smaller when
the false negative rate is 0.01 and greater

**find optimal FPR**

Anderson's recommendation for being confident about not erroneously
identifying unrelated individuals as related pairs is to require that
the FPR be about 10 to 100 times smaller than the reciprocal of the
number of comparisons--

For FRANz_run1, I'm not including any pedigree info so FRANz will be
comparing every individual to every other individual. The number of
pairs being compared can thus be calculated as follows:

```{r}
countComparisons <- function(n) {
  totalComparisons = (n*(n-1))/2
  return(totalComparisons)
}

lwed_compCount <- countComparisons(72) # 2556 LWED
simp_compCount <- countComparisons(55) # 1485 SIMP

# now calculate max FPR to shoot for--
lwed_fprGoal <- 0.1*lwed_compCount^(-1) # 0.00003912363
simp_fprGoal <- 0.1*simp_compCount^(-1) # 0.00006734007
```

Let's look back at our FNR/FPR estimates--

```{r}
# LWED
lwed_PO_is_test <- mc_sample_simple(lwed_Qs,
                               nu = "PO",
                               de = "U",
                               lambda_stars = seq(-2, 16, by = 0.1))
lwed_PO_is_test

# logl ratio that gives largest FPR smaller than fprGoal = 8.1
lwed_PO_is_test %>%
  filter(FPR > lwed_fprGoal) %>%
  arrange(FPR) %>%
  dplyr::slice(1)

# SIMP
simp_PO_is_test <- mc_sample_simple(simp_Qs,
                               nu = "PO",
                               de = "U",
                               lambda_stars = seq(-2, 16, by = 0.1))
simp_PO_is_test

# logl ratio that gives largest FPR smaller than fprGoal = 7.5
simp_PO_is_test %>%
  filter(FPR > simp_fprGoal) %>%
  arrange(FPR) %>%
  dplyr::slice(1)
```

The optimal logl ratio cutoffs for LWED = 8.1 and SIMP = 7.5 -- both,
however, come with rather high FNR values (0.529 and 0.591,
respectively). So these logl cutoffs would give us decent confidence in
our assignments while having a higher chance of ditching true PO pairs.

------------------------------------------------------------------------

Alternatively, we could choose a logl cutoff that optimizes the
trade-off between false negative and false positive errors by
identifying the logl ratio (LOD) that maximizes the product of the two
(following McPhee et al. 2014;
<https://github.com/mvmcphee/Auke_sockeye_F1_RP/blob/main/scripts/SAUKE_F1_Rscript_01_CKMRsim.R>)

```{r}
# LWED
lwed_PO_is_test2 <- mc_sample_simple(lwed_Qs,
                                 nu = "PO",
                                 de = "U",
                                 lambda_stars = seq(-2, 16, by = 0.1))
lwed_PO_is_test2

ggplot(data = lwed_PO_is_test2) +
  (aes(x = Lambda_star, y = (FNR*FPR))) +
  geom_point() +
  scale_x_continuous(limits = c(-2, 16), breaks = seq(-2, 16, by = 1)) +
  theme_Publication()

# value that maximizes FNR*FPR = 3 (fnr 0.0535; fpr 0.00270)
lwed_PO_is_test2 %>%
  mutate(fnrFPR = FNR * FPR) %>%
  relocate(fnrFPR) %>%
  arrange(desc(fnrFPR)) %>%
  dplyr::slice(1)

# SIMP
simp_PO_is_test2 <- mc_sample_simple(simp_Qs,
                                 nu = "PO",
                                 de = "U",
                                 lambda_stars = seq(-2, 16, by = 0.1))
simp_PO_is_test2

ggplot(data = simp_PO_is_test2) +
  (aes(x = Lambda_star, y = (FNR*FPR))) +
  geom_point() +
  scale_x_continuous(limits = c(-2, 16), breaks = seq(-2, 16, by = 1)) +
  theme_Publication()

# value that maximizes FNR*FPR = 2.3
simp_PO_is_test2 %>%
  mutate(fnrFPR = FNR * FPR) %>%
  relocate(fnrFPR) %>%
  arrange(desc(fnrFPR)) %>%
  dplyr::slice(1)
```

# 5 Summary stats

Print basic info

```{r}
lwed_gen
```

Print number of alleles per locus

```{r}
table(lwed_gen$loc.fac)
```

Print sample size for each site (NA)

```{r}
summary(lwed_gen$pop)
```

Print mean allelic richness per site across all loci (NA)

```{r}
allelic.richness(genind2hierfstat(lwed_gen))$Ar %>%
  apply(MARGIN = 2, FUN = mean) %>%
  round(digits = 3)
```

INBREEDING COEFFICIENT (Fis)

```{r}

```

# X Nf, Nm

## Some background

-   Jolly-Seber (JS) model - focused on estimating abundance parameters
    (e.g., pop growth, recruitment, and abundance)
    -   explicit definitions of process by which unmarked animals are
        newly captured and marked; assumptions about process allows est
        of recruitment and pop sizes
    -   assumed that unmarked indivs in pop have same prob of capture as
        marked indiv in pop (i.e., newly captured unmarked animals are a
        random sample of all unmarked animals in pop)
-   Cormack-Jolly-Seber (CJS) models - focused on estimating survival
    rates (but not abundance)
    -   no assumptions made about how newly marked animals are obtained
    -   subsequent process of recovering marked animals in CJS is
        conditional upon animal being released alive at first encounter;
        survival and catchability refer only to those marked animals

Based on Iijima (2020), I want:

-   **capture-recapture model for open population**, where an "open
    population" is one in which the number of individuals can change via
    natural mortality, birth, and migration during the study period -
    may also benefit from **robust design** where multiple surveys are
    conducted under a constant condition
-   1st hierarchy of sampling in which variables can change across
    sampling periods
-   2nd hierarchy of sampling in which things should remain relatively
    stable

And based on Sandercock & Murray (2020), I may also want:

-   a multistate model, in which detections are coded as dynamic
    categorical states that potentially change b/t consecutive occasions

**Resources:** - James Paterson has a
blog(<https://jamesepaterson.github.io/jamespatersonblog/2020-07-26_jolly_seber_models.html>)
with a lot of really great info on how to use mark & recapture data

## About the model

We can estimate *N* for FRANz by estimating abundance via the POPAN
formulation (Schwarz & Arnason 1996) of the Jolly-Seber mark-recapture
model. Note that Jolly-Seber models make the following assumptions (from
Schwarz & Arnason 2021 in Program MARK book):

1)  indivs retain their tags throughout experiment
2)  tags are read properly
3)  sampling is instantaneous
4)  survival probabilities are same for all indivs (marked and unmarked)
    b/t each pair of sampling occasions (homogeneous survival)
5)  catchability is same for all indivs (marked and unmarked) at each
    sampling event (homogeneous catchability) ((most crucial assumption
    for JS models))
6)  study area is constant - if study area changes over time, then pop
    size may change w/changing size of study area

The POPAN formulation modifies the parameterization of JS slightly by
postulating 1) the existence of a *super-population* containing all of
the indivs that would ever be born to the population and 2) parameter
b_i representing the probability that an indiv from the hypothetical
super-population would enter the population b/t occasion i and i + 1.

## Create capture histories

Just using 2009-2019 data to align w/geno data

```{r}
capHist <- capData %>%
  # filter to 2009-2019 only
  arrange(rowID) %>%
  filter(as.numeric(rowID) < 614) %>%
  
  select(animalID, captureDate) %>%
  filter(animalID != "UNK") %>%
  mutate(
    captureYear = as.numeric(str_sub(captureDate, 1, 4))
  ) %>%
  select(-captureDate) %>%
  mutate(
    detect = 1
  ) %>%
  
  # format as capture history (notes below are from James Paterson blog)
  
  # remove duplicates, which may occur when individuals are caught multiple times in an event
  # For example, your event may be a year and an individual may be caught multiple times in a year.
  distinct() %>%
  
  # spread out data. The fill = 0 adds rows for combinations of id and event where individuals were not observerd
  spread(captureYear, detect, fill = 0) %>% 
  
  # For every individual....
  group_by(animalID) %>%
  # Paste together 0's and 1's
  # Unite is similar to paste. Here we are pasting the strings together from the second column (first capture event)
  # to the last capture event ("tail(names(.),1)").
  # we don't want any characters separating 0's and 1's, so we use: sep = ""
  unite("ch", 2:tail(names(.),1), sep = "") %>%
  
  # add sex (as factor)
  merge(., capData[, c("animalID", "sex", "species")], by = "animalID") %>%
  distinct() %>%
  mutate(
    sex = as.factor(sex)
  )

# species/sex subsets
capHist_lwedF <- capHist %>%
  filter(species == "LWED") %>%
  filter(sex == "F")
capHist_lwedM <- capHist %>%
  filter(species == "LWED") %>%
  filter(sex == "M")

capHist_simpF <- capHist %>%
  filter(species == "SIMP") %>%
  filter(sex == "F")
capHist_simpM <- capHist %>%
  filter(species == "SIMP") %>%
  filter(sex == "M")
```

## Run model

**Load 'marked'**

Using package 'marked' (make sure RMark is detached first! too many
overlapping function names)

```{r}
#detach("package:RMark", unload=TRUE)
library(marked)
```

**1. Process data**

```{r}
# lwed
lwedF.js.proc <- process.data(capHist_lwedF,
                              model = "JS")
lwedM.js.proc <- process.data(capHist_lwedM,
                              model = "JS")

# simp
simpF.js.proc <- process.data(capHist_simpF,
                              model = "JS")
simpM.js.proc <- process.data(capHist_simpM,
                              model = "JS")
```

**2. Make design data**

```{r}
lwedF.js.ddl <- make.design.data(lwedF.js.proc)
lwedM.js.ddl <- make.design.data(lwedM.js.proc)

simpF.js.ddl <- make.design.data(simpF.js.proc)
simpM.js.ddl <- make.design.data(simpM.js.proc)
```

**3. Fit models**

```{r}
fit.js.models <- function(procData, designData) {
  # Phi formulas
  Phi.dot <- list(formula=~1)
  Phi.time <- list(formula=~time)
  # p formulas
  p.dot <- list(formula=~1)
  # pent formulas. pent estimates MUST SUM to 1 (for each group).
  # This is constained using a Multinomial Logit link
  pent.time <- list(formula=~time)
  pent.dot <- list(formula=~1)
  # Nsuper formulas. Don't confuse "N" from model with predicted population size
  N.dot <- list(formula=~1)
  cml <- create.model.list(c("Phi","p", "pent", "N"))
  results <- crm.wrapper(cml,
                         data = procData,
                         ddl = designData,
                         external = FALSE,
                         accumulate = FALSE,
                         hessian = TRUE)
  
  return(results)
}
```

**3. Run function**

```{r}
# run function
lwedF.js.models <- fit.js.models(lwedF.js.proc, lwedF.js.ddl)
lwedM.js.models <- fit.js.models(lwedM.js.proc, lwedM.js.ddl)

simpF.js.models <- fit.js.models(simpF.js.proc, simpF.js.ddl)
simpM.js.models <- fit.js.models(simpM.js.proc, simpM.js.ddl)

# view models
lwedF.js.models
lwedM.js.models

simpF.js.models
simpM.js.models
```

**4. Estimate real parameters**

We can look at the top model estimates:

```{r}
# pull top model
lwedF.js.models[[1]]
lwedM.js.models[[1]]

simpF.js.models[[1]]
simpM.js.models[[1]]
```

But these estimates \^\^ are not on probability scale (or in individuals
for N) (e.g. Phi, p on logit scale, pent on mlogit scale) - so we need
to use the *predict* function to estimate the real parameters:

```{r}
# predict real values
lwedF.js.predicted <- predict(lwedF.js.models[[1]])
lwedM.js.predicted <- predict(lwedM.js.models[[1]])

simpF.js.predicted <- predict(simpF.js.models[[1]])
simpM.js.predicted <- predict(simpM.js.models[[1]])

# look at predictions of real parameters
lwedF.js.predicted
lwedM.js.predicted

simpF.js.predicted
simpM.js.predicted

# what these parameters are saying (just lwedF for an example):
lwedF.js.predicted$Phi$estimate # 0.64 survival b/t capture events
lwedF.js.predicted$p$estimate # 0.84 detection prob
lwedF.js.predicted$pent # 0.09 prob of entry for each capture event
lwedF.js.predicted$N$estimate # ~ 5 unmarked indiv
```

## Estimate pop size

(from Paterson blog): There is no direct estimate of population size in
the model. The estimate of "N" in the model output is for the number of
unmarked individuals in the superpopulation.

To estimate population size, we can derive it using the model estimates:

```         
N_t+1 = (N_t * phi_t) + (N_super * pent_t)
```

In words, this equation says that the **population size** is the
**previous population size** times the **survival rate** (to calculate
how many individuals survive) plus the **number of new individuals**
from births and immigration.

To get the inital population size, we estimate pent_1 by the constraint:

```         
pent_1 = ` - SUM(pent_2..t)
```

Abundance (N) is derived from the estimated parameters We will estimate
population size at each time by making a dataframe of estimates and
calculating N We will use the predicted estimates from the
top-performing model (in this case: "dipper.js.predicted")

```{r}
# number of capture events
n_capEvents <- nchar("00011000000")

# number of survival estimates
nrow(lwedF.js.predicted$pent)

# number of captured indivs
lwedF_nCap <- nrow(capData_2009to2019_firstEntry[capData_2009to2019_firstEntry$species == "LWED" & capData_2009to2019_firstEntry$sex == "F",])
lwedM_nCap <- nrow(capData_2009to2019_firstEntry[capData_2009to2019_firstEntry$species == "LWED" & capData_2009to2019_firstEntry$sex == "M",])

simpF_nCap <- nrow(capData_2009to2019_firstEntry[capData_2009to2019_firstEntry$species == "SIMP" & capData_2009to2019_firstEntry$sex == "F",])
simpM_nCap <- nrow(capData_2009to2019_firstEntry[capData_2009to2019_firstEntry$species == "SIMP" & capData_2009to2019_firstEntry$sex == "M",])

lwedF_N.derived <- data.frame(
  # 11 capture events
  occ = c(1:11),
  # 10 survival estimates all the same
  Phi = c(rep(lwedF.js.predicted$Phi$estimate, 10), NA),
  # Nsuper estimate + number of marked animals
  Nsuper = rep(lwedF.js.predicted$N$estimate + lwedF_nCap, n_capEvents),
  # sum of all pent must be 1
  pent = c(1-sum(lwedF.js.predicted$pent$estimate),
           lwedF.js.predicted$pent$estimate))

# Set-up empty vector for calculating N
lwedF_N.derived$N <- NA

# The inital population size (N[1]) = Nsuper * (1 - sum(all other pent estimates))
# This is because of the link function for estimating pent.
# The sum of all pent parameters MUST equal 1 (therefore, one less must be estimated)
lwedF_N.derived$N[1] <- (lwedF_N.derived$Nsuper[1] * lwedF_N.derived$pent[1])

# Subsequent population sizes are estimated by calculating surviving individuals (N[t-1] * Phi[t]), and
# Adding new births (Nsuper * pent[t])
for(i in 2:nrow(lwedF_N.derived)){
  lwedF_N.derived$N[i] <- (lwedF_N.derived$N[i-1]*lwedF_N.derived$Phi[i-1]) + (lwedF_N.derived$Nsuper[i] * lwedF_N.derived$pent[i])
}

# Look at what we did
lwedF_N.derived
```

\^\^I'm pretty sure what this is giving is the number of indivs PER YEAR
vs. Nsuper = number of indivs across ALL YEARS in the pop

We have now derived the population size (N) using the estimates form our
Jolly-Seber model. But, how do we estimate standard error and confidence
intervals?

```{r}
# and get estimates of unmarked individuals
lwedF_js_unmarkEst <- round(lwedF.js.predicted$N$estimate)
lwedM_js_unmarkEst <- round(lwedM.js.predicted$N$estimate)

simpF_js_unmarkEst <- round(simpF.js.predicted$N$estimate)
simpM_js_unmarkEst <- round(simpM.js.predicted$N$estimate)
```

## Assign Nsuper

Using output from JS models for estimates of unmarked females & males
for each species added to capData numbers:

```{r}
# number of captured indivs
lwed_n_capF <- nrow(capData_2009to2019_firstEntry[capData_2009to2019_firstEntry$species == "LWED" & capData_2009to2019_firstEntry$sex == "F",])
lwed_n_capM <- nrow(capData_2009to2019_firstEntry[capData_2009to2019_firstEntry$species == "LWED" & capData_2009to2019_firstEntry$sex == "M",])

simp_n_capF <- nrow(capData_2009to2019_firstEntry[capData_2009to2019_firstEntry$species == "SIMP" & capData_2009to2019_firstEntry$sex == "F",])
simp_n_capM <- nrow(capData_2009to2019_firstEntry[capData_2009to2019_firstEntry$species == "SIMP" & capData_2009to2019_firstEntry$sex == "M",])

lwed_n_capF # 68
lwed_n_capM # 77

simp_n_capF # 48
simp_n_capM # 54

# calculate super-population totals
lwed_Nf <- lwed_n_capF +
  lwedF_js_unmarkEst
lwed_Nm <- lwed_n_capM +
  lwedM_js_unmarkEst

simp_Nf <- simp_n_capF +
  simpF_js_unmarkEst
simp_Nm <- simp_n_capM +
  simpM_js_unmarkEst

lwed_Nf # 73
lwed_Nm # 82

simp_Nf # 52
simp_Nm # 60
```

# FRANz

Following Shedd et al GitHub
1_RRS_Hogan_13_15.Rmd<https://github.com/krshedd/Relative-fitness-of-Pink-Salmon/blob/main/analysis/1_RRS_Hogan_13_15.Rmd>

## Create input files

**Subset genos to samples with \>= 70% genotyping success**

### Prep files

**LWED**

```{r}
lwed_genos_franz_temp <- lwed_genos %>%
  column_to_rownames("sampleID") %>%
  select(-c(animalID, sex, birthYear_est, pop, captureYear))

lwed_genos_franz <- names(lwed_genos_franz_temp) %>%
  map_dfc(~ lwed_genos_franz_temp %>%
            select(all_of(.x)) %>%
            separate(.x,
                     into = paste0(.x, c("a", "b")),
                     sep = ",")
          ) %>%
  
  # add sampleID_franz, sex, & animalID (animalID is just for merging birthYear)
  rownames_to_column("sampleID") %>%
  merge(., sampleRef[, c("sampleID", "sampleID_franz", "animalID", "sex")], by = "sampleID") %>%
  relocate(c(sampleID_franz, sex)) %>%
  
  # add birthYear
  merge(., birthAssignments[, c("animalID", "birthYear_est")], by = "animalID", all.x = T) %>%
  dplyr::rename("birthYear" = "birthYear_est") %>%
  relocate(birthYear, .before = "sex") %>%
  
  # add deathYear
  merge(., deathAssignments[, c("animalID", "deathYear_est")], by = "animalID", all.x = T) %>%
  dplyr::rename("deathYear" = "deathYear_est") %>%
  relocate(deathYear, .before = "sex") %>%
  
  # replace allele letters with numbers (required by FRANz)
  mutate_at(vars(starts_with(c("INDID", "LWED", "SIMP"))), ~ str_replace(., "A", "4")) %>%
  mutate_at(vars(starts_with(c("INDID", "LWED", "SIMP"))), ~ str_replace(., "T", "7")) %>%
  mutate_at(vars(starts_with(c("INDID", "LWED", "SIMP"))), ~ str_replace(., "C", "3")) %>%
  mutate_at(vars(starts_with(c("INDID", "LWED", "SIMP"))), ~ str_replace(., "G", "6")) %>%
  arrange(sampleID) %>%
  select(-animalID, -sampleID)

# export
write.csv(lwed_genos_franz, "./DISSERTATION/ch2_demography/00_data/lwed_genos_franz.csv", row.names = F)
```

**SIMP**

```{r}
simp_genos_franz_temp <- simp_genos %>%
  column_to_rownames("sampleID") %>%
  select(-c(animalID, sex, birthYear_est, pop, captureYear))

simp_genos_franz <- names(simp_genos_franz_temp) %>%
  map_dfc(~ simp_genos_franz_temp %>%
            select(all_of(.x)) %>%
            separate(.x,
                     into = paste0(.x, c("a", "b")),
                     sep = ",")
          ) %>%
  
  # add sampleID_franz, sex, & animalID (animalID is just for merging birthYear)
  rownames_to_column("sampleID") %>%
  merge(., sampleRef[, c("sampleID", "sampleID_franz", "animalID", "sex")], by = "sampleID") %>%
  relocate(c(sampleID_franz, sex)) %>%
  
  # add birthYear
  merge(., birthAssignments[, c("animalID", "birthYear_est")], by = "animalID", all.x = T) %>%
  dplyr::rename("birthYear" = "birthYear_est") %>%
  relocate(birthYear, .before = "sex") %>%
  
  # add deathYear
  merge(., deathAssignments[, c("animalID", "deathYear_est")], by = "animalID", all.x = T) %>%
  dplyr::rename("deathYear" = "deathYear_est") %>%
  relocate(deathYear, .before = "sex") %>%
  
  # replace allele letters with numbers (required by FRANz)
  mutate_at(vars(starts_with(c("INDID", "LWED", "SIMP"))), ~ str_replace(., "A", "4")) %>%
  mutate_at(vars(starts_with(c("INDID", "LWED", "SIMP"))), ~ str_replace(., "T", "7")) %>%
  mutate_at(vars(starts_with(c("INDID", "LWED", "SIMP"))), ~ str_replace(., "C", "3")) %>%
  mutate_at(vars(starts_with(c("INDID", "LWED", "SIMP"))), ~ str_replace(., "G", "6")) %>%
  arrange(sampleID) %>%
  select(-animalID, -sampleID)

# export
write.csv(simp_genos_franz, "./DISSERTATION/ch2_demography/00_data/simp_genos_franz.csv", row.names = F)
```

### .csv to .dat

FRANz provides a perl script (csv.pl) to convert .csv files to .dat
files. The manual says that you can also go to a GUI at
<http://legacy.bioinf.uni-leipzig.de/Software/FRANz/CSV_Import.html>,
but I haven't gotten it to work (it just loads forever).

FRANz has examples of how to use csv.pl in their "man" file; I've edited
their example slightly so that the output automatically creates a new
file.

NOTE that column IDs start with 0.

How to put output into file:
<https://askubuntu.com/questions/420981/how-do-i-save-terminal-output-to-a-file>

**NOTE** - use csv_modified.pl; line 164 has an error in the original
script -- original has "\$death_col = $data->[$death_col];", but should
be "\$death = $data->[$death_col];"

```{r}
system2("perl",
        args="./project_scripts/franz/csv_modified.pl --in ./DISSERTATION/ch2_demography/00_data/lwed_genos_franz.csv --alleles_per_col 1 --has_header --missing_allele 'NA' --birth_col 1 --death_col 2 --sex_col 3 --data_col 4 > ./DISSERTATION/ch2_demography/00_data/lwed_genos_franz.dat")

system2("perl",
        args="./project_scripts/franz/csv_modified.pl --in ./DISSERTATION/ch2_demography/00_data/simp_genos_franz.csv --alleles_per_col 1 --has_header --missing_allele 'NA' --birth_col 1 --death_col 2 --sex_col 3 --data_col 4 > ./DISSERTATION/ch2_demography/00_data/simp_genos_franz.dat")
```

## Sample counts

Quick overview of who all is included:

LWED 1 F 38 (out of Nf = 73) 2 M 34 (out of Nm = 82)

SIMP 1 F 27 (out of Nf = 52) 2 M 28 (out of Nm = 60)

```{r}
lwed_genos_franz %>%
  group_by(sex) %>%
  summarise(count = n())

sum(!is.na(lwed_genos_franz$birthYear)) # 29
sum(is.na(lwed_genos_franz$birthYear)) # 43


simp_genos_franz %>%
  group_by(sex) %>%
  summarise(count = n())

sum(!is.na(simp_genos_franz$birthYear)) # 28
sum(is.na(simp_genos_franz$birthYear)) # 27
```

## Assign Nf & Nm

From FRANz manual: *N* is the number of candidate female (*Nf*) and male
(*Nm*) parents in the population, equal to the sum of the average number
of sampled (*n*) and unsampled (*N-n*) breeding females and males in the
population (respectively). If unset, then estimated jointly with the
pedigree. For parentage inference N=Nf=Nm (NOT Nf+Nm), but one can also
specify Nf and Nm instead of N if these numbers differ. If N isnot
known, use Nmax instead.

The FRANz manual suggests that if a good estimate of Nf/Nm is possible,
specifying these variables is preferable to Nfmax/Nmmax (the maximum
number of candidate female/male parents in the population). As such, I'm
using the sex-specific abundance estimates calculated in
ch2_demography.Rmd (currently in ch2_relatedness.Rmd tho, need to move
it) based on capData records.

```         
lwed_Nf = 73
lwed_Nm = 82

simp_Nf = 52
simp_Nm = 60
```

## franz_run1

### Run script

```{bash}
# lwed
cd /home/rachelvoyt/Documents/UT-Grad/Development/repos/tamGenetics_primatesPeru/DISSERTATION/ch2_demography/02_results/lwed/franz_run1

FRANz --femrepro 2:30 --malerepro 2:30 --Nf 73 --Nm 82 --fullsibtest --fullsibparental --updatefreqs --poutformat 2 ./../../../00_data/lwed_genos_franz.dat

# simp
cd ./../../simp/franz_run1

FRANz --femrepro 2:30 --malerepro 2:30 --Nf 52 --Nm 60 --fullsibtest --fullsibparental --updatefreqs --poutformat 2 ./../../../00_data/simp_genos_franz.dat
```

FRANz parameters are defined as follows: --femrepro and --malerepro
specify the age ranges in which females and males can reproduce

--Nf and --Nm specify the number of candidate female and male parents in
the population, which is the sum of the average number of sampled (n)
and unsampled (N-n) breeding females/males in the population. The FRANz
manual suggests that if a good estimate of Nf/Nm is possible, specifying
these variables is preferable to Nfmax/Nmmax (the maximum number of
candidate female/male parents in the population)

--fullsibtest specifies that FRANz should run the fullsib heuristic (as
described in Riester et al. 2009)

--fullsibparental specifies that FRANz should also detect fullsibs in
the parent generation

--updatefreqs specifies that FRANz should update allele frequencies
using MCMC sampling

--poutformat 2 specifies that parentage output files should include all
with positive LOD (vs. just the most likely parentages)

**NOTE** that FRANz automatically sets --mintyped loci to (1+loci)/2;
this also defines the min number of common typed loci for a pair of
indivs. Other defaults include:

-   typingerror = 0.01

### Assess results

#### Cumulative exclusion probability

LWED: Cumulative exclusion probability when 1 to 7 fullsibs are
genotyped First Parent : 0.9971977 0.9998257 0.9999910 0.9999988
0.9999994 0.9999995 0.9999995 Second Parent : 0.9999621 0.9999996
1.0000000 1.0000000 1.0000000 1.0000000 1.0000000 Parent Pair :
0.9999999 1.0000000 1.0000000 1.0000000 1.0000000 1.0000000 1.0000000

SIMP: Cumulative exclusion probability when 1 to 7 fullsibs are
genotyped First Parent : 0.9925770 0.9992520 0.9999363 0.9999884
0.9999937 0.9999941 0.9999941 Second Parent : 0.9998635 0.9999972
0.9999999 1.0000000 1.0000000 1.0000000 1.0000000 Parent Pair :
0.9999994 1.0000000 1.0000000 1.0000000 1.0000000 1.0000000 1.0000000

#### PO assignments

FRANz colnames are not my favorite - set new ones:

```{r}
colnames_parentage <- c("offspring",
                        "lociTyped_offspring",
                        "parent1",
                        "lociTyped_parent1",
                        "parent2",
                        "lociTyped_parent2",
                        "LOD",
                        "posterior",
                        "commonLociTyped",
                        "mismatches",
                        "n_f",
                        "n_m",
                        "pairLOD_parent1",
                        "pairLOD_parent2",
                        "posterior_parent1",
                        "posterior_parent2",
                        "parentage_mlPedigree")
```

**PO assignments**

```{r}
lwed_po_run1_file <- read_csv("./DISSERTATION/ch2_demography/02_results/lwed/franz_run1/parentage.csv") %>%
  separate('Posterior Parent 2', c("pp2", "mlPed"), sep = ",") %>%
  `colnames<-`(colnames_parentage) %>%
  mutate(
    rowID = row_number(),
    poType = case_when(
      !is.na(parent1) & !is.na(parent2) ~ "triple",
      !is.na(parent1) & is.na(parent2) ~ "dyad",
      is.na(parent1) & !is.na(parent2) ~ "dyad",
      .default = NA
      ),
    poType = case_when(
      !is.na(parentage_mlPedigree) ~ str_c(poType, "_ml"),
      .default = poType
      )
    ) %>%
  relocate(c(rowID, poType))

simp_po_run1_file <- read_csv("./DISSERTATION/ch2_demography/02_results/simp/franz_run1/parentage.csv") %>%
  separate('Posterior Parent 2', c("pp2", "mlPed"), sep = ",") %>%
  `colnames<-`(colnames_parentage) %>%
  mutate(franzRun = "run1") %>%
  mutate(
    rowID = row_number(),
    poType = case_when(
      !is.na(parent1) & !is.na(parent2) ~ "triple",
      !is.na(parent1) & is.na(parent2) ~ "dyad",
      is.na(parent1) & !is.na(parent2) ~ "dyad",
      .default = NA
      ),
    poType = case_when(
      !is.na(parentage_mlPedigree) ~ str_c(poType, "_ml"),
      .default = poType
      )
    ) %>%
  relocate(c(rowID, poType))
```

**Take a peek at parentage posterior probabilities**

Definitely a lot on the low side, but this makes sense since I opted to
include all pairs w/a positive LOD score in the output (vs. most likely
PO pairs)

```{r}
lwed_po_run1_file %>%
  filter(!is.na(parent1)) %>%
  ggplot(aes(x = posterior)) +
  geom_histogram(breaks = seq(0, 1, 0.01)) +
  ggtitle("Histogram of FRANz posterior probabilities for parentage assignments") +
  theme_bw()

simp_po_run1_file %>%
  filter(!is.na(parent1)) %>%
  ggplot(aes(x = posterior)) +
  geom_histogram(breaks = seq(0, 1, 0.01)) +
  ggtitle("Histogram of FRANz posterior probabilities for parentage assignments") +
  theme_bw()
```

**PO checks**

```{r}
# reload genos
lwed_genos_franz <- read.csv("./DISSERTATION/ch2_demography/00_data/lwed_genos_franz.csv")
simp_genos_franz <- read.csv("./DISSERTATION/ch2_demography/00_data/simp_genos_franz.csv")

# LWED
lwed_po_run1 <- lwed_po_run1_file %>%
  #filter(posterior > 0.9) %>%
  select(rowID, poType, offspring, parent1, parent2) %>%
  pivot_longer(-c(rowID, poType, offspring),
               names_to = "parentID",
               values_to = "parent") %>%
  na.omit() %>%
  
  # add birthYear and sex
  merge(., lwed_genos_franz[, c("sampleID_franz", "birthYear", "sex")], by.x = "parent", by.y = "sampleID_franz", all.x = T) %>%
  merge(., lwed_genos_franz[, c("sampleID_franz", "birthYear", "sex")], by.x = "offspring", by.y = "sampleID_franz", all.x = T, suffixes = c("_par", "_off")) %>%
  
  # add groups
  mutate(
    animalID_par = gsub(".*_([0-9]+)$", "\\1", parent),
    animalID_off = gsub(".*_([0-9]+)$", "\\1", offspring)
    ) %>%
  
  mutate(
    temp = as.numeric(birthYear_off) - 1
    ) %>%
  merge(., capData_2009to2019[, c("animalID", "captureYear", "groupName")], by.x = c("animalID_par", "temp"), by.y = c("animalID", "captureYear"), all.x = T) %>%
  select(-temp) %>%
  
  merge(., capData_2009to2019[, c("animalID", "captureYear", "groupName")], by.x = c("animalID_off", "birthYear_off"), by.y = c("animalID", "captureYear"), all.x = T, suffixes = c("_par", "_off")) %>%
  
  # same groupName?
  mutate(
    groupMatch = groupName_par == groupName_off
  ) %>%
  
  # add parity for F
  merge(., parityAssigments_firstParous[, c("animalID", "parityYear")], by.x = "animalID_par", by.y = "animalID", all.x = T) %>%
  mutate(
    diff_birth.parity = as.numeric(birthYear_off) - as.numeric(parityYear)
  ) %>%
  
  relocate(parentID, poType, parent, offspring, animalID_par, animalID_off, sex_par, sex_off, diff_birth.parity, groupMatch, birthYear_par, parityYear, birthYear_off, groupName_par, groupName_off)


temp <- lwed_po_run1 %>%
  filter(sex_par == "F")



# SIMP
simp_po_run1 <- simp_po_run1_file %>%
  filter(posterior > 0.9) %>%
  select(offspring, parent1, parent2) %>%
  pivot_longer(-offspring,
               names_to = "parentID",
               values_to = "parent") %>%
  select(-parentID) %>%
  na.omit() %>%
  
  # add birthYear and sex
  merge(., simp_genos_franz[, c("sampleID_franz", "birthYear", "sex")], by.x = "parent", by.y = "sampleID_franz", all.x = T) %>%
  merge(., simp_genos_franz[, c("sampleID_franz", "birthYear", "sex")], by.x = "offspring", by.y = "sampleID_franz", all.x = T, suffixes = c("_par", "_off")) %>%
  
  # add groups
  mutate(
    animalID_par = gsub(".*_([0-9]+)$", "\\1", parent),
    animalID_off = gsub(".*_([0-9]+)$", "\\1", offspring)
    ) %>%
  
  mutate(
    temp = as.numeric(birthYear_off) - 1
    ) %>%
  merge(., capData_2009to2019[, c("animalID", "captureYear", "groupName")], by.x = c("animalID_par", "temp"), by.y = c("animalID", "captureYear"), all.x = T) %>%
  select(-temp) %>%
  
  merge(., capData_2009to2019[, c("animalID", "captureYear", "groupName")], by.x = c("animalID_off", "birthYear_off"), by.y = c("animalID", "captureYear"), all.x = T, suffixes = c("_par", "_off")) %>%
  
  # same groupName?
  mutate(
    groupMatch = groupName_par == groupName_off
  ) %>%
  
  # add parity for F
  merge(., parityAssigments_firstParous[, c("animalID", "parityYear")], by.x = "animalID_par", by.y = "animalID", all.x = T) %>%
  mutate(
    diff_birth.parity = as.numeric(birthYear_off) - as.numeric(parityYear)
  ) %>%
  
  relocate(parent, offspring, animalID_par, animalID_off, sex_par, sex_off, diff_birth.parity, groupMatch, birthYear_par, parityYear, birthYear_off, groupName_par, groupName_off)
```

**PO pedigree.dat**

Just filtering by groupMatch == T; filtering by diff_birth.parity gives
too few I think

LWED: F = 5, M = 5 SIMP: F = 3, M = 1

```{r}
lwed_po_run1_forPed <- lwed_po_run1 %>%
  filter(groupMatch == TRUE) %>%
  mutate(franzPed = str_c(parent, offspring))

simp_po_run1_forPed <- simp_po_run1 %>%
  filter(groupMatch == TRUE) %>%
  mutate(franzPed = str_c(parent, offspring))

# overview of parent assignment numbers for pedigree.dat
lwed_po_run1_forPed %>%
  group_by(sex_par) %>%
  summarise(count = n())

simp_po_run1_forPed %>%
  group_by(sex_par) %>%
  summarise(count = n())
```

Reformat and export as .dat files

```{r}
nrow(lwed_genos_franz) # 72 lwed samples
nrow(simp_genos_franz) # 55 simp samples

lwed_ped_forRun2 <- c(lwed_genos_franz$sampleID_franz,
                      lwed_po_run1_forPed$franzPed) %>%
  as.data.frame() %>%
  dplyr::rename("72" = ".")

simp_ped_forRun2 <- c(simp_genos_franz$sampleID_franz,
                      simp_po_run1_forPed$franzPed) %>%
  as.data.frame() %>%
  dplyr::rename("55" = ".")

# export
write.table(lwed_ped_forRun2, file = "./DISSERTATION/ch2_demography/00_data/lwed_poPed_forRun2.dat", quote = F, row.names = F)

write.table(simp_ped_forRun2, file = "./DISSERTATION/ch2_demography/00_data/simp_poPed_forRun2.dat", quote = F, row.names = F)
```

#### SIB assignments

```{r}
lwed_sibs_run1 <- read_franzSibs("./DISSERTATION/ch2_demography/02_results/lwed/franz_run1/siblings.txt") %>%
  merge(., twinList[, c("sampleID_franz", "twinSet")], by.x = "sample1", by.y = "sampleID_franz", all.x = T) %>%
  merge(., twinList[, c("sampleID_franz", "twinSet")], by.x = "sample2", by.y = "sampleID_franz", all.x = T, suffixes = c("_1", "_2")) %>%
  mutate(
    twinSetMatch = twinSet_1 == twinSet_2
  ) %>%
  relocate(sample1)


simp_sibs_run1 <- read_franzSibs("./DISSERTATION/ch2_demography/02_results/simp/franz_run1/siblings.txt") %>%
  merge(., twinList[, c("sampleID_franz", "twinSet")], by.x = "sample1", by.y = "sampleID_franz", all.x = T) %>%
  merge(., twinList[, c("sampleID_franz", "twinSet")], by.x = "sample2", by.y = "sampleID_franz", all.x = T, suffixes = c("_1", "_2")) %>%
  mutate(
    twinSetMatch = twinSet_1 == twinSet_2
  ) %>%
  relocate(sample1)
```

## franz_run2

### Run script

```{bash}
# lwed
cd /home/rachelvoyt/Documents/UT-Grad/Development/repos/tamGenetics_primatesPeru/DISSERTATION/ch2_demography/02_results/lwed/franz_run2

FRANz --femrepro 2:30 --malerepro 2:30 --Nf 73 --Nm 82 --pedigreein ./../../../00_data/lwed_poPed_forRun2.dat --fullsibtest --fullsibparental --updatefreqs --poutformat 2 ./../../../00_data/lwed_genos_franz.dat

# simp
cd ./../../simp/franz_run2

FRANz --femrepro 2:30 --malerepro 2:30 --Nf 52 --Nm 60 --pedigreein ./../../../00_data/simp_poPed_forRun2.dat --fullsibtest --fullsibparental --updatefreqs --poutformat 2 ./../../../00_data/simp_genos_franz.dat
```

### Assess results

#### PO assignments

**PO assignments**

```{r}
lwed_po_run2_file <- read_csv("./DISSERTATION/ch2_demography/02_results/lwed/franz_run2/parentage.csv") %>%
  separate('Posterior Parent 2', c("pp2", "mlPed"), sep = ",") %>%
  `colnames<-`(colnames_parentage) %>%
  mutate(franzRun = "run2")

simp_po_run2_file <- read_csv("./DISSERTATION/ch2_demography/02_results/simp/franz_run2/parentage.csv") %>%
  separate('Posterior Parent 2', c("pp2", "mlPed"), sep = ",") %>%
  `colnames<-`(colnames_parentage) %>%
  mutate(franzRun = "run2")
```

**PO checks**

```{r}
# LWED
lwed_po_run2 <- lwed_po_run2_file %>%
  filter(posterior > 0.9) %>%
  select(offspring, parent1, parent2) %>%
  pivot_longer(-offspring,
               names_to = "parentID",
               values_to = "parent") %>%
  select(-parentID) %>%
  na.omit() %>%
  
  # add birthYear and sex
  merge(., lwed_genos_franz[, c("sampleID_franz", "birthYear", "sex")], by.x = "parent", by.y = "sampleID_franz", all.x = T) %>%
  merge(., lwed_genos_franz[, c("sampleID_franz", "birthYear", "sex")], by.x = "offspring", by.y = "sampleID_franz", all.x = T, suffixes = c("_par", "_off")) %>%
  
  # add groups
  mutate(
    animalID_par = gsub(".*_([0-9]+)$", "\\1", parent),
    animalID_off = gsub(".*_([0-9]+)$", "\\1", offspring)
    ) %>%
  
  mutate(
    temp = as.numeric(birthYear_off) - 1
    ) %>%
  merge(., capData_2009to2019[, c("animalID", "captureYear", "groupName")], by.x = c("animalID_par", "temp"), by.y = c("animalID", "captureYear"), all.x = T) %>%
  select(-temp) %>%
  
  merge(., capData_2009to2019[, c("animalID", "captureYear", "groupName")], by.x = c("animalID_off", "birthYear_off"), by.y = c("animalID", "captureYear"), all.x = T, suffixes = c("_par", "_off")) %>%
  
  # same groupName?
  mutate(
    groupMatch = groupName_par == groupName_off
  ) %>%
  
  # add parity for F
  merge(., parityAssigments_firstParous[, c("animalID", "parityYear")], by.x = "animalID_par", by.y = "animalID", all.x = T) %>%
  mutate(
    diff_birth.parity = as.numeric(birthYear_off) - as.numeric(parityYear)
  ) %>%
  
  relocate(parent, offspring, animalID_par, animalID_off, sex_par, sex_off, diff_birth.parity, groupMatch, birthYear_par, parityYear, birthYear_off, groupName_par, groupName_off)


# SIMP
simp_po_run2 <- simp_po_run2_file %>%
  filter(posterior > 0.9) %>%
  select(offspring, parent1, parent2) %>%
  pivot_longer(-offspring,
               names_to = "parentID",
               values_to = "parent") %>%
  select(-parentID) %>%
  na.omit() %>%
  
  # add birthYear and sex
  merge(., simp_genos_franz[, c("sampleID_franz", "birthYear", "sex")], by.x = "parent", by.y = "sampleID_franz", all.x = T) %>%
  merge(., simp_genos_franz[, c("sampleID_franz", "birthYear", "sex")], by.x = "offspring", by.y = "sampleID_franz", all.x = T, suffixes = c("_par", "_off")) %>%
  
  # add groups
  mutate(
    animalID_par = gsub(".*_([0-9]+)$", "\\1", parent),
    animalID_off = gsub(".*_([0-9]+)$", "\\1", offspring)
    ) %>%
  
  mutate(
    temp = as.numeric(birthYear_off) - 1
    ) %>%
  merge(., capData_2009to2019[, c("animalID", "captureYear", "groupName")], by.x = c("animalID_par", "temp"), by.y = c("animalID", "captureYear"), all.x = T) %>%
  select(-temp) %>%
  
  merge(., capData_2009to2019[, c("animalID", "captureYear", "groupName")], by.x = c("animalID_off", "birthYear_off"), by.y = c("animalID", "captureYear"), all.x = T, suffixes = c("_par", "_off")) %>%
  
  # same groupName?
  mutate(
    groupMatch = groupName_par == groupName_off
  ) %>%
  
  # add parity for F
  merge(., parityAssigments_firstParous[, c("animalID", "parityYear")], by.x = "animalID_par", by.y = "animalID", all.x = T) %>%
  mutate(
    diff_birth.parity = as.numeric(birthYear_off) - as.numeric(parityYear)
  ) %>%
  
  relocate(parent, offspring, animalID_par, animalID_off, sex_par, sex_off, diff_birth.parity, groupMatch, birthYear_par, parityYear, birthYear_off, groupName_par, groupName_off)
```

```{r}
# number of parents assigned to each offspring
temp <- lwed_po_run2 %>%
  group_by(offspring) %>%
  summarise(count = n()) %>%
  as.data.frame() %>%
  mutate(count = str_c("p", count))
table(temp$count)

temp2 <- simp_po_run2 %>%
  group_by(offspring) %>%
  summarise(count = n()) %>%
  as.data.frame() %>%
  mutate(count = str_c("p", count))
table(temp2$count)

# rep skew stuff
lwed_po_run2 %>%
  filter(sex_par == "F") %>%
  group_by(parent) %>%
  summarise(count = n())

lwed_po_run2 %>%
  filter(sex_par == "M") %>%
  group_by(parent) %>%
  summarise(count = n())


simp_po_run2 %>%
  filter(sex_par == "F") %>%
  group_by(parent) %>%
  summarise(count = n())

simp_po_run2 %>%
  filter(sex_par == "M") %>%
  group_by(parent) %>%
  summarise(count = n())
```

#### SIB assignments

```{r}
lwed_sibs_run2 <- read_franzSibs("./DISSERTATION/ch2_demography/02_results/lwed/franz_run2/siblings.txt") %>%
  merge(., twinList[, c("sampleID_franz", "twinSet")], by.x = "sample1", by.y = "sampleID_franz", all.x = T) %>%
  merge(., twinList[, c("sampleID_franz", "twinSet")], by.x = "sample2", by.y = "sampleID_franz", all.x = T, suffixes = c("_1", "_2")) %>%
  mutate(
    twinSetMatch = twinSet_1 == twinSet_2
  ) %>%
  relocate(sample1)


simp_sibs_run2 <- read_franzSibs("./DISSERTATION/ch2_demography/02_results/simp/franz_run2/siblings.txt") %>%
  merge(., twinList[, c("sampleID_franz", "twinSet")], by.x = "sample1", by.y = "sampleID_franz", all.x = T) %>%
  merge(., twinList[, c("sampleID_franz", "twinSet")], by.x = "sample2", by.y = "sampleID_franz", all.x = T, suffixes = c("_1", "_2")) %>%
  mutate(
    twinSetMatch = twinSet_1 == twinSet_2
  ) %>%
  relocate(sample1)
```

# Plots

Want to make a Cleveland dot plot

```{r}
lwed_clevFig <- presenceMatrix %>%
  filter(species == "LWED") %>%
  mutate(temp = captureYear_last - captureYear_first) %>%
  filter(temp > 0) %>%
  arrange(desc(temp)) %>%
  mutate(animalID = factor(animalID, animalID)) %>%
  ggplot() +
  geom_segment(aes(x = 0,
                   xend = temp,
                   y = animalID,
                   yend = animalID,
                   color = sex)) +
  labs(x = "", y = "") +
  theme_bw() +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank())
lwed_clevFig
```

# SCRAPS

## SambaR

```{r}
source("/home/rachelvoyt/programs/SambaR-master/SAMBAR_v1.10.txt")
getpackages(mylib = "/home/rachelvoyt/R/x86_64-pc-linux-gnu-library/4.4")
conflicted::conflicts_prefer(base::`%in%`)

devtools::install_github("pievos101/PopGenome")

install.packages("PopGenome")


# Install the necessary Bioconductor packages
install.packages("devtools")
install.packages("BiocManager")
BiocManager::install("SNPRelate")

# Install dartRverse (dartRverse) & core (dartR.base, dartR.data)
install.packages("dartRverse")
library(dartRverse)
dartRverse::dartRverse_install()
library(dartR)

temp2 <- gi2gl(lwed_gen) # supply genind
```

```{r}
lwed_genos <- read.table("./05_tamRun5/03_run5GTscore/fullSet_polyGenResults_singleSNP_0x.txt", header = T, na.strings = "0") %>%
  # subset to lwed hair
  select(md[md$species == "LWED" & md$sampleType == "hair", "sampleID"]) %>%
  # reformat loci & subset to indid only
  `rownames<-`(sub('[_][^_]+$', '', rownames(.))) %>%
  filter(!str_detect(rownames(.), "SEXID|SPECIESID|SIMP")) %>%
  # filter to samples w/decent geno success
  #select(where(~sum(!is.na(.x))/length(.x) >= 0.7)) %>%
  t() %>%
  as.data.frame() %>%
  rownames_to_column("sampleID")

lwed_genind <- adegenet::df2genind(X = lwed_genos[,c(2:155)],
                                sep = ",",
                                ind.names = lwed_genos$sampleID,
                                NA.char = "NA",
                                ploidy = 2,
                                type = "codom")

lwed_genlight <- gi2gl(lwed_genind)
setwd("./DISSERTATION/ch2_demography/02_results/")

lwed_sambar <- genlight2sambar(genlight_object = "lwed_genlight", do_confirm = T, major = lwed_genlight$other[[1]], minor = lwed_genlight$other[[2]])

head(snps,5)
```

## COLONY

```{r}
devtools::install_github("jonesor/rcolony")
library(rcolony)

build.colony.input(wd = getwd(), )
```

# STRUCTURE

```{r}
devtools::install_github("nicholasjclark/STRUCTURE.popgen")
```
