---
title: "ch2_create_capHistories"
author: "Rachel Voyt"
date: '`r Sys.Date()`'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Overview

Quick little script to create capture histories for use in demography analyses.

**Resources:** - James Paterson has a
blog(<https://jamesepaterson.github.io/jamespatersonblog/07_creatingcapturehistories>)
with a lot of really great info on how to use mark & recapture data -- some background info from this site:

"Capture histories" are binary strings to describe when individuals were
found in the study.

For example, a capture history of "101" describes a study where an
individual was:

-   caught on the first event,
-   not observed on the second event, and
-   observed on the third event.

Data from mark-recapture and mark-resight studies are not usually
recorded or stored in this format. Often data are stored in 'long-form'
where each capture or resighting is a row. Plus, events when an
individual are not re-sighted do not usually have rows.

# Packages

```{r}
library(tidyverse)
```

# Data

Ditch those w/unknown animalIDs

```{r}
capData <- read.csv("./DISSERTATION/ch2_demography/00_data/01_demoData_clean/captureData_byIndividual_v6.csv") %>%
  filter(animalID != "UNK")
```

# Create capture histories

```{r}
capHist <- capData %>%
  select(animalID, captureDate) %>%
  filter(animalID != "UNK") %>%
  mutate(
    captureYear = as.numeric(str_sub(captureDate, 1, 4))
  ) %>%
  select(-captureDate) %>%
  mutate(
    detect = 1
  ) %>%
  
  # format as capture history (notes below are from James Paterson blog)
  
  # remove duplicates, which may occur when individuals are caught multiple times in an event
  # For example, your event may be a year and an individual may be caught multiple times in a year.
  distinct() %>%
  
  # spread out data. The fill = 0 adds rows for combinations of id and event where individuals were not observerd
  spread(captureYear, detect, fill = 0) %>% 
  
  # For every individual....
  group_by(animalID) %>%
  # Paste together 0's and 1's
  # Unite is similar to paste. Here we are pasting the strings together from the second column (first capture event)
  # to the last capture event ("tail(names(.),1)").
  # we don't want any characters separating 0's and 1's, so we use: sep = ""
  unite("ch", 2:tail(names(.),1), sep = "") %>%
  
  # add some metadata (as factor)
  merge(., capData[, c("animalID", "sex", "species")], by = "animalID") %>%
  distinct() %>%
  mutate(
    animalID = as.numeric(animalID),
    species = as.factor(species),
    sex = as.factor(sex),
    ch = as.character(ch)
  ) %>%
  relocate(animalID, species, sex) %>%
  arrange(animalID)
```

# Export

```{r}
write.csv(capHist, "./DISSERTATION/ch2_demography/00_data/01_demoData_clean/captureHistories_2009to2023.csv", row.names = F)
```

