---
title: "ch2_popGenReport"
author: "Rachel Voyt"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1 Overview

asdf

# 2 Packages

```{r}
library(adegenet)
library(conflicted) # deal with conflicting package issues
library(PopGenReport)
library(tidyverse)
```

### Troubleshooting

```{r}
install.packages("/home/rachelvoyt/programs/PopGenReport_3.1.tar.gz", repos = NULL, type = "source")
```

**Issue #1:**

Install failed, the base error seems to be related to gdal--

"configure: error: gdal-config not found or not executable."

Folks on stackoverflow (https://stackoverflow.com/questions/12141422/error-gdal-config-not-found-while-installing-r-dependent-packages-whereas-gdal) found that running the below helped--

```
sudo apt install libgdal-dev
```

**Issue #2:**

Dealing with Issue #1 seemed to help with most of the errors, but 'dismo' still won't install, as per the error message below--

"ERROR: dependency ‘dismo’ is not available for package ‘PopGenReport’"

Try to install 'dismo' manually--

```{r}
install.packages("dismo")
```

Now try PopGenReport again-- SUCCESS!

```{r}
install.packages("/home/rachelvoyt/programs/PopGenReport_3.1.tar.gz", repos = NULL, type = "source")
```

# 3 testing

## 3.1 prep

```{r}
lwed_genoSuccess_byLocus <- read.table("./05_tamRun5/03_run5GTscore/fullSet_polyGenResults_singleSNP_0x.txt", header = T, na.strings = "0") %>%
  # subset to lwed hair
  dplyr::select(md[md$species == "LWED" & md$sampleType == "hair", "sampleID"]) %>%
  # reformat loci & subset to indid only
  `rownames<-`(sub('[_][^_]+$', '', rownames(.))) %>%
  dplyr::filter(!str_detect(rownames(.), "SEXID|SPECIESID|SIMP")) %>%
  
  # remove 0% genoSuccess samples
  dplyr::select(where(~sum(!is.na(.x))/length(.x) != 0)) %>%
  
  mutate(
    genoSuccess = rowSums(!is.na(.))/( rowSums(is.na(.)) + rowSums(!is.na(.)) )
  )

mean(lwed_genoSuccess_byLocus$genoSuccess) # 0.516543
median(lwed_genoSuccess_byLocus$genoSuccess) # 0.5912698

ggplot(lwed_genoSuccess_byLocus, aes(x = genoSuccess)) +
  geom_histogram(binwidth = 0.01)


lwed_genoSuccess_bySample <- read.table("./05_tamRun5/03_run5GTscore/fullSet_polyGenResults_singleSNP_0x.txt", header = T, na.strings = "0") %>%
  # subset to lwed hair
  dplyr::select(md[md$species == "LWED" & md$sampleType == "hair", "sampleID"]) %>%
  # reformat loci & subset to indid only
  `rownames<-`(sub('[_][^_]+$', '', rownames(.))) %>%
  dplyr::filter(!str_detect(rownames(.), "SEXID|SPECIESID|SIMP")) %>%
  t() %>%
  as.data.frame() %>%
  mutate(
    genoSuccess = rowSums(!is.na(.))/( rowSums(is.na(.)) + rowSums(!is.na(.)) )
  )

mean(lwed_genoSuccess_bySample$genoSuccess) # 0.4750687
median(lwed_genoSuccess_bySample$genoSuccess) # 0.5437956
min(lwed_genoSuccess_bySample$genoSuccess)

ggplot(lwed_genoSuccess_bySample, aes(x = genoSuccess)) +
  geom_histogram(binwidth = 0.01)
```


## 3.2 df2genind

```{r}
md <- read.csv("./metadataReconciliation/tamRun5_metadata_v5.csv")

lwed_genos <- read.table("./05_tamRun5/03_run5GTscore/fullSet_polyGenResults_singleSNP_0x.txt", header = T, na.strings = "0") %>%
  # subset to lwed hair
  dplyr::select(md[md$species == "LWED" & md$sampleType == "hair", "sampleID"]) %>%
  # reformat loci & subset to indid only
  `rownames<-`(sub('[_][^_]+$', '', rownames(.))) %>%
  dplyr::filter(!str_detect(rownames(.), "SEXID|SPECIESID|SIMP")) %>%
  
  # FILTER
  ## remove 0% genoSuccess samples
  dplyr::select(where(~sum(!is.na(.x))/length(.x) != 0)) %>%
  ## remove <= 50% genoSuccess loci
  mutate(
    propNA = rowSums(is.na(.))/( rowSums(is.na(.)) + rowSums(!is.na(.)) )
    ) %>%
  dplyr::filter(propNA <= 0.5) %>%
  dplyr::select(-propNA) %>%
  # remove <= 50% genoSuccess samples
  dplyr::select(where(~sum(!is.na(.x))/length(.x) >= 0.5)) %>%
  
  t() %>%
  as.data.frame() %>%
  rownames_to_column("ind") %>%
  mutate(pop = "lwed",
         lat = NA,
         long = NA) %>%
  relocate(ind, pop, lat, long)

write.csv(lwed_genos, "./DISSERTATION/ch2_demography/00_data/lwed_genos_forPGR.csv", row.names = F)

lwed_genind <- read.genetable("./DISSERTATION/ch2_demography/00_data/lwed_genos_forPGR.csv", ind = 1, pop = 2, lat = 3, long = 4, oneColPerAll = F, sep = ",", ploidy = 2)

#lwed_genind <- adegenet::df2genind(X = lwed_genos[,c(2:155)],
#                                   sep = ",",
#                                   ind.names = lwed_genos$ind,
#                                   pop = lwed_genos$pop,
#                                   NA.char = "NA",
#                                   ploidy = 2,
#                                   type = "codom")
```


```{r}
lwed_out1 <- popgenreport(lwed_genind, path.pgr = "/home/rachelvoyt/Documents/UT-Grad/Development/repos/tamGenetics_primatesPeru/DISSERTATION/ch2_demography/02_results/lwed", mk.complete = T, foldername = "popGenReport")
```

