---
title: "tamGenetics_wgs"
author: "Rachel Voyt"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1 Overview

Note that VCF files are 1-based

## RESOURCES:

-   tabix <https://www.htslib.org/doc/tabix.html>

-   vcftools <https://vcftools.github.io/links.html>

-   GATK tool index <https://gatk.broadinstitute.org/hc/en-us/sections/27007892463131-4-6-0-0-Current?page=2#articles>

-   samtools how to extract stuff from VCFs <https://samtools.github.io/bcftools/howtos/query.html>

-   bcftools documentation <https://samtools.github.io/bcftools/bcftools.html#query>

-   eriqande book <https://eriqande.github.io/eca-bioinf-handbook/handle-vcf.html#bcftools>

-   reddit post <https://www.reddit.com/r/bioinformatics/comments/jbw72n/comment/g94sge1/> - youtube vid <https://www.youtube.com/watch?v=7W7hrWNyCaM>

-   biostars <https://www.biostars.org/p/298361/>

-   biostars (less helpful) <https://www.biostars.org/p/9576200/>

-   pysam maybe helpful code bit? <https://github.com/pysam-developers/pysam/issues/469>

-   <https://speciationgenomics.github.io/filtering_vcfs/>

# 2 Packages

## R packages

```{r}
library(janitor)
library(readxl)
library(tidyverse)
```

## command line packages

Following <https://eriqande.github.io/eca-bioinf-handbook/working-on-remote-servers.html#miniconda>

Install Miniconda if not already installed (install in home directory! makes things easier)

```{bash}
# start in your home directory and do the following:
mkdir conda_install

cd conda_install/

wget https://repo.anaconda.com/miniconda/Miniconda3-latest-Linux-x86_64.sh

chmod u+x Miniconda3-latest-Linux-x86_64.sh 

./Miniconda3-latest-Linux-x86_64.sh 
```

Then install mamba

```{bash}
conda install mamba -n base -c conda-forge
```

Install software into bioinformatics environment

```{bash}
mamba create -n bioinf -c bioconda bcftools bwa samtools
```

# 3 Data

## 3.1 Metadata

```{r}
md <- read.csv("./samFiles/tamGenetics_wgsMetadata.csv") %>%
  mutate(
    species = case_when(
      species == "Leontocebus weddelli" ~ "lwed",
      species == "Saguinus imperator" ~ "simp"
    ),
    sex = case_when(
      sex == "Female" ~ "F",
      sex == "Male" ~ "M"
    ),
    animalID_md = str_c(species, sex, animalID, sep = "_")
  ) %>%
  relocate(sampleID, animalID, animalID_md)
```

## 3.2 Loci info

```{r}
lociInfo <- read.csv("./project_data/master_lociInfo.csv") %>%
  mutate(
    snpID = str_c(chr_name, ":", snpPos_inChr)
  )
```

# 4 Get VCF genos

## 4.1 Setup

```{bash}
# 1) activate bioinf environment
conda activate bioinf

# 2) get in correct directory
cd /home/rachelvoyt/Documents/UT-Grad/Development/repos/tamGenetics_primatesPeru/DISSERTATION/ch1_panelDev/00_dataRaw
```

## 4.2 Create loci ref file

### Check format in VCF files

```{bash}
# 3) get CHROM and POS for each row of VCF file to check format
bcftools query -f '%CHROM\t%POS\n' interval_1.vcf.gz
```

### .txt file for bcftools

```{r}
lociList_bcftools <- lociInfo %>%
  select(chr_name, snpPos_inChr) %>%
  arrange(chr_name, snpPos_inChr)

write.table(lociList_bcftools, "./06_tamRun_wgs/00_data/lociList_bcftools.txt", quote = FALSE, row.names = FALSE, col.names = FALSE, sep = "\t")
```

### .list file for GATK SelectVariants

Use lociInfo to create .list file for GATK SelectVariants

```{r}
lociList_gatk <- lociInfo %>%
  select(chr_num, startPos, endPos) %>%
  mutate(
    temp = str_c(chr_num, ":", startPos, "-", endPos)
  ) %>%
  select(temp)

write.table(lociList_gatk, "./06_tamRun_wgs/00_data/lociList_gatk.list", quote = FALSE, row.names = FALSE, col.names = FALSE)
```

## 4.3 Extract genos from VCF

They literally have an example of what I want on the bcftools documentation page. wowww. <https://samtools.github.io/bcftools/bcftools.html#query>

### Shell script

Wrote a shell script that can access each VCF file on Sam's server and call bcftools query for each:

\*\*note that there are files on the server up to interval_24, but we only have loci on chromosomes 1-23 (maybe saddlebacks and emps only have 23 chromosomes while the common marmoset has 24?? dunno)

```{bash}
#!/bin/bash

# definitions
base_url="https://export.uppmax.uu.se/snic2022-6-144/share/tmp_dir_vcfs/"
loci_list="/home/rachelvoyt/Documents/UT-Grad/Development/repos/tamGenetics_primatesPeru/06_tamRun_wgs/00_data/lociList_bcftools.txt"
output_dir="/home/rachelvoyt/Documents/UT-Grad/Development/repos/tamGenetics_primatesPeru/06_tamRun_wgs/00_data/00_vcfData/"

# loop function
for i in {1..23}
do 
    # define url for individual vcf file
    vcf_url="$base_url/interval_${i}.vcf.gz"

    # define output for individual vcf file
    output_file="$output_dir/interval_${i}_genos.txt"

    # extract genos for loci in list, output to txt file
    bcftools query -H -f '%CHROM\t%POS\t%REF\t%ALT[\t%GT\t%DP\t%GQ]\n' -R "$loci_list" "$vcf_url" > "$output_file"

    # print message indicating file has been processed
    echo "Processed interval_${i}.vcf.gz -> $output_file"
done
```

bcftools variable definitions (more details at <https://samtools.github.io/bcftools/bcftools.html#query>):

-H, --print-header -f, --format -R, --regions-file

Notes on output: /: genotype unphased \|: genotype phased

GQ (Integer): Conditional genotype quality, encoded as a phred quality − 10log_10 p(genotype call is wrong, conditioned on the site's being variant).

DP (Integer): Read depth at this position for this sample

### Run script

Should still be in bioinf environment

```{bash}
# voila!
./get_vcfGenos.sh
```

# 6 Assemble base genos

## 6.1 Create combo df

In read.table, be sure to include comment.char = "" otherwise it will see "\#[1]CHROM" and ditch the headers

```{r}
# define directory w/vcfGeno txt files
vcfGenos_dir <- "./06_tamRun_wgs/00_data/00_vcfData"

# list text files in the directory
vcfGenos_dflist <- lapply(list.files(path = vcfGenos_dir, pattern = "*.txt", full.names = TRUE), read.table, sep = "\t", header = T, comment.char = "")

# create single df
vcfGenos_df_messy <- do.call(rbind, vcfGenos_dflist)
```

## 6.2 Reformat

Have records for 196 loci (doesn't include the 12 SPECIESID loci bc these are fixed in each species I guess?)

**NOTE** bcftools query returned an extra variant site (CM018917.1-101190598; REF = ATG, ALT = .); this likely happened bc it's very close to one of the variant sites that *is* in the loci list (CM018917.1-101190600). Chatgpt says possible reasons for this are:

1.  Variant Proximity: Some tools might not limit results strictly to the requested position but also include nearby variants due to overlap or when querying within a small window.

2.  Indexing in the VCF File: The way the VCF file is structured or indexed could cause variants that are very close in position to be reported together, especially if they are part of a complex event or if querying pulls nearby regions.

**NOTE2** LWED_254 is listed with REF = T and ALT = A,G -- in my lociInfo file, I have that this locus has alleles A,G. Maybe the T ref allele is based on SIMP samples??

```{r}
# clean up df
vcfGenos_df_temp <- vcfGenos_df_messy %>%
  # create snpID 
  unite(snpID, c(X..1.CHROM, X.2.POS), sep = ":", remove = T) %>%
  # get loci names
  merge(., lociInfo[, c("snpID", "locus")], by = "snpID", all = T) %>%
  relocate(locus) %>%
  arrange(locus) %>%
  # ditch speciesID loci (automatically ditches that extra variant too)
  filter(!str_detect(locus, "SPECIES")) %>%
  select(-snpID)

vcfGenos_df_refAlt <- vcfGenos_df_temp %>%
  select(locus, X.3.REF, X.4.ALT) %>%
  dplyr::rename("ref" = "X.3.REF",
                "alt" = "X.4.ALT") %>%
  # some loci have multiple alt alleles!
  separate(alt, into = c("alt1", "alt2"), sep = ",")

vcfGenos_df_clean <- vcfGenos_df_temp %>%
  select(-X.3.REF, -X.4.ALT) %>%
  column_to_rownames("locus") %>%
  t() %>%
  as.data.frame() %>%
  rownames_to_column("temp") %>%
  # remove the prefix "X.." or "X." from the strings
  mutate(temp = gsub("^X\\.\\.1\\.|^X\\.[1-8]?[0-9]\\.", "", temp)) %>%
  # split into two columns
  separate(temp, into = c("sampleID", "fmt"), sep = "\\.") %>%
  merge(., md[, c("sampleID", "animalID_md")], by = "sampleID", all.x = T) %>%
  relocate(animalID_md) %>%
  pivot_longer(!c(animalID_md, sampleID, fmt),
               names_to = "locus",
               values_to = "value") %>%
  pivot_wider(id_cols = c(animalID_md, sampleID, locus),
              names_from = fmt,
              values_from = value) %>%
  # add ref/alt back in -- some loci have multiple alt alleles!
  merge(., vcfGenos_df_refAlt, by = "locus", all.x = T) %>%
  
  
  # adjust data types & rename
  mutate(
    DP = as.numeric(DP),
    GQ = as.numeric(GQ)
  ) %>%
  dplyr::rename("vcf_sampleID" = "sampleID",
                "vcf_geno" = "GT",
                "genoDepth" = "DP",
                "genoQual" = "GQ")
```

## 6.3 Filter genos

Following Ravinet and Meiers Speciation and Population Genomics guide <https://speciationgenomics.github.io/filtering_vcfs/> section on "Variant based statistics" (this is what Sam said she used in her thesis)

**Visualize data:**

```{r}
# read depth
mean(vcfGenos_df_clean$genoDepth) # 32.80141

ggplot(vcfGenos_df_clean, aes(genoDepth)) +
  geom_density(fill = "dodgerblue1", color = "black", alpha = 0.3) +
  theme_light()

# geno quality
mean(vcfGenos_df_clean$genoQual, na.rm = T) # 88.38947

ggplot(vcfGenos_df_clean, aes(genoQual)) +
  geom_density(fill = "dodgerblue1", color = "black", alpha = 0.3) +
  theme_light()
```

**Filter & assign letter genos**

Notes from Ravinet and Meiers website-- - min depth cutoff: "10x is a good rule of thumb as a minimum cutoff for read depth, although if we wanted to be conservative, we could go with 15x." - max depth cutoff: "Usually a good rule of thumb is something the mean depth x 2"

```{r}
minDepth_cutoff = 15
maxDepth_cutoff = mean(vcfGenos_df_clean$genoDepth)*2
qual_cutoff = 30

vcfGenos_filtered <- vcfGenos_df_clean %>%
  filter(genoDepth >= minDepth_cutoff & genoDepth <= maxDepth_cutoff) %>%
  filter(genoQual >= qual_cutoff) %>%
  # assign letter geno to each allele
  mutate(
    temp1 = str_sub(vcf_geno, 1, 1),
    temp1 = case_when(
      temp1 == "0" ~ ref,
      temp1 == "1" ~ alt1,
      temp1 == "2" ~ alt2
    ),
    
    temp2 = str_sub(vcf_geno, -1, -1),
    temp2 = case_when(
      temp2 == "0" ~ ref,
      temp2 == "1" ~ alt1,
      temp2 == "2" ~ alt2
    )
  ) %>%
  # create geno col
  mutate(
    geno = str_c(temp1, temp2, sep = ",")
  ) %>%
  relocate(
    animalID_md, locus, geno, vcf_sampleID, ref, alt1, alt2, genoDepth, genoQual, temp1, temp2
  ) %>%
  arrange(locus, animalID_md)
```

## 6.4 Export genos

```{r}
vcfGenos_toExport <- vcfGenos_filtered %>%
  # mark as tissue sample geno
  mutate(animalID_md = str_c(animalID_md, "_t")) %>%
  select(animalID_md, locus, geno) %>%
  pivot_wider(id_cols = locus,
              names_from = animalID_md,
              values_from = geno)

write.csv(vcfGenos_toExport, "./06_tamRun_wgs/02_results/tamGenetics_wgs_genos.csv", row.names = F)
```