---
title: "ch2_fitness"
author: "Rachel Voyt"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1 Overview



# 2 Packages

```{r}
library(adegenet)
library(ggExtra)
library(lubridate)
library(PopGenReport) # FAILED
library(poppr)
library(tidyverse)

source("project_scripts/ggplot_theme_Publication-2.R") # for pretty ggplot themes
```

# 3 Data

## metadata

```{r}
md <- read.csv("./metadataReconciliation/tamRun5_metadata_v5.csv")

sampleRef <- read.csv("./project_data/sampleID_reference.csv")
```

## capData

```{r}
capData <- read.csv("./DISSERTATION/ch2_demographics/00_data/01_demoData_clean/captureData_byIndividual_v6.csv") %>%
  mutate(
    captureYear = str_sub(captureDate, 1, 4),
    captureYear = as.numeric(captureYear)
  ) %>%
  # adjust rowID 1 ageClass to "UNK"
  mutate(
    ageClass = case_when(
      rowID == 1 ~ "UNK",
      .default = ageClass
    )
  )

capData_firstEntry <- capData[match(unique(capData$animalID), capData$animalID),] %>%
  select(rowID, captureDate, animalID, ageClass, animalName1, animalName2, groupName, species, sex, notes_MD, notes_RV)
```

## birth/death year

```{r}
birthAssignments <- read.csv("./DISSERTATION/ch2_demographics/00_data/01_demoData_clean/birthAssignments_capData_tamRun5_v3.csv") %>%
  select(animalID, birthYear_est) %>%
  na.omit() %>%
  distinct() %>%
  arrange(animalID)

deathAssignments <- read.csv("./DISSERTATION/ch2_demographics/00_data/01_demoData_clean/deathAssignments_capData_tamRun5_v1.csv") %>%
  select(animalID, deathYear_est) %>%
  na.omit() %>%
  distinct() %>%
  arrange(animalID)
```

## parityAssignments

Using "parity2" -- meaning either left or right (vs avg) meets threshold

```{r}
source("./project_scripts/ravo_ch2Scripts.R")

parityAssignments <- get_parityStatus(capData_file = capData) %>%
  select(animalID, captureDate, parity2) %>%
  dplyr::rename("parity" = "parity2") %>%
  mutate(
    captureYear = str_sub(captureDate, 1, 4)
  ) %>%
  select(-captureDate) %>%
  distinct()

parityAssigments_firstParous <- parityAssignments[match(unique(parityAssignments$animalID), parityAssignments$animalID),] %>%
  filter(parity == "parous") %>%
  dplyr::rename("parityYear" = "captureYear")
```

## likelyTwins

created in ch2_demographics.Rmd

```{r}
likelyTwins <- read.csv("./DISSERTATION/ch2_demographics/00_data/01_demoData_clean/twinList_tamRun5_6June2024.csv")

twinList <- likelyTwins %>%
  select(twin1, twin2, twin3) %>%
  arrange(twin1) %>%
  mutate(twinSet = str_c("twinSet", row_number())) %>%
  pivot_longer(-twinSet,
               names_to = "twinID",
               values_to = "animalID") %>%
  na.omit() %>%
  merge(., sampleRef[, c("animalID", "sampleID_franz", "seqRun", "sampleType")], by = "animalID", all.x = T) %>%
  filter(seqRun == "run5") %>%
  filter(sampleType == "hair") %>%
  relocate(twinSet) %>%
  arrange(twinSet)
```

## genos

```{r}
genos0x <- read.table("./05_tamRun5/03_run5GTscore/fullSet_polyGenResults_singleSNP_0x.txt", header = T, na.strings = "0")
```

# FRANz

Following Shedd et al GitHub 1_RRS_Hogan_13_15.Rmd<https://github.com/krshedd/Relative-fitness-of-Pink-Salmon/blob/main/analysis/1_RRS_Hogan_13_15.Rmd>

## Create input files

**Subset genos to samples with >= 70% genotyping success**

### Prep files

**LWED**

```{r}
# gtscore0x genos w/>= 70% genoSuccess
lwed_genos0x <- genos0x %>%
  # subset to lwed hair
  select(md[md$species == "LWED" & md$sampleType == "hair", "sampleID"]) %>%
  # reformat loci & subset to indid only
  `rownames<-`(sub('[_][^_]+$', '', rownames(.))) %>%
  filter(!str_detect(rownames(.), "SEXID|SPECIESID|SIMP")) %>%
  # filter to samples w/decent geno success
  select(where(~sum(!is.na(.x))/length(.x) >= 0.7)) %>%
  t() %>%
  as.data.frame()

lwed_genos0x_franz <- names(lwed_genos0x) %>%
  map_dfc(~ lwed_genos0x %>%
            select(all_of(.x)) %>%
            separate(.x,
                     into = paste0(.x, c("a", "b")),
                     sep = ",")
          ) %>%
  
  # add sampleID_franz, sex, & animalID (animalID is just for merging birthYear)
  rownames_to_column("sampleID") %>%
  merge(., sampleRef[, c("sampleID", "sampleID_franz", "animalID", "sex")], by = "sampleID") %>%
  relocate(c(sampleID_franz, sex)) %>%
  
  # add birthYear
  merge(., birthAssignments[, c("animalID", "birthYear_est")], by = "animalID", all.x = T) %>%
  dplyr::rename("birthYear" = "birthYear_est") %>%
  relocate(birthYear, .before = "sex") %>%
  
  # add deathYear
  merge(., deathAssignments[, c("animalID", "deathYear_est")], by = "animalID", all.x = T) %>%
  dplyr::rename("deathYear" = "deathYear_est") %>%
  relocate(deathYear, .before = "sex") %>%
  
  # replace allele letters with numbers (required by FRANz)
  mutate_at(vars(starts_with(c("INDID", "LWED"))), ~ str_replace(., "A", "4")) %>%
  mutate_at(vars(starts_with(c("INDID", "LWED"))), ~ str_replace(., "T", "7")) %>%
  mutate_at(vars(starts_with(c("INDID", "LWED"))), ~ str_replace(., "C", "3")) %>%
  mutate_at(vars(starts_with(c("INDID", "LWED"))), ~ str_replace(., "G", "6")) %>%
  arrange(sampleID) %>%
  select(-animalID, -sampleID)

# export
write.csv(lwed_genos0x_franz, "./DISSERTATION/ch2_demographics/00_data/lwed_genos0x_franz.csv", row.names = F)
```

**SIMP**

```{r}
# gtscore0x genos w/>= 70% genoSuccess
simp_genos0x <- genos0x %>%
  # subset to simp hair
  select(md[md$species == "SIMP" & md$sampleType == "hair", "sampleID"]) %>%
  # reformat loci & subset to indid only
  `rownames<-`(sub('[_][^_]+$', '', rownames(.))) %>%
  filter(!str_detect(rownames(.), "SEXID|SPECIESID|LWED")) %>%
  # filter to samples w/decent geno success
  select(where(~sum(!is.na(.x))/length(.x) >= 0.7)) %>%
  t() %>%
  as.data.frame()

simp_genos0x_franz <- names(simp_genos0x) %>%
  map_dfc(~ simp_genos0x %>%
            select(all_of(.x)) %>%
            separate(.x,
                     into = paste0(.x, c("a", "b")),
                     sep = ",")
          ) %>%
  
  # add sampleID_franz, sex, & animalID (animalID is just for merging birthYear)
  rownames_to_column("sampleID") %>%
  merge(., sampleRef[, c("sampleID", "sampleID_franz", "animalID", "sex")], by = "sampleID") %>%
  relocate(c(sampleID_franz, sex)) %>%
  
  # add birthYear
  merge(., birthAssignments[, c("animalID", "birthYear_est")], by = "animalID", all.x = T) %>%
  dplyr::rename("birthYear" = "birthYear_est") %>%
  relocate(birthYear, .before = "sex") %>%
  
  # add deathYear
  merge(., deathAssignments[, c("animalID", "deathYear_est")], by = "animalID", all.x = T) %>%
  dplyr::rename("deathYear" = "deathYear_est") %>%
  relocate(deathYear, .before = "sex") %>%
  
  # replace allele letters with numbers (required by FRANz)
  mutate_at(vars(starts_with(c("INDID", "SIMP"))), ~ str_replace(., "A", "4")) %>%
  mutate_at(vars(starts_with(c("INDID", "SIMP"))), ~ str_replace(., "T", "7")) %>%
  mutate_at(vars(starts_with(c("INDID", "SIMP"))), ~ str_replace(., "C", "3")) %>%
  mutate_at(vars(starts_with(c("INDID", "SIMP"))), ~ str_replace(., "G", "6")) %>%
  arrange(sampleID) %>%
  select(-animalID, -sampleID)

# export
write.csv(simp_genos0x_franz, "./DISSERTATION/ch2_demographics/00_data/simp_genos0x_franz.csv", row.names = F)
```

### .csv to .dat

FRANz provides a perl script (csv.pl) to convert .csv files to .dat
files. The manual says that you can also go to a GUI at
<http://legacy.bioinf.uni-leipzig.de/Software/FRANz/CSV_Import.html>,
but I haven't gotten it to work (it just loads forever).

FRANz has examples of how to use csv.pl in their "man" file; I've edited
their example slightly so that the output automatically creates a new
file.

NOTE that column IDs start with 0.

How to put output into file:
<https://askubuntu.com/questions/420981/how-do-i-save-terminal-output-to-a-file>

**NOTE** - use csv_modified.pl; line 164 has an error in the original
script -- original has "\$death_col = $data->[$death_col];", but should
be "\$death = $data->[$death_col];"

```{r}
system2("perl",
        args="./project_scripts/franz/csv_modified.pl --in ./DISSERTATION/ch2_demographics/00_data/lwed_genos0x_franz.csv --alleles_per_col 1 --has_header --missing_allele 'NA' --birth_col 1 --death_col 2 --sex_col 3 --data_col 4 > ./DISSERTATION/ch2_demographics/00_data/lwed_genos0x_franz.dat")

system2("perl",
        args="./project_scripts/franz/csv_modified.pl --in ./DISSERTATION/ch2_demographics/00_data/simp_genos0x_franz.csv --alleles_per_col 1 --has_header --missing_allele 'NA' --birth_col 1 --death_col 2 --sex_col 3 --data_col 4 > ./DISSERTATION/ch2_demographics/00_data/simp_genos0x_franz.dat")
```

## Assign Nf & Nm

From FRANz manual: *N* is the number of candidate female (*Nf*) and male (*Nm*) parents in the population, equal to the sum of the average number of sampled (*n*) and unsampled (*N-n*) breeding females and males in the population (respectively). If unset, then estimated jointly with the pedigree. For parentage inference N=Nf=Nm (NOT Nf+Nm), but one can also specify Nf and Nm instead of N if these numbers differ. If N isnot known, use Nmax instead.

The FRANz manual suggests that if a good estimate of Nf/Nm is possible, specifying these variables is preferable to Nfmax/Nmmax (the maximum number of candidate female/male parents in the population). As such, I'm using the sex-specific abundance estimates calculated in ch2_demography.Rmd (currently in ch2_relatedness.Rmd tho, need to move it) based on capData records.

```
lwed_Nf = 73
lwed_Nm = 82

simp_Nf = 52
simp_Nm = 60
```

## franz_run1

### Run script

```{bash}
# lwed
cd /home/rachelvoyt/Documents/UT-Grad/Development/repos/tamGenetics_primatesPeru/DISSERTATION/ch2_demographics/02_results/lwed/franz_run1

FRANz --femrepro 2:30 --malerepro 2:30 --Nf 73 --Nm 82 --fullsibtest --fullsibparental --updatefreqs ./../../../00_data/lwed_genos0x_franz.dat

# simp
cd ./../../simp/franz_run1

FRANz --femrepro 2:30 --malerepro 2:30 --Nf 52 --Nm 60 --fullsibtest --fullsibparental --updatefreqs ./../../../00_data/simp_genos0x_franz.dat
```

FRANz parameters are defined as follows:
--femrepro and --malerepro specify the age ranges in which females and males can reproduce

--Nf and --Nm specify the number of candidate female and male parents in the population, which is the sum of the average number of sampled (n) and unsampled (N-n) breeding females/males in the population. The FRANz manual suggests that if a good estimate of Nf/Nm is possible, specifying these variables is preferable to Nfmax/Nmmax (the maximum number of candidate female/male parents in the population)

--fullsibtest specifies that FRANz should run the fullsib heuristic (as described in Riester et al. 2009)

--fullsibparental specifies that FRANz should also detect fullsibs in the parent generation

--updatefreqs specifies that FRANz should update allele frequencies using MCMC sampling

**NOTE** that FRANz automatically sets --mintyped loci to (1+loci)/2; this also defines the min number of common typed loci for a pair of indivs. Other defaults include:

-   typingerror = 0.01

### Assess results

#### Cumulative exclusion probability

LWED: Cumulative exclusion probability when 1 to 7 fullsibs are genotyped
  First Parent              : 0.9999997  1.0000000  1.0000000  1.0000000  1.0000000  1.0000000  1.0000000 
  Second Parent             : 1.0000000  1.0000000  1.0000000  1.0000000  1.0000000  1.0000000  1.0000000 
  Parent Pair               : 1.0000000  1.0000000  1.0000000  1.0000000  1.0000000  1.0000000  1.0000000 
  
SIMP: Cumulative exclusion probability when 1 to 7 fullsibs are genotyped
  First Parent              : 0.9999993  1.0000000  1.0000000  1.0000000  1.0000000  1.0000000  1.0000000 
  Second Parent             : 1.0000000  1.0000000  1.0000000  1.0000000  1.0000000  1.0000000  1.0000000 
  Parent Pair               : 1.0000000  1.0000000  1.0000000  1.0000000  1.0000000  1.0000000  1.0000000 

#### PO assignments

FRANz colnames are not my favorite - set new ones:

```{r}
colnames_parentage <- c("offspring",
                        "lociTyped_offspring",
                        "parent1",
                        "lociTyped_parent1",
                        "parent2",
                        "lociTyped_parent2",
                        "LOD",
                        "posterior",
                        "commonLociTyped",
                        "mismatches",
                        "n_f",
                        "n_m",
                        "pairLOD_parent1",
                        "pairLOD_parent2",
                        "posterior_parent1",
                        "posterior_parent2",
                        "parentage_mlPedigree")
```

**PO assignments**

```{r}
lwed_po_run1_file <- read.csv("./DISSERTATION/ch2_demographics/02_results/lwed/franz_run1/parentage.csv", na.strings = c("", "NA")) %>%
  rownames_to_column("temp") %>%
  `colnames<-`(colnames_parentage) %>%
  mutate(franzRun = "run1")

simp_po_run1_file <- read.csv("./DISSERTATION/ch2_demographics/02_results/simp/franz_run1/parentage.csv", na.strings = c("", "NA")) %>%
  rownames_to_column("temp") %>%
  `colnames<-`(colnames_parentage) %>%
  mutate(franzRun = "run1")
```

**Take a peek at parentage posterior probabilities**

```{r}
lwed_po_run1_file %>%
  filter(!is.na(parent1)) %>%
  ggplot(aes(x = posterior)) +
  geom_histogram(breaks = seq(0, 1, 0.01)) +
  ggtitle("Histogram of FRANz posterior probabilities for parentage assignments") +
  theme_bw()

simp_po_run1_file %>%
  filter(!is.na(parent1)) %>%
  ggplot(aes(x = posterior)) +
  geom_histogram(breaks = seq(0, 1, 0.01)) +
  ggtitle("Histogram of FRANz posterior probabilities for parentage assignments") +
  theme_bw()
```

**PO checks**

```{r}
# LWED
lwed_po_run1 <- lwed_po_run1_file %>%
  filter(posterior > 0.9) %>%
  select(offspring, parent1, parent2) %>%
  pivot_longer(-offspring,
               names_to = "parentID",
               values_to = "parent") %>%
  select(-parentID) %>%
  na.omit() %>%
  
  # add birthYear and sex
  merge(., lwed_genos0x_franz[, c("sampleID_franz", "birthYear", "sex")], by.x = "parent", by.y = "sampleID_franz", all.x = T) %>%
  merge(., lwed_genos0x_franz[, c("sampleID_franz", "birthYear", "sex")], by.x = "offspring", by.y = "sampleID_franz", all.x = T, suffixes = c("_par", "_off")) %>%
  
  # add groups
  mutate(
    animalID_par = gsub(".*_([0-9]+)$", "\\1", parent),
    animalID_off = gsub(".*_([0-9]+)$", "\\1", offspring)
    ) %>%
  
  mutate(
    temp = as.numeric(birthYear_off) - 1
    ) %>%
  merge(., capData[, c("animalID", "captureYear", "groupName")], by.x = c("animalID_par", "temp"), by.y = c("animalID", "captureYear"), all.x = T) %>%
  select(-temp) %>%
  
  merge(., capData[, c("animalID", "captureYear", "groupName")], by.x = c("animalID_off", "birthYear_off"), by.y = c("animalID", "captureYear"), all.x = T, suffixes = c("_par", "_off")) %>%
  
  # same groupName?
  mutate(
    groupMatch = groupName_par == groupName_off
  ) %>%
  
  # add parity for F
  merge(., parityAssigments_firstParous[, c("animalID", "parityYear")], by.x = "animalID_par", by.y = "animalID", all.x = T) %>%
  mutate(
    diff_birth.parity = as.numeric(birthYear_off) - as.numeric(parityYear)
  ) %>%
  
  relocate(parent, offspring, animalID_par, animalID_off, sex_par, sex_off, diff_birth.parity, groupMatch, birthYear_par, parityYear, birthYear_off, groupName_par, groupName_off)


# SIMP
simp_po_run1 <- simp_po_run1_file %>%
  filter(posterior > 0.9) %>%
  select(offspring, parent1, parent2) %>%
  pivot_longer(-offspring,
               names_to = "parentID",
               values_to = "parent") %>%
  select(-parentID) %>%
  na.omit() %>%
  
  # add birthYear and sex
  merge(., simp_genos0x_franz[, c("sampleID_franz", "birthYear", "sex")], by.x = "parent", by.y = "sampleID_franz", all.x = T) %>%
  merge(., simp_genos0x_franz[, c("sampleID_franz", "birthYear", "sex")], by.x = "offspring", by.y = "sampleID_franz", all.x = T, suffixes = c("_par", "_off")) %>%
  
  # add groups
  mutate(
    animalID_par = gsub(".*_([0-9]+)$", "\\1", parent),
    animalID_off = gsub(".*_([0-9]+)$", "\\1", offspring)
    ) %>%
  
  mutate(
    temp = as.numeric(birthYear_off) - 1
    ) %>%
  merge(., capData[, c("animalID", "captureYear", "groupName")], by.x = c("animalID_par", "temp"), by.y = c("animalID", "captureYear"), all.x = T) %>%
  select(-temp) %>%
  
  merge(., capData[, c("animalID", "captureYear", "groupName")], by.x = c("animalID_off", "birthYear_off"), by.y = c("animalID", "captureYear"), all.x = T, suffixes = c("_par", "_off")) %>%
  
  # same groupName?
  mutate(
    groupMatch = groupName_par == groupName_off
  ) %>%
  
  # add parity for F
  merge(., parityAssigments_firstParous[, c("animalID", "parityYear")], by.x = "animalID_par", by.y = "animalID", all.x = T) %>%
  mutate(
    diff_birth.parity = as.numeric(birthYear_off) - as.numeric(parityYear)
  ) %>%
  
  relocate(parent, offspring, animalID_par, animalID_off, sex_par, sex_off, diff_birth.parity, groupMatch, birthYear_par, parityYear, birthYear_off, groupName_par, groupName_off)
```

**PO pedigree.dat**

Just filtering by groupMatch == T; filtering by diff_birth.parity gives too few I think

LWED: F = 7, M = 6
SIMP: F = 4, M = 4

```{r}
lwed_po_run1_forPed <- lwed_po_run1 %>%
  filter(groupMatch == TRUE) %>%
  mutate(franzPed = str_c(parent, offspring))

simp_po_run1_forPed <- simp_po_run1 %>%
  filter(groupMatch == TRUE) %>%
  mutate(franzPed = str_c(parent, offspring))

# overview of parent assignment numbers for pedigree.dat
lwed_po_run1_forPed %>%
  group_by(sex_par) %>%
  summarise(count = n())

simp_po_run1_forPed %>%
  group_by(sex_par) %>%
  summarise(count = n())
```

Reformat and export as .dat files

```{r}
nrow(lwed_genos0x_franz) # 59 lwed samples
nrow(simp_genos0x_franz) # 41 simp samples

lwed_ped_forRun2 <- c(lwed_genos0x_franz$sampleID_franz,
                      lwed_po_run1_forPed$franzPed) %>%
  as.data.frame() %>%
  dplyr::rename("59" = ".")

simp_ped_forRun2 <- c(simp_genos0x_franz$sampleID_franz,
                      simp_po_run1_forPed$franzPed) %>%
  as.data.frame() %>%
  dplyr::rename("41" = ".")

# export
write.table(lwed_ped_forRun2, file = "./DISSERTATION/ch2_demographics/00_data/lwed_poPed_forRun2.dat", quote = F, row.names = F)

write.table(simp_ped_forRun2, file = "./DISSERTATION/ch2_demographics/00_data/simp_poPed_forRun2.dat", quote = F, row.names = F)
```

#### SIB assignments

```{r}
colnames_sibs <- c("Genotype1", "Genotype2", "FS", "PO", "HS", "pV_PO", "pV_HS", "pV_U", "Common_Parents", "Common_Loci")

lwed_sibs_run1_file <- readLines("./DISSERTATION/ch2_demographics/02_results/lwed/franz_run1/siblings.txt")

lwed_sibs_run1_file1 <- lwed_sibs_run1_file[24:28] %>%
  as.data.frame() %>%
  dplyr::rename("temp" = ".") %>%
  mutate(
    temp = gsub(" ", "", temp),
    temp = str_sub(temp, 2, -2)
    ) %>%
  separate(temp, into = colnames_sibs, sep = "\\|") %>%
  mutate(dyadStatus = "accepted")
  
lwed_sibs_run1_file2 <- lwed_sibs_run1_file[35:53] %>%
  as.data.frame() %>%
  dplyr::rename("temp" = ".") %>%
  mutate(
    temp = gsub(" ", "", temp),
    temp = str_sub(temp, 2, -2)
    ) %>%
  separate(temp, into = colnames_sibs, sep = "\\|") %>%
  mutate(dyadStatus = "rejected")

lwed_sibs_run1 <- rbind(lwed_sibs_run1_file1, lwed_sibs_run1_file2) %>%
  merge(., twinList[, c("sampleID_franz", "twinSet")], by.x = "Genotype1", by.y = "sampleID_franz", all.x = T) %>%
  merge(., twinList[, c("sampleID_franz", "twinSet")], by.x = "Genotype2", by.y = "sampleID_franz", all.x = T, suffixes = c("_1", "_2")) %>%
  mutate(
    twinSetMatch = twinSet_1 == twinSet_2
  )
```


## franz_run2

### Run script

```{bash}
# lwed
cd /home/rachelvoyt/Documents/UT-Grad/Development/repos/tamGenetics_primatesPeru/DISSERTATION/ch2_demographics/02_results/lwed/franz_run2

FRANz --femrepro 2:30 --malerepro 2:30 --Nf 73 --Nm 82 --pedigreein ./../../../00_data/lwed_poPed_forRun2.dat --fullsibtest --fullsibparental --updatefreqs ./../../../00_data/lwed_genos0x_franz.dat

# simp
cd ./../../simp/franz_run2

FRANz --femrepro 2:30 --malerepro 2:30 --Nf 52 --Nm 60 --pedigreein ./../../../00_data/simp_poPed_forRun2.dat --fullsibtest --fullsibparental --updatefreqs ./../../../00_data/simp_genos0x_franz.dat
```

LWED warnings:
Warning: Locus 78 not informative.
Warning: Locus 84 not informative.
Warning: Locus 89 not informative.
Warning: Locus 91 not informative.
Warning: Locus 130 not informative.
Warning: Locus 131 not informative.
Warning: Locus 132 not informative.
Warning: Locus 133 not informative.
Warning: Locus 135 not informative.
Warning: Locus 139 not informative.
Warning: Locus 142 not informative.
Warning: Locus 144 not informative.
Warning: Locus 147 not informative.
Warning: Locus 149 not informative.

SIMP warnings:
Warning: Locus 71 not informative.
Warning: Locus 78 not informative.
Warning: Locus 121 not informative.
Warning: Locus 125 not informative.
Warning: Locus 132 not informative.
Warning: Locus 136 not informative.
Warning: Locus 137 not informative.
Warning: Locus 142 not informative.
Warning: Locus 145 not informative.

### Assess results

