---
title: "ch1_panelPerformance"
author: "Rachel Voyt"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1 Overview

# 2 Packages

```{r}
library(poppr)
library(tidyverse)

# for PopGenUtils
library(devtools)
install_github("nikostourvas/PopGenUtils")
library("PopGenUtils")
```

# 3 Data formatting

```{r}
md <- read.csv("./metadataReconciliation/tamRun5_metadata_v5.csv")

# LWED n = 42
lwed_genos <- read.table("./05_tamRun5/03_run5GTscore/fullSet_polyGenResults_singleSNP_0x.txt", header = T, na.strings = "0") %>%
  # subset to lwed hair
  select(md[md$species == "LWED" & md$sampleType == "hair", "sampleID"]) %>%
  # reformat loci & subset to indid only
  `rownames<-`(sub('[_][^_]+$', '', rownames(.))) %>%
  filter(!str_detect(rownames(.), "SEXID|SPECIESID|SIMP")) %>%
  # remove samples w/>20% missing data
  select(where(~sum(!is.na(.x))/length(.x) >= 0.8)) %>%
  t() %>%
  as.data.frame() %>%
  rownames_to_column("sampleID")

# genind format genos
lwed_genos_genind <- adegenet::df2genind(X = lwed_genos[,c(2:155)],
                                         sep = ",",
                                         ind.names = lwed_genos$sampleID,
                                         NA.char = "NA",
                                         ploidy = 2,
                                         type = "codom")

# SIMP n = 24
simp_genos <- read.table("./05_tamRun5/03_run5GTscore/fullSet_polyGenResults_singleSNP_0x.txt", header = T, na.strings = "0") %>%
  # subset to lwed hair
  select(md[md$species == "SIMP" & md$sampleType == "hair", "sampleID"]) %>%
  # reformat loci & subset to indid only
  `rownames<-`(sub('[_][^_]+$', '', rownames(.))) %>%
  filter(!str_detect(rownames(.), "SEXID|SPECIESID|LWED")) %>%
  # remove samples w/>20% missing data
  select(where(~sum(!is.na(.x))/length(.x) >= 0.8)) %>%
  t() %>%
  as.data.frame() %>%
  rownames_to_column("sampleID")

# genind format genos
simp_genos_genind <- adegenet::df2genind(X = simp_genos[,c(2:155)],
                                         sep = ",",
                                         ind.names = simp_genos$sampleID,
                                         NA.char = "NA",
                                         ploidy = 2,
                                         type = "codom")
```

# poppr::genotype_curve

A genotype accumulation curve is a tool that allows you to assess how much power you have to discriminate between unique individuals given a random sample of _n_ loci.

```{r}
# plot
lwed_genoCurve <- genotype_curve(gen = genos_lwed_genind,
                                 sample = 1000,
                                 plot = T,
                                 drop = T,
                                 dropna = T)
lwed_genoCurve <- last_plot()

lwed_genoCurve +
  geom_smooth()

# get data
lwed_genoCurve_data <- lwed_genoCurve$data %>%
  group_by(NumLoci) %>%
  summarize(
    MinMLG = min(MLG, na.rm = TRUE),
    MaxMLG = max(MLG, na.rm = TRUE),
    MedianMLG = median(MLG, na.rm = TRUE)
  )
```

Definitions of parameters:
sample = i; the number of times loci will be resampled without replacement
plot = T; will return plotted genotype curve, if F will return matrix
drop = T; monomorphic loci will be removed before analysis
dropna = T; NAs will be ignored when determining if locus is monomorphic

Output includes NumLoci (number of loci) vs. MLG (multilocus genotypes)

# PopGenUtils::pid_permute

## Calculations

```{r}
lwed_pidPerm <- pid_permute(obj = lwed_genos_genind, nrep = 1000)

simp_pidPerm <- pid_permute(obj = simp_genos_genind, nrep = 1000)
```

## Analysis

```{r}
# LWED
lwed_pidPerm_data <- lwed_pidPerm$median_values

## PID meets 0.0001 threshold at 12 loci
lwed_pid.thresh <- as.numeric(lwed_pidPerm_data[which(lwed_pidPerm_data$PID<0.0001, arr.ind=TRUE)[1],][[1]])

## PIDsibs meets 0.0001 threshold at 23 loci
lwed_pidSibs.thresh <- as.numeric(lwed_pidPerm_data[which(lwed_pidPerm_data$PIDsibs<0.0001, arr.ind=TRUE)[1],][[1]])

# SIMP
simp_pidPerm_data <- simp_pidPerm$median_values

## PID meets 0.0001 threshold at 12 loci
simp_pid.thresh <- as.numeric(simp_pidPerm_data[which(simp_pidPerm_data$PID<0.0001, arr.ind=TRUE)[1],][[1]])

## PIDsibs meets 0.0001 threshold at 24 loci
simp_pidSibs.thresh <- as.numeric(simp_pidPerm_data[which(simp_pidPerm_data$PIDsibs<0.0001, arr.ind=TRUE)[1],][[1]])
```

## Plots

```{r}
# LWED
lwed_pidPerm_fig <- lwed_pidPerm$plot +
  scale_x_discrete(limits = as.character(1:25)) +
  geom_vline(xintercept = lwed_pid.thresh, color = "#386cb0") +
  geom_vline(xintercept = lwed_pidSibs.thresh, color = "#fdb462") +
  #geom_hline(yintercept = 0.0001) +
  #scale_y_continuous(limits = c(0, 0.001)) +
  scale_colour_Publication() +
  theme_Publication() +
  labs(title = "LWED PID + PIDsibs")

lwed_pidPerm_fig

# SIMP
simp_pidPerm_fig <- simp_pidPerm$plot +
  scale_x_discrete(limits = as.character(1:25)) +
  geom_vline(xintercept = simp_pid.thresh, color = "#386cb0") +
  geom_vline(xintercept = simp_pidSibs.thresh, color = "#fdb462") +
  #geom_hline(yintercept = 0.0001) +
  #scale_y_continuous(limits = c(0, 0.001)) +
  scale_colour_Publication() +
  theme_Publication() +
  labs(title = "SIMP PID + PIDsibs")

simp_pidPerm_fig
```




