---
title: "tamGenetics_wgs"
author: "Rachel Voyt"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1 Overview

# 2 Packages

## command line packages

Install Miniconda if not already installed (install in home directory! makes things easier)

```{bash}
# start in your home directory and do the following:
mkdir conda_install
cd conda_install/
wget https://repo.anaconda.com/miniconda/Miniconda3-latest-Linux-x86_64.sh
chmod u+x Miniconda3-latest-Linux-x86_64.sh 
./Miniconda3-latest-Linux-x86_64.sh 
```

Install mamba

```{bash}
conda install mamba -n base -c conda-forge
```

Install software into bioinformatics environment

```{bash}
mamba create -n bioinf -c bioconda bcftools bwa samtools
```

## R packages

```{r}
library(tidyverse)
```

# 3 Data

There's some kind of issue with locus "SEXID_gatk_filtered_new_non_hom_intervals_CM018939.1_332662-333103_chosen_consensus_modified" -- based on some digging around, I'm *fairly* certain that 333103 is the pre-bp3error stopPos while 333110 is the updated stopPos. 

```{r}
panel_v3 <- read.table("./project_data/primerProbeFile_panelv3_fullSet.txt", header = T)

# pretty sure this is Sam's file w/the loci info after bp3error interval updates
samMasterSNP <- read.csv("./primers/26Sep2022_master_SNP_with_replacement_SNPs.csv") %>%
  mutate(
    snpID = str_c("chr", chrom_num, ":", start, "-", stop)
  ) %>%
  relocate(snpID)

lociInfo <- read.csv("./project_data/seqNames_to_shortNames.csv") %>%
  select(primerName3, seqID3) %>%
  dplyr::rename("Locus" = "primerName3") %>%
  filter(Locus %in% panel_v3$Locus) %>%
  # extract chr number
  mutate(
    temp1 = sub(".*_(CM[0-9]+\\.[0-9]+).*", "\\1", seqID3),
    temp2 = str_sub(temp1, 7, 8),
    chr = as.numeric(temp2) - 16,
    chr = str_c("chr", chr)
  ) %>%
  select(-contains("temp")) %>%
  # extract start/end numbers
  mutate(
    beginPos = sub(".*CM[0-9]+\\.[0-9]+[-:.\\_]([0-9]+)-([0-9]+).*", "\\1", seqID3),
    endPos = sub(".*CM[0-9]+\\.[0-9]+[-:.\\_]([0-9]+)-([0-9]+).*", "\\2", seqID3)
  ) %>%
  merge(., panel_v3[, c("Locus", "SNPpos")], by = "Locus") %>%
  mutate(
    snpID = str_c(chr, ":", beginPos, "-", endPos)
  ) %>%
  # update chr23:332662-333103 to 333110
  mutate(
    snpID = case_when(
      snpID == "chr23:332662-333103" ~ "chr23:332662-333110",
      .default = snpID
    ),
    endPos = case_when(
      snpID == "chr23:332662-333110" ~ "333110",
      .default = endPos
    )
  ) %>%
  merge(., samMasterSNP[, c("snpID", "SNP")], by = "snpID", all.x = T)
```

# 4 Create .list file

Use lociInfo to create .list file for GATK SelectVariants 

```{r}
listFile <- lociInfo %>%
  select(chr, beginPos, endPos) %>%
  mutate(
    temp = str_c(chr, ":", beginPos, "-", endPos)
  )

write.table(listFile, "./DISSERTATION/ch1_panelDev/snpList.list", quote = FALSE, row.names = FALSE, col.names = FALSE)
```

# 5 Extract genos from VCF

```{bash}
# example
bcftools query -i 'ID=@NOD2SNPS.txt' -f'[CHROM POS ID REF ALT QUAL %SAMPLE\n][%CHROM %POS %ID %REF %ALT %QUAL %GT\n]' "../Raw Files/annotated.genotypes.vcf" | sed 's|0/1|HET|g' | sed 's|0/0|REF|g' | sed 's|1/1|HOM|g' | sed 's|R/.|HET|g' > next.txt

# mine:
bcftools query -i 'ID=@NOD2SNPS.txt' -f'[CHROM POS ID REF ALT QUAL %SAMPLE\n][%CHROM %POS %ID %REF %ALT %QUAL %GT\n]' "../Raw Files/annotated.genotypes.vcf" | sed 's|0/1|HET|g' | sed 's|0/0|REF|g' | sed 's|1/1|HOM|g' | sed 's|R/.|HET|g' > next.txt
```




.vcf.gz = vcf files
.tbi = indexed file

First need to install miniconda/mamba -- follow https://eriqande.github.io/eca-bioinf-handbook/working-on-remote-servers.html#miniconda



How to access VCF files on remote server (from youtube video https://www.youtube.com/watch?v=7W7hrWNyCaM):

1. Copy VCF file link

2. Assign the link a variable name for easier manipulation:

```{bash}
# assign variable name
link=https://export.uppmax.uu.se/snic2022-6-144/share/tmp_dir_vcfs/interval_1.vcf.gz

# $ indicates that we want the actual value
bcftools query -l $link
```



RESOURCES

reddit thread talking about doing what I think I want to do? https://www.reddit.com/r/bioinformatics/comments/jbw72n/retrieve_samples_with_specific_snps_from_a_vcf/ 