---
title: "xtnMasterLists"
author: "Rachel Voyt"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Overview

This document is to reformat the tamGenetics extraction master lists, with the output used to update the current extraction master lists for blood, fecal, and hair samples on Quip.

Final column headings:

-   sampleLocated
-   animalID
-   sampleRemaining
-   sampleType
    -   blood
    -   fecal
    -   hair
    -   xtnNeg
-   sampleStorage
    -   longmire
    -   rnaLater
    -   envelope
-   sampleID
    -   longmire tube \#
    -   rnaLater tube \#
    -   env \#
-   sampleNotes
-   lysateRemaining
-   lysisDate
-   lysisType
    -   tube
    -   plate
-   lysisLoc
-   lysisRow
-   lysisCol
-   lysisVol_ul
-   lysis_prokProp
-   lysisDuration
-   lysisProtocol
-   xtnRemaining
-   xtnDate
-   xtnYear
-   xtnType
-   xtnID
-   xtnLoc
-   eltn
-   xtnRow
-   xtnCol
-   xtnProtocol
-   xtnNotes
-   qubit_ng.ul
-   qubitDate
-   pico_ng.ul
-   picoDate
-   purity_260.280
-   purity_260.230
-   purityDate
-   fieldNotes
-   captureDate
-   captureYear
-   group
-   animalName
-   species
-   sex

# Packages

```{r}
library(tidyverse)
library(stringr)
```

# Xtns 27 July 2023

## Blood samples

```{r}
xtnMaster_blood <- read.csv("./extractions/tamGenetics_extractionsBlood_master_21July2023.csv", na.strings = c("", "NA"))

xtnMaster_blood_e1 <- xtnMaster_blood %>%
  mutate(
    sampleLocated = case_when(
      animalID == "BLANK" ~ NA,
      .default = "yes"
      ),
    animalID = case_when(
      animalID == "BLANK" ~ str_c("xtnNeg_p", xtnPlate, ".", xtnRow, xtnCol),
      .default = animalID
      ),
    sampleRemaining = "yes",
    sampleType = case_when(
      str_detect(animalID, "xtnNeg") ~ "xtnNeg",
      .default = "blood"
      ),
    sampleStorage = case_when(
      str_detect(animalID, "xtnNeg") ~ NA,
      .default = "longmire"
      ),
    sampleID = case_when(
      str_detect(animalID, "xtnNeg") ~ NA,
      .default = str_c("lm_", sampleTube)
      ),
    sampleNotes = 
      case_when(
        !is.na(generalNotes) ~ str_c(sampleColor, "; ", generalNotes),
        is.na(generalNotes) ~ sampleColor,
        .default = sampleColor
      ),
    lysateRemaining = "no",
    lysisDate = lysisDate,
    lysisType = case_when(
      str_detect(lysisPlate, "tube") ~ "tube",
      .default = "plate"
      ),
    lysisLoc = case_when(
      str_detect(lysisPlate, "tube") ~ str_c("t", gsub("\\s*\\([^\\)]+\\)", "", lysisPlate)),
      .default = str_c("p", lysisPlate)
      ),
    lysisRow = lysisRow,
    lysisCol = lysisCol,
    lysisVol_ul = 400 + lysisProk_ul,
    lysis_prokProp = round(lysisProk_percent/100, 2),
    lysisDuration = case_when(
      str_detect(lysisDuration, "'") ~ gsub("'", "", lysisDuration),
      str_detect(lysisDuration, "‘") ~ gsub("‘", "", lysisDuration)
    ),
    lysisProtocol = 
      case_when(
        str_detect(animalID, "xtnNeg") ~ NA,
        .default = "laLab_prok"
        ),
    lysisNotes = 
      case_when(
        lysisProtocol == "laLab_prok" ~ "Used protocol for longmire-stored blood sample ProK lysis from 2018-09-15-2018-10-25 Barcoding extractions SEPTEMBER",
        .default = NA
      ),
    xtnRemaining = "yes",
    xtnDate = xtnDate,
    xtnYear = str_sub(xtnDate, start = -4, end = length(xtnDate)),
    xtnType = case_when(
      !is.na(xtnTube) ~ "tube",
      .default = "plate"
      ),
    xtnLoc = case_when(
      !is.na(xtnTube) ~ str_c("t", xtnTube),
      .default = str_c("p", xtnPlate)
      ),
    eltn = "e1",
    xtnRow = xtnRow,
    xtnCol = xtnCol,
    xtnID = case_when(
      xtnType == "plate" ~ str_c(sampleType, "_", xtnLoc, ".", xtnRow, xtnCol, "_", eltn),
      .default = str_c(sampleType, "_", xtnLoc, "_", eltn)
      ),
    xtnProtocol = 
      case_when(
        xtnProtocol == "OT2" ~ "zymoViralMagbead_ot2",
        str_detect(xtnProtocol, "manual") ~ "zymoViralMagbead_manual",
        .default = xtnProtocol
        ),
    xtnNotes = NA,
    qubit_ng.ul = case_when(
      yieldType == "qubit" ~ ng.ul_E1,
      .default = NA
      ),
    qubitDate = case_when(
      !is.na(qubit_ng.ul) ~ yieldDate,
      .default = NA
      ),
    pico_ng.ul = case_when(
      yieldType == "picogreen" ~ ng.ul_E1,
      .default = NA
      ),
    picoDate = case_when(
      !is.na(pico_ng.ul) ~ yieldDate,
      .default = NA
      ),
    purity_260.280 = NA,
    purity_260.230 = NA,
    purityDate = NA,
    fieldNotes = case_when(
      notesField == "NULL" ~ NA,
      notesField == "BLANK" ~ NA,
      .default = notesField
      ),
    captureDate = case_when(
      str_detect(animalID, "xtnNeg") ~ NA,
      .default = captureDate
      ),
    captureYear = case_when(
      str_detect(animalID, "xtnNeg") ~ NA,
      .default = str_sub(captureDate, start = -4, end = -1)
    ),
    group = case_when(
      str_detect(animalID, "xtnNeg") ~ NA,
      .default = groupName
      ),
    animalName = case_when(
      str_detect(animalID, "xtnNeg") ~ NA,
      .default = animalName
      ),
    species = case_when(
      str_detect(animalID, "xtnNeg") ~ NA,
      species == "SFUS" ~ "LWED",
      species == "SIMP" ~ "SIMP"
      ),
    sex = case_when(
      str_detect(animalID, "xtnNeg") ~ NA,
      .default = sex
      )) %>%
  mutate("animalLog_microchip" = NA,
         "animalLog_group" = NA,
         "animalLog_name" = NA) %>%
  select(c("sampleLocated", "animalID", "sampleRemaining", "sampleType", "sampleStorage", "sampleID", "sampleNotes", "lysateRemaining", "lysisDate", "lysisType", "lysisLoc", "lysisRow", "lysisCol", "lysisVol_ul", "lysisDuration", "lysis_prokProp", "lysisProtocol", "lysisNotes", "xtnRemaining", "xtnDate", "xtnYear", "xtnType", "xtnID", "xtnLoc", "eltn", "xtnRow", "xtnCol", "xtnProtocol", "xtnNotes", "qubit_ng.ul", "qubitDate", "pico_ng.ul", "picoDate", "purity_260.280", "purity_260.230", "purityDate", "fieldNotes", "captureDate", "captureYear", "group", "animalName", "species", "sex", "animalLog_microchip", "animalLog_group", "animalLog_name"))

xtnMaster_blood_e2 <- xtnMaster_blood %>%
  mutate(
    sampleLocated = case_when(
      animalID == "BLANK" ~ NA,
      .default = "yes"
      ),
    animalID = case_when(
      animalID == "BLANK" ~ str_c("xtnNeg_p", xtnPlate, ".", xtnRow, xtnCol),
      .default = animalID
      ),
    sampleRemaining = "yes",
    sampleType = case_when(
      str_detect(animalID, "xtnNeg") ~ "xtnNeg",
      .default = "blood"
      ),
    sampleStorage = case_when(
      str_detect(animalID, "xtnNeg") ~ NA,
      .default = "longmire"
      ),
    sampleID = case_when(
      str_detect(animalID, "xtnNeg") ~ NA,
      .default = str_c("lm_", sampleTube)
      ),
    sampleNotes = 
      case_when(
        !is.na(generalNotes) ~ str_c(sampleColor, "; ", generalNotes),
        is.na(generalNotes) ~ sampleColor,
        .default = sampleColor
      ),
    lysateRemaining = "no",
    lysisDate = lysisDate,
    lysisType = case_when(
      str_detect(lysisPlate, "tube") ~ "tube",
      .default = "plate"
      ),
    lysisLoc = case_when(
      str_detect(lysisPlate, "tube") ~ str_c("t", gsub("\\s*\\([^\\)]+\\)", "", lysisPlate)),
      .default = str_c("p", lysisPlate)
      ),
    lysisRow = lysisRow,
    lysisCol = lysisCol,
    lysisVol_ul = 400 + lysisProk_ul,
    lysis_prokProp = round(lysisProk_percent/100, 2),
    lysisDuration = case_when(
      str_detect(lysisDuration, "'") ~ gsub("'", "", lysisDuration),
      str_detect(lysisDuration, "‘") ~ gsub("‘", "", lysisDuration)
    ),
    lysisProtocol = 
      case_when(
        str_detect(animalID, "xtnNeg") ~ NA,
        .default = "laLab_prok"
        ),
    lysisNotes = 
      case_when(
        lysisProtocol == "laLab_prok" ~ "Used protocol for longmire-stored blood sample ProK lysis from 2018-09-15-2018-10-25 Barcoding extractions SEPTEMBER",
        .default = NA
      ),
    xtnRemaining = "yes",
    xtnDate = xtnDate,
    xtnYear = str_sub(xtnDate, start = -4, end = length(xtnDate)),
    xtnType = case_when(
      !is.na(xtnTube) ~ "tube",
      .default = "plate"
      ),
    xtnLoc = case_when(
      !is.na(xtnTube) ~ str_c("t", xtnTube),
      .default = str_c("p", xtnPlate)),
    eltn = "e2",
    xtnRow = xtnRow,
    xtnCol = xtnCol,
    xtnID = case_when(
      xtnType == "plate" ~ str_c(sampleType, "_", xtnLoc, ".", xtnRow, xtnCol, "_", eltn),
      .default = str_c(sampleType, "_", xtnLoc, "_", eltn)
      ),
    xtnProtocol = 
      case_when(
        xtnProtocol == "OT2" ~ "zymoViralMagbead_ot2",
        str_detect(xtnProtocol, "manual") ~ "zymoViralMagbead_manual",
        .default = xtnProtocol
        ),
    xtnNotes = NA,
    qubit_ng.ul = case_when(
      yieldType == "qubit" ~ ng.ul_E2,
      .default = NA
      ),
    qubitDate = case_when(
      !is.na(qubit_ng.ul) ~ yieldDate,
      .default = NA
      ),
    pico_ng.ul = case_when(
      yieldType == "picogreen" ~ ng.ul_E2,
      .default = NA
      ),
    picoDate = case_when(
      !is.na(pico_ng.ul) ~ yieldDate,
      .default = NA
      ),
    purity_260.280 = NA,
    purity_260.230 = NA,
    purityDate = NA,
    fieldNotes = case_when(
      notesField == "NULL" ~ NA,
      notesField == "BLANK" ~ NA,
      .default = notesField
      ),
    captureDate = case_when(
      str_detect(animalID, "xtnNeg") ~ NA,
      .default = captureDate
      ),
    captureYear = case_when(
      str_detect(animalID, "xtnNeg") ~ NA,
      .default = str_sub(captureDate, start = -4, end = -1)
    ),
    group = case_when(
      str_detect(animalID, "xtnNeg") ~ NA,
      .default = groupName
      ),
    animalName = case_when(
      str_detect(animalID, "xtnNeg") ~ NA,
      .default = animalName
      ),
    species = case_when(
      str_detect(animalID, "xtnNeg") ~ NA,
      species == "SFUS" ~ "LWED",
      species == "SIMP" ~ "SIMP"
      ),
    sex = case_when(
      str_detect(animalID, "xtnNeg") ~ NA,
      .default = sex
      )) %>%
  mutate("animalLog_microchip" = NA,
         "animalLog_group" = NA,
         "animalLog_name" = NA) %>%
  select(c("sampleLocated", "animalID", "sampleRemaining", "sampleType", "sampleStorage", "sampleID", "sampleNotes", "lysateRemaining", "lysisDate", "lysisType", "lysisLoc", "lysisRow", "lysisCol", "lysisVol_ul", "lysisDuration", "lysis_prokProp", "lysisProtocol", "lysisNotes", "xtnRemaining", "xtnDate", "xtnYear", "xtnType", "xtnID", "xtnLoc", "eltn", "xtnRow", "xtnCol", "xtnProtocol", "xtnNotes", "qubit_ng.ul", "qubitDate", "pico_ng.ul", "picoDate", "purity_260.280", "purity_260.230", "purityDate", "fieldNotes", "captureDate", "captureYear", "group", "animalName", "species", "sex", "animalLog_microchip", "animalLog_group", "animalLog_name"))

xtnMaster_blood_updated27July2023 <- rbind(xtnMaster_blood_e1, xtnMaster_blood_e2)
```

Export

```{r}
write.csv(xtnMaster_blood_updated27July2023, "./extractions/tamGenetics_xtnsBlood_master_updated27July2023.csv", row.names = F)
```

## Fecal samples

```{r}
xtnMaster_fecal <- read.csv("./extractions/tamGenetics_extractionsFecal_master_21July2023.csv", na.strings = c("", "NA"))

xtnMaster_fecal_updated27July2023 <- xtnMaster_fecal %>%
  mutate(
    sampleLocated = "yes",
    animalID = animalID,
    sampleRemaining = "yes",
    sampleType = "fecal",
    sampleStorage = "rnaLater",
    sampleID = str_c("rnaL_", sampleTube),
    sampleNotes = generalNotes,
    lysateRemaining = "unk",
    lysisDate = lysisDate,
    lysisType = "tube",
    lysisLoc = str_c("t", lysisTube),
    lysisRow = NA,
    lysisCol = NA,
    lysisVol_ul = 950+20,
    lysis_prokProp = 20/lysisVol_ul,
    lysisDuration = case_when(
      lysisDuration_h == "3" ~ "03:00",
      lysisDuration_h == "20.5" ~ "20:30"
    ),
    lysisProtocol = 
      case_when(
        lysisDate == "07/08/2022" ~ "zymoFecalSoil_a_manual",
        lysisDate == "07/12/2022" ~ "zymoFecalSoil_a_manual",
        lysisDate == "08/10/2022" ~ "zymoFecalSoil_a_manual",
        lysisDate == "03/21/2023" ~ "zymoFecalSoil_b_manual"
        ),
    lysisNotes = NA,
    xtnRemaining = "yes",
    xtnDate = xtnDate,
    xtnYear = str_sub(xtnDate, start = -4, end = length(xtnDate)), xtnType = "tube",
    xtnLoc = str_c("t", xtnTube),
    eltn = tolower(elution),
    xtnRow = NA,
    xtnCol = NA,
    xtnID = case_when(
      xtnType == "plate" ~ str_c(sampleType, "_", xtnLoc, ".", xtnRow, xtnCol, "_", eltn),
      .default = str_c(sampleType, "_", xtnLoc, "_", eltn)
      ),
    xtnProtocol = 
      case_when(
        lysisDate == "07/08/2022" ~ "zymoFecalSoil_a",
        lysisDate == "07/12/2022" ~ "zymoFecalSoil_a",
        lysisDate == "08/10/2022" ~ "zymoFecalSoil_a",
        lysisDate == "03/21/2023" ~ "zymoFecalSoil_b"
        ),
    xtnNotes = NA,
    qubit_ng.ul = case_when(
      yieldType == "qubit" & eltn == "e1" ~ ng.ul_E1,
      .default = NA
      ),
    qubitDate = case_when(
      !is.na(qubit_ng.ul) ~ yieldDate,
      .default = NA
      ),
    pico_ng.ul = case_when(
      yieldType == "qubit" & eltn == "e1" ~ ng.ul_E1,
      .default = NA
      ),
    picoDate = case_when(
      !is.na(pico_ng.ul) ~ yieldDate,
      .default = NA
      ),
    purity_260.280 = A260.A280,
    purity_260.230 = A260.A230,
    purityDate = purityDate,
    fieldNotes = NA,
    captureDate = "unk",
    captureYear = "unk",
    group = groupName,
    animalName = animalName,
    species = case_when(
      str_detect(animalID, "xtnNeg") ~ NA,
      species == "SFUS" ~ "LWED",
      species == "SIMP" ~ "SIMP"
      ),
    sex = sex) %>%
  mutate("animalLog_microchip" = NA,
         "animalLog_group" = NA,
         "animalLog_name" = NA) %>%
  select(c("sampleLocated", "animalID", "sampleRemaining", "sampleType", "sampleStorage", "sampleID", "sampleNotes", "lysateRemaining", "lysisDate", "lysisType", "lysisLoc", "lysisRow", "lysisCol", "lysisVol_ul", "lysisDuration", "lysis_prokProp", "lysisProtocol", "lysisNotes", "xtnRemaining", "xtnDate", "xtnYear", "xtnType", "xtnID", "xtnLoc", "eltn", "xtnRow", "xtnCol", "xtnProtocol", "xtnNotes", "qubit_ng.ul", "qubitDate", "pico_ng.ul", "picoDate", "purity_260.280", "purity_260.230", "purityDate", "fieldNotes", "captureDate", "captureYear", "group", "animalName", "species", "sex", "animalLog_microchip", "animalLog_group", "animalLog_name"))
```

Export

```{r}
write.csv(xtnMaster_fecal_updated27July2023, "./extractions/tamGenetics_xtnsFecal_master_updated27July2023.csv", row.names = F)
```

## Hair samples

NOTE - the format for one of the capture dates was messed up; looked up capture date in capture files and adjusted below

```{r}
xtnMaster_hair <- read.csv("./extractions/tamGenetics_extractionsHair_master_21July2023.csv", na.strings = c("", "NA")) %>%
  mutate(xtnTube = gsub("\\s+", "", xtnTube))

xtnMaster_hair_e1 <- xtnMaster_hair %>%
  filter(is.na(elution)|elution != "E2") %>%
  mutate(
    hairType = case_when(
      hairType == "NULL" ~ "unk; assume plucked",
      hairType == "UNK" ~ "unk; assume plucked",
      .default = hairType
      )
    ) %>%
  mutate(
    sampleLocated = case_when(
      animalID == "BLANK" ~ NA,
      .default = located.
      ),
    animalID = case_when(
      animalID == "BLANK" ~ str_c("xtnNeg_p", xtnPlate, ".", xtnRow, xtnCol),
      .default = animalID
      ),
    sampleRemaining = case_when(
      hairDepleted. == "yes" ~ "no",
      hairDepleted. == "no" ~ "yes",
      .default = NA
    ),
    sampleType = case_when(
      str_detect(animalID, "xtnNeg") ~ "xtnNeg",
      .default = "hair"
    ),
    sampleStorage = case_when(
      str_detect(animalID, "xtnNeg") ~ NA,
      .default = "envelope"
    ),
    sampleID = str_c("env_", envelope.),
    sampleNotes = case_when(
      str_detect(hairType, "unk") ~ generalNotes,
      !str_detect(hairType, "unk") ~ str_c(hairType, "; ", generalNotes)
      ),
    lysateRemaining = NA,
    lysisDate = case_when(
      str_detect(animalID, "xtnNeg") ~ NA,
      .default = lysisDate
      ),
    lysisType = case_when(
      is.na(lysisDate) ~ NA,
      .default = "tube"
    ),
    lysisLoc = case_when(
      str_detect(animalID, "xtnNeg") ~ NA,
      str_detect(xtnTube, "A,B") ~ NA,
      .default = str_c("t", lysisTube)
      ),
    lysisRow = NA,
    lysisCol = NA,
    lysisVol_ul = 
      case_when(
        str_detect(animalID, "xtnNeg") ~ NA,
        lysisProtocol == "A" ~ "200",
        lysisProtocol == "B" ~ "200",
        lysisProtocol == "C" ~ "250",
        lysisProtocol == "D" ~ "250",
        lysisProtocol == "E" ~ "130",
        lysisProtocol == "E*" ~ "130",
        .default = NA
        ),
    lysis_prokProp = 
      case_when(
        str_detect(animalID, "xtnNeg") ~ NA,
        lysisProtocol == "A" ~ as.character(20/200),
        lysisProtocol == "B" ~ as.character(10/200),
        lysisProtocol == "C" ~ as.character(12.5/250),
        lysisProtocol == "D" ~ as.character(20/250),
        lysisProtocol == "E" ~ as.character(20/130),
        lysisProtocol == "E*" ~ as.character(20/130),
        .default = NA
      ),
    lysisDuration = case_when(
      str_detect(animalID, "xtnNeg") ~ NA,
      str_detect(xtnTube, "A,B") ~ NA,
      str_detect(lysisDuration, "'") ~ gsub("'", "", lysisDuration),
      str_detect(lysisDuration, "‘") ~ gsub("‘", "", lysisDuration),
      .default = NA
    ),
    lysisProtocol = 
      case_when(
        str_detect(animalID, "xtnNeg") ~ NA,
        lysisProtocol == "A" ~ "zymoSTB_a",
        lysisProtocol == "B" ~ "zymoSTB_b",
        lysisProtocol == "C" ~ "zymoSTB_c",
        lysisProtocol == "D" ~ "zymoSTB_d",
        lysisProtocol == "E" ~ "omegaTL_a",
        lysisProtocol == "E*" ~ "omegaTL_b",
        .default = lysisProtocol
        ),
    lysisNotes = case_when(
        str_detect(animalID, "xtnNeg") ~ NA,
        lysisProtocol == "zymoSTB_a" ~ "~4-6 hairs; 80 ul water, 90 ul Solid Tissue Buffer, 10 ul 1M DTT, 20 ul ProK",
        lysisProtocol == "zymoSTB_b" ~ "~15-20 hairs; 90 ul water, 90 ul Solid Tissue Buffer, 10 ul 1M DTT, 10 ul ProK",
        lysisProtocol == "zymoSTB_c" ~ "~10 hairs; 112.5 ul water, 112.5 ul Solid Tissue Buffer, 12.5 ul 1M DTT, 12.5 ul ProK",
        lysisProtocol == "zymoSTB_d" ~ "~10 hairs; 110 ul water, 110 ul Solid Tissue Buffer, 10 ul 1M DTT, 20 ul ProK",
        lysisProtocol == "omegaTL_a" ~ "~10 hairs; 100 ul TL Buffer, 10 ul 1M DTT, 20 ul ProK",
        lysisProtocol == "omegaTL_b" ~ "~10 hairs; 100 ul TL Buffer, 10 ul 1M DTT, 20 ul ProK; *in my lab notebook I have protocol B steps listed, yet the lysis mix calculator for protocol E - I'm pretty certain I used protocol E here, but wanted to note this discrepancy just in case",
        .default = NA
        ),
    xtnDate = xtnDate,
    xtnRemaining = case_when(
      str_detect(xtnStatus, "OUT") ~ "no",
      str_detect(xtnStatus, "nonfxnl") ~ "yes, but nonfxnl",
      !is.na(xtnDate) ~ "yes"
      ),
    xtnYear = str_sub(xtnDate, start = -4, end = length(xtnDate)),
    xtnType = case_when(
      is.na(xtnDate) ~ NA,
      !is.na(xtnTube) ~ "tube",
      .default = "plate"
      ),
    xtnLoc = case_when(
      str_detect(xtnTube, "A,B") ~ str_c("tx", gsub("A,B", "A", xtnTube)),
      str_detect(xtnTube, "WBO") ~ str_c("tx", xtnTube),
      !is.na(xtnTube) ~ str_c("t", xtnTube),
      .default = str_c("p", xtnPlate)
      ),
    eltn = case_when(
      !is.na(xtnDate) ~ "e1"
      ),
    xtnRow = xtnRow,
    xtnCol = xtnCol,
    xtnID = case_when(
      xtnType == "plate" ~ str_c(sampleType, "_", xtnLoc, ".", xtnRow, xtnCol, "_", eltn),
      .default = str_c(sampleType, "_", xtnLoc, "_", eltn)
      ),
    xtnProtocol = 
      case_when(
        xtnProtocol == "A" ~ "zymoViralMagbead_a_manual",
        xtnProtocol == "B" ~ "omegaBloodTissue_a_manual",
        xtnProtocol == "C" ~ "zymoViralMagbead_b_manual",
        xtnProtocol == "D" ~ "zymoGenomicDNA_a_manual",
        xtnProtocol == "E" ~ "zymoMagbeadPlus_a_manual",
        xtnProtocol == "F" ~ "zymoViralMagbead_a_ot2",
        xtnProtocol == "G" ~ "omegaBloodTissue_a_ot2",
        .default = xtnProtocol
      ),
    xtnNotes = case_when(
      !is.na(xtnStatus) ~ str_c(xtnNotes, "; ", xtnStatus),
      .default = xtnNotes
      ),
    qubit_ng.ul = qubit_ng.ul_e1,
    qubitDate = qubitDate,
    pico_ng.ul = NA,
    picoDate = NA,
    purity_260.280 = ndrop.260.280,
    purity_260.230 = ndrop.260.230,
    purityDate = ndropDate,
    fieldNotes = NA,
    captureDate = case_when(
      str_detect(animalID, "Neg") ~ NA,
      animalID == 56 & sampleLocated == "yes" ~ "07/18/2016",
      .default = captureDate
      ),
    captureYear = case_when(
      str_detect(animalID, "Neg") ~ NA,
      .default = str_sub(captureDate, start = -4, end = -1)
    ),
    group = case_when(
      str_detect(animalID, "Neg") ~ NA,
      .default = groupName
      ),
    animalName = case_when(
      str_detect(animalID, "Neg") ~ NA,
      .default = animalName
      ),
    species = case_when(
      str_detect(animalID, "Neg") ~ NA,
      species == "SFUS" ~ "LWED",
      species == "SIMP" ~ "SIMP"
      ),
    sex = case_when(
      str_detect(animalID, "Neg") ~ NA,
      .default = sex
      )) %>%
  dplyr::rename("animalLog_microchip" = "AnimalLog_microchip",
         "animalLog_group" = "AnimalLog_group",
         "animalLog_name" = "AnimalLog_name") %>%
  select(c("sampleLocated", "animalID", "sampleRemaining", "sampleType", "sampleStorage", "sampleID", "sampleNotes", "lysateRemaining", "lysisDate", "lysisType", "lysisLoc", "lysisRow", "lysisCol", "lysisVol_ul", "lysisDuration", "lysis_prokProp", "lysisProtocol", "lysisNotes", "xtnRemaining", "xtnDate", "xtnYear", "xtnType", "xtnID", "xtnLoc", "eltn", "xtnRow", "xtnCol", "xtnProtocol", "xtnNotes", "qubit_ng.ul", "qubitDate", "pico_ng.ul", "picoDate", "purity_260.280", "purity_260.230", "purityDate", "fieldNotes", "captureDate", "captureYear", "group", "animalName", "species", "sex", "animalLog_microchip", "animalLog_group", "animalLog_name"))

xtnMaster_hair_e2 <- xtnMaster_hair %>%
  filter(is.na(elution)|elution != "E2") %>%
  mutate(
    hairType = case_when(
      hairType == "NULL" ~ "unk; assume plucked",
      hairType == "UNK" ~ "unk; assume plucked",
      .default = hairType
      )
    ) %>%
  mutate(
    sampleLocated = case_when(
      animalID == "BLANK" ~ NA,
      .default = located.
      ),
    animalID = case_when(
      animalID == "BLANK" ~ str_c("xtnNeg_p", xtnPlate, ".", xtnRow, xtnCol),
      .default = animalID
      ),
    sampleRemaining = case_when(
      hairDepleted. == "yes" ~ "no",
      hairDepleted. == "no" ~ "yes",
      .default = NA
    ),
    sampleType = case_when(
      str_detect(animalID, "xtnNeg") ~ "xtnNeg",
      .default = "hair"
    ),
    sampleStorage = case_when(
      str_detect(animalID, "xtnNeg") ~ NA,
      .default = "envelope"
    ),
    sampleID = str_c("env_", envelope.),
    sampleNotes = case_when(
      str_detect(hairType, "unk") ~ generalNotes,
      !str_detect(hairType, "unk") ~ str_c(hairType, "; ", generalNotes)
      ),
    lysateRemaining = "unk",
    lysisDate = case_when(
      str_detect(animalID, "xtnNeg") ~ NA,
      .default = lysisDate
      ),
    lysisType = case_when(
      is.na(lysisDate) ~ NA,
      .default = "tube"
    ),
    lysisLoc = case_when(
      str_detect(animalID, "xtnNeg") ~ NA,
      str_detect(xtnTube, "A,B") ~ NA,
      .default = str_c("t", lysisTube)
      ),
    lysisRow = NA,
    lysisCol = NA,
    lysisVol_ul = 
      case_when(
        str_detect(animalID, "xtnNeg") ~ NA,
        lysisProtocol == "A" ~ "200",
        lysisProtocol == "B" ~ "200",
        lysisProtocol == "C" ~ "250",
        lysisProtocol == "D" ~ "250",
        lysisProtocol == "E" ~ "130",
        lysisProtocol == "E*" ~ "130",
        .default = lysisProtocol
        ),
    lysis_prokProp = case_when(
      str_detect(animalID, "xtnNeg") ~ NA,
      lysisProtocol == "A" ~ as.character(20/200),
      lysisProtocol == "B" ~ as.character(10/200),
      lysisProtocol == "C" ~ as.character(12.5/250),
      lysisProtocol == "D" ~ as.character(20/250),
      lysisProtocol == "E" ~ as.character(20/130),
      lysisProtocol == "E*" ~ as.character(20/130),
      .default = lysisProtocol
      ),
    lysisDuration = case_when(
      str_detect(animalID, "xtnNeg") ~ NA,
      str_detect(xtnTube, "A,B") ~ NA,
      str_detect(lysisDuration, "'") ~ gsub("'", "", lysisDuration),
      str_detect(lysisDuration, "‘") ~ gsub("‘", "", lysisDuration)
    ),
    lysisProtocol = case_when(
      str_detect(animalID, "xtnNeg") ~ NA,
      lysisProtocol == "A" ~ "zymoSTB_a",
      lysisProtocol == "B" ~ "zymoSTB_b",
      lysisProtocol == "C" ~ "zymoSTB_c",
      lysisProtocol == "D" ~ "zymoSTB_d",
      lysisProtocol == "E" ~ "omegaTL_a",
      lysisProtocol == "E*" ~ "omegaTL_b",
      .default = lysisProtocol
      ),
    lysisNotes = case_when(
        str_detect(animalID, "xtnNeg") ~ NA,
        lysisProtocol == "zymoSTB_a" ~ "~4-6 hairs; 80 ul water, 90 ul Solid Tissue Buffer, 10 ul 1M DTT, 20 ul ProK",
        lysisProtocol == "zymoSTB_b" ~ "~15-20 hairs; 90 ul water, 90 ul Solid Tissue Buffer, 10 ul 1M DTT, 10 ul ProK",
        lysisProtocol == "zymoSTB_c" ~ "~10 hairs; 112.5 ul water, 112.5 ul Solid Tissue Buffer, 12.5 ul 1M DTT, 12.5 ul ProK",
        lysisProtocol == "zymoSTB_d" ~ "~10 hairs; 110 ul water, 110 ul Solid Tissue Buffer, 10 ul 1M DTT, 20 ul ProK",
        lysisProtocol == "omegaTL_a" ~ "~10 hairs; 100 ul TL Buffer, 10 ul 1M DTT, 20 ul ProK",
        lysisProtocol == "omegaTL_b" ~ "~10 hairs; 100 ul TL Buffer, 10 ul 1M DTT, 20 ul ProK; *in my lab notebook I have protocol B steps listed, yet the lysis mix calculator for protocol E - I'm pretty certain I used protocol E here, but wanted to note this discrepancy just in case",
        .default = NA
        ),
    xtnDate = xtnDate,
    xtnRemaining = case_when(
      str_detect(xtnStatus, "OUT") ~ "no",
      str_detect(xtnStatus, "nonfxnl") ~ "yes, but nonfxnl",
      !is.na(xtnDate) ~ "yes"
      ),
    xtnYear = str_sub(xtnDate, start = -4, end = length(xtnDate)),
    xtnType = case_when(
      is.na(xtnDate) ~ NA,
      !is.na(xtnTube) ~ "tube",
      .default = "plate"
      ),
    xtnLoc = case_when(
      str_detect(xtnTube, "A,B") ~ str_c("tx", gsub("A,B", "B", xtnTube)),
      str_detect(xtnTube, "WBO") ~ str_c("tx", xtnTube),
      !is.na(xtnTube) ~ str_c("t", xtnTube),
      .default = str_c("p", xtnPlate)
      ),
    eltn = case_when(
      !is.na(xtnDate) ~ "e2"
      ),
    xtnRow = xtnRow,
    xtnCol = xtnCol,
    xtnID = case_when(
      xtnType == "plate" ~ str_c(sampleType, "_", xtnLoc, ".", xtnRow, xtnCol, "_", eltn),
      .default = str_c(sampleType, "_", xtnLoc, "_", eltn)
      ),
    xtnProtocol = 
      case_when(
        xtnProtocol == "A" ~ "zymoViralMagbead_a_manual",
        xtnProtocol == "B" ~ "omegaBloodTissue_a_manual",
        xtnProtocol == "C" ~ "zymoViralMagbead_b_manual",
        xtnProtocol == "D" ~ "zymoGenomicDNA_a_manual",
        xtnProtocol == "E" ~ "zymoMagbeadPlus_a_manual",
        xtnProtocol == "F" ~ "zymoViralMagbead_a_ot2",
        xtnProtocol == "G" ~ "omegaBloodTissue_a_ot2",
        .default = xtnProtocol
      ),
    xtnNotes = case_when(
      !is.na(xtnStatus) ~ str_c(xtnNotes, "; ", xtnStatus),
      .default = xtnNotes
      ),
    qubit_ng.ul = qubit_ng.ul_e2,
    qubitDate = qubitDate,
    pico_ng.ul = NA,
    picoDate = NA,
    purity_260.280 = ndrop.260.280,
    purity_260.230 = ndrop.260.230,
    purityDate = ndropDate,
    fieldNotes = NA,
    captureDate = case_when(
      str_detect(animalID, "xtnNeg") ~ NA,
      animalID == 56 & sampleLocated == "yes" ~ "07/18/2016",
      .default = captureDate
      ),
    captureYear = case_when(
      str_detect(animalID, "Neg") ~ NA,
      .default = str_sub(captureDate, start = -4, end = -1)
    ),
    group = case_when(
      str_detect(animalID, "xtnNeg") ~ NA,
      .default = groupName
      ),
    animalName = case_when(
      str_detect(animalID, "xtnNeg") ~ NA,
      .default = animalName
      ),
    species = case_when(
      str_detect(animalID, "xtnNeg") ~ NA,
      species == "SFUS" ~ "LWED",
      species == "SIMP" ~ "SIMP"
      ),
    sex = case_when(
      str_detect(animalID, "xtnNeg") ~ NA,
      .default = sex
      )) %>%
  dplyr::rename("animalLog_microchip" = "AnimalLog_microchip",
         "animalLog_group" = "AnimalLog_group",
         "animalLog_name" = "AnimalLog_name") %>%
  select(c("sampleLocated", "animalID", "sampleRemaining", "sampleType", "sampleStorage", "sampleID", "sampleNotes", "lysateRemaining", "lysisDate", "lysisType", "lysisLoc", "lysisRow", "lysisCol", "lysisVol_ul", "lysisDuration", "lysis_prokProp", "lysisProtocol", "lysisNotes", "xtnRemaining", "xtnDate", "xtnYear", "xtnType", "xtnID", "xtnLoc", "eltn", "xtnRow", "xtnCol", "xtnProtocol", "xtnNotes", "qubit_ng.ul", "qubitDate", "pico_ng.ul", "picoDate", "purity_260.280", "purity_260.230", "purityDate", "fieldNotes", "captureDate", "captureYear", "group", "animalName", "species", "sex", "animalLog_microchip", "animalLog_group", "animalLog_name"))

xtnMaster_hair_updated27July2023 <- rbind(xtnMaster_hair_e1, xtnMaster_hair_e2)
```

Export

```{r}
write.csv(xtnMaster_hair_updated27July2023, "./extractions/tamGenetics_xtnsHair_master_updated27July2023.csv", row.names = F)
```

# Xtns remaining after August 2023

## Import xtn records post-onsite checks & post-30x3

First import xtn records updated after on-site checks & 30x3 run

```{r}
xtnsAll_updated_postOnsiteChecks_post30x3 <- read.csv("./extractions/tamGenetics_xtnsAll_updated_postOnsiteChecks_post30x3.csv")
```

## Xtns depleted

Then update xtns depleted during the tamGenetics prep (which used xtns from all blood & hair samples) - xtns listed were either completely depleted or have a remaining volume that's too low to be usable unless rehydrated

```{r}
xtnsOut_postTamGenetics_Aug2023 <- data.frame(
  xtnID = c(
            "blood_t700_e1",
            "blood_t1275_e1",
            "hair_p3.C5_e1",
            "hair_p3.F6_e1",
            "hair_p3.F6_e2",
            "hair_p3.H10_e1",
            "hair_t76_e1",
            "hair_p3.D10_e1",
            "hair_p3.F12_e1",
            "hair_p4.F3_e1",
            "hair_p4.F3_e2",
            "hair_p4.H1_e1",
            "hair_p4.H1_e2",
            "hair_t16_e1",
            "hair_t16_e2",
            "hair_tT11.e1",
            "hair_t110_e1",
            "hair_tx23A_e1",
            "hair_tx23A_e2",
            "hair_ptamTiger1.A1_e1",
            "hair_ptamTiger1.B1_e1",
            "hair_ptamTiger2.A2_e1",
            "hair_t55_e1",
            "hair_t1_e1",
            "hair_tT1.11_e2",
            "hair_tT1.13_e1",
            "hair_tT1.14_e1",
            "hair_t2_e1",
            "hair_t49_e1",
            "hair_t51_e1",
            "hair_t7_e1"
            )) %>%
  merge(., xtnsAll_updated_postOnsiteChecks_post30x3[, c("animalID", "xtnID")], by = "xtnID", all.x = T) %>%
  mutate(xtnRemaining = "no") %>%
  select(c("animalID", "xtnID", "xtnRemaining"))
```

## Xtns mistakenly listed as depleted

There were also a few xtns that were listed as depleted, but still have volume - update those here:

```{r}
xtnsUpdate_asPresent <- data.frame(
  xtnID = c(
            "hair_p3.C9_e2",
            "hair_p3.A11_e2",
            "hair_t6_e1",
            "hair_t6_e2"
            )) %>%
  merge(., xtnsAll_updated_postOnsiteChecks_post30x3[, c("animalID", "xtnID")], by = "xtnID", all.x = T) %>%
  mutate(xtnRemaining = "yes") %>%
  select(c("animalID", "xtnID", "xtnRemaining"))
```

## Update xtn records

```{r}
xtnsAll_updated_postTamGenetics_Aug2023 <- xtnsAll_updated_postOnsiteChecks_post30x3 %>%
  mutate(
    xtnRemaining = case_when(
      xtnID %in% xtnsOut_postTamGenetics_Aug2023$xtnID ~ "no",
      xtnID %in% xtnsUpdate_asPresent$xtnID ~ "yes",
      .default = xtnRemaining
      ))

xtnsBlood_updated_postTamGenetics_Aug2023 <- xtnsAll_updated_postTamGenetics_Aug2023 %>%
  filter(sampleType == "blood")

xtnsFecal_updated_postTamGenetics_Aug2023 <- xtnsAll_updated_postTamGenetics_Aug2023 %>%
  filter(sampleType == "fecal")

xtnsHair_updated_postTamGenetics_Aug2023 <- xtnsAll_updated_postTamGenetics_Aug2023 %>%
  filter(sampleType == "hair")
```

Export

```{r}
write.csv(xtnsAll_updated_postTamGenetics_Aug2023, "./extractions/tamGenetics_xtnsBloodFecalHair_master_updated22Aug2023.csv", row.names = F)

write.csv(xtnsBlood_updated_postTamGenetics_Aug2023, "./extractions/tamGenetics_xtnsBlood_master_updated22Aug2023.csv", row.names = F)

write.csv(xtnsFecal_updated_postTamGenetics_Aug2023, "./extractions/tamGenetics_xtnsFecal_master_updated22Aug2023.csv", row.names = F)

write.csv(xtnsHair_updated_postTamGenetics_Aug2023, "./extractions/tamGenetics_xtnsHair_master_updated22Aug2023.csv", row.names = F)
```


