---
title: |
    | "How-to guide for GT-seq data analysis" 
    | "Part 2: Amplicon read counting & genotyping with GTscore" 
author: "Rachel Voyt"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
editor_options:
  markdown:
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1 Overview

Part 2 of the tamGenetics GT-seq how-to guide goes over downloading, interleaving, and performing quality checks for sequencing results from both Illumina and Nanopore platforms.

If you don't yet have your own sequencing results, proceed to the Quality Checks section and use the toy data available in 00_toyRun_seqs and 01_toyRun_interleaved.

# 2 Preparation

## 2.1 Packages

### R packages

```{r, eval = FALSE}
source("./GTscore_sourceScripts/GTscore_modified.R") # NOTE- added an option for "\\[ATGC\\]"="N" (lines 616 and 617) to replace brackets in probe

if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
BiocManager::install("Biostrings")

library(gsubfn)
library(multiqc)
library(phylotools)
library(TidyMultiqc)
library(tidyverse)
```

### Perl and modules

The GTscore pipeline includes several Perl scripts as well as Perl
modules "Algorithm::Combinatorics" and "Excel-Writer-XLSX". Use the
scripts below to install on Linux/Mac computers (more on how to install
Perl modules [here](http://www.cpan.org/modules/INSTALL.html)).

A couple notes--

-   It used to be possible to install Perl through Conda (conda install
    -c conda-forge perl), but this doesn't seem to work well anymore
-   Be sure to install both Perl and the modules in the same place
    (e.g., both in the root directory)
-   You can also just use "sudo cpan Algorithm::Combinatorics" instead
    of using the cpanimus command "cpanm" - I had to switch to using the
    regular "cpan" command because cpanm started putting my modules in
    odd places

```{bash, eval = FALSE}
# Install perl
curl -L http://xrl.us/installperlnix | bash

# Install cpanminus (makes installing perl modules easier)
cpan App::cpanminus

# Install necessary perl modules
sudo cpanm Algorithm::Combinatorics
sudo cpanm Excel::Writer::XLSX
## OR
sudo cpan Algorithm::Combinatorics
sudo cpan Excel::Writer::XLSX
```

I also recently had issues with installing perl modules with "cpanm" --
installing with "cpan" instead solved things. Generally speaking, perl
continues to be a massive pain in new fun ways every time I open this up
to re-run something.

## 2.2 Input files

The GTscore pipeline requires two types of input files: 1) a file
containing all sample file names, and 2) a file containing primer-probe
information. Because our dataset includes two different species with
some species-specific loci, I also found it helpful to create separate
files for each so that we have a more accurate read count and genotype
rates for each sample and locus.

Note that if your dataset includes multiple species, you will need a
sample metadata file to create the GTscore sample files.

### 2.4.1 Metadata

Metadata can include whatever information you wish, but at a minimum
should contain the following:

-   Sample sequence ID
-   Sample file name (sample sequence ID .fastq)
-   Individual ID
-   Sample type
-   Species
-   Sex

An example metadata file is below:

```{r, eval = FALSE}
toyRun_md <- read.csv("./toyRun/toyRun_metadata.csv")
```

### 2.4.2 Sample files

Sample file prep is split into Illumina and Nanopore workflows since
Nanopore output has previously included a few extra things that needed
to be filtered out.

**NOTE** - regardless of sequencing platform, I recommend adjusting
sample names so that they do NOT include characters such as periods,
underscores, dashes, etc. The GTscore perl scripts tend to switch these
around, which then requires formatting adjustments within the R script
(nothing hard, but a bit of a pain). I haven't taken the time to dig
into the perl scripts themselves to adjust them, so at this point it's
easiest to name samples (or change sample names) so that they don't
include anything aside from text and numbers.

#### ILLUMINA

##### All samples

Create file for all samples in bash

```{bash, eval = FALSE}
# Navigate to 01_seqrunInterleaved directory
cd ./toyPipeline/toyRun/01_toyRun_interleaved/

# Create sample file
for i in *fastq; do echo $i; done > ./../03_toyRun_gtscore/fullSet_sampleFiles.txt
```

Load into R

```{r, eval = FALSE}
sf_fullSet <- read.table("./toyRun/03_toyRun_gtscore/fullSet_sampleFiles.txt")
```

##### Species subsets

Here we're creating separate sample files for each species.

**NOTE** that we're including the negative controls in both species subsets -- e.g., "negative control #1" would be listed in the LWED sample file as well as the SIMP sample file.

```{r, eval = FALSE}
# Create subsets
lwed <- filter(md, is.na(species) | !species == "SIMP")
simp <- filter(md, is.na(species) | !species == "LWED")

sf_lwed <- sf_fullSet %>%
  filter(V1 %in% lwed$sampleFile)

sf_simp <- sf_fullSet %>%
  filter(V1 %in% simp$sampleFile)
```

##### Export sample files

Follow the file export scripts below to ensure that sample files are in
the correct format for the GTscore pipeline.

```{r, eval = FALSE}
write.table(sf_fullSet, file = "fullSet_sampleFiles.txt", sep = "\t", row.names = F, col.names = F, quote = F)

write.table(sf_LWED, file = "lwed_sampleFiles.txt", sep = "\t", row.names = F, col.names = F, quote = F)

write.table(sf_SIMP, file = "simp_sampleFiles.txt", sep = "\t", row.names = F, col.names = F, quote = F)
```

#### NANOPORE

Nanopore data may require an extra step when creating sample files,
since previous MinION output sequences included barcodes that weren't
supposed to be there. As such, I opted to create an original set of
sample files (including all MinION output), then subset that to include
only the barcodes that were used for the samples, and then create
separate files by species.

##### All barcodes

Easiest to do this part in bash--

```{bash, eval = FALSE}
for i in bar*; do echo $i; done > sampleFiles_original.txt
```

##### Barcodes of interest

First step is to clean up the metadata file - note that the script below
also includes a step to remove duplicates, as some sample entries in the
metadata file from a previous MinION run were listed multiple times
(might not be necessary for other runs).

```{r, eval = FALSE}
md <- read.csv("METADATA_FILE.csv") %>%
  filter(!is.na(Barcode)) %>%
  mutate(Barcode = str_replace(Barcode, "BC", "barcode")) %>%
  mutate(sampleID = paste(Barcode, "trimmed.fastq", sep = "_")) %>%
  mutate(Field.id.a = str_replace(Field.id.a, "Leontocebus weddeli", "LWED")) %>%
  mutate(Field.id.a = str_replace(Field.id.a, "Saguinus imperator", "SIMP")) %>%
  dplyr::rename(species = Field.id.a) %>%
  distinct()
```

Then we can subset the original sample files to only the barcodes of
interest.

```{r, eval = FALSE}
# Read in original list of sample files
sf_original <- read.table("sampleFiles_original.txt")

# Subset to barcodes of interst only
sf_fullSet <- sf_original %>%
  filter(V1 %in% md$sampleID)
```

##### Species subsets

Here again I'm including the negative controls in the sample files for
both species subsets.

```{r, eval = FALSE}
# Create subsets
lwed <- filter(md, is.na(species) | !species == "SIMP")
simp <- filter(md, is.na(species) | !species == "LWED")

sf_LWED <- sf_fullSet %>%
  filter(V1 %in% lwed$sampleID)

sf_SIMP <- sf_fullSet %>%
  filter(V1 %in% simp$sampleID)
```

##### Export sample files

Follow the file export scripts below to ensure that sample files are in
the correct format for the GTscore pipeline.

```{r, eval = FALSE}
write.table(sf_fullSet, file = "sampleFiles_fullSet.txt", sep = "\t", row.names = F, col.names = F, quote = F)
write.table(sf_LWED, file = "sampleFiles_LWED.txt", sep = "\t", row.names = F, col.names = F, quote = F)
write.table(sf_SIMP, file = "sampleFiles_SIMP.txt", sep = "\t", row.names = F, col.names = F, quote = F)
```

### 2.4.3 Primer-probe files

Each primer-probe file needs to contain the following:

1.  Locus - locus name
2.  Ploidy - in our case, all are diploid so ploidy = 2
3.  SNPpos - position of SNP in the amplicon, assuming that the first
    base = 0
4.  Allele1 - one of the options for the SNP
5.  Allele2 - the other option for the SNP
6.  Probe1 - an 8 nt sequence overlapping the SNP; contains Allele1
7.  Probe2 - same as Probe1, but contains Allele2
8.  Primer - the forward primer sequence for each locus

#### Toy data

Here I'm loading in primer-probe files that I created for previous runs.
I've divided the primer-probe files into three sets, one with all
samples (contains all loci), one for LWED samples only (excludes
SIMP-specific loci) and one for SIMP samples only (excludes
LWED-specific loci).

```{r, eval = FALSE}
pp_fullSet <- read.table("./toyRun/fullSet_primerProbeFile.txt", header = T)
pp_lwed <- read.table("./toyRun/lwed_primerProbeFile.txt", header = T)
pp_simp <- read.table("./toyRun/simp_primerProbeFile.txt", header = T)
```

#### Create your own

If you need to create these files from scratch, you'll need the
following: **TBD**

```{python}
# amplicon pos
from Bio import SeqIO
import csv
import pandas as pd
import re

primers_file="/proj/sllstore2017021/nobackup/SAM/flanking_regions/final_merged_fastas/19.02.2022_fastas_for_merging/simulating_seqs_to_troubleshoot_gt_seq/primers.tsv"
fasta_new_snps="/proj/sllstore2017021/nobackup/SAM/flanking_regions/final_merged_fastas/19.02.2022_fastas_for_merging/simulating_seqs_to_troubleshoot_gt_seq/concatenated_ALL_19.02.2022_adding_brakets_replace_snps.fasta"
probe_file="/proj/sllstore2017021/nobackup/SAM/flanking_regions/final_merged_fastas/19.02.2022_fastas_for_merging/simulating_seqs_to_troubleshoot_gt_seq/LocusInfo.csv"

primers = pd.read_csv(primers_file, sep = '\t')

def reverse_complement(dna):
    complement = {'A':'T','C':'G','G':'C','T':'A'}
    return ''.join([complement[base] for base in dna[::-1]])

pos_file = "/proj/sllstore2017021/webexport/first_geno_run/pos_file_rachel/pos_amplicon.tsv"

with open(pos_file, 'w') as output:
    output.write("SNP_name" + "\t" + "bp3_index" + "\t" + "snp_base" + "\t" + "pos_in_amplicon" + "\t" + "length_amplicon" + "\t" + "amplicon" + "\n")
    with open(fasta_new_snps) as f:
        SeqDict = {}
        pos_dict = {}
        for record in SeqIO.parse(f, "fasta"):
            sequence = str(record.seq).replace('<','').replace('>','')
            snp_name = str(record.id)
            bp3_index = snp_name.split('_')[-1]
            for index, row in primers.iterrows():
                forward_p=str(row["forward"])
                reverse_p=str(reverse_complement(row["reverse"]))
                if forward_p in sequence and reverse_p in sequence:
                    amplicon = re.search('{}(.*?){}'.format(forward_p, reverse_p), sequence).group()
                    if '[' in amplicon and ']' in amplicon:
                        start = amplicon.index('[')
                        # the following will mean the position is 0-based (start counting at 0)
                        # if you would rather have it 1-based (start at 1), replace the 1 with 2
                        pos = start + 1
                        snp = amplicon[pos]
                        length = len(amplicon) - 2
                        if snp_name not in pos_dict:
                            pos_dict[snp_name] = snp_name, bp3_index, str(snp), str(pos)
                            output.write(snp_name + "\t" + str(bp3_index) + "\t" + str(snp) + "\t" + str(pos) + "\t" + str(length) + "\t" + amplicon + "\n")
```
