---
title: "tamRun4_filePrep"
author: "Rachel Voyt"
date: "`r Sys.Date()`"
output: html_document
---

# 1 Overview

Doc is to prep files for tamRun4 GTscore pipeline and genoAnalyses so they're not so cluttered

# 2 Packages

```{r}
library(janitor)
library(tidyverse)
```

# 3 Metadata

For this run, we need to create the metadata files from scratch. We can use the extraction lists from the tamGenetics_ot2InputFiles.Rmd and merge it with the full extraction list to add metadata.

#### Extractions included in run

```{r}
tamRun4_xtns <- read.csv("./04_tamRun4/tamRun4_xtnsUsed.csv") %>%
  select(c(wellID, xtnID)) %>%
  mutate(xtnID = 
           case_when(
             xtnID == "tH2O.1" ~ "fecal_pcr1Neg.1",
             xtnID == "tH2O.2" ~ "fecal_pcr1Neg.2",
             .default = xtnID
           ))
```

#### Add metadata

```{r}
xtnsAll_updated_postTamGenetics_Aug2023 <- read.csv("./extractions/tamGenetics_xtnsBloodFecalHair_master_updated22Aug2023.csv") %>%
  dplyr::rename("sampleName" = "sampleID") %>%
  mutate(
    xtnID = 
      case_when(
        str_detect(xtnID, "Neg") & xtnLoc %in% c("p1", "p2") ~ str_c("blood_", xtnID),
        str_detect(xtnID, "Neg") & xtnLoc %in% c("p3", "p4") ~ str_c("hair_", xtnID),
        .default = xtnID
      )
  )

tamRun4_sampleSheet <- read.csv("./04_tamRun4/00_illumina/ill_tamRun4_30x3_sampleSheet.csv") %>%
  filter(str_detect(X.Header., "tamRun4")) %>%
  dplyr::rename("sampleID" = "X.Header.",
         "plate" = "X",
         "well" = "X.1",
         "i7_index" = "X.2",
         "i7_seq" = "X.3",
         "i5_index" = "X.4",
         "i5_seq" = "X.5") %>%
  select(c("sampleID", "plate", "well", "i7_index", "i5_index")) %>%
  mutate(
    well = gsub("01", "1", well),
    well = gsub("02", "2", well),
    well = gsub("03", "3", well),
    well = gsub("04", "4", well),
    well = gsub("05", "5", well),
    well = gsub("06", "6", well),
    well = gsub("07", "7", well),
    well = gsub("08", "8", well),
    well = gsub("09", "9", well),
  ) %>%
  mutate(wellID = row_number())

md_tamRun4 <- merge(tamRun4_xtns, tamRun4_sampleSheet, by = "wellID") %>%
  select(-wellID) %>%
  merge(., xtnsAll_updated_postTamGenetics_Aug2023, by = "xtnID", all.x = T) %>%
  mutate(sampleID = gsub("_","-", sampleID)) %>%
  arrange(plate, as.numeric(str_sub(well, 2, -1)), str_sub(well, 1, 1)) %>%
  mutate(sampleFile = paste0(sampleID, ".fastq"),
         sampleID = gsub("-", "\\.", sampleID),
         sampleType = 
           case_when(
             xtnID == "fecal_pcr1Neg.1" ~ "pcr1Neg",
             xtnID == "fecal_pcr1Neg.2" ~ "pcr1Neg",
             str_detect(xtnID, "Neg") ~ "xtnNeg",
             .default = sampleType
           ),
         animalID = case_when(
           xtnID == "fecal_pcr1Neg.1" ~ "pcr1Neg.1",
           xtnID == "fecal_pcr1Neg.2" ~ "pcr1Neg.2",
           .default = animalID
         )) %>%
  select(c("sampleID", "sampleFile", "i5_index", "i7_index", "well", "animalID", "sampleName", "sampleType", "sampleNotes", "lysisDate", "lysisType", "lysisLoc", "lysisRow", "lysisCol", "lysisVol_ul", "lysisDuration", "lysis_prokProp", "lysisProtocol", "lysisNotes", "xtnID", "xtnDate", "xtnYear", "xtnType", "xtnLoc", "eltn", "xtnRow", "xtnCol", "xtnProtocol", "xtnNotes", "qubit_ng.ul", "qubitDate", "pico_ng.ul", "picoDate", "purity_260.280", "purity_260.230", "purityDate", "fieldNotes", "captureDate", "captureYear", "group", "animalName", "species", "sex", "animalLog_microchip", "animalLog_group", "animalLog_name")) %>%
  arrange(sampleID) %>%
  mutate(
    sampleID_unique = case_when(
      str_detect(sampleType, "Neg") ~ animalID,
    .default = str_c(animalID, "_", sampleType)
  )) %>%
  relocate(sampleID_unique)

# remove extra fecal t30
md_tamRun4 <- md_tamRun4[-64,]

write.csv(md_tamRun4, "./04_tamRun4/tamRun4_metadata_ill.csv", row.names = F)
```

## clean md

```{r, eval=FALSE}
# Illumina samples
md_ill_samplesAll <- read.csv("./04_tamRun4/tamRun4_metadata_ill.csv") %>%
  dplyr::rename("sampleID_ill" = "sampleID",
                "sampleFile_ill" = "sampleFile")

# Nanopore samples
md_ont_samplesAll <- read.csv("./04_tamRun4/tamRun4_metadata_ont.csv") %>%
  dplyr::rename("sampleID_ont" = "sampleID",
                "sampleFile_ont" = "sampleFile")

# Metadata w/both Illumina and Nanopore sampleIDs/sampleFiles
md_illONT_samplesAll <- merge(md_ill_samplesAll, md_ont_samplesAll[, c("sampleID_unique", "sampleID_ont", "sampleFile_ont")], by = "sampleID_unique") %>%
  relocate(c(sampleID_ont, sampleFile_ont), .after = sampleFile_ill) %>%
  arrange(sampleID_ill) %>%
  select(-c(animalLog_microchip, animalLog_group, animalLog_name))

# EXPORT
write.csv(md_illONT_samplesAll, "./04_tamRun4/tamRun4_metadata_illONT.csv", row.names = F)
```

## add sampleID_md

```{r}
md_illONT_samplesAll <- read.csv("./04_tamRun4/tamRun4_metadata_illONT.csv")

tamRun4_md <- md_illONT_samplesAll %>%
  mutate(
    sampleID_ill_md = case_when(
      str_detect(sampleID_unique, "xtnNeg") ~ str_c("s", str_sub(sampleID_ill, -3, -1), "_", "xtnNeg_xtnNeg_xtnNeg"),
      str_detect(sampleID_unique, "pcr1Neg") ~ str_c("s", str_sub(sampleID_ill, -3, -1), "_", "pcr1Neg_pcr1Neg_pcr1Neg"),
      .default = str_c("s", str_sub(sampleID_ill, -3, -1), "_", tolower(species), "_", animalID, "_", sampleType)
    )
  ) %>%
  relocate(sampleID_ill_md, .after = sampleID_ill)

write.csv(tamRun4_md, "./04_tamRun4/tamRun4_metadata_illONT.csv", row.names = F)
```

