---
title: "GTscorePipeline_tamRun5"
author: "Rachel Voyt"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1 Packages

```{r load packages, include=FALSE}
source("./GTscore_sourceScripts/GTscore_modified.R") # NOTE- added N=ATGC to script for later analyses (not included in original)
library(gsubfn)
library(phylotools)
library(TidyMultiqc)
library(tidyverse)
if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager", version = "3.16")
BiocManager::install("Biostrings")
```

GTscore also requires Perl and the packages "Algorithm::Combinatorics" and "Excel-Writer-XLSX". Perform the following in terminal to install on Linux. NOTE -- be sure to install both perl and the modules in root. See how to install perl packages [here](http://www.cpan.org/modules/INSTALL.html).

NOTE - the old way of installing perl through conda (conda install -c conda-forge perl) doesn't play nice anymore; better to install directly from perl:

```{bash, eval = FALSE}
# Install perl
curl -L http://xrl.us/installperlnix | bash

# Install cpanminus (makes installing perl modules easier)
cpan App::cpanminus

# Install necessary perl modules
sudo cpan Algorithm::Combinatorics
sudo cpan Excel::Writer::XLSX
```

I also recently had issues with installing perl modules with "cpanm" -- installing with "cpan" instead solved things. Generally speaking, perl continues to be a massive pain in new fun ways every time I open this up to re-run something.

# 2 Overview

## 2.1 About the pipeline

This pipeline is to analyze data from tamGenetics_run5 and includes all steps needed to analyze MiSeq data output, including how to 1) download MiSeq sequencing files, 2) do quality checks, 3) create the necessary files for analysis in GTscore, and 4) run the GTscore pipeline itself.

The GTscore pipeline was originally created by Garrett McKinney (see the associated github [here](https://github.com/gjmckinney/GTscore)). Here, I combine the scripts included in his original "GTscore_pipeline.R" as well as his original "README.md" - there's a good bit of overlap between these files, but a few scripts exist in one but not the other. I've also included separate analyses for each species, since not all not all loci are informative for both species.

## 2.2 About the data

The sequencing data used in this pipeline is from the full set of hair and blood samples and includes data from two MiSeq runs of this library. There are a total of **440 samples**, including xxx hair samples (xxx LWED, xxx SIMP), xxx blood samples (xxx LWED, xxx SIMP), and 9 negative controls.

**NOTE** that of these samples, 3 were not actually added to PCR1 (xtn volume was out). Treat these as negatives when analysing the data:
*   plate 3 well F6
*   plate 3 well B12
*   plate 3 well C1

### Metadata

**Quick-load metadata**

The metadata file loaded below has been updated to reflect discrepancies related to duplicate sample numbers; see metadataReconciliation.Rmd for details.

```{r}
md_tamRun5_v5 <- read.csv("./metadataReconciliation/tamRun5_metadata_v5.csv")
```

### Original md file creation

For this run, we need to create the metadata files from scratch. We can use the extraction lists from the tamGenetics_ot2InputFiles.Rmd and merge it with the full extraction list to add metadata.

##### Extractions included in run

```{r}
# all xtns
xtnsAll_updated_postOnsiteChecks_post30x3 <- read.csv("./extractions/tamGenetics_xtnsAll_updated_postOnsiteChecks_post30x3.csv")

# p1 (96 total)
tamRun5_xtns_p1 <- read.csv("./robotProtocols/inputFiles/tamGenetics_pcr1setup_forOT2_p12.csv") %>%
  filter(str_detect(ot2set, "set1|set2")) %>%
  select(c("reagent", "dest_well")) %>%
  dplyr::rename("xtnID" = "reagent") %>%
  mutate(
    plate = 1
  )

# p2 (96 total)
tamRun5_xtns_p2 <- read.csv("./robotProtocols/inputFiles/tamGenetics_pcr1setup_forOT2_p12.csv") %>%
  filter(str_detect(ot2set, "set4|set5")) %>%
  select(c("reagent", "dest_well")) %>%
  dplyr::rename("xtnID" = "reagent") %>%
  mutate(
    plate = 2
  )

# strip 1 (6 total)
tamRun5_xtns_s1 <- xtnsAll_updated_postOnsiteChecks_post30x3 %>%
  filter(sampleType == "blood") %>%
  filter(str_detect(lysisLoc, "t")) %>%
  filter(eltn == "e1") %>%
  arrange(as.numeric(str_sub(xtnLoc, 2, -1))) %>%
  select(xtnID) %>%
  mutate(
    dest_well = c("C7", "D7", "E7", "F7", "G7", "H7"),
    plate = 5
  )

# p3 (96 total)
## NOTE that hair_p3.B12 (well B12), hair_p3.F6 (well F6), and hair_t16 (well C1) were OUT; treat as negatives for analyses
tamRun5_xtns_p3_manualTransfers <- data.frame(
  xtnID = c("hair_p3.A11_e2", "hair_p3.C9_e2", "hair_p3.B12_e1"),
  dest_well = c("A11", "C9", "B12")
)

tamRun5_xtns_p3 <- read.csv("./robotProtocols/inputFiles/tamGenetics_pcr1_p3ab.csv") %>%
  filter(str_detect(ot2set, "set1|set2")) %>%
  filter(dest_slot == 2) %>%
  select(c("reagent", "dest_well")) %>%
  dplyr::rename("xtnID" = "reagent") %>%
  rbind(., tamRun5_xtns_p3_manualTransfers) %>%
  mutate(
    plate = 3
  )

# p4 (96 total)
tamRun5_xtns_p4 <- read.csv("./robotProtocols/inputFiles/tamGenetics_pcr1_p4ab.csv") %>%
  filter(str_detect(ot2set, "set1")) %>%
  select(c("reagent", "dest_well")) %>%
  dplyr::rename("xtnID" = "reagent") %>%
  mutate(
    plate = 4
  )

# p5 (50 total)
tamRun5_xtns_p5 <- read.csv("./robotProtocols/inputFiles/tamGenetics_pcr1_p5ab.csv") %>%
  filter(str_detect(ot2set, "set1|set2|set3")) %>%
  select(c("reagent", "dest_well")) %>%
  dplyr::rename("xtnID" = "reagent") %>%
  mutate(
    plate = 5
  )

# all negatives (9 total)
tamRun5_xtns_allNegs <- rbind(
  tamRun5_xtns_p1,
  tamRun5_xtns_p2,
  tamRun5_xtns_s1,
  tamRun5_xtns_p3,
  tamRun5_xtns_p4,
  tamRun5_xtns_p5
) %>%
  filter(str_detect(xtnID, "Neg"))

# all samples (430 total)
tamRun5_xtns_allSamples <- rbind(
  tamRun5_xtns_p1,
  tamRun5_xtns_p2,
  tamRun5_xtns_s1,
  tamRun5_xtns_p3,
  tamRun5_xtns_p4,
  tamRun5_xtns_p5
) %>%
  filter(!str_detect(xtnID, "Neg"))
```

##### Add metadata

**NOTE** that sex for tamRun5.302 is missing in metadata; should be M -- adjust below:

```{r}
xtnsAll_updated_postTamGenetics_Aug2023 <- read.csv("./extractions/tamGenetics_xtnsBloodFecalHair_master_updated22Aug2023.csv")

tamRun5_sampleSheet <- read.csv("./05_tamRun5/M83285346-300V2.csv") %>%
  filter(str_detect(X.Header., "tamRun5")) %>%
  dplyr::rename("sampleID" = "X.Header.",
         "plate" = "X",
         "well" = "X.1",
         "i7_index" = "X.2",
         "i7_seq" = "X.3",
         "i5_index" = "X.4",
         "i5_seq" = "X.5") %>%
  select(c("sampleID", "plate", "well", "i7_index", "i5_index")) %>%
  mutate(
    plate = gsub("p", "", plate),
    well = gsub("01", "1", well),
    well = gsub("02", "2", well),
    well = gsub("03", "3", well),
    well = gsub("04", "4", well),
    well = gsub("05", "5", well),
    well = gsub("06", "6", well),
    well = gsub("07", "7", well),
    well = gsub("08", "8", well),
    well = gsub("09", "9", well),
  )

md_tamRun5 <- rbind(
  tamRun5_xtns_p1,
  tamRun5_xtns_p2,
  tamRun5_xtns_s1,
  tamRun5_xtns_p3,
  tamRun5_xtns_p4,
  tamRun5_xtns_p5
) %>%
  dplyr::rename("well" = "dest_well") %>%
  merge(., xtnsAll_updated_postTamGenetics_Aug2023, by = "xtnID", all.x = T) %>%
  dplyr::rename("sampleName" = "sampleID") %>%
  arrange(plate, as.numeric(str_sub(well, 2, -1)), str_sub(well, 1, 1)) %>%
  merge(., tamRun5_sampleSheet, by.x = c("plate", "well"), by.y = c("plate", "well"), all.x = T) %>%
  mutate(sampleID = gsub("_","-", sampleID)) %>%
  mutate(sampleFile = paste0(sampleID, ".fastq")) %>%
  select(c("sampleID", "sampleFile", "i5_index", "i7_index", "well", "animalID", "sampleName", "sampleType", "sampleNotes", "lysisDate", "lysisType", "lysisLoc", "lysisRow", "lysisCol", "lysisVol_ul", "lysisDuration", "lysis_prokProp", "lysisProtocol", "lysisNotes", "xtnDate", "xtnYear", "xtnType", "xtnLoc", "eltn", "xtnRow", "xtnCol", "xtnProtocol", "xtnNotes", "qubit_ng.ul", "qubitDate", "pico_ng.ul", "picoDate", "purity_260.280", "purity_260.230", "purityDate", "fieldNotes", "captureDate", "captureYear", "group", "animalName", "species", "sex", "animalLog_microchip", "animalLog_group", "animalLog_name")) %>%
  arrange(sampleID) %>%
  mutate(
    sex = 
      case_when(
        sampleID == "tamRun5.302" ~ "M",
        .default = sex
    )
  )
```

Export
```{r}
write.csv(md_tamRun5, "./05_tamRun5/03_run5GTscore/tamRun5_metadata.csv", row.names = F)
```

# 3 Prep sequencing results

## 3.1 Download sequences from BaseSpace

To download all sample files, it's easiest to use the BaseSpace Command Line Interface (CLI).

To install, follow the scripts provided here: <https://developer.basespace.illumina.com/docs/content/documentation/cli/cli-overview>

Note that even if it's already installed, you may need to re-authenticate via the command below. Running it will give a link; click to open in a browser and complete authentication.

```{bash, eval = F}
bs auth
```

Once authentication is complete, you can download the desired files as follows:

```{bash, eval = F}
# Check Project IDs (outputs Project IDs for tamRun3_L1Raquel and Unindexed Reads)
bs list projects

# Sample sequences (tamRun3_L1Raquel, tamRun3b_L1)
bs download project -i 384884910 -o /home/rachelvoyt/Documents/UT-Grad/Development/repos/tamGenetics_primatesPeru/03_tamRun3/00_run3Seqs --extension=fastq.g]

# Unindexed Reads
bs download project -i 374582211 -o /home/rachelvoyt/Documents/UT-Grad/Development/repos/tamGenetics_primatesPeru/03_tamRun3/00_run3Seqs/unindexedReads --extension=fastq.gz
```

## 3.2 Extract fast.gz files

```{bash Extract sequencing results, eval = FALSE}
# Navigate to the directory containing the .fastq.gz files from the sequencing run
cd 00_run5Seqs

# Move all .gz files out of their individual folders and into the current directory & remove empty folders
mv */* ./
find -empty -delete

# Extract all files
gunzip -r .
```

## 3.3 Merge the two runs

```{bash}
# Trim filenames to make them more manageable
for f in *R1_001.fastq; do g="${f%%_S*}_R1.fastq"; mv "${f}" "${g}"; done
for f in *R2_001.fastq; do g="${f%%_S*}_R2.fastq"; mv "${f}" "${g}"; done

# Edit first run filenames to be tamRun5a (repeat for runb if needed)
rename 's/tamRun5/tamRun5a/' *.fastq

# Create tamRun5ab merged fastq files
for i in tamRun5a-*; do 
    f=${i##*tamRun5a}; 
    cat $i tamRun5b$f > tamRun5cat$f; 
done

# Check to see that total runs in tamRun5cat matches tamRun5a + tamRun5b
for i in tamRun5a-*; do 
    f=${i##*tamRun5a}; 
    
    # Calculate the number of reads in the concatenated file
    r1=$(( $(cat tamRun5cat$f | wc -l) / 4 ));
    echo "${i%_*} concatenated $r1";
    
    # Calculate the number of reads in the separate files and sum them
    r2=$(( $(cat $i | wc -l) / 4 + $(cat tamRun5b$f | wc -l) / 4 ));
    echo "${i%_*} separated $r2";
    
    # Calculate the difference
    diff=$(( r1 - r2 ));
    echo "${i%_*} diff $diff";
done
```

## 3.4 Interleave files

To prep for the GTscore pipeline, we need to combine reads 1 & 2 for each sample into a single interleaved file - leave a set of non-interleaved files in the original folder though to keep for fastq quality checks.

Note that we have interleaved files for tamRun3, tamRun3b, and the merged version of the two.

```{bash Interleave fasta seqs, eval = FALSE}
# Interleave reads 1 & 2 for each sample and place in a new folder
cd ./../

for i in ./00_run5Seqs/*_R1_001.fastq; do name=$(basename $i _R1_001.fastq); reformat.sh in=00_run5Seqs/${name}_R1_001.fastq in2=00_run5Seqs/${name}_R2_001.fastq out=02_run5Interleaved/${name}_interleaved.fastq; done
```

# 4 Quality checks

To get sequence quality scores for each sample, we can use 'fastqc' paired with 'MultiQC' -- 'fastp' is better for paired-end reads, but doesn't play nice when it comes to getting quality scores from the MultiQC report. Note that we'll be running quality checks on reads 1 and 2 separately (vs. interleaved).

## 4.1 fastqc + multiqc

*Note*: use "\--interactive" to force plots to tell you which line belongs to which sample, otherwise it will remove this option since there are so many samples in this run

```{bash eval = F}
# Activate fastqc environment
cd ../01_run5QualityChecks
conda activate fastqc-env

# Run fastqc & move fastqc files to quality checks folder
for i in ../00_run5Seqs/tamRun5a*; do fastqc $i; done
mv ../00_run5Seqs/*fastqc* .

for i in ../00_run5Seqs/tamRun5b*; do fastqc $i; done
mv ../00_run5Seqs/*fastqc* .

# Activate multiqc environment
conda activate multiqc-env

# Run MultiQC on all files within the quality checks folder
multiqc --interactive .
```

## 4.2 TidyMultiqc

To convert the MultiQC into something more usable in R, we can use the package 'TidyMultiqc', which converts the 'multiqc_data.json' file into tidy data frames.

Load FastQC MultiQC report to R

```{r}
MultiQCfastQCpath <- file.path("/home/rachelvoyt/Documents/UT-Grad/Development/repos/tamGenetics_primatesPeru/05_tamRun5/01_run5QualityChecks/multiqc_data", "multiqc_data.json")

MultiQCfastQC <- TidyMultiqc::load_multiqc(MultiQCfastQCpath, sections = c("general", "raw"))
```

### 4.2.1 Obtain median sequence quality scores

For the analyses in this section, I followed [this vignette](https://cran.r-project.org/web/packages/TidyMultiqc/vignettes/TidyMultiqc.html) provided by the makers of the TidyMultiqc package.

MultiQC reports do not provide a numerical summary statistic for read quality; they only have mapping quality and pass/fails for the per-base sequence quality. We instead need to pull this data from one of the plots -- I'm using the "Per Sequence Quality Scores" plot here:

```{r}
df <- TidyMultiqc::load_multiqc(
  MultiQCfastQCpath, 
  sections = 'plot',
  plots = "fastqc_per_sequence_quality_scores_plot")
```

This provides a nested data frame as a set of x, y pairs. As it's a histogram plot, we know that the `x` value is the quality score, and `y` is the number of times that score has been counted.

We can use tidyr to unnest the data, HistDat to create a HistDat object for each group, then purr to map each plot data frame into a row of summary statistics:

```{r}
df_unNest <- df %>%
  dplyr::mutate(
    purrr::map_dfr(plot.fastqc_per_sequence_quality_scores_plot, function(plot_df){
      hist = HistDat::HistDat(vals=plot_df$x, counts = plot_df$y)
      list(
        mean_qc = mean(hist),
        median_qc = median(hist),
        max_qc = max(hist)
      )
    }),
    plot.fastqc_per_sequence_quality_scores_plot = NULL
  )
```

This dataframe has separate rows for reads 1 & 2; we can rearrange that to make it easier to work with for our purposes. Since we're most interested in median quality, we'll just pull that value for each sample.

```{r}
seqQC_r1 <- df_unNest %>%
  select(c("metadata.sample_id", "median_qc")) %>%
  filter(grepl("R1", metadata.sample_id)) %>%
  rename(medianQC_r1 = median_qc) %>%
  mutate(sampleNo = substr(metadata.sample_id, start = 9, stop = 11))
seqQC_r2 <- df_unNest %>%
  select(c("metadata.sample_id", "median_qc")) %>%
  filter(grepl("R2", metadata.sample_id)) %>%
  rename(medianQC_r2 = median_qc) %>%
  mutate(sampleNo = substr(metadata.sample_id, start = 9, stop = 11))
seqQC <- merge(seqQC_r1, seqQC_r2, by = "sampleNo") %>%
  select(c("sampleNo", "medianQC_r1", "medianQC_r2")) %>%
  arrange(sampleNo)
```

Let's just take a peek at the distribution of sequence quality scores for now; we'll use it in our other analyses later on:

```{r}
ggplot(seqQC, aes(medianQC_r1)) +
  geom_bar(stat = "count") +
  theme_bw()

ggplot(seqQC, aes(medianQC_r2)) +
  geom_bar(stat = "count") +
  theme_bw()
```

# 5 Create primer-probe & sample files

Prior to running the GTscore pipeline itself, we need to make a few files, including 1) sample files 2) primer-probe files for each species. I'm separating them out like this so that we have more accurate read counts and genotype rates for each sample and locus.

## 5.1 Sample files

We need three sets of sample files total: 1) all samples 2) LWED samples only + all negatives 3) SIMP samples only + all negatives.

### 5.1.1 All samples

Create file for all samples in bash

```{bash}
cd ./../02_run5Interleaved

for i in tamRun5a*; do echo $i; done > ./../03_run5GTscore/run5a_sampleFiles_fullSet.txt

for i in tamRun5b*; do echo $i; done > ./../03_run5GTscore/run5b_sampleFiles_fullSet.txt

for i in tamRun5cat*; do echo $i; done > ./../03_run5GTscore/run5cat_sampleFiles_fullSet.txt
```

Load into R

```{r}
sf_fullSet_a <- read.table("./05_tamRun5/03_run5GTscore/run5a_sampleFiles_fullSet.txt")
sf_fullSet_b <- read.table("./05_tamRun5/03_run5GTscore/run5b_sampleFiles_fullSet.txt")
sf_fullSet_cat <- read.table("./05_tamRun5/03_run5GTscore/run5cat_sampleFiles_fullSet.txt")
```

### 5.1.2 Species subsets

Use md file to create sample file species subsets. Note that we're including the negative controls in both species subsets 

```{r}
# Create subsets
lwed <- filter(md_tamRun5_v5, is.na(species) | !species == "SIMP")
simp <- filter(md_tamRun5_v5, is.na(species) | !species == "LWED")

sf_lwed_a <- sf_fullSet_a %>%
  filter(V1 %in% lwed$sampleFile)
sf_lwed_b <- sf_fullSet_b %>%
  filter(V1 %in% lwed$sampleFile)
sf_lwed_cat <- sf_fullSet_cat %>%
  filter(V1 %in% lwed$sampleFile)

sf_simp_a <- sf_fullSet_a %>%
  filter(V1 %in% simp$sampleFile)
sf_simp_b <- sf_fullSet_b %>%
  filter(V1 %in% simp$sampleFile)
sf_simp_cat <- sf_fullSet_cat %>%
  filter(V1 %in% simp$sampleFile)

# Export sample files
write.table(sf_lwed_a, file = "./05_tamRun5/03_run5GTscore/run5a_sampleFiles_lwed.txt", sep = "\t", row.names = F, col.names = F, quote = F)
write.table(sf_lwed_b, file = "./05_tamRun5/03_run5GTscore/run5b_sampleFiles_lwed.txt", sep = "\t", row.names = F, col.names = F, quote = F)
write.table(sf_lwed_cat, file = "./05_tamRun5/03_run5GTscore/run5cat_sampleFiles_lwed.txt", sep = "\t", row.names = F, col.names = F, quote = F)

write.table(sf_simp_a, file = "./05_tamRun5/03_run5GTscore/run5a_sampleFiles_simp.txt", sep = "\t", row.names = F, col.names = F, quote = F)
write.table(sf_simp_b, file = "./05_tamRun5/03_run5GTscore/run5b_sampleFiles_simp.txt", sep = "\t", row.names = F, col.names = F, quote = F)
write.table(sf_simp_cat, file = "./05_tamRun5/03_run5GTscore/run5cat_sampleFiles_simp.txt", sep = "\t", row.names = F, col.names = F, quote = F)
```

## 5.2 Primer probe files

Each primer-probe file contains the following:

1.  Locus - locus name
2.  Ploidy - in our case, all are diploid so ploidy = 2
3.  SNPpos - position of SNP in the amplicon, assuming that the first base = 0
4.  Allele1 - one of the options for the SNP
5.  Allele2 - the other option for the SNP
6.  Probe1 - a sequence containing bases 8 upstream and 8 downstream from SNP (including SNP); contains Allele1
7.  Probe2 - same as Probe1, but contains Allele2
8.  Primer - the forward primer sequence for each locus

### 5.2.1 Load in primer-probe files

Here I'm loading in primer-probe files that I created for previous runs, then filtering them to reflect only those in SNP primer pool version 3. I've also divided the primer-probe files into three sets, one with all loci, one for LWED loci (excludes SIMP-specific loci) and one for SIMP loci (excludes LWED-specific loci).

```{r}
primerList_v3 <- read.csv("./primers/03_lociChoices/tamGenetics_primerList_v3.csv")

pp_fullSet <- read.table("./03_tamRun3/03_run3GTscore/primerProbeFile_fullSet.txt", header = T) %>%
  filter(Locus %in% primerList_v3$locus)

pp_lwed <- pp_fullSet %>%
  filter(!str_detect(Locus, "SIMP"))

pp_simp <- pp_fullSet %>%
  filter(!str_detect(Locus, "LWED"))

# Export
write.table(pp_fullSet, "./05_tamRun5/03_run5GTscore/primerProbeFileV3_fullSet.txt", sep = "\t", row.names = F, col.names = T, quote = F)

write.table(pp_lwed, "./05_tamRun5/03_run5GTscore/primerProbeFileV3_LWED.txt", sep = "\t", row.names = F, col.names = T, quote = F)

write.table(pp_simp, "./05_tamRun5/03_run5GTscore/primerProbeFileV3_SIMP.txt", sep = "\t", row.names = F, col.names = T, quote = F)
```

# 5 Count reads for amplicons: No coverage cutoff

Once the files are created, we can proceed with the first step in the GTscore pipeline - counting amplicon reads. We'll be doing doing this for the full set of data (all samples, all loci) as well as the species subsets of the data to allow for more accurate analyses of loci and sample performance.

## 5.1 tamRun5a

(primer-probe file from project_data is the same as in tamRun5 files; just wanted the shorter path)

```{r count reads for amplicons, eval=FALSE}
# all samples, all loci (includes negatives)
system2("perl",
        args="./project_scripts/GTscore_AmpliconReadCounter_modified.pl -p ./primerProbeFile_panelv3_fullSet.txt --files ./05_tamRun5/03_run5GTscore/run5a_sampleFiles_fullSet.txt --inDir ./05_tamRun5/02_run5Interleaved/ --outDir ./05_tamRun5/03_run5GTscore/ --prefix run5a_fullSet_")

# pos samples (no negs), dual loci
system2("perl",
        args="./project_scripts/GTscore_AmpliconReadCounter_modified.pl -p ./project_data/primerProbeFile_panelv3_dual.txt --files ./05_tamRun5/03_run5GTscore/run5a_sampleFiles_pos.txt --inDir ./05_tamRun5/02_run5Interleaved/ --outDir ./05_tamRun5/03_run5GTscore/ --prefix run5a_posDual_")

# lwed samples, lwed + dual loci




# LWED
system2("perl",
        args="./project_scripts/GTscore_AmpliconReadCounter_modified.pl -p ./05_tamRun5/03_run5GTscore/primerProbeFileV3_LWED.txt --files ./05_tamRun5/03_run5GTscore/run5a_sampleFiles_lwed.txt --inDir ./05_tamRun5/02_run5Interleaved/ --outDir ./05_tamRun5/03_run5GTscore/ --prefix run5a_lwed_")

# SIMP
system2("perl",
        args="./project_scripts/GTscore_AmpliconReadCounter_modified.pl -p ./05_tamRun5/03_run5GTscore/primerProbeFileV3_SIMP.txt --files ./05_tamRun5/03_run5GTscore/run5a_sampleFiles_simp.txt --inDir ./05_tamRun5/02_run5Interleaved/ --outDir ./05_tamRun5/03_run5GTscore/ --prefix run5a_simp_")
```

AmpliconReadCounter.pl outputs a LocusTable file and an AlleleReads file for single-SNP and haplotype data, plus two summary files.

The default names for these files are as follows, with LWED and SIMP appended to files associated with species-specific analyses:

-   LocusTable_singleSNPs.txt - locus name, ploidy, and alleles for each SNP
-   AlleleReads_singleSNPs.txt - counts for each SNP allele (rows are loci, columns are individuals)
-   LocusTable_haplotypes.txt - same as above (doesn't matter for us since we're only using single SNPs)
-   AlleleReads_haplotypes.txt - same as above (doesn't matter for us since we're only using single SNPs)
-   GTscore_individualSummary.txt - counts by individual of total reads, off-target reads, primer-only reads, and primer probe reads
-   GTscore_locusSummary.txt - counts by locus of primer reads and primer probe reads

# 6 Genotyping

Genotyping by sequencing generally requires a minimum of 10x coverage to be considered reliable, though others implement even higher cutoffs (e.g., Bootsma et al. (2020), who also used the GTscore PolyGen script for genotyping, added a 30x cutoff). However, the GTscore genotyping script doesn't include this cutoff, and will provide genotypes with as little as one read per allele for heterozygous calls and 3 reads (vs. 0) for a homozygous call.

As such, I'm adding a filtering step here to try both 10x and 30x cutoffs, accomplished by recoding all loci with \<10 or \<30 reads to "0" so that loci with lower than 10x/30x coverage will not receive genotypes.

## 6.1 0x coverage

### Full set genos

```{r}
#load locus table and 0x allele reads file
fullSet_singleSNP_locusTable <- read.delim("./05_tamRun5/03_run5GTscore/fullSet_LocusTable_singleSNPs.txt", header = TRUE, stringsAsFactors = FALSE)

fullSet_singleSNP_alleleReads_0x <- read.delim("./05_tamRun5/03_run5GTscore/fullSet_AlleleReads_singleSNPs.txt", header = TRUE, row.names = 1, stringsAsFactors = FALSE)

#generate singleSNP genotypes using the polyGen algorithm, adjust "0" formatting
fullSet_polyGenResults_singleSNP_0x <- polyGen(fullSet_singleSNP_locusTable, fullSet_singleSNP_alleleReads_0x)

# write results
write.table(fullSet_polyGenResults_singleSNP_0x, "./05_tamRun5/03_run5GTscore/fullSet_polyGenResults_singleSNP_0x.txt", quote = FALSE, sep = "\t")
```

## 6.2 10x coverage

### Recode loci \<10x coverage

Step 1: Recode loci with \<10x coverage in the full set of allele read counts

```{r}
# Read in allele counts for the full dataset
readCounts_original <- read.table("./05_tamRun5/03_run5GTscore/fullSet_AlleleReads_singleSNPs.txt")

# Create read counts copy for 10x coverage
readCounts10x <- readCounts_original

# Set up a function to sum the read counts per allele for each locus, using package gsubfn
repl <- function(x) gsubfn("(\\d+),(\\d+)", ~ as.numeric(x) + as.numeric(y), paste(x))

# Then apply the function to readCounts to sum each set of allele reads for each locus
readCounts_sum10x <- replace(readCounts10x, TRUE, lapply(readCounts10x, repl)) %>%
  mutate(across(everything(),as.numeric))

# Recode <10x loci with "0"
readCounts10x[readCounts_sum10x < 10] <- "0,0"
```

Step 2: Create lwed & simp subsets & create new AlleleReads_singleSNPs files

```{r}
# Reformat lists with . vs. -
lwed_point <- lwed %>%
  mutate(sampleID = gsub("-", "\\.", sampleID)) %>%
  select(sampleID) %>%
  pull(., sampleID)
simp_point <- simp %>%
  mutate(sampleID = gsub("-", "\\.", sampleID)) %>%
  select(sampleID) %>%
  pull(., sampleID)

# Get lists of lwed- and simp-specific loci
lwed_loci <- read.delim("./05_tamRun5/03_run5GTscore/LWED_LocusTable_singleSNPs.txt",header=TRUE,stringsAsFactors=FALSE)
simp_loci <- read.delim("./05_tamRun5/03_run5GTscore/SIMP_LocusTable_singleSNPs.txt",header=TRUE,stringsAsFactors=FALSE)

readCounts10x_lwed <- readCounts10x %>%
  select(any_of(lwed_point)) %>%
  rownames_to_column("Locus") %>%
  filter(Locus %in% lwed_loci$Locus_ID) %>%
  column_to_rownames("Locus")
readCounts10x_simp <- readCounts10x %>%
  select(any_of(simp_point)) %>%
  rownames_to_column("Locus") %>%
  filter(Locus %in% simp_loci$Locus_ID) %>%
  column_to_rownames("Locus")
  
# Export new 10x AlleleReads_singleSNPs files
write.table(readCounts10x,"./05_tamRun5/03_run5GTscore/fullSet_AlleleReads_singleSNPs_10x.txt",quote=FALSE,sep="\t")
write.table(readCounts10x_lwed,"./05_tamRun5/03_run5GTscore/LWED_AlleleReads_singleSNPs_10x.txt",quote=FALSE,sep="\t")
write.table(readCounts10x_simp,"./05_tamRun5/03_run5GTscore/SIMP_AlleleReads_singleSNPs_10x.txt",quote=FALSE,sep="\t")
```

Genotyping is accomplished using the polyGen function. The genotyping algorithm is described in McKinney et al. 2018 and is a maximum likelihood algorithm capable of genotyping any number of alleles and ploidy per locus. This allows genoyping of single SNPs as well as microhaplotypes, and loci with elevated ploidy.

Two arguments are required for polyGen, the locusTable and alleleReads files output by AmpliconReadCounter.

Optional arguments for polyGen are:

-   p_thresh - threshold p-value for likelihood ratio test (default 0.05)
-   epsilon - error rate for genotyping model (default 0.01)

**NOTE** that only primer probe reads are used in genotyping!

Note also that I'm making three genotype files, one for all loci vs. all samples and two species-specific sets - this allows our genotyping metrics to be more accurate.

### Full set genos

```{r}
#load locus table and 10x allele reads file
fullSet_singleSNP_locusTable <- read.delim("./05_tamRun5/03_run5GTscore/fullSet_LocusTable_singleSNPs.txt", header = TRUE, stringsAsFactors = FALSE)
fullSet_singleSNP_alleleReads_10x <- read.delim("./05_tamRun5/03_run5GTscore/fullSet_AlleleReads_singleSNPs_10x.txt", header = TRUE, row.names = 1, stringsAsFactors = FALSE)

#generate singleSNP genotypes using the polyGen algorithm, adjust "0" formatting
fullSet_polyGenResults_singleSNP_10x <- polyGen(fullSet_singleSNP_locusTable, fullSet_singleSNP_alleleReads_10x)

# write results
write.table(fullSet_polyGenResults_singleSNP_10x, "./05_tamRun5/03_run5GTscore/fullSet_polyGenResults_singleSNP_10x.txt", quote = FALSE, sep = "\t")
```

### Species subsets

#### LWED

```{r Genotyping}
#load locus table and 10x allele reads file
LWED_singleSNP_locusTable <- read.delim("./05_tamRun5/03_run5GTscore/LWED_LocusTable_singleSNPs.txt", header = TRUE, stringsAsFactors = FALSE)
LWED_singleSNP_alleleReads_10x <- read.delim("./05_tamRun5/03_run5GTscore/LWED_AlleleReads_singleSNPs_10x.txt", header = TRUE, row.names = 1, stringsAsFactors = FALSE)

#generate singleSNP genotypes using the polyGen algorithm
LWED_polyGenResults_singleSNP_10x <- polyGen(LWED_singleSNP_locusTable, LWED_singleSNP_alleleReads_10x)

#write results
write.table(LWED_polyGenResults_singleSNP_10x,"./05_tamRun5/03_run5GTscore/LWED_polyGenResults_singleSNP_10x.txt", quote = FALSE, sep = "\t")
```

#### SIMP

```{r}
#load locus table and 10x allele reads file
SIMP_singleSNP_locusTable <- read.delim("./05_tamRun5/03_run5GTscore/SIMP_LocusTable_singleSNPs.txt", header = TRUE, stringsAsFactors = FALSE)
SIMP_singleSNP_alleleReads_10x <- read.delim("./05_tamRun5/03_run5GTscore/SIMP_AlleleReads_singleSNPs_10x.txt", header = TRUE, row.names = 1, stringsAsFactors = FALSE)

#generate singleSNP genotypes using the polyGen algorithm
SIMP_polyGenResults_singleSNP_10x <- polyGen(SIMP_singleSNP_locusTable, SIMP_singleSNP_alleleReads_10x)

#write results
write.table(SIMP_polyGenResults_singleSNP_10x,"./05_tamRun5/03_run5GTscore/SIMP_polyGenResults_singleSNP_10x.txt", quote = FALSE, sep = "\t")
```

## 6.3 30x coverage

### Recode loci \<30x coverage

Step 1: Recode loci with \<30x coverage in the full set of allele read counts

```{r}
# Read in allele counts for the full dataset
readCounts_original <- read.table("./05_tamRun5/03_run5GTscore/fullSet_AlleleReads_singleSNPs.txt")
readCounts30x <- readCounts_original

# Set up a function to sum the read counts per allele for each locus, using package gsubfn
repl <- function(x) gsubfn("(\\d+),(\\d+)", ~ as.numeric(x) + as.numeric(y), paste(x))

# Then apply the function to readCounts to sum each set of allele reads for each locus
readCounts_sum30x <- replace(readCounts30x, TRUE, lapply(readCounts30x, repl)) %>%
  mutate(across(everything(),as.numeric))

# Recode <30x loci with "0"
readCounts30x[readCounts_sum30x < 30] <- "0,0"
```

Step 2: Create lwed & simp subsets & create new AlleleReads_singleSNPs files

```{r}
# Reformat lists with . vs. -
lwed_point <- lwed %>%
  mutate(sampleID = gsub("-", "\\.", sampleID)) %>%
  select(sampleID) %>%
  pull(., sampleID)
simp_point <- simp %>%
  mutate(sampleID = gsub("-", "\\.", sampleID)) %>%
  select(sampleID) %>%
  pull(., sampleID)

# Get lists of lwed- and simp-specific loci
lwed_loci <- read.delim("./05_tamRun5/03_run5GTscore/LWED_LocusTable_singleSNPs.txt",header=TRUE,stringsAsFactors=FALSE)
simp_loci <- read.delim("./05_tamRun5/03_run5GTscore/SIMP_LocusTable_singleSNPs.txt",header=TRUE,stringsAsFactors=FALSE)

readCounts30x_lwed <- readCounts30x %>%
  select(any_of(lwed_point)) %>%
  rownames_to_column("Locus") %>%
  filter(Locus %in% lwed_loci$Locus_ID) %>%
  column_to_rownames("Locus")
readCounts30x_simp <- readCounts30x %>%
  select(any_of(simp_point)) %>%
  rownames_to_column("Locus") %>%
  filter(Locus %in% simp_loci$Locus_ID) %>%
  column_to_rownames("Locus")
  
# Export new AlleleReads_singleSNPs files
write.table(readCounts30x,"./05_tamRun5/03_run5GTscore/fullSet_AlleleReads_singleSNPs_30x.txt",quote=FALSE,sep="\t")
write.table(readCounts30x_lwed,"./05_tamRun5/03_run5GTscore/LWED_AlleleReads_singleSNPs_30x.txt",quote=FALSE,sep="\t")
write.table(readCounts30x_simp,"./05_tamRun5/03_run5GTscore/SIMP_AlleleReads_singleSNPs_30x.txt",quote=FALSE,sep="\t")
```

Genotyping is accomplished using the polyGen function. The genotyping algorithm is described in McKinney et al. 2018 and is a maximum likelihood algorithm capable of genotyping any number of alleles and ploidy per locus. This allows genoyping of single SNPs as well as microhaplotypes, and loci with elevated ploidy.

Two arguments are required for polyGen, the locusTable and alleleReads files output by AmpliconReadCounter.

Optional arguments for polyGen are:

-   p_thresh - threshold p-value for likelihood ratio test (default 0.05)
-   epsilon - error rate for genotyping model (default 0.01)

**NOTE** that only primer probe reads are used in genotyping!

Note also that I'm making three genotype files, one for all loci vs. all samples and two species-specific sets - this allows our genotyping metrics to be more accurate.

### Full set genos

```{r}
#load locus table and 10x allele reads file
fullSet_singleSNP_locusTable <- read.delim("./05_tamRun5/03_run5GTscore/fullSet_LocusTable_singleSNPs.txt", header = TRUE, stringsAsFactors = FALSE)
fullSet_singleSNP_alleleReads_30x <- read.delim("./05_tamRun5/03_run5GTscore/fullSet_AlleleReads_singleSNPs_30x.txt", header = TRUE, row.names = 1, stringsAsFactors = FALSE)

#generate singleSNP genotypes using the polyGen algorithm, adjust "0" formatting
fullSet_polyGenResults_singleSNP_30x <- polyGen(fullSet_singleSNP_locusTable, fullSet_singleSNP_alleleReads_30x)

# write results
write.table(fullSet_polyGenResults_singleSNP_30x, "./05_tamRun5/03_run5GTscore/fullSet_polyGenResults_singleSNP_30x.txt",quote = FALSE, sep = "\t")
```

### Species subsets

#### LWED

```{r Genotyping}
#load locus table and 30x allele reads file
LWED_singleSNP_locusTable <- read.delim("./05_tamRun5/03_run5GTscore/LWED_LocusTable_singleSNPs.txt", header = TRUE, stringsAsFactors = FALSE)
LWED_singleSNP_alleleReads_30x <- read.delim("./05_tamRun5/03_run5GTscore/LWED_AlleleReads_singleSNPs_30x.txt", header = TRUE, row.names = 1, stringsAsFactors = FALSE)

#generate singleSNP genotypes using the polyGen algorithm
LWED_polyGenResults_singleSNP_30x <- polyGen(LWED_singleSNP_locusTable, LWED_singleSNP_alleleReads_30x)

#write results
write.table(LWED_polyGenResults_singleSNP_30x, "./05_tamRun5/03_run5GTscore/LWED_polyGenResults_singleSNP_30x.txt", quote = FALSE, sep = "\t")
```

#### SIMP

```{r}
#load locus table and 30x allele reads file
SIMP_singleSNP_locusTable <- read.delim("./05_tamRun5/03_run5GTscore/SIMP_LocusTable_singleSNPs.txt", header = TRUE, stringsAsFactors = FALSE)
SIMP_singleSNP_alleleReads_30x <- read.delim("./05_tamRun5/03_run5GTscore/SIMP_AlleleReads_singleSNPs_30x.txt", header = TRUE, row.names = 1, stringsAsFactors = FALSE)

#generate singleSNP genotypes using the polyGen algorithm
SIMP_polyGenResults_singleSNP_30x <- polyGen(SIMP_singleSNP_locusTable, SIMP_singleSNP_alleleReads_30x)

#write results
write.table(SIMP_polyGenResults_singleSNP_30x, "./05_tamRun5/03_run5GTscore/SIMP_polyGenResults_singleSNP_30x.txt", quote = FALSE, sep = "\t")
```

# 7 Data summaries

## 7.1 Locus summaries

### 0x summaries

```{r}
fullSet_singleSNP_alleleReads_0x <- read.delim("./05_tamRun5/03_run5GTscore/fullSet_AlleleReads_singleSNPs.txt", header = TRUE, row.names = 1, stringsAsFactors = FALSE)

fullSet_singleSNP_locusTable <- read.delim("./05_tamRun5/03_run5GTscore/fullSet_LocusTable_singleSNPs.txt", header = TRUE, stringsAsFactors = FALSE)

# polygenResults from above

#summarize single SNP results
fullSet_singleSNP_summary_0x <- summarizeGTscore(fullSet_singleSNP_alleleReads_0x, fullSet_singleSNP_locusTable, fullSet_polyGenResults_singleSNP_0x)

#write results
write.table(fullSet_singleSNP_summary_0x, "./05_tamRun5/03_run5GTscore/fullSet_singleSNP_summary_0x.txt", quote = FALSE, sep = "\t", row.names = FALSE)
```

### 10x summaries

Locus summary metrics are based on data without filtering for 10x/30x coverage EXCEPT for genotype rates.

#### Summarize single SNP results for loci

The summarizeGTscore command generates summary data for each locus in table form. The summary data includes genotype rate, average read depth, minor (least frequent) allele frequency, major (most frequent) allele frequency, alleles per locus, and frequency per allele. Minor allele frequency is a common metric for filtering loci that are likely to be uninformative for population genetics; however, loci with haplotype alleles may have an allele with very low frequency but still have appreciable frequency at multiple other alleles. Because of this, the major allele frequency is included in output, as well as the observed frequencies for all alleles at a given locus.

```{r}
# non-filtered singleSNP_alleleReads
fullSet_singleSNP_alleleReads <- read.delim("./05_tamRun5/03_run5GTscore/fullSet_AlleleReads_singleSNPs.txt", header = TRUE, row.names = 1, stringsAsFactors = FALSE)

LWED_singleSNP_alleleReads <- read.delim("./05_tamRun5/03_run5GTscore/LWED_AlleleReads_singleSNPs.txt", header = TRUE, row.names = 1, stringsAsFactors = FALSE)

SIMP_singleSNP_alleleReads <- read.delim("./05_tamRun5/03_run5GTscore/SIMP_AlleleReads_singleSNPs.txt", header = TRUE, row.names = 1, stringsAsFactors = FALSE)

#summarize single SNP results
fullSet_singleSNP_summary_10x <- summarizeGTscore(fullSet_singleSNP_alleleReads, fullSet_singleSNP_locusTable, fullSet_polyGenResults_singleSNP_10x)

LWED_singleSNP_summary_10x <- summarizeGTscore(LWED_singleSNP_alleleReads, LWED_singleSNP_locusTable, LWED_polyGenResults_singleSNP_10x)

SIMP_singleSNP_summary_10x <- summarizeGTscore(SIMP_singleSNP_alleleReads, SIMP_singleSNP_locusTable, SIMP_polyGenResults_singleSNP_10x)

#write results
write.table(fullSet_singleSNP_summary_10x, "./05_tamRun5/03_run5GTscore/fullSet_singleSNP_summary_10x.txt", quote = FALSE, sep = "\t", row.names = FALSE)
write.table(LWED_singleSNP_summary_10x, "./05_tamRun5/03_run5GTscore/LWED_singleSNP_summary_10x.txt", quote = FALSE, sep = "\t", row.names = FALSE)
write.table(SIMP_singleSNP_summary_10x, "./05_tamRun5/03_run5GTscore/SIMP_singleSNP_summary_10x.txt", quote = FALSE, sep = "\t", row.names = FALSE)
```

#### Generate locus summary files with single SNP summaries + GTscore locus summaries created earlier

I also found it helpful to create a "master" locus summary file - this way, we have a single file for all loci that provides performance metrics that are specific to that locus's species. This file includes data from both the locus summaries and single SNP summaries created earlier.

In doing so, this means that...

-   values for LWED-specific loci reflect performance with LWED individuals only
-   values for SIMP-specific loci reflect performance with SIMP individuals only
-   all other loci reflect performance with LWED and SIMP individuals combined

```{r}
# Load locus summaries (not filtered for coverage)
ls <- read.table("./05_tamRun5/03_run5GTscore/fullSet_GTscore_locusSummary.txt", header = T, sep = "\t")
lwed_ls <- read.table("./05_tamRun5/03_run5GTscore/LWED_GTscore_locusSummary.txt", header = T, sep = "\t")
simp_ls <- read.table("./05_tamRun5/03_run5GTscore/SIMP_GTscore_locusSummary.txt", header = T, sep = "\t")

# Separate locus summaries and singleSNP summaries into results from shared and species-specific loci analyses
## Locus summaries
ls1 <- ls %>%
  filter(!str_detect(Locus, "LWED|SIMP"))
lwed1_ls <- lwed_ls %>%
  filter(str_detect(Locus, "LWED"))
simp1_ls <- simp_ls %>%
  filter(str_detect(Locus, "SIMP"))

## SingleSNP summaries
ss_10x <- fullSet_singleSNP_summary_10x %>%
  filter(!str_detect(Locus_ID, "LWED|SIMP"))
lwed_ss_10x <- LWED_singleSNP_summary_10x %>%
  filter(str_detect(Locus_ID, "LWED"))
simp_ss_10x <- SIMP_singleSNP_summary_10x %>%
  filter(str_detect(Locus_ID, "SIMP"))

# Recombine locus summaries & single snp summary files, then merge the two
ls_recombine <- rbind(ls1, lwed1_ls, simp1_ls)

ss_recombine_10x <- rbind(ss_10x, lwed_ss_10x, simp_ss_10x)
ss_recombine_10x$Locus_ID <- sub("_[^_]+$", "\\1", ss_recombine_10x$Locus_ID)

ls_ss_10x <- merge(ls_recombine, ss_recombine_10x, by.x = "Locus", by.y = "Locus_ID") 

# Export
write.csv(ls_ss_10x, "./05_tamRun5/03_run5GTscore/summaryFiles/master_locusSummary_10x.csv", row.names = F)
```

#### Plots

plot genotype rate

```{r locus genotype rate, warning=FALSE}
ggplot()+geom_histogram(data=ls_ss_10x,aes(x=GenotypeRate),binwidth=0.03)+xlim(-0.01,1.01)+
  labs(title="Locus Genotype Rate", x="Genotype Rate", y="Count")+
  theme_bw()+theme(plot.title=element_text(hjust=0.5),plot.subtitle=element_text(hjust=0.5))
```

plot average read depth for single SNP data

```{r locus read depth}
ggplot()+geom_histogram(data=ls_ss_10x,aes(x=AvgReadDepth),binwidth=1)+
  labs(title="Average Read Depth per SNP", x="Average Read Depth", y="Count")+
  theme_bw()+theme(plot.title=element_text(hjust=0.5),plot.subtitle=element_text(hjust=0.5))
```

plot genotype rate relative to average depth

```{r locus genotype rate vs read depth}
ggplot()+geom_point(data=ls_ss_10x,aes(x=AvgReadDepth,y=GenotypeRate))+ylim(0,1)+
  labs(title="Genotype Rate vs Average Depth per SNP", x="Average Depth", y="Genotype Rate")+
  theme_bw()+theme(plot.title=element_text(hjust=0.5),plot.subtitle=element_text(hjust=0.5))
```

plot distribution of minor allele frequency

```{r histogram of MAF, warning=FALSE}
ggplot()+geom_histogram(data=ls_ss_10x,aes(x=minAF),binwidth=0.01)+
  labs(title="Minor Allele Frequency Single SNP", x="Minor Allele Frequency", y="Count")+
  theme_bw()+theme(plot.title=element_text(hjust=0.5),plot.subtitle=element_text(hjust=0.5))
```

plot distribution of major allele frequency

```{r histogram of MajAF, warning=FALSE}
ggplot()+geom_histogram(data=ls_ss_10x,aes(x=majAF),binwidth=0.01)+
  labs(title="Major Allele Frequency Single SNP", x="Major Allele Frequency", y="Count")+
  theme_bw()+theme(plot.title=element_text(hjust=0.5),plot.subtitle=element_text(hjust=0.5))
```

### 30x summaries

Locus summary metrics are based on data without filtering for 10x/30x coverage EXCEPT for genotype rates.

#### Summarize single SNP results for loci

The summarizeGTscore command generates summary data for each locus in table form. The summary data includes genotype rate, average read depth, minor (least frequent) allele frequency, major (most frequent) allele frequency, alleles per locus, and frequency per allele. Minor allele frequency is a common metric for filtering loci that are likely to be uninformative for population genetics; however, loci with haplotype alleles may have an allele with very low frequency but still have appreciable frequency at multiple other alleles. Because of this, the major allele frequency is included in output, as well as the observed frequencies for all alleles at a given locus.

```{r}
# non-filtered singleSNP_alleleReads
fullSet_singleSNP_alleleReads <- read.delim("./05_tamRun5/03_run5GTscore/fullSet_AlleleReads_singleSNPs.txt", header = TRUE, row.names = 1, stringsAsFactors = FALSE)

LWED_singleSNP_alleleReads <- read.delim("./05_tamRun5/03_run5GTscore/LWED_AlleleReads_singleSNPs.txt", header = TRUE, row.names = 1, stringsAsFactors = FALSE)

SIMP_singleSNP_alleleReads <- read.delim("./05_tamRun5/03_run5GTscore/SIMP_AlleleReads_singleSNPs.txt", header = TRUE, row.names = 1, stringsAsFactors = FALSE)

#summarize single SNP results
fullSet_singleSNP_summary_30x <- summarizeGTscore(fullSet_singleSNP_alleleReads, fullSet_singleSNP_locusTable, fullSet_polyGenResults_singleSNP_30x)

LWED_singleSNP_summary_30x <- summarizeGTscore(LWED_singleSNP_alleleReads, LWED_singleSNP_locusTable, LWED_polyGenResults_singleSNP_30x)

SIMP_singleSNP_summary_30x <- summarizeGTscore(SIMP_singleSNP_alleleReads, SIMP_singleSNP_locusTable, SIMP_polyGenResults_singleSNP_30x)

#write results
write.table(fullSet_singleSNP_summary_30x, "./05_tamRun5/03_run5GTscore/fullSet_singleSNP_summary_30x.txt", quote = FALSE, sep = "\t", row.names = FALSE)
write.table(LWED_singleSNP_summary_30x, "./05_tamRun5/03_run5GTscore/LWED_singleSNP_summary_30x.txt", quote = FALSE, sep = "\t", row.names = FALSE)
write.table(SIMP_singleSNP_summary_30x, "./05_tamRun5/03_run5GTscore/SIMP_singleSNP_summary_30x.txt", quote = FALSE, sep = "\t", row.names = FALSE)
```

#### Generate locus summary files with single SNP summaries + GTscore locus summaries created earlier

I also found it helpful to create a "master" locus summary file - this way, we have a single file for all loci that provides performance metrics that are specific to that locus's species. This file includes data from both the locus summaries and single SNP summaries created earlier.

In doing so, this means that...

-   values for LWED-specific loci reflect performance with LWED individuals only
-   values for SIMP-specific loci reflect performance with SIMP individuals only
-   all other loci reflect performance with LWED and SIMP individuals combined

```{r}
# Load locus summaries (not filtered for coverage)
ls <- read.table("./05_tamRun5/03_run5GTscore/fullSet_GTscore_locusSummary.txt", header = T, sep = "\t")
lwed_ls <- read.table("./05_tamRun5/03_run5GTscore/LWED_GTscore_locusSummary.txt", header = T, sep = "\t")
simp_ls <- read.table("./05_tamRun5/03_run5GTscore/SIMP_GTscore_locusSummary.txt", header = T, sep = "\t")

# Separate locus summaries and singleSNP summaries into results from shared and species-specific loci analyses
## Locus summaries
ls1 <- ls %>%
  filter(!str_detect(Locus, "LWED|SIMP"))
lwed1_ls <- lwed_ls %>%
  filter(str_detect(Locus, "LWED"))
simp1_ls <- simp_ls %>%
  filter(str_detect(Locus, "SIMP"))

## SingleSNP summaries
ss_30x <- fullSet_singleSNP_summary_30x %>%
  filter(!str_detect(Locus_ID, "LWED|SIMP"))
lwed_ss_30x <- LWED_singleSNP_summary_30x %>%
  filter(str_detect(Locus_ID, "LWED"))
simp_ss_30x <- SIMP_singleSNP_summary_30x %>%
  filter(str_detect(Locus_ID, "SIMP"))

# Recombine locus summaries & single snp summary files, then merge the two
ls_recombine <- rbind(ls1, lwed1_ls, simp1_ls)

ss_recombine_30x <- rbind(ss_30x, lwed_ss_30x, simp_ss_30x)
ss_recombine_30x$Locus_ID <- sub("_[^_]+$", "\\1", ss_recombine_30x$Locus_ID)

ls_ss_30x <- merge(ls_recombine, ss_recombine_30x, by.x = "Locus", by.y = "Locus_ID") 

# Export
write.csv(ls_ss_30x, "./05_tamRun5/03_run5GTscore/summaryFiles/master_locusSummary_30x.csv", row.names = F)
```

#### Plots

plot genotype rate

```{r locus genotype rate, warning=FALSE}
ggplot()+geom_histogram(data=ls_ss_30x,aes(x=GenotypeRate),binwidth=0.03)+xlim(-0.01,1.01)+
  labs(title="Locus Genotype Rate", x="Genotype Rate", y="Count")+
  theme_bw()+theme(plot.title=element_text(hjust=0.5),plot.subtitle=element_text(hjust=0.5))
```

plot average read depth for single SNP data

```{r locus read depth}
ggplot()+geom_histogram(data=ls_ss_30x,aes(x=AvgReadDepth),binwidth=1)+
  labs(title="Average Read Depth per SNP", x="Average Read Depth", y="Count")+
  theme_bw()+theme(plot.title=element_text(hjust=0.5),plot.subtitle=element_text(hjust=0.5))
```

plot genotype rate relative to average depth

```{r locus genotype rate vs read depth}
ggplot()+geom_point(data=ls_ss_30x,aes(x=AvgReadDepth,y=GenotypeRate))+ylim(0,1)+
  labs(title="Genotype Rate vs Average Depth per SNP", x="Average Depth", y="Genotype Rate")+
  theme_bw()+theme(plot.title=element_text(hjust=0.5),plot.subtitle=element_text(hjust=0.5))
```

plot distribution of minor allele frequency

```{r histogram of MAF, warning=FALSE}
ggplot()+geom_histogram(data=ls_ss_30x,aes(x=minAF),binwidth=0.01)+
  labs(title="Minor Allele Frequency Single SNP", x="Minor Allele Frequency", y="Count")+
  theme_bw()+theme(plot.title=element_text(hjust=0.5),plot.subtitle=element_text(hjust=0.5))
```

plot distribution of major allele frequency

```{r histogram of MajAF, warning=FALSE}
ggplot()+geom_histogram(data=ls_ss_30x,aes(x=majAF),binwidth=0.01)+
  labs(title="Major Allele Frequency Single SNP", x="Major Allele Frequency", y="Count")+
  theme_bw()+theme(plot.title=element_text(hjust=0.5),plot.subtitle=element_text(hjust=0.5))
```

## 7.2 Sample summaries

Sample summary metrics are based on data without 10x/30x coverage cutoffs for everything EXCEPT genotype rate.

### 10x summaries

load sample summaries from AmpliconReadCounter (not filtered for coverage)

```{r}
# Full set
fullSet_GTscore_individualSummary <- read.delim("./05_tamRun5/03_run5GTscore/fullSet_GTscore_individualSummary.txt", header = TRUE, stringsAsFactors = FALSE) %>%
  mutate(Sample = gsub("-", "\\.", Sample))

# LWED
LWED_GTscore_individualSummary <-  read.delim("./05_tamRun5/03_run5GTscore/LWED_GTscore_individualSummary.txt", header = TRUE, stringsAsFactors = FALSE) %>%
  mutate(Sample = gsub("-", "\\.", Sample))

# SIMP
SIMP_GTscore_individualSummary <- read.delim("./05_tamRun5/03_run5GTscore/SIMP_GTscore_individualSummary.txt", header = TRUE, stringsAsFactors = FALSE) %>%
  mutate(Sample = gsub("-", "\\.", Sample))  
```

#### Summarize single SNP summaries for samples

The summarizeSamples function changes the "-" to a "." -- be sure to change this back.

```{r}
fullSet_singleSNP_sampleSummary_10x <- summarizeSamples(fullSet_polyGenResults_singleSNP_10x, fullSet_singleSNP_alleleReads_10x)

write.csv(fullSet_singleSNP_sampleSummary_10x, "./05_tamRun5/03_run5GTscore/fullSet_singleSNP_sampleSummary_10x.csv", row.names = F)

LWED_singleSNP_sampleSummary_10x <- summarizeSamples(LWED_polyGenResults_singleSNP_10x, LWED_singleSNP_alleleReads_10x)

SIMP_singleSNP_sampleSummary_10x <- summarizeSamples(SIMP_polyGenResults_singleSNP_10x, SIMP_singleSNP_alleleReads_10x)
```

#### Combine AmpliconReadCounter individual summary with GTscore sample summary data

```{r}
# make new copies & add column to note which loci-set the values are based on
LWED_gtIndivSummary_10x <- LWED_GTscore_individualSummary %>%
  mutate(lociSet = "LWED") %>%
  merge(., LWED_singleSNP_sampleSummary_10x, by.x = "Sample", by.y = "sample")

SIMP_gtIndivSummary_10x <- SIMP_GTscore_individualSummary %>%
  mutate(lociSet = "SIMP") %>%
  merge(., SIMP_singleSNP_sampleSummary_10x, by.x = "Sample", by.y = "sample")

# Combine LWED & SIMP to form one, adjust sample name
lwedSIMP_gtIndivSummary_10x <- rbind(LWED_gtIndivSummary_10x, SIMP_gtIndivSummary_10x) %>%
  mutate(
    sampleFile = gsub("\\.", "-", Sample),
    sampleFile = paste0(sampleFile, ".fastq")
         )
```

#### Create master sample summary file & add metadata

```{r}
# Merge metadata & fullSet sample summaries
master_sampleSummary_10x <- merge(lwedSIMP_gtIndivSummary_10x, md_tamRun5, by = "sampleFile")

# Export
write.csv(master_sampleSummary_10x, "./05_tamRun5/03_run5GTscore/summaryFiles/master_sampleSummary_10x.csv", row.names = F)
```

#### Plots

plot histogram of genotype rate

```{r histogram of genotype rate, warning=FALSE}
ggplot()+geom_histogram(data=master_sampleSummary_10x,aes(x=GenotypeRate),binwidth=0.01)+xlim(-0.01,1.01)+
  labs(title="Sample Genotype Rate (10x)", x="Genotype Rate", y="Count")+
  theme_bw()+theme(plot.title=element_text(hjust=0.5),plot.subtitle=element_text(hjust=0.5))
```

plot histogram of heterozygosity

```{r histogram of heterozygosity, warning=FALSE}
ggplot()+geom_histogram(data=master_sampleSummary_10x,aes(x=Heterozygosity),binwidth=0.03)+xlim(-0.01,1.01)+
  labs(title="Sample Heterozygosity", x="Heterozygosity", y="Count")+
  theme_bw()+theme(plot.title=element_text(hjust=0.5),plot.subtitle=element_text(hjust=0.5))
```

plot genotype rate vs primer probe reads

```{r genotype rate vs primer probe reads}
#dashed line added at 90% genotype rate, this is not a strict threshold, just a goal to aim for
ggplot(data=master_sampleSummary_10x,aes(x=Primer.Probe.Reads,y=GenotypeRate,color=sampleType))+
  geom_point(stat = "identity")+
  geom_smooth(method = "loess")+
  labs(title="Genotype Rate vs Total Reads per Sample", x="Primer Probe Reads", y="Genotype Rate")+
  theme_bw()+theme(plot.title=element_text(hjust=0.5),plot.subtitle=element_text(hjust=0.5))+
  geom_hline(yintercept=0.5,lty="dashed")
```

Samples with unusually high heterozygosity may be contaminated or have elevated ploidy. [[Given chimerism, however, this isn't super helpful for blood samples]]

plot heterozygosity vs primer probe reads

```{r heterozygosity vs primer probe reads}
ggplot()+geom_point(data=master_sampleSummary_10x,aes(x=Primer.Probe.Reads,y=Heterozygosity))+
  labs(title="Heterozygosity vs Total Reads per Sample", x="Primer Probe Reads", y="Heterozygosity")+
  theme_bw()+theme(plot.title=element_text(hjust=0.5),plot.subtitle=element_text(hjust=0.5))+
  geom_hline(yintercept=0.3, lty="dashed")
```

### 30x summaries

load sample summaries from AmpliconReadCounter (not filtered for coverage)

```{r}
# Full set
fullSet_GTscore_individualSummary <- read.delim("./05_tamRun5/03_run5GTscore/fullSet_GTscore_individualSummary.txt", header = TRUE, stringsAsFactors = FALSE) %>%
  mutate(Sample = gsub("-", "\\.", Sample))

# LWED
LWED_GTscore_individualSummary <-  read.delim("./05_tamRun5/03_run5GTscore/LWED_GTscore_individualSummary.txt", header = TRUE, stringsAsFactors = FALSE) %>%
  mutate(Sample = gsub("-", "\\.", Sample))

# SIMP
SIMP_GTscore_individualSummary <- read.delim("./05_tamRun5/03_run5GTscore/SIMP_GTscore_individualSummary.txt", header = TRUE, stringsAsFactors = FALSE) %>%
  mutate(Sample = gsub("-", "\\.", Sample))  
```

#### Summarize single SNP summaries for samples

The summarizeSamples function changes the "-" to a "." -- be sure to change this back.

```{r}
fullSet_singleSNP_sampleSummary_30x <- summarizeSamples(fullSet_polyGenResults_singleSNP_30x, fullSet_singleSNP_alleleReads_30x)

write.csv(fullSet_singleSNP_sampleSummary_30x, "./05_tamRun5/03_run5GTscore/fullSet_singleSNP_sampleSummary_30x.csv", row.names = F)

LWED_singleSNP_sampleSummary_30x <- summarizeSamples(LWED_polyGenResults_singleSNP_30x, LWED_singleSNP_alleleReads_30x)

SIMP_singleSNP_sampleSummary_30x <- summarizeSamples(SIMP_polyGenResults_singleSNP_30x, SIMP_singleSNP_alleleReads_30x)
```

#### Combine AmpliconReadCounter individual summary with GTscore sample summary data

```{r}
# make new copies & add column to note which loci-set the values are based on
LWED_gtIndivSummary_30x <- LWED_GTscore_individualSummary %>%
  mutate(lociSet = "LWED") %>%
  merge(., LWED_singleSNP_sampleSummary_30x, by.x = "Sample", by.y = "sample")

SIMP_gtIndivSummary_30x <- SIMP_GTscore_individualSummary %>%
  mutate(lociSet = "SIMP") %>%
  merge(., SIMP_singleSNP_sampleSummary_30x, by.x = "Sample", by.y = "sample")

# Combine LWED & SIMP to form one, adjust sample name
lwedSIMP_gtIndivSummary_30x <- rbind(LWED_gtIndivSummary_30x, SIMP_gtIndivSummary_30x) %>%
  mutate(
    sampleFile = gsub("\\.", "-", Sample),
    sampleFile = paste0(sampleFile, ".fastq")
         )
```

#### Create master sample summary file & add metadata

```{r}
# Merge metadata & fullSet sample summaries
master_sampleSummary_30x <- merge(lwedSIMP_gtIndivSummary_30x, md_tamRun5, by = "sampleFile")

# Export
write.csv(master_sampleSummary_30x, "./05_tamRun5/03_run5GTscore/summaryFiles/master_sampleSummary_30x.csv", row.names = F)
```

#### Plots

plot histogram of genotype rate

```{r histogram of genotype rate, warning=FALSE}
ggplot()+geom_histogram(data=master_sampleSummary_30x,aes(x=GenotypeRate),binwidth=0.01)+xlim(-0.01,1.01)+
  labs(title="Sample Genotype Rate (30x)", x="Genotype Rate", y="Count")+
  theme_bw()+theme(plot.title=element_text(hjust=0.5),plot.subtitle=element_text(hjust=0.5))
```

plot histogram of heterozygosity

```{r histogram of heterozygosity, warning=FALSE}
ggplot()+geom_histogram(data=master_sampleSummary_30x,aes(x=Heterozygosity),binwidth=0.03)+xlim(-0.01,1.01)+
  labs(title="Sample Heterozygosity", x="Heterozygosity", y="Count")+
  theme_bw()+theme(plot.title=element_text(hjust=0.5),plot.subtitle=element_text(hjust=0.5))
```

plot genotype rate vs primer probe reads

```{r genotype rate vs primer probe reads}
#dashed line added at 90% genotype rate, this is not a strict threshold, just a goal to aim for
ggplot(data=master_sampleSummary_30x,aes(x=Primer.Probe.Reads,y=GenotypeRate,color=sampleType))+
  geom_point(stat = "identity")+
  geom_smooth(method = "loess")+
  labs(title="Genotype Rate vs Total Reads per Sample", x="Primer Probe Reads", y="Genotype Rate")+
  theme_bw()+theme(plot.title=element_text(hjust=0.5),plot.subtitle=element_text(hjust=0.5))+
  geom_hline(yintercept=0.5,lty="dashed")
```

Samples with unusually high heterozygosity may be contaminated or have elevated ploidy. [[Given chimerism, however, this isn't super helpful for blood samples]]

plot heterozygosity vs primer probe reads

```{r heterozygosity vs primer probe reads}
ggplot()+geom_point(data=master_sampleSummary_30x,aes(x=Primer.Probe.Reads,y=Heterozygosity))+
  labs(title="Heterozygosity vs Total Reads per Sample", x="Primer Probe Reads", y="Heterozygosity")+
  theme_bw()+theme(plot.title=element_text(hjust=0.5),plot.subtitle=element_text(hjust=0.5))+
  geom_hline(yintercept=0.3, lty="dashed")
```

# 8 Quality control

## 8.1 Identify duplicate samples

The IDduplicateSamples function does all pairwise comparisons of samples and outputs two metrics:

-   proportionCommon - The proportion of loci that had genotypes for both samples in a pair
-   proportionMatch - The proportion of loci that have identical genotypes in the sample pair

```{r identify duplicate samples,eval=FALSE}
#convert missing genotypes "0" to NA
polyGenResults_singleSNP_NA<-fullSet_polyGenResults_singleSNP
polyGenResults_singleSNP_NA[polyGenResults_singleSNP_NA=="0"]<-NA

#compare all samples, for each comparison get proportion of loci that have genotypes
#in both samples (proportionCommon) and proportion of shared loci that have identical
#genotypes (proportionMatch).
#Samples with a high proportionCommon and high proportionMatch are likely duplicates
polyGenResults_dupTest<-IDduplicateSamples(polyGenResults_singleSNP_NA)

write.table(polyGenResults_dupTest,"./05_tamRun5/03_run5GTscore/run5_polyGenResults_dupTest.txt",quote=FALSE,sep="\t",row.names=FALSE)
```

```{r duplicate sample results}
polyGenResults_dupTest<-read.delim("./05_tamRun5/03_run5GTscore/run5_polyGenResults_dupTest.txt",stringsAsFactors=FALSE)
head(polyGenResults_dupTest)
```

Plot results of IDduplicateSamples

McKinney says that he generally sets thresholds of proportionMatch \> 0.8 and proportionCommon \> 0.75 to identify duplicate sample pairs, but the appropriate thresholds will depend on the marker set being used and the relatedness among samples. In this plot, he notes that it is possible that individuals at the lower range of identical genotypes (\~0.85) are relatives rather than duplicated samples.

```{r plot IDduplicateSamples results}
#potential duplicates are proportionMatch>0.8 and proportionCommon>0.75
#feel free to adjust these thresholds as needed
matchThresh=0.8
commonThresh=0.75 #switched to 0, looks like some loci were successfully genotyped for one sample and some for the other

#plot results
ggplot()+geom_point(data=polyGenResults_dupTest,aes(x=proportionCommon,y=proportionMatch))+
  geom_segment(aes(x=commonThresh,xend=1,y=matchThresh,yend=matchThresh),lty="dashed")+
  geom_segment(aes(x=commonThresh,xend=commonThresh,y=matchThresh,yend=1),lty="dashed")
```

Identify duplicate sample pairs with thresholds set previously

```{r}
#filter to potentially duplicated samples using thresholds above
polyGenResults_dupTest %>% filter(proportionMatch>=matchThresh,proportionCommon>=commonThresh)
```

## 8.2 Identify contaminated samples

Contaminated samples can be identified through elevated heterozygosity. NOTE- this isn't going to be helpful for us given the blood chimerism.

```{r identify contaminated samples (heterozygosity)}
#plot heterozygosity vs genotype rate per sample
#samples with unusually high heterozygosity relative to others are candidates for contamination
#the sample above the dashed line in this example is likely contaminated
ggplot()+geom_point(data=master_sampleSummary,aes(x=GenotypeRate,y=Heterozygosity))+
  labs(title="Heterozygosity vs Genotype Rate per Sample", x="Genotype Rate", y="Heterozygosity")+
  theme_bw()+theme(plot.title=element_text(hjust=0.5),plot.subtitle=element_text(hjust=0.5))+
  geom_hline(yintercept=0.30, lty="dashed")
```

Set heterozygosity treshold to identify likely contaminated sample

```{r}
#identify likely contaminated sample
contaminatedSample<-master_sampleSummary %>% filter(Heterozygosity>0.30) %>% pull(Sample)
contaminatedSample
```

Contaminated samples can also be identified by an elevated contamination score. In this example, samples with a contamination score \> 0.3 are known to be contaminated.

```{r }
#plot histogram of contamination score
ggplot()+geom_histogram(data=master_sampleSummary,aes(x=conScore),binwidth=0.02)+geom_vline(xintercept=0.3,lty="dashed")

```

Set contamination score threshold to identify contaminated samples

```{r}
#identify likely contaminated samples
contaminatedSamples2<-master_sampleSummary %>% filter(conScore>=0.3) %>% pull(Sample)
contaminatedSamples2
```

Plot genotype scatterplots for each sample

```{r}
#Scatter Plots can show evidence of contamination or elevated ploidy
plotGenotypes_sample(fullSet_singleSNP_locusTable, fullSet_singleSNP_alleleReads, fullSet_polyGenResults_singleSNP, type='scatter', savePlot="Y", saveDir="./scatterPlots_sample")
#look at plots for samples, particularly for putatively contaminated samples identified by high heterozygosity
```

## 8.3 Identify poor quality loci to remove

Flag SNPs with \< 50% genotype rate as candidates for removal. Will need to do this with LWED and SIMP samples separately, then recombine.

```{r}
poorQualitySNPs <- ls_ss %>% filter(GenotypeRate<0.5) %>% mutate(Locus=as.character(Locus)) %>% pull(Locus)

View(poorQualitySNPs)

pq_lwed <- data.frame(name=poorQualitySNPs_LWED) %>%
  mutate(species="LWED")
pq_simp<- data.frame(name=poorQualitySNPs_SIMP) %>%
  mutate(species="SIMP")

pq_both <- rbind(pq_lwed, pq_simp)
View(pq_both)

pq_both2 <- pq_both %>%
  group_by(name) %>%
  mutate(count=n()) %>%
  distinct(name, .keep_all = T) %>%
  within(species[count == '2'] <- "both")
View(pq_both2)

# Export file
write.csv(pq_both2, "./summaryFiles/poorQualitySNPs_fullSet.csv", row.names = F)
```

## 8.4 Plots

Ratio & Scatter Plots can show evidence of systemic issues in read counting (off-target reads, biases) or elevated ploidy

Plot genotype ratio & scatterplots for each locus

```{r}
#Allele Ratio Plots  
plotGenotypes(singleSNP_locusTable, singleSNP_alleleReads, polyGenResults_singleSNP, type='ratio', savePlot="Y", saveDir="ratioPlots_loci")
#Scatter Plots 
plotGenotypes(singleSNP_locusTable, singleSNP_alleleReads, polyGenResults_singleSNP, type='scatter', savePlot="Y", saveDir="scatterPlots_loci")
```

Plot histogram of contamination score for loci. An arbitrary score of 0.3 was set to identify problematic loci, but a lower or higher threshold may be more suitable. Visual examination of loci with elevated scores is recommended. In cases where there are very few heterozygous individuals, a high contamination score may be not truly reflect locus performance..

```{r, warning=FALSE}
#loci with high contamination scores should have scatterplots examined to ensure genotypes look accurate
fullSet_singleSNP_summary %>% ggplot(data=.) + geom_histogram(aes(x=conScore),binwidth=0.01)+geom_vline(xintercept=0.3)
```

Plot contamination score vs minor allele frequency. Loci with both a high contamination score and high minor allele frequency are likely to be problematic.

```{r, warning=FALSE}
fullSet_singleSNP_summary %>% ggplot(data=.) + geom_point(aes(x=conScore,y=minAF))
```

Get list of loci with high contamination score

```{r}
singleSNP_summary %>% filter(conScore>0.3)
```

Look at scatterplots for loci with high contamination score. See orignal GTscore README.Rmd for examples of a paralog, locus w/likely off-target sequence, locus w/high contamination score but may be fine.

# 9 Export genotype files

When exporting files, you can either export the full results and remove any poor quality samples or loci afterwards, or you can use whitelists or blacklists to filter during export. Both the exportGenepop and exportRubias functions have the following options: sampleWhitelist, locusWhitelist, sampleBlacklist, locusBlacklist. The whitelists are a list of samples or loci to retain, the blacklists are lists of samples or loci to discard.

## 9.1 Genepop format



export genepop format - poor quality loci removed

```{r}
#example with poor quality loci and contaminated sample removed
exportGenepop(polyGenResults_singleSNP,singleSNP_locusTable,
              locusBlacklist=poorQualitySNPs,
              filename="polyGenResults_singleSNP_genepop_filtered.txt")
```

## 9.2 Rubias format

export Rubias format with poor quality loci and contaminated sample removed

```{r}
exportRubias(polyGenResults_singleSNP,singleSNP_locusTable,sampleMetaData,
             locusBlacklist=poorQualitySNPs,
             filename="polyGenResults_singleSNP_rubias_filtered.txt")
exportRubias(polyGenResults_singleSNP_LWED,singleSNP_locusTable_LWED,sampleMetaData,
             locusBlacklist=poorQualitySNPs_LWED,
             filename="polyGenResults_singleSNP_rubias_filtered_LWED.txt")
exportRubias(polyGenResults_singleSNP_SIMP,singleSNP_locusTable_SIMP,sampleMetaData,
             locusBlacklist=poorQualitySNPs_SIMP,
             filename="polyGenResults_singleSNP_rubias_filtered_SIMP.txt")
```

# X Locus diagnostics

Diagnostic functions are provided to help in identifying patterns of sequence variation for loci.  This can assist in identifying SNPs that were not accounted for in the initial probe design and in adjusting the in-silico probes to exclude off-target sequence from genotyping.  

There are two approaches used by GTscore: First, plotting the number of nucleotide mismatches by position in the sequence data relative to the reference sequence for each locus, and second, performing sequence alignments using the program MSA (Bodenhofer et al. 2015).  Plotting the mismatches by position is fast and easy to interpret except in the case of indels.  The MSA alignments take longer to complete but facilitate the visualization of indels.

The first step for both methods is to get the sequences that aligned to each locus, this can be done based on primer alignments or primer-probe alignments using the script matchReads.pl.

### Sequence alignment to each locus (Perl)

Input flags for this script are:
* p    a tab delimited file containing primer/probe information for each locus
* --files  a text file containing a list of .fastq sequence files to count reads from
* --matchType the match type chosen for retaining sequences for each locus. Options are primer to retain sequences that match the primer for a locus or primerProbe to retain sequences that match both the primer and probe.  primerProbe is the default if no option is specified.

Optional flags:
* --prefix  optional prefix for output file names

You can run matchReads.pl from the command line using the following example:
perl matchReads.pl -p primerProbeFile.txt --files sampleFiles.txt --matchType primerProbe [--prefix]

Alternatively, you can also run it through R:

```{r match reads, eval=FALSE}
#primer match
system2("perl",
        args="./GTscore_sourceScripts/matchReads_modified.pl -p ./05_tamRun5/03_run5GTscore/primerProbeFileV3_fullSet.txt --files ./05_tamRun5/03_run5GTscore/sampleFiles_fullSet.txt --inDir ./05_tamRun5/02_run5Interleaved/ --outDir ./05_tamRun5/03_run5GTscore/ --matchType primer --prefix fullSet")

#system2("perl",
#        args="matchReads.pl -p primerProbeFile.txt --files sampleFiles_LWED.txt --matchType primer --prefix LWED_")

#system2("perl",
#        args="matchReads.pl -p primerProbeFile.txt --files sampleFiles_SIMP.txt --matchType primer --prefix SIMP_")

#primer AND probe match
system2("perl",
        args="./GTscore_sourceScripts/matchReads_modified.pl -p ./05_tamRun5/03_run5GTscore/primerProbeFileV3_fullSet.txt --files ./05_tamRun5/03_run5GTscore/sampleFiles_fullSet.txt --inDir ./05_tamRun5/02_run5Interleaved/ --outDir ./05_tamRun5/03_run5GTscore/ --matchType primerProbe --prefix fullSet")


#system2("perl",
#        args="matchReads.pl -p primerProbeFile.txt --files sampleFiles_LWED.txt --matchType primerProbe --prefix LWED_")

#system2("perl",
#        args="matchReads.pl -p primerProbeFile.txt --files sampleFiles_SIMP.txt --matchType primerProbe --prefix SIMP_")
```

The default output file names are matchReads_primerAligned.txt or matchReads_primerProbeAligned.txt depending on which match type was chosen.

### Create reference amplicon sequences file

To conduct locus diagnostics, we need both a text file and fasta file of reference amplicon sequences. The sequences are already in Sam's pos_amplicon.csv file, we just need to get it in the right format:

Text file first (made v1 in 00_tamRun1_optimization; now updating to reflect updated locus shortnames):
```{r create reference sequences file}
refSeqs_v2 <- read.delim("./00_tamRun1_optimization/ampliconRefSeqs.txt", header = TRUE, stringsAsFactors = FALSE) %>%
  mutate(
    Locus = gsub("\\..*", "", Locus),
    Locus = case_when(
      Locus == "SEXID_218" ~ "SEXID_SIMP_218",
      Locus == "SEXID_203" ~ "SEXID_SIMP_203",
      Locus == "SEXID_198" ~ "SEXID_SIMP_198",
      Locus == "SEXID_197" ~ "SEXID_SIMP_197",
      Locus == "SEXID_208" ~ "SEXID_LWED_208",
      Locus == "SEXID_195" ~ "SEXID_LWED_195",
      .default = Locus
    )
  ) %>%
  filter(Locus %in% pp_fullSet$Locus)


# Export
write.table(refSeqs_v2, "./05_tamRun5/03_run5GTscore/ampliconRefSeqs_v2.txt", sep= "\t", row.names = F, quote = F)
```

Then fasta file:
```{r create amplicon fasta file}
library(phylotools)

amplicon_txt <- read.table("./05_tamRun5/03_run5GTscore/ampliconRefSeqs_v2.txt", header = T) %>%
  dplyr::rename(c(seq.name=Locus, seq.text=refSeq))
View(amplicon_txt)

dat2fasta(amplicon_txt, outfile = "./05_tamRun5/03_run5GTscore/ampliconRefSeqs_v2.fasta")
```

### MSA alignment (GTscore.R)

MSA alignment of the matched reads can be done directly on the output from matchReads.pl using the alignMatchedSeqs command in GTscore. Probe sequence is incorporated into the alignment to visualize how probes may be affecting read counting, and reference sequence can optionally be included to facilitate comparison of the observed sequence with the target locus. The probe sequence is obtained from the primer-probe file originally used by AmpliconReadCounter.pl; the reference sequence file has a two column format where the first column is the locus name and the second column is the reference sequence for the locus amplicon. MSA will take a very long time to run if too many sequences are included for alignment so a minimum read threshold (minReads) is included as an option in the alignment command (suggest begin with minReads=20). In addition, a maximum of 100 unique sequences (ranked by number of reads) will be aligned.  The maximum number of reads allowed for alignment can be changed using the maxAlignedSeqs option. Be sure the required perl package has been installed (Excel-Writer-XLSX).

**NOTE** to run correctly, need to use the updated GTscore_test2.R script PLUS need to only include seqs for which refSeqs are available

**START HERE**
-   need to do this again but species-specific
-   then check for additional variants w/in amplicons & add to pipeline

```{r MSA alignment}
#load reference sequences in table format
refSeqs <- read.delim("./05_tamRun5/03_run5GTscore/ampliconRefSeqs_v2.txt", header = TRUE, stringsAsFactors = FALSE)
head(refSeqs)

#load primer probe file
primerProbes <- read.delim("./05_tamRun5/03_run5GTscore/primerProbeFileV3_fullSet.txt", header = TRUE, stringsAsFactors = FALSE) %>%
  # filter to only those w/refSeq
  filter(Locus %in% refSeqs$Locus)
head(primerProbes)

#primer aligned reads
primerMatchedReads <- read.delim("./05_tamRun5/03_run5GTscore/fullSet_matchedReads_primerAligned.txt", header = TRUE, stringsAsFactors = FALSE) %>%
  # filter to only those w/refSeq
  filter(Locus %in% refSeqs$Locus)
head(primerMatchedReads)

source("./GTscore_sourceScripts/GTscore_test2.R") # IT WORKS

alignMatchedSeqs(referenceSeqs = refSeqs,
                 primerProbes = primerProbes,
                 matchedReads = primerMatchedReads,
                 minReads = 20,
                 maxAlignedSeqs = 100,
                 type = "primer",
                 saveDir = "./05_tamRun5/03_run5GTscore/MSA_primerMatched")

#primer probe aligned reads
primerProbeMatchedReads <- read.delim("./05_tamRun5/03_run5GTscore/fullSet_matchedReads_primerProbeAligned.txt", header = TRUE, stringsAsFactors = FALSE)

alignMatchedSeqs(referenceSeqs = refSeqs,
                 primerProbes = primerProbes,
                 matchedReads = primerProbeMatchedReads,
                 minReads = 20,
                 maxAlignedSeqs = 100,
                 type = "primerProbe",
                 saveDir = "./05_tamRun5/03_run5GTscore/MSA_primerProbeMatched")
```

Excel files are output for each locus within the directory specified by savDir. The first row is a consensus sequence based on the sequence alignments. This is followed by the probes, reference sequence if included, and then the sequence alignments.

### Plotting mismatches by position (Perl and GTscore.R)

#### Calculate sequence mismatches by position (Perl)

Plotting the number of mismatches by position requires first comparing each matched sequence against the reference sequence and summing the number of mismatches for each nucleotide position.  This is done with the perl script seqMismatchPositions.pl.  

Input flags for this script are 
* --amplicon    a fasta file containing the reference amplicon sequence for each locus
* --matchedSeqs  a tab delimited file containing sequences that matched each locus.  This file is generated by matchReads.pl
* --matchType the match type chosen for retaining sequences for each locus.  Options are primer to retain sequences that match the primer for a locus or primerProbe to retain sequences that match both the primer and probe.  primerProbe is the default if no option is specified.

Optional flags
* --prefix  optional prefix for output file names

You can run matchReads.pl from the command line using the following example:
perl seqMismatchPositions.pl --amplicon ampliconRefSeqs.fasta --matchedSeqs matchedReads_primerAligned.txt  --matchType primer [--prefix]

Alternatively, you can also run it through R:

```{r sequence mismatches, eval=FALSE}
# primer only
system2("perl",
        args="./GTscore_sourceScripts/seqMismatchPositions_modified.pl --amplicon ./05_tamRun5/03_run5GTscore/ampliconRefSeqs_v2.fasta --matchedSeqs ./05_tamRun5/03_run5GTscore/fullSet_matchedReads_primerAligned.txt --outDir ./05_tamRun5/03_run5GTscore/ --matchType primer --prefix fullSet_")

# primerProbe
system2("perl",
        args="./GTscore_sourceScripts/seqMismatchPositions_modified.pl --amplicon ./05_tamRun5/03_run5GTscore/ampliconRefSeqs_v2.fasta --matchedSeqs ./05_tamRun5/03_run5GTscore/fullSet_matchedReads_primerProbeAligned.txt --outDir ./05_tamRun5/03_run5GTscore/ --matchType primerProbe --prefix fullSet_")
```

The default output file names are mismatchPositions_primer.txt or mismatchPositions_primerProbe.txt depending on which match type was chosen.  The output files will be used by the summarizeMismatches function in GTscore to generate plots of mismatches by position for each locus.

#### Plot of sequence mismatches by position (GTscore.R)

If throws an error, use .rs.restartR()

```{r}
#load results from seqMismatchPositions.pl
#primer matched sequences
mismatchPositionData_primer<-read.delim("./05_tamRun5/03_run5GTscore/fullSet_mismatchPositions_primer.txt", header=TRUE, stringsAsFactors=FALSE)

#primer probe matched sequences
mismatchPositionData_primerProbe<-read.delim("./05_tamRun5/03_run5GTscore/fullSet_mismatchPositions_primerProbe.txt", header=TRUE, stringsAsFactors=FALSE)


#generate plots of mismatches by position
#primer matched sequences
summarizeMismatches(mismatchPositionData_primer, saveDir = "./05_tamRun5/03_run5GTscore/mismatchPositionPlots_primer")

#primer probe matched sequences
summarizeMismatches(mismatchPositionData_primerProbe,saveDir="./05_tamRun5/03_run5GTscore/mismatchPositionPlots_primerProbe")
```
