---
title: "tamGenetics_paper3_relatedness"
author: "Rachel Voyt"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1 Overview

This document contains the relatedness analyses for the tamarin
demographics paper (paper 3) using genotypes generated from tamRun5
MiSeq data.

# 2 Potential packages

## 2.1 All potential

There are a ton of different options for estimating relatedness,
parentage, and sibships, many of which I've outlined below. After going
through all of these, it seems like FRANz would be the best option - it
seems to be fairly commonly used, and importantly, it can do both
parentage and sibship (unlike CERVUS, which just does parentage) and can
account for multiple generations (unlike COLONY), missing data and
genotyping error (unlike Sequoia), and can incorporate prior knowledge
if available (e.g., age, sex, etc.).

**Packages:**

-   [apparent](https://bmcbioinformatics.biomedcentral.com/articles/10.1186/s12859-019-2662-3)
    -   pros:
        -   infers both parentage and sibships
        -   requires no prior info about family structure
    -   cons:
        -   exclusion-based method; uses homozygous genos only
-   [CERVUS](http://www.fieldgenetics.com/pages/aboutCervus_Overview.jsp)
    (R version: [CeRvus](https://github.com/irmoodie/ceRvus))
    -   most popular parentage software
    -   parentage estimates only (no sibships)
    -   does not support multi-generational pedigrees
-   [CLAPPER](https://journals.plos.org/plosgenetics/article?id=10.1371/journal.pgen.1006963)
    -   "composite likelihood approach to pedigree reconstruction"
    -   focuses on the pedigree, which is not equivalent to set of all
        pairwise relationships
    -   supports multi-generational pedigrees (up to 5 generations),
        polygynous reproduction, allows many missing indivs in the
        sample
    -   assumes that all indiv are outbred & pedigrees don't create
        cycles, except in the case of full-sibs
-   [COLONY](https://www.zsl.org/about-zsl/resources/software/colony)
    -   pros:
        -   infers parentage and sibships
        -   infers clones or duplicated individuals
        -   accounts for typing errors
    -   cons:
        -   only allows up to two-generation sample
    -   NOTE - Flanagan & Jones 2018 supp material notes that "COLONY2.0
        is more computationally efficient than the previous
        full-likelihood version of COLONY but suffers a modest cost in
        accuracy."
-   [FRANz](https://github.com/lima1/franzpedigree) ([Riester et al.
    2009](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC2722992/))
    -   general:
        -   MCMC-based
    -   pros:
        -   infers both parentage and sibships
        -   supports multi-generational pedigrees
        -   can incorporate prior knowledge if available (software will
            automatically use this to create lists of candidate parents
            internally)
        -   accounts for missing data, typing errors, estimation of
            number of unsampled candidate parents
    -   examples:
        -   [Ekblom et al.
            2021](https://link.springer.com/article/10.1007/s12686-021-01208-5)
            -- this paper also shows how relationships change
            w/decreasing number of SNPs; includes cases of inbreeding
        -   [Giangregorio et al.
            2023](https://air.uniud.it/bitstream/11390/1259544/1/NC-53-105_article-86739_en_2.pdf) -
            confirmed FRANz parentage assignments w/COLONY (all matched)
-   [popkin](https://cran.r-project.org/web/packages/popkin/vignettes/popkin.html)
    ([paper
    link](https://journals.plos.org/plosgenetics/article?id=10.1371/journal.pgen.1009241))
    -   estimates Fst and kinship for arbitrary population structures
    -   seems like it accounts for overlapping generations?
    -   seems more for population-level relatedness vs. individuals
        though
-   [sequoia](https://cran.r-project.org/web/packages/sequoia/vignettes/vignette-main.html)
    -   pros:
        -   both parentage and sibship
        -   supports multi-generational pedigrees
        -   likelihood-based method, uses both homozygous & heterozygous
            genos
    -   cons:
        -   doesn't work well if a lot of data is missing
        -   requires 100s of SNPs for accurate sibship estimates
-   [skater](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC8844523/)
    -   for further downstream processing and manipulation of SNP-based
        kinship analysis results
-   [SNPpit](https://github.com/eriqande/snppit)
    -   general:
        -   MCMC simulations
    -   pros:
        -   one of fastest parentage assignment programs
    -   cons:
        -   parentage only
        -   only assigned parent pairs; not appropriate for assignment
            of single parents unless one is already known

**Comparison papers:**

-   parentage analysis review
    -   [Flanagan & Jones
        2018](https://onlinelibrary.wiley.com/doi/abs/10.1111/mec.14988?casa_token=YDbHdVfhO6cAAAAA%3ASCDFf-lvIZt_G0Bk72WKLR9rDx8sL3ZAM2VCbV5RppaMrhk0n_VStIpWn1JZ1JCfPtMWE1dWHTio5g) +
        [supplementary
        material](https://onlinelibrary.wiley.com/action/downloadSupplement?doi=10.1111%2Fmec.14988&file=mec14988-sup-0002-AppendixS2.pdf)
-   CERVUS vs. COLONY
    -   [Karaket & Poompuang
        2012](https://www.sciencedirect.com/science/article/pii/S0044848611008611?casa_token=BigERUkYmeAAAAAA:YWnNgJjilSLG2Fs8Oted1_NRn5xgX-OgspYJQXXchSkzVT4TMBtoyVdCWNdjaS1VvlQh7DOPiQ) -
        used microsats; results suggest COLONY was more effective

**Other helpful papers:**

-   [Raval
    2021](https://eprints.hud.ac.uk/id/eprint/35631/1/RAVAL%20-%20THESIS.pdf) -
    dissertation; has helpful info on pedigree reconstruction methods &
    note on rep success as a measure of fitness

something to keep in mind:

"Noninvasive and minimally invasive samples such as hair samples are
prone to genotyping errors due to low DNA quality and quantity (Carroll
et al., 2018; Pompanon, Bonin, Bellemain, & Taberlet, 2005)" from
[Escoda et al.
2018](https://onlinelibrary.wiley.com/doi/10.1111/1755-0998.12967)
(Quantitative analysis of connectivity in populations of a semi‐aquatic
mammal using kinship categories and network assortativity) \< seems like
a helpful paper ((they used PRIMUS and VCF2LR softwares for kinship
estimates))

## 2.2 FRANz info

(Huisman 2017) - for sequoia, but compares w/FRANz &
opposite-homozygosity-exclusion methods ((AR = assignment rates; ER =
error rate)) - maximizes total likelihood by calculating for each set of
candidate relatives the likelihoods under many possible alternative
relationships; implicit to KINSHIP (Goodnight & Queller 1999) & for
parentage assignment in FRANz - finds that relatedness estimates for
sequoia-reconstructed pedigree was more strongly related to true
relatedness than for franz-reconstructed pedigree - with this, sequoia
error rates are also a lot lower than franz for assignments - when
interest is solely in parentage assignment, sequoia performs
intermediately b/t OH-methods & franz - notes that franz explicitly
deals w/genotyping errors & makes use of birth year, death year, & sex,
but is less conservative than sequoia when this life-history info is
lacking for some indivs - says that FRANz performs clustering of
full-sibs only to support parentage assignment & in a way that's less
integrated than SEQUOIA - SEQUOIA can't use candidate parent list, but
these are explicitly used by FRANz - uncertainty around birth year
estimates not accounted for by SEQUOIA

Jones & Manseau (2022) - "Programs COLONY (Jones and Wang, 2010) and
FRANz (Riester et al., 2009) implement Bayesian maximum likelihood
methods to infer both parentage and siblingship simultaneously
(Almudevar, 2007; Cowell, 2009; Almudevar and LaCombe, 2012), and these
parent-offspring relationships can then be used to construct pedigree
networks (as in Salvi et al., 2014; McFarlane et al., 2018)." (p. 4)

Shedd et al () - "We used FRANz because likelihood- and Bayesian- based
parentage analyses have been shown to perform better than
exclusion-based techniques (Anderson & Ng, 2014; Harrison et al., 2013;
Hauser et al., 2011; Jones et al., 2010; Steele et al., 2013).
Additionally, a full- probability Bayesian model for pedigree
reconstruction is better suited for studies that are not able to sample
all potential parents and offspring because the model accounts for
unsampled parents and can use sibships among sampled individuals to
infer parental genotypes from offspring and fill out sparse pedigrees
(Jones et al., 2010; Riester et al., 2009)" - "we followed code from
Baetscher et al. (2018) to use the CKMRsim R package
(<https://github.com/eriqande/CKMRsim>) to evaluate the power of our SNP
panel to accurately make parent-- offspring and full- and half-sibling
assignments"

# 3 Packages

## 3.1 R packages

```{r}
library(lubridate)
library(tidyverse)
```

## 3.2 FRANz

### SourceForge (latest stable source code)

The SourceForge script set works, but was last updated in 2013 -- there
are some updates on the Github page that occurred as late as 2020. Given
that I wasn't able to get the Github version to compile, however, this
seems like the best bet.

One error that I caught: - csv.pl has a typo in line 164; the original
has "death_col =", but should be "death ="; the github version looks
like it was corrected (switched it out for this version)

Step 1: Download the latest stable source code from SourceForge.net
([link](http://sourceforge.net/projects/franzpedigree/files/))

Step 2: Unzip

```{bash}
tar -xvfz FRANz-2.0.0.tar.gz
cd FRANz-2.0.0
```

Step 3: Compile

```{bash}
./configure 
make check  (optional, may take a while)
sudo make install # need sudo or won't install correctly
```

Access manual:

```{bash}
# in terminal
man FRANz

# create pdf copy
man -t FRANz | ps2pdf - FRANz_manual.pdf
```

### Github (newest code)

The Github page for FRANz has the newest code, with updates as late as
2020 (vs. 2013 for SourceForge). However, it looks like there are some
backwards compatability issues related to autoconf/autoreconf, where
FRANz uses an older version. Seems like too much of a pain to get it to
work, so going to go with SourceForge version for now.

Step 1: Download the newest code from
[GitHub](https://github.com/lima1/franzpedigree)

Step 2: Unzip

```{bash}
tar xvfz lima1-franzpedigree-28b6d43.tar.gz
cd lima1-franzpedigree-28b6d43
```

Step 3: Compile [[doesn't work; autoreconf error]]

```{bash}
autoreconf (only necessary if you use the latest code from github)
./configure 
make check  (optional, may take a while)
sudo make install # need sudo or won't install correctly
```

# x thinking spot

How to differentiate b/t parent-offspring vs. sibling relationship when
both individuals in a dyad were captured as adults/subadults?

let's also look at geno success across hair samples

```{r}
genoSuccess_byLocus_lwedHair_indidOnly <- genos0x_tamRun5_lwedHair_indidOnly %>%
  t() %>%
  as.data.frame() %>%
  mutate_all(na_if, "0,0")

genoSuccess_byLocus_lwedHair_indidOnly$totalGenos <- rowSums(!is.na(genoSuccess_byLocus_lwedHair_indidOnly))

genoSuccess_byLocus_lwedHair_indidOnly <- genoSuccess_byLocus_lwedHair_indidOnly %>%
  rownames_to_column("locus") %>%
  select(locus, totalGenos) %>%
  mutate(
    propGenos = totalGenos/138
  )
```

# 4 Data

## 4.1 md_run5 & sampleLists

In tamRun5, we have the following for hair samples:

-   LWED
    -   n_total = 137 \>\> 134 minus dups (n_f = 64, n_m = 70)
-   SIMP
    -   n_total = 100 \>\> 99 minus dups (n_f = 47, n_m = 52)

Be sure to use the latest updated metadata file from
metadataReconciliation.Rmd - this md file also includes alternate
animalIDs & associated metadata for all samples with capData metadata
issues.

**Datasets**

```{r}
# full set
md_run5 <- read.csv("./metadataReconciliation/tamRun5_metadata_v5.csv") %>%
  # add unique sampleID w/animalID + sampleType
  mutate(
    sampleID_unique = str_c(animalID, sampleType, sep = "_"),
    sampleID_franz = str_c(animalID, sampleID, sep = "_"),
    # add seqID b/c now have some duplicate animalID/sampleTypes
    sampleID_franz = gsub("tamRun5.", "seq", sampleID_franz)
    ) %>%
  # pad sampleID_franz so all are exactly 10 characters (required by FRANz)
  mutate(
    sampleID_franz = str_pad(sampleID_franz, 10, pad = "_", side = "left")
  ) %>%
  relocate(c(sampleID_unique, sampleID_franz))

# hair
md_run5_hair <- md_run5 %>%
  filter(sampleType == "hair")

md_run5_lwedHair <- md_run5_hair %>%
  filter(species == "LWED")

md_run5_simpHair <- md_run5_hair %>%
  filter(species == "SIMP")

# simplified sample lists
sampleList_lwedHair <- md_run5_lwedHair %>%
  select(c("sampleID_unique", "sampleID_franz", "sampleID"))

sampleList_simpHair <- md_run5_simpHair %>%
  select(c("sampleID_unique", "sampleID_franz", "sampleID"))
```

**Metrics**

```{r}
## n = 4 dups (animalID 191_simp, 200_lwed, 26_lwed, 27_lwed)
md_run5_hair %>%
  get_dupes(animalID)

# minus dups, have lwedF = 64, lwedM = 70, simpF = 47, simpM = 52
md_run5_hair %>%
  select(animalID, species, sex) %>%
  distinct() %>%
  group_by(species, sex) %>%
  summarise(
    count = n()
  )

## n = 46 previously extracted (lwed = 36, simp = 10)
temp <- md_run5_hair %>%
  filter(str_detect(xtnLoc, "tx"))

table(temp$species)
```

## 4.2 capData

Import latest cleaned version of capData_byIndiv (v5)

From 2009 to 2019, we have the following:

-   LWED
    -   n = 145 total captured (n_f = 68, n_m = 77)
-   SIMP
    -   n = 102 total captured (n_f = 48, n_m = 54)

**Datasets**

```{r}
# full set
capData_byIndiv <- read.csv("./paper3_demographics/01_dataOrganization/02_dataCleaning/captureData_byIndividual_v5.csv") %>%
  mutate(
    captureYear = str_sub(captureDate, 1, 4),
    captureYear = as.numeric(captureYear)
  ) %>%
  # adjust rowID 1 ageClass to "UNK"
  mutate(
    ageClass = case_when(
      rowID == 1 ~ "UNK",
      .default = ageClass
    )
  )

# up to 2019 captures only
## fullSet
capData_2009to2019_byIndiv <- capData_byIndiv %>%
  filter(as.numeric(rowID) <= 613) %>%
  # ditch indiv w/UNK animalID (should be "2", but pending response from Gideon for what to do w/this record) %>%
  filter(rowID != 611)

## species subsets
capData_2009to2019_lwed <- capData_2009to2019_byIndiv %>%
  filter(species == "LWED")

capData_2009to2019_simp <- capData_2009to2019_byIndiv %>%
  filter(species == "SIMP")

# filter to first capture for each animalID
capData_2009to2019_firstEntry <- capData_2009to2019_byIndiv[match(unique(capData_2009to2019_byIndiv$animalID), capData_2009to2019_byIndiv$animalID),] %>%
  select(rowID, captureDate, animalID, ageClass, animalName1, animalName2, groupName, species, sex, notes_MD, notes_RV)
```

**Metrics**

species sex count 1 LWED F 68 2 LWED M 77 3 SIMP F 48 4 SIMP M 54

```{r}
# lwed 145; simp 102
table(capData_2009to2019_firstEntry$species)

# lwedF = 68, lwedM = 77, simpF = 48, simpM = 54
capData_2009to2019_firstEntry %>%
  group_by(species, sex) %>%
  summarise(
    count = n()
  ) %>%
  as.data.frame()
```

## 4.3 Birth/death data

Load in latest birthDeath file from
/paper3_demographics/tamGenetics_paper3_dataOrganization

```{r}
birthYear_capData_run5 <- read.csv("./paper3_demographics/01_dataOrganization/02_dataCleaning/birthAssignments_capData_tamRun5_v3.csv")

birthYear_run5_hair <- birthYear_capData_run5 %>%
  select(animalID, birthYear_est) %>%
  merge(., md_run5_hair[, c("sampleID_franz", "animalID")], by = "animalID", all.x = T) %>%
  na.omit() %>%
  distinct()

deathYear_capData_run5 <- read.csv("./paper3_demographics/01_dataOrganization/02_dataCleaning/deathAssignments_capData_tamRun5_v1.csv")
```

## 4.4 Loci lists

Below are the loci sets that I'll be using for kinship analyses. These
include ONLY the INDID loci, including general and species-specific
INDID loci.

Each species has 154 INDID loci, including both general and
species-specific INDID loci.

```{r}
lociList_lwed <- read.table("./05_tamRun5/03_run5GTscore/primerProbeFileV3_LWED.txt", sep = "\t", header = T) %>%
  select(Locus) %>%
  filter(!str_detect(Locus, "SEXID|SPECIESID")) %>%
  pull()

lociList_simp <- read.table("./05_tamRun5/03_run5GTscore/primerProbeFileV3_SIMP.txt", sep = "\t", header = T) %>%
  select(Locus) %>%
  filter(!str_detect(Locus, "SEXID|SPECIESID")) %>%
  pull()

length(lociList_lwed)
length(lociList_simp)
```

## 4.5 Genotypes

### 0x cutoff

**tamRun5 plus tamRun3 hair samples**

```{r}
genos0x_run5.3_hair <- read.table("./sandbox/tamRun5_plus_tamRun3/fullSet_tamRun5_plus_tamRun3_hair_polyGenResults_0x.txt") %>%
  rownames_to_column("locus") %>%
  mutate(locus = sub('[_][^_]+$', '', locus)) %>%
  column_to_rownames("locus") %>%
  t() %>%
  as.data.frame() %>%
  mutate(across(everything(), ~ gsub("0", "0,0", .)))

genos0x_run5.3_lwedHair_indidOnly <- genos0x_run5.3_hair %>%
  filter(rownames(.) %in% sampleList_lwedHair$sampleID) %>%
  select(all_of(lociList_lwed))

genos0x_run5.3_simpHair_indidOnly <- genos0x_run5.3_hair %>%
  filter(rownames(.) %in% sampleList_simpHair$sampleID) %>%
  select(all_of(lociList_simp))
```

**tamRun5**

```{r}
# geno file w/o coverage cutoff
genos0x_run5 <- read.table("./05_tamRun5/03_run5GTscore/fullSet_polyGenResults_singleSNP_0x.txt", header = T) %>%
  rownames_to_column("locus") %>%
  mutate(locus = sub('[_][^_]+$', '', locus)) %>%
  column_to_rownames("locus") %>%
  t() %>%
  as.data.frame() %>%
  mutate(across(everything(), ~ gsub("0", "0,0", .)))

## Species subsets + hair + INDID/LWED-SIMP only
genos0x_run5_lwedHair_indidOnly <- genos0x_run5 %>%
  filter(rownames(.) %in% sampleList_lwedHair$sampleID) %>%
  select(all_of(lociList_lwed))

genos0x_run5_simpHair_indidOnly <- genos0x_run5 %>%
  filter(rownames(.) %in% sampleList_simpHair$sampleID) %>%
  select(all_of(lociList_simp))
```

**tamRun4**

```{r}
genos0x_tamRun4 <- read.table("./04_tamRun4/00_illumina/03_run4GTscore/ill_fullSet_polyGenResults_singleSNP_0x.txt", header = T) %>%
  rownames_to_column("locus") %>%
  mutate(locus = sub('[_][^_]+$', '', locus)) %>%
  column_to_rownames("locus") %>%
  t() %>%
  as.data.frame() %>%
  mutate(across(everything(), ~ gsub("0", "0,0", .))) 
```

### 10x cutoff

```{r}
# original 10x genotype file (reformatted a bit)
genos10x_tamRun5 <- read.table("./05_tamRun5/03_run5GTscore/fullSet_polyGenResults_singleSNP_10x.txt", header = T) %>%
  rownames_to_column("locus") %>%
  mutate(locus = sub('[_][^_]+$', '', locus)) %>%
  column_to_rownames("locus") %>%
  t() %>%
  as.data.frame() %>%
  mutate(across(everything(), ~ gsub("0", "0,0", .)))
```

## 4.6 Expected assignments

**Juveniles**

From the capture data, there were a total of n = 123 juveniles observed
from 2009-2019 and could thus be assigned an estimated year/month of
birth. We have tamRun5 hair samples for n = 116 out of those 123
juveniles, including n = 63 LWED and n = 53 SIMP. Note that tamRun5 does
have blood samples for an additional n = 7 LWED juveniles, but no hair
samples (animalIDs 29, 70, 72, 169, 189, 229, and 233).

**Likely mothers**

We also have notes on likely mother-offspring relationships - 113 =
mother of 1/79 (simp) - Watsa (2013) p. 193: F11/76 gave birth to
infants [7/101] in 2009-Nov; twins were first nursed by mother for 109
days/3.6 months w/last 30 days attempted weaning; from day 109-234 (125
days/4.2 months) F10/117 allonursed infants

#### tamRun5 juv hair samples

```{r}
# LWED
## n = 63 LWED tamRun5 hair samples w/birthYear_est
md_run5_hair %>%
  filter(species == "LWED") %>%
  filter(animalID %in% birthYear_knownOnly$animalID) %>%
  nrow()

## n = 7 LWED tamRun5 blood samples w/birthYear_est & no hair sample
md_run5_blood %>%
  filter(species == "LWED") %>%
  filter(!animalID %in% md_run5_hair$animalID) %>%
  filter(animalID %in% birthYear_knownOnly$animalID) %>%
  nrow()

# SIMP
## n = 53 SIMP tamRun5 hair samples w/birthYear_est
md_run5_hair %>%
  filter(species == "SIMP") %>%
  filter(animalID %in% birthYear_knownOnly$animalID) %>%
  nrow()

## n = 0 SIMP tamRun5 blood samples w/birthYear_est & no hair sample
md_tamRun5_bloodSamples %>%
  filter(species == "SIMP") %>%
  filter(!animalID %in% md_tamRun5_hairSamples$animalID) %>%
  filter(animalID %in% birthYear_knownOnly$animalID) %>%
  nrow()
```

#### tamRun5 hair vs. blood numbers

```{r}
# 57 hair samples w/o corresponding blood sample
md_tamRun5_hair_noBloodMatch <- md_tamRun5_hairSamples %>%
  filter(!animalID %in% md_tamRun5_bloodSamples$animalID) %>%
  arrange(as.numeric(animalID))

# 10 blood samples w/o corresponding hair sample
md_tamRun5_blood_noHairMatch <- md_tamRun5_bloodSamples %>%
  filter(!animalID %in% md_tamRun5_hairSamples$animalID) %>%
  arrange(as.numeric(animalID))
```

## 4.7 Parity

How to figure out if assignments are reasonable? Start with offspring
whose birth years are known + assigned mothers, and look at the
following: 1) Is assigned mother in the group during offspring's
birthYear? What about year before birthYear? 2) Does assigned mother
have nipple measurements that year?

My df of females vs. parity based on nipple length:

***may need to change this; can't really determine parityYear unless
there are records from the previous year saying that indiv was
primiparous***

```{r}
capData_parity <- capData_byIndiv %>%
  filter(sex == "F") %>%
  select(animalID, animalName1, animalName2, species, captureDate, ageClass, nippleL_length, nippleR_length) %>%
  arrange(as.numeric(animalID), captureDate) %>%
  # calculate mean nipple length
  mutate(
    nippleL_length = case_when(
      str_detect(nippleL_length, "<") ~ "0",
      str_detect(nippleR_length, "small") ~ "0",
      .default = nippleL_length
    ),
    nippleR_length = case_when(
      str_detect(nippleR_length, "<") ~ "0",
      str_detect(nippleR_length, "small") ~ "0",
      .default = nippleR_length
    ),
    nippleL_length = as.numeric(nippleL_length),
    nippleR_length = as.numeric(nippleR_length)
  ) %>%
  rowwise() %>%
  mutate(
    nippleMean = mean(c_across(c(nippleL_length, nippleR_length)), na.rm = T),
    nippleMean = round(nippleMean, digits = 2)
  ) %>%
  # assign parity
  mutate(
    parity = case_when(
      species == "LWED" & nippleMean >= 3 ~ "parous",
      species == "SIMP" & nippleMean >= 4 ~ "parous",
      .default = NA
    ),
    parity2 = case_when(
      species == "LWED" & nippleL_length >= 3 ~ "parous",
      species == "LWED" & nippleR_length >= 3 ~ "parous",
      species == "SIMP" & nippleL_length >= 4 ~ "parous",
      species == "SIMP" & nippleR_length >= 4 ~ "parous",
      .default = NA
    )
  ) %>%
  arrange(as.numeric(animalID), captureDate) %>%
  filter(!is.na(parity2))

capData_parity_firstEntry <- capData_parity[match(unique(capData_parity$animalID), capData_parity$animalID),] %>%
  mutate(
    parityYear = str_sub(captureDate, 1, 4)#,
    #parityYear = as.numeric(parityYear) - 1 # since would've conceived year prior
  ) %>%
  merge(., birthYear_run5_knownOnly[, c("animalID", "birthYear_est")], by = "animalID", all.x = T) %>%
  mutate(
    parityAge = as.numeric(parityYear) - as.numeric(birthYear_est)
  ) %>%
  # add sampleID_franz
  merge(., md_run5_hair[, c("animalID", "sampleID_franz")], by = "animalID", all.x = T)
```

-   animalID 41/WPO notes in 2012 capture say nipples indicate has given
    birth before; animalID 136 born in 2014
-   RPS/34 noted as might still be nursing in 2017-06-11

# 5 FRANz

## 5.1 Prep

### femrepro, malerepro

Setting as 2:30 based on lit (see chapter2 for details)

### N & Nmax

#### About

**N (Nf, Nm):** For paternity inference, N is the number of candidate
fathers in the population (default auto) [NMCP01]. This is the sum of
the average number of sampled (n) and unsampled (N-n) breeding males in
the population. If unset, then estimated jointly with the pedigree. For
parentage inference N = Nf = Nm (NOT Nf + Nm), but one can also specify
Nf and Nm instead of N if these numbers differ. If N is not known, use
Nmax instead.

FRANz manual also notes the following-- IMPORTANT: if you have a good
estimate of the number of unsampled candidate parents, use the --N
instead of the --Nmax options. See also the section FRANz RUNS FOREVER.
If both --N or --Nmax are omitted, then a complete sampling is assumed
and the pedigree that maximizes the mendelian segregation probabilities
is returned.

**Nmax (Nfmax, Nmmax):** Maximum number of candidate fathers/mothers in
the population. This is the estimated upper limit of N when N is not
known. FRANz will then incorporate the uncertainity of N in the pedigree
reconstruction. We need this limit to avoid that the Markov Chain
converges to a very high N, which would result in an empty pedigree.
When mothers are unknown and the numbers of males and females differ, we
can again specify Nfmax and Nmmax.

#### JS model

We can estimate *N* for FRANz by estimating abundance via the POPAN
formulation (Schwarz & Arnason 1996) of the Jolly-Seber mark-recapture
model. Note that Jolly-Seber models make the following assumptions (from
Schwarz & Arnason 2021 in Program MARK book):

1)  indivs retain their tags throughout experiment
2)  tags are read properly
3)  sampling is instantaneous
4)  survival probabilities are same for all indivs (marked and unmarked)
    b/t each pair of sampling occasions (homogeneous survival)
5)  catchability is same for all indivs (marked and unmarked) at each
    sampling event (homogeneous catchability) ((most crucial assumption
    for JS models))
6)  study area is constant - if study area changes over time, then pop
    size may change w/changing size of study area

The POPAN formulation modifies the parameterization of JS slightly by
postulating 1) the existence of a *super-population* containing all of
the indivs that would ever be born to the population and 2) parameter
b_i representing the probability that an indiv from the hypothetical
super-population would enter the population b/t occasion i and i + 1.

##### Create capture histories

I've already played with this a bit in tamGenetics_paper3_demography for
the full 2009-2023 dataset, but since I only have genetics samples from
2009-2019, I'm going to re-run the models again here for this subset of
the data.

```{r}
capHist_2009to2019 <- capData_byIndiv_v5 %>%
  # ensure order is correct
  arrange(rowID) %>%
  
  # filter to 2009-2019 only
  filter(as.numeric(rowID) < 614) %>%
  
  select(animalID, captureDate) %>%
  filter(animalID != "UNK") %>%
  mutate(
    captureYear = as.numeric(str_sub(captureDate, 1, 4))
  ) %>%
  select(-captureDate) %>%
  mutate(
    detect = 1
  ) %>%
  
  # format as capture history (notes below are from James Paterson blog)
  
  # remove duplicates, which may occur when individuals are caught multiple times in an event
  # For example, your event may be a year and an individual may be caught multiple times in a year.
  distinct() %>%
  
  # spread out data. The fill = 0 adds rows for combinations of id and event where individuals were not observerd
  spread(captureYear, detect, fill = 0) %>% 
  
  # For every individual....
  group_by(animalID) %>%
  # Paste together 0's and 1's
  # Unite is similar to paste. Here we are pasting the strings together from the second column (first capture event)
  # to the last capture event ("tail(names(.),1)").
  # we don't want any characters separating 0's and 1's, so we use: sep = ""
  unite("ch", 2:tail(names(.),1), sep = "") %>%
  
  # add sex (as factor)
  merge(., capData_byIndiv_v5[, c("animalID", "sex")], by = "animalID") %>%
  distinct() %>%
  mutate(
    sex = as.factor(sex)
  )


# species/sex subsets
## lwed
capHist_2009to2019_lwed <- capHist_2009to2019 %>%
  filter(animalID %in% capData_2009to2019_lwed$animalID)

capHist_2009to2019_lwedF <- capHist_2009to2019_lwed %>%
  filter(sex == "F") %>%
  select(-sex)
capHist_2009to2019_lwedM <- capHist_2009to2019_lwed %>%
  filter(sex == "M") %>%
  select(-sex)

## simp
capHist_2009to2019_simp <- capHist_2009to2019 %>%
  filter(animalID %in% capData_2009to2019_simp$animalID)

capHist_2009to2019_simpF <- capHist_2009to2019_simp %>%
  filter(sex == "F") %>%
  select(-sex)
capHist_2009to2019_simpM <- capHist_2009to2019_simp %>%
  filter(sex == "M") %>%
  select(-sex)
```

##### Run model

**Load 'marked'**

Using package 'marked' (make sure RMark is detached first! too many
overlapping function names)

```{r}
detach("package:RMark", unload=TRUE)
library(marked)
```

###### Full population

**LWED**

```{r}
# Jolly-Seber models (POPAN formulation) are open population models, and 
# can be used to estimate abundance by including two more parameters than the CJS

# Additional parameters:
# Nsuper (or "superpopulation") = total number of individuals available to enter population throughout study
# pent ("probability of entry") =  the rate at which individuals enter the population from Nsuper (via births and immigration)

# WARNING: there is no adequate GOF tests for Jolly-Seber models. 
# One common method: Test equivalent structure of CJS model with R2ucare (previous tutorials).

# This tests *some* assumptions of Phi and p.
# Jolly-Seber models have an additional assumption:
# marked AND unmarked animals have same p (R2ucare doesn't test this)
# This assumption is required to estimate total abundance (sum of marked and unmarked animals in population)

# First, process data (Notice model = "JS", previous version = "CJS"); "collapses data" by distinct rows
lwed2009to2019.js.proc <- process.data(capHist_2009to2019_lwed,
                                       model = "JS",
                                       groups = "sex")

# Second, make design data (from processed data)
lwed2009to2019.js.ddl <- make.design.data(lwed2009to2019.js.proc)

fit.js.lwed2009to2019.models <- function(){
  # Phi formulas
  Phi.dot <- list(formula=~1)
  Phi.time <- list(formula=~time)
  # p formulas
  p.dot <- list(formula=~1)
  # pent formulas. pent estimates MUST SUM to 1 (for each group).
  # This is constained using a Multinomial Logit link
  pent.time <- list(formula=~time)
  pent.sex <- list(formula=~sex)
  pent.dot <- list(formula=~1)
  # Nsuper formulas. Don't confuse "N" from model with predicted population size
  N.sex <- list(formula=~sex)
  N.dot <- list(formula=~1)
  cml <- create.model.list(c("Phi","p", "pent", "N"))
  results <- crm.wrapper(cml,
                         data = lwed2009to2019.js.proc,
                         ddl = lwed2009to2019.js.ddl,
                         external = FALSE,
                         accumulate = FALSE,
                         hessian = TRUE)
  
  return(results)
}

# Run function
lwed2009to2019.js.models <- fit.js.lwed2009to2019.models()

# Display model table
lwed2009to2019.js.models
```

Look at estimates of top model (row number on left of model table, or
using name)

```{r}
lwed2009to2019.js.models[[1]]  # or dipper.js.models[["Phi.dot.p.dot.pent.dot.N.dot"]] or dipper.js.models$Phi.dot.p.dot.pent.dot.N.dot
```

Estimate number of unmarked indivs

```{r}
# The estimates above are not on probability scale (or in individuals for N)
# (e.g. Phi, p on logit scale, pent on mlogit scale)
# Predict (real) values using top model
lwed2009to2019.js.predicted <- predict(lwed2009to2019.js.models[[1]]) # [[1]] just calls the model row according to the model table.

# Look at predictions of real parameters
lwed2009to2019.js.predicted
```

Output shows:

-   survival b/t capture events = 0.69
-   detection prob = 0.78
-   prob of entry = 0.07 each capture event
-   number of unmarked indiv is about 8, making the super population
    \~7 + 145 marked indiv = \~152

**SIMP**

```{r}
# Jolly-Seber models (POPAN formulation) are open population models, and 
# can be used to estimate abundance by including two more parameters than the CJS

# Additional parameters:
# Nsuper (or "superpopulation") = total number of individuals available to enter population throughout study
# pent ("probability of entry") =  the rate at which individuals enter the population from Nsuper (via births and immigration)

# WARNING: there is no adequate GOF tests for Jolly-Seber models. 
# One common method: Test equivalent structure of CJS model with R2ucare (previous tutorials).

# This tests *some* assumptions of Phi and p.
# Jolly-Seber models have an additional assumption:
# marked AND unmarked animals have same p (R2ucare doesn't test this)
# This assumption is required to estimate total abundance (sum of marked and unmarked animals in population)

# First, process data (Notice model = "JS", previous version = "CJS"); "collapses data" by distinct rows
simp2009to2019.js.proc <- process.data(capHist_2009to2019_simp,
                                       model = "JS",
                                       groups = "sex")

# Second, make design data (from processed data)
simp2009to2019.js.ddl <- make.design.data(simp2009to2019.js.proc)

fit.js.simp2009to2019.models <- function(){
  # Phi formulas
  Phi.dot <- list(formula=~1)
  Phi.time <- list(formula=~time)
  # p formulas
  p.dot <- list(formula=~1)
  # pent formulas. pent estimates MUST SUM to 1 (for each group).
  # This is constained using a Multinomial Logit link
  pent.time <- list(formula=~time)
  pent.sex <- list(formula=~sex)
  pent.dot <- list(formula=~1)
  # Nsuper formulas. Don't confuse "N" from model with predicted population size
  N.sex <- list(formula=~sex)
  N.dot <- list(formula=~1)
  cml <- create.model.list(c("Phi","p", "pent", "N"))
  results <- crm.wrapper(cml,
                         data = simp2009to2019.js.proc,
                         ddl = simp2009to2019.js.ddl,
                         external = FALSE,
                         accumulate = FALSE,
                         hessian = TRUE)
  
  return(results)
}

# Run function
simp2009to2019.js.models <- fit.js.simp2009to2019.models()

# Display model table
simp2009to2019.js.models
```

Look at estimates of top model (row number on left of model table, or
using name)

```{r}
simp2009to2019.js.models[[1]]  # or dipper.js.models[["Phi.dot.p.dot.pent.dot.N.dot"]] or dipper.js.models$Phi.dot.p.dot.pent.dot.N.dot
```

Estimate number of unmarked indivs

```{r}
# The estimates above are not on probability scale (or in individuals for N)
# (e.g. Phi, p on logit scale, pent on mlogit scale)
# Predict (real) values using top model
simp2009to2019.js.predicted <- predict(simp2009to2019.js.models[[1]]) # [[1]] just calls the model row according to the model table.

# Look at predictions of real parameters
simp2009to2019.js.predicted 
```

Output shows: - survival b/t capture events = 0.75 - detection prob =
0.72 - prob of entry = 0.08 each capture event - number of unmarked
indiv is about 8, making the super populat \~6 + 102 marked indiv =
\~108

###### Females

**LWED**

```{r}
# 1. process data
lwedF2009to2019.js.proc <- process.data(capHist_2009to2019_lwedF,
                                        model = "JS")

# 2. make design data
lwedF2009to2019.js.ddl <- make.design.data(lwedF2009to2019.js.proc)

fit.js.lwedF2009to2019.models <- function(){
  # Phi formulas
  Phi.dot <- list(formula=~1)
  Phi.time <- list(formula=~time)
  # p formulas
  p.dot <- list(formula=~1)
  # pent formulas. pent estimates MUST SUM to 1 (for each group).
  # This is constained using a Multinomial Logit link
  pent.time <- list(formula=~time)
  pent.dot <- list(formula=~1)
  # Nsuper formulas. Don't confuse "N" from model with predicted population size
  N.dot <- list(formula=~1)
  cml <- create.model.list(c("Phi","p", "pent", "N"))
  results <- crm.wrapper(cml,
                         data = lwedF2009to2019.js.proc,
                         ddl = lwedF2009to2019.js.ddl,
                         external = FALSE,
                         accumulate = FALSE,
                         hessian = TRUE)
  
  return(results)
}

# Run function
lwedF2009to2019.js.models <- fit.js.lwedF2009to2019.models()

# Display model table
lwedF2009to2019.js.models
```

Look at estimates of top model (row number on left of model table, or
using name)

```{r}
lwedF2009to2019.js.models[[1]]  # or dipper.js.models[["Phi.dot.p.dot.pent.dot.N.dot"]] or dipper.js.models$Phi.dot.p.dot.pent.dot.N.dot
```

Estimate number of unmarked indivs

```{r}
# The estimates above are not on probability scale (or in individuals for N)
# (e.g. Phi, p on logit scale, pent on mlogit scale)
# Predict (real) values using top model
lwedF2009to2019.js.predicted <- predict(lwedF2009to2019.js.models[[1]]) # [[1]] just calls the model row according to the model table.

# Look at predictions of real parameters
lwedF2009to2019.js.predicted
```

**SIMP**

```{r}
# 1. process data
simpF2009to2019.js.proc <- process.data(capHist_2009to2019_simpF,
                                        model = "JS")

# 2. make design data
simpF2009to2019.js.ddl <- make.design.data(simpF2009to2019.js.proc)

fit.js.simpF2009to2019.models <- function(){
  # Phi formulas
  Phi.dot <- list(formula=~1)
  Phi.time <- list(formula=~time)
  # p formulas
  p.dot <- list(formula=~1)
  # pent formulas. pent estimates MUST SUM to 1 (for each group).
  # This is constained using a Multinomial Logit link
  pent.time <- list(formula=~time)
  pent.dot <- list(formula=~1)
  # Nsuper formulas. Don't confuse "N" from model with predicted population size
  N.dot <- list(formula=~1)
  cml <- create.model.list(c("Phi","p", "pent", "N"))
  results <- crm.wrapper(cml,
                         data = simpF2009to2019.js.proc,
                         ddl = simpF2009to2019.js.ddl,
                         external = FALSE,
                         accumulate = FALSE,
                         hessian = TRUE)
  
  return(results)
}

# Run function
simpF2009to2019.js.models <- fit.js.simpF2009to2019.models()

# Display model table
simpF2009to2019.js.models
```

Look at estimates of top model (row number on left of model table, or
using name)

```{r}
simpF2009to2019.js.models[[1]]  # or dipper.js.models[["Phi.dot.p.dot.pent.dot.N.dot"]] or dipper.js.models$Phi.dot.p.dot.pent.dot.N.dot
```

Estimate number of unmarked indivs

```{r}
# The estimates above are not on probability scale (or in individuals for N)
# (e.g. Phi, p on logit scale, pent on mlogit scale)
# Predict (real) values using top model
simpF2009to2019.js.predicted <- predict(simpF2009to2019.js.models[[1]]) # [[1]] just calls the model row according to the model table.

# Look at predictions of real parameters
simpF2009to2019.js.predicted 
```

###### Males

**LWED**

```{r}
# 1. process data
lwedM2009to2019.js.proc <- process.data(capHist_2009to2019_lwedM,
                                        model = "JS")

# 2. make design data
lwedM2009to2019.js.ddl <- make.design.data(lwedM2009to2019.js.proc)

fit.js.lwedM2009to2019.models <- function(){
  # Phi formulas
  Phi.dot <- list(formula=~1)
  Phi.time <- list(formula=~time)
  # p formulas
  p.dot <- list(formula=~1)
  # pent formulas. pent estimates MUST SUM to 1 (for each group).
  # This is constained using a Multinomial Logit link
  pent.time <- list(formula=~time)
  pent.dot <- list(formula=~1)
  # Nsuper formulas. Don't confuse "N" from model with predicted population size
  N.dot <- list(formula=~1)
  cml <- create.model.list(c("Phi","p", "pent", "N"))
  results <- crm.wrapper(cml,
                         data = lwedM2009to2019.js.proc,
                         ddl = lwedM2009to2019.js.ddl,
                         external = FALSE,
                         accumulate = FALSE,
                         hessian = TRUE)
  
  return(results)
}

# Run function
lwedM2009to2019.js.models <- fit.js.lwedM2009to2019.models()

# Display model table
lwedM2009to2019.js.models
```

Look at estimates of top model (row number on left of model table, or
using name)

```{r}
lwedM2009to2019.js.models[[1]]  # or dipper.js.models[["Phi.dot.p.dot.pent.dot.N.dot"]] or dipper.js.models$Phi.dot.p.dot.pent.dot.N.dot
```

Estimate number of unmarked indivs

```{r}
# The estimates above are not on probability scale (or in individuals for N)
# (e.g. Phi, p on logit scale, pent on mlogit scale)
# Predict (real) values using top model
lwedM2009to2019.js.predicted <- predict(lwedM2009to2019.js.models[[1]]) # [[1]] just calls the model row according to the model table.

# Look at predictions of real parameters
lwedM2009to2019.js.predicted
```

**SIMP**

```{r}
# 1. process data
simpM2009to2019.js.proc <- process.data(capHist_2009to2019_simpM,
                                        model = "JS")

# 2. make design data
simpM2009to2019.js.ddl <- make.design.data(simpM2009to2019.js.proc)

fit.js.simpM2009to2019.models <- function(){
  # Phi formulas
  Phi.dot <- list(formula=~1)
  Phi.time <- list(formula=~time)
  # p formulas
  p.dot <- list(formula=~1)
  # pent formulas. pent estimates MUST SUM to 1 (for each group).
  # This is constained using a Multinomial Logit link
  pent.time <- list(formula=~time)
  pent.dot <- list(formula=~1)
  # Nsuper formulas. Don't confuse "N" from model with predicted population size
  N.dot <- list(formula=~1)
  cml <- create.model.list(c("Phi","p", "pent", "N"))
  results <- crm.wrapper(cml,
                         data = simpM2009to2019.js.proc,
                         ddl = simpM2009to2019.js.ddl,
                         external = FALSE,
                         accumulate = FALSE,
                         hessian = TRUE)
  
  return(results)
}

# Run function
simpM2009to2019.js.models <- fit.js.simpM2009to2019.models()

# Display model table
simpM2009to2019.js.models
```

Look at estimates of top model (row number on left of model table, or
using name)

```{r}
simpM2009to2019.js.models[[1]]  # or dipper.js.models[["Phi.dot.p.dot.pent.dot.N.dot"]] or dipper.js.models$Phi.dot.p.dot.pent.dot.N.dot
```

Estimate number of unmarked indivs

```{r}
# The estimates above are not on probability scale (or in individuals for N)
# (e.g. Phi, p on logit scale, pent on mlogit scale)
# Predict (real) values using top model
simpM2009to2019.js.predicted <- predict(simpM2009to2019.js.models[[1]]) # [[1]] just calls the model row according to the model table.

# Look at predictions of real parameters
simpM2009to2019.js.predicted 
```

#### Assign N

Number of hair samples in tamRun5 vs. numbers from capData:

-   LWED
    -   n_capData = 145 marked individuals (n_fem = 68, n_male = 77)
    -   n_hair = 137 (females = 64; males = 70)
-   SIMP
    -   n_capData = 102 (n_fem = 48, n_male = 54)
    -   n_hair = 99 (females = 47, males = 52)

Using output from JS models for estimates of unmarked females & males
for each species added to capData numbers

```{r}
# model estimates of unmarked indivs:
js_lwedF <- lwedF2009to2019.js.predicted$N$estimate # N_est = 4.828369
js_lwedM <- lwedM2009to2019.js.predicted$N$estimate # N_est = 5.438029

js_simpF <- simpF2009to2019.js.predicted$N$estimate # N_est = 3.775646
js_simpM <- simpM2009to2019.js.predicted$N$estimate # N_est = 6.274664

# marked indivs from capData:
n_lwedF <- capData_2009to2019_firstEntry %>%
  filter(species == "LWED") %>%
  filter(sex == "F") %>%
  nrow() # 68
n_lwedM <- capData_2009to2019_firstEntry %>%
  filter(species == "LWED") %>%
  filter(sex == "M") %>%
  nrow() # 77

n_simpF <- capData_2009to2019_firstEntry %>%
  filter(species == "SIMP") %>%
  filter(sex == "F") %>%
  nrow() # 48
n_simpM <- capData_2009to2019_firstEntry %>%
  filter(species == "SIMP") %>%
  filter(sex == "M") %>%
  nrow() # 54

round(n_lwedF + js_lwedF) # 73 LWED_F superpop (Nf)
round(n_lwedM + js_lwedM) # 82 LWED_M superpop (Nm)

round(n_simpF + js_simpF) # 52 SIMP_F superpop (Nf)
round(n_simpM + js_simpM) # 60 SIMP_M superpop (Nm)
```

## 5.3 pedigreein

Since tamarins are cooperative breeders and females sometimes allonurse
infants, I don't think it's a good idea to include any "known"
relationships, even mother-offspring. Instead, I'll confirm likely
mother-offspring relationships with FRANz, then use this to create a
pedigree file with the confirmed mother-offspring relationships for a
second run of FRANz.

## 5.4 typingerror

\*\*should include this for franz_run1; franz_run2 can use the typing
error that it automatically calculates after I provide it w/known
relationships

Maybe good to do a few tests of different error rates-- 0.01, 0.03,
0.05, 0.1

**About**

From FRANz manual: Rate of typing error. If some parent-offspring
relationships are known (with --pedigreein), then the observed
mismatches in these relationships are used to estimate the typing error.
The minimum number of relationships is calculated with the standard
statistical methods (z=1.96, 20% accepted error). If the observed number
is too low, we assume that the error rate is constant across loci and
divide the required number by the number of loci. If this number is not
reached (or no relationships are known), the default error rate of 0.01
is used. The minimum estimated error rate is 0.005.

### qc_sampleDiff

Genotype differences for:

-   Same individual
-   Same sampleType
-   Same seqRun
-   Different samples

#### Data

In tamRun5, we have n = 3 individuals with duplicate hair samples and n
= 7 individuals with duplicate blood samples, with two samples for each
individual. We can use the GTscore duptest results (run for both 0x and
10x cutoff genos in metadataReconciliation.Rmd section 6.2)

**duptest results**

```{r}
dupTest_tamRun5_0x <- read.table("./metadataReconciliation/dupTest_polyGenResults_0x.txt", header = T) %>%
  dplyr::rename("sampleID1" = "Sample1",
                "sampleID2" = "Sample2")

dupTest_tamRun5_10x <- read.table("./metadataReconciliation/dupTest_polyGenResults_10x.txt", header = T) %>%
  dplyr::rename("sampleID1" = "Sample1",
                "sampleID2" = "Sample2")
```

**Duplicate samples in tamRun5**

```{r}
dupHair_tamRun5_sampleList <- md_tamRun5_hairSamples[duplicated(md_tamRun5_hairSamples$sampleID_unique)|duplicated(md_tamRun5_hairSamples$sampleID_unique, fromLast = TRUE),] %>%
  arrange(sampleID_unique) %>%
  mutate(
    temp = rep(c("sampleID1", "sampleID2"), 3)
  ) %>%
  select(temp, sampleID, sampleID_unique) %>%
  pivot_wider(names_from = temp,
              values_from = sampleID) %>%
  arrange(sampleID1)

dupBlood_tamRun5_sampleList <- md_tamRun5_bloodSamples[duplicated(md_tamRun5_bloodSamples$sampleID_unique)|duplicated(md_tamRun5_bloodSamples$sampleID_unique, fromLast = TRUE),] %>%
  arrange(sampleID_unique) %>%
  mutate(
    temp = rep(c("sampleID1", "sampleID2"), 7)
  ) %>%
  select(temp, sampleID, sampleID_unique) %>%
  pivot_wider(names_from = temp,
              values_from = sampleID) %>%
  arrange(sampleID1)
```

#### Results

Of the original n = 3 duplicate hair samples and n = 7 duplicate blood
samples, n = 1 duplicate hair pairs and n = 6 duplicate blood pairs
could be used to assess genotype consistency using both the 0x and 10x
coverage cutoff genotypes. The other duplicate sample pairs had under 15
common genotypes, so couldn't be used reliably.

**Results are as follows:**

Blood pairs:

-   0x cutoff (n = 6 blood pairs with \> 115 commonGenotypes)
    -   mean_discrepancyRate = 0.0288
    -   min_discrepancyRate = 0.00613
    -   max_discrepancyRate = 0.0696
-   10x cutoff (n = 6 blood pairs with \> 115 commonGenotypes)
    -   mean_discrepancyRate = 0.0208

    -   min_discrepancyRate = 0

    -   max_discrepancyRate = 0.0625

For the n = 1 hair pair (200_hair):

-   0x cutoff
    -   discrepancyRate = 0.0199
    -   commmonGenos = 151
    -   matchedGenos = 148
    -   propCommon = 0.73
    -   propMatched = 0.98
-   10x cutoff
    -   discrepancyRate = 0

    -   commmonGenos = 120

    -   matchedGenos = 120

    -   propCommon = 0.58

    -   propMatched = 1.0

**0x cutoff**

```{r}
qc_tamRun5_0x_sampleDiffs_bloodHair <- rbind(dupHair_tamRun5_sampleList,
                                             dupBlood_tamRun5_sampleList) %>%
  merge(., dupTest_tamRun5_0x, by = c("sampleID1", "sampleID2"), all.x = T) %>%
  mutate(
    discrepancyRate = 1 - proportionMatch
  )

# blood sample data
qc_tamRun5_0x_sampleDiffs_bloodData <- qc_tamRun5_0x_sampleDiffs_bloodHair %>%
  filter(str_detect(sampleID_unique, "blood")) %>%
  filter(commonGenotypes > 115)

mean(qc_tamRun5_0x_sampleDiffs_bloodData$discrepancyRate) # 0.0288
min(qc_tamRun5_0x_sampleDiffs_bloodData$discrepancyRate) # 0.930
max(qc_tamRun5_0x_sampleDiffs_bloodData$discrepancyRate) # 0.994

## blood sample data w/o 27_blood
qc_tamRun5_0x_sampleDiffs_bloodData_minus27blood <- qc_tamRun5_0x_sampleDiffs_bloodHair %>%
  filter(str_detect(sampleID_unique, "blood")) %>%
  filter(sampleID_unique != "27_blood") %>%
  filter(commonGenotypes > 100)

mean(qc_tamRun5_0x_sampleDiffs_bloodData_minus27blood$commonGenotypes) # 156
mean(qc_tamRun5_0x_sampleDiffs_bloodData_minus27blood$proportionCommon) # 0.75

mean(qc_tamRun5_0x_sampleDiffs_bloodData_minus27blood$matchedGenotypes) # 152.8
mean(qc_tamRun5_0x_sampleDiffs_bloodData_minus27blood$proportionMatch) # 0.98 - ok maybe not that much higher lol (only increases by 0.008)

min(qc_tamRun5_0x_sampleDiffs_bloodData_minus27blood$proportionMatch) # 0.969
max(qc_tamRun5_0x_sampleDiffs_bloodData_minus27blood$proportionMatch) # 0.994

mean(qc_tamRun5_0x_sampleDiffs_bloodData_minus27blood$discrepancyRate) # 0.0205

############################################

# hair sample data (only 1 record; 200_hair)
qc_tamRun5_0x_sampleDiffs_hairData <- qc_tamRun5_0x_sampleDiffs_bloodHair %>%
  filter(str_detect(sampleID_unique, "hair")) %>%
  filter(commonGenotypes > 100)
qc_tamRun5_0x_sampleDiffs_hairData

# commmonGenos = 151
# matchedGenos = 148

# propCommon = 0.73
# propMatched = 0.98

# discrepancyRate = 0.0199
```

**10x cutoff**

```{r}
qc_tamRun5_10x_sampleDiffs_bloodHair <- rbind(dupHair_tamRun5_sampleList,
                                              dupBlood_tamRun5_sampleList) %>%
  merge(., dupTest_tamRun5_10x, by = c("sampleID1", "sampleID2"), all.x = T) %>%
  mutate(
    discrepancyRate = 1 - proportionMatch
  )

# blood sample data
qc_tamRun5_10x_sampleDiffs_bloodData <- qc_tamRun5_10x_sampleDiffs_bloodHair %>%
  filter(str_detect(sampleID_unique, "blood")) %>%
  filter(commonGenotypes > 115)

mean(qc_tamRun5_10x_sampleDiffs_bloodData$discrepancyRate) # 0.0208
min(qc_tamRun5_10x_sampleDiffs_bloodData$discrepancyRate) # 0
max(qc_tamRun5_10x_sampleDiffs_bloodData$discrepancyRate) # 0.0625

## blood sample data w/o 27_blood
qc_tamRun5_10x_sampleDiffs_bloodData_minus27blood <- qc_tamRun5_10x_sampleDiffs_bloodHair %>%
  filter(str_detect(sampleID_unique, "blood")) %>%
  filter(sampleID_unique != "27_blood") %>%
  filter(commonGenotypes > 100)

mean(qc_tamRun5_10x_sampleDiffs_bloodData_minus27blood$commonGenotypes) # 129.4
mean(qc_tamRun5_10x_sampleDiffs_bloodData_minus27blood$proportionCommon) # 0.62

mean(qc_tamRun5_10x_sampleDiffs_bloodData_minus27blood$matchedGenotypes) # 127.8
mean(qc_tamRun5_10x_sampleDiffs_bloodData_minus27blood$proportionMatch) # 0.99 (still only increases by 0.008)

min(qc_tamRun5_10x_sampleDiffs_bloodData_minus27blood$proportionMatch) # 0.976
max(qc_tamRun5_10x_sampleDiffs_bloodData_minus27blood$proportionMatch) # 1.0

mean(qc_tamRun5_10x_sampleDiffs_bloodData_minus27blood$discrepancyRate) # 0.0125

############################################

# hair sample data (only 1 record; 200_hair)
qc_tamRun5_10x_sampleDiffs_hairData <- qc_tamRun5_10x_sampleDiffs_bloodHair %>%
  filter(str_detect(sampleID_unique, "hair")) %>%
  filter(commonGenotypes > 100)
qc_tamRun5_10x_sampleDiffs_hairData

# commmonGenos = 120
# matchedGenos = 120

# propCommon = 0.58
# propMatched = 1

# discrepancyRate = 0
```

### qc_seqRunDiff

Genotype differences for:

-   Same individual
-   Same sample
-   Different seqRun

Using tamRun5 vs. tamRun4 Illumina results to compare.

#### Data

```{r}
qc_seqRunDiff_sampleList <- md_tamRun5 %>%
  select(sampleID, sampleID_unique, sampleName) %>%
  merge(., md_tamRun4[, c("sampleID_unique", "sampleName", "sampleID_ill")], by = c("sampleID_unique", "sampleName")) %>%
  dplyr::rename("sampleID_run5" = "sampleID",
                "sampleID_run4" = "sampleID_ill")
```

Need to run polygen for matched tamRun5/4 samples, then duptest

**polygen**

0x cutoff

```{r}
#load locus table and 0x allele reads file
fullSet_singleSNP_locusTable <- read.delim("./05_tamRun5/03_run5GTscore/fullSet_LocusTable_singleSNPs.txt", header = TRUE, stringsAsFactors = FALSE)

alleleReads_tamRun4_0x <- read.delim("./04_tamRun4/00_illumina/03_run4GTscore/ill_fullSet_AlleleReads_singleSNPs.txt", header = TRUE, row.names = 1, stringsAsFactors = FALSE) %>%
  select(qc_seqRunDiff_sampleList$sampleID_run4) %>%
  rownames_to_column("locus")

alleleReads_tamRun4v5_0x <- read.delim("./05_tamRun5/03_run5GTscore/fullSet_AlleleReads_singleSNPs.txt", header = TRUE, row.names = 1, stringsAsFactors = FALSE) %>%
  select(qc_seqRunDiff_sampleList$sampleID_run5) %>%
  rownames_to_column("locus") %>%
  merge(., alleleReads_tamRun4_0x, by = "locus") %>%
  column_to_rownames("locus")

#generate singleSNP genotypes using the polyGen algorithm, adjust "0" formatting
polygen_tamRun4v5_0x <- polyGen(fullSet_singleSNP_locusTable, alleleReads_tamRun4v5_0x)
```

10x cutoff

```{r}
#load locus table and 0x allele reads file
fullSet_singleSNP_locusTable <- read.delim("./05_tamRun5/03_run5GTscore/fullSet_LocusTable_singleSNPs.txt", header = TRUE, stringsAsFactors = FALSE)

alleleReads_tamRun4_10x <- read.delim("./04_tamRun4/00_illumina/03_run4GTscore/ill_fullSet_AlleleReads_singleSNPs_10x.txt", header = TRUE, row.names = 1, stringsAsFactors = FALSE) %>%
  select(qc_seqRunDiff_sampleList$sampleID_run4) %>%
  rownames_to_column("locus")

alleleReads_tamRun4v5_10x <- read.delim("./05_tamRun5/03_run5GTscore/fullSet_AlleleReads_singleSNPs_10x.txt", header = TRUE, row.names = 1, stringsAsFactors = FALSE) %>%
  select(qc_seqRunDiff_sampleList$sampleID_run5) %>%
  rownames_to_column("locus") %>%
  merge(., alleleReads_tamRun4_10x, by = "locus") %>%
  column_to_rownames("locus")

#generate singleSNP genotypes using the polyGen algorithm, adjust "0" formatting
polygen_tamRun4v5_10x <- polyGen(fullSet_singleSNP_locusTable, alleleReads_tamRun4v5_10x)
```

**duptest**

```{r}
source("./GTscore_sourceScripts/GTscore_modified.R") # NOTE- added N=ATGC to script for later analyses (not included in original)
```

0x cutoff

```{r identify duplicate samples,eval=FALSE}
#convert missing genotypes "0" to NA
polygen_tamRun4v5_0x_NA <- polygen_tamRun4v5_0x
polygen_tamRun4v5_0x_NA[polygen_tamRun4v5_0x_NA == "0"] <- NA

#compare all samples, for each comparison get proportion of loci that have genotypes
#in both samples (proportionCommon) and proportion of shared loci that have identical
#genotypes (proportionMatch).
#Samples with a high proportionCommon and high proportionMatch are likely duplicates
dupTest_tamRun5v4_0x <- IDduplicateSamples(polygen_tamRun4v5_0x_NA)

write.table(dupTest_tamRun5v4_0x,"./paper3_demographics/franz/dupTest_polyGenResults_tamRun4v5_0x.txt", quote = FALSE, sep = "\t", row.names = FALSE)
```

10x cutoff

```{r identify duplicate samples,eval=FALSE}
#convert missing genotypes "0" to NA
polygen_tamRun4v5_10x_NA <- polygen_tamRun4v5_10x
polygen_tamRun4v5_10x_NA[polygen_tamRun4v5_10x_NA == "0"] <- NA

#compare all samples, for each comparison get proportion of loci that have genotypes
#in both samples (proportionCommon) and proportion of shared loci that have identical
#genotypes (proportionMatch).
#Samples with a high proportionCommon and high proportionMatch are likely duplicates
dupTest_tamRun5v4_10x <- IDduplicateSamples(polygen_tamRun4v5_10x_NA)

write.table(dupTest_tamRun5v4_10x,"./paper3_demographics/franz/dupTest_polyGenResults_tamRun4v5_10x.txt", quote = FALSE, sep = "\t", row.names = FALSE)
```

#### Results

**Results are as follows:**

Blood samples:

-   0x cutoff (n = 24 blood pairs with \> 115 commonGenotypes)
    -   mean_discrepancyRate = 0.0335
    -   min_discrepancyRate = 0.00581
    -   max_discrepancyRate = 0.0828
-   10x cutoff (n = 22 blood pairs with \> 115 commonGenotypes)
    -   mean_discrepancyRate = 0.195

    -   min_discrepancyRate = 0

    -   max_discrepancyRate = 0.690

Hair samples:

-   0x cutoff (n = 14 hair pairs with \> 115 commonGenotypes)
    -   mean_discrepancyRate = 0.0587
    -   min_discrepancyRate = 0.00685
    -   max_discrepancyRate = 0.161
-   10x cutoff (n = 5 hair pairs with \> 115 commonGenotypes)
    -   mean_discrepancyRate = 0.0334

    -   min_discrepancyRate = 0

    -   max_discrepancyRate = 0.090

**0x cutoff**

```{r}
# only has NAs; disregard
temp1 <- qc_seqRunDiff_sampleList %>%
  merge(., dupTest_tamRun5v4_0x, by.x = c("sampleID_run4", "sampleID_run5"), by.y = c("Sample1", "Sample2"), all.x = T)

# use this one
temp2 <- qc_seqRunDiff_sampleList %>%
  merge(., dupTest_tamRun5v4_0x, by.x = c("sampleID_run5", "sampleID_run4"), by.y = c("Sample1", "Sample2"), all.x = T)

qc_seqRunDiffs_0x <- temp2

# blood sample data
qc_seqRunDiffs_0x_bloodData <- qc_seqRunDiffs_0x %>%
  filter(str_detect(sampleID_unique, "blood")) %>%
  filter(commonGenotypes > 115) %>% # (115 is 75% of 154)
  mutate(
    discrepancyRate = 1 - proportionMatch
  )

mean(qc_seqRunDiffs_0x_bloodData$discrepancyRate) # 0.0335
min(qc_seqRunDiffs_0x_bloodData$discrepancyRate) # 0.00581
max(qc_seqRunDiffs_0x_bloodData$discrepancyRate) # 0.0828

############################################

# hair sample data (only 1 record; 200_hair)
qc_seqRunDiffs_0x_hairData <- qc_seqRunDiffs_0x %>%
  filter(str_detect(sampleID_unique, "hair")) %>%
  filter(commonGenotypes > 115) %>%
  mutate(
    discrepancyRate = 1 - proportionMatch
  )

mean(qc_seqRunDiffs_0x_hairData$discrepancyRate) # 0.0587
min(qc_seqRunDiffs_0x_hairData$discrepancyRate) # 0.00685
max(qc_seqRunDiffs_0x_hairData$discrepancyRate) # 0.161
```

```{r}
test <- dupTest_tamRun5v4_0x %>%
  filter(commonGenotypes > 100) %>%
  filter(Sample1 == "tamRun4.082")

test2 <- dupTest_tamRun5v4_0x %>%
  filter(commonGenotypes > 100) %>%
  filter(Sample2 == "tamRun4.082")

md_tamRun5_v5
md_tamRun4
```

tamRun4.082 50_hair has % match as follows

tamRun4.017 50_blood 0.930

tamRun5.097 50_blood 0.9245 tamRun5.352 50_hair 0.8387 why so unhappy!

**10x cutoff**

```{r}
# only has NAs; disregard
temp1 <- qc_seqRunDiff_sampleList %>%
  merge(., dupTest_tamRun5v4_10x, by.x = c("sampleID_run4", "sampleID_run5"), by.y = c("Sample1", "Sample2"), all.x = T)

# use this one
temp2 <- qc_seqRunDiff_sampleList %>%
  merge(., dupTest_tamRun5v4_10x, by.x = c("sampleID_run5", "sampleID_run4"), by.y = c("Sample1", "Sample2"), all.x = T)

qc_seqRunDiffs_10x <- temp2

# blood sample data
qc_seqRunDiffs_10x_bloodData <- qc_seqRunDiffs_10x %>%
  filter(str_detect(sampleID_unique, "blood")) %>%
  filter(commonGenotypes > 115) %>% # (115 is 75% of 154)
  mutate(
    discrepancyRate = 1 - proportionMatch
  )

mean(qc_seqRunDiffs_10x_bloodData$discrepancyRate) # 0.0195
min(qc_seqRunDiffs_10x_bloodData$discrepancyRate) # 0
max(qc_seqRunDiffs_10x_bloodData$discrepancyRate) # 0.0690

############################################

# hair sample data (only 1 record; 200_hair)
qc_seqRunDiffs_10x_hairData <- qc_seqRunDiffs_10x %>%
  filter(str_detect(sampleID_unique, "hair")) %>%
  filter(commonGenotypes > 115) %>%
  mutate(
    discrepancyRate = 1 - proportionMatch
  )

mean(qc_seqRunDiffs_10x_hairData$discrepancyRate) # 0.0334
min(qc_seqRunDiffs_10x_hairData$discrepancyRate) # 0
max(qc_seqRunDiffs_10x_hairData$discrepancyRate) # 0.090
```

## 5.5 fullsib options

From franz.man:

Fullsib relationships are very informative and reduce the candidate
parents tremendously. FRANz can identify highly probable fullsibs in t
he data with the --fullsibtest option. If it is very unlikely that your
data contains many fullsibs, you should not turn this on. False
positives can decrease the accuracy of the reconstruction. As we have
already said, true positives can greatly enhance the accuracy, but if
there are no fullsibs, you can only lose.

This feature is still experimental and the default parameters are very
arbitrary. The number of false negatives will be high, as this feature
is designed to only support the parentage inference. If fullsib
inference is your main interest, then you have to set reasonable
parameters manually.

Important: --Nmax and --fullsib test together will currently produce
biased estimates of N when many fullsibs are detected. Use --N instead
or compare the influence of --fullsibtest on the estimated sampling
rates.

**--fullsibtest** - Detect siblings. This turns our fullsib heuristic as
described in [RSK09] on or off (default --nofull-sibtest). Please read
the section FULLSIB before turning this on!

**-- fullsibparental** - Detect fullsib also in the parental generation?

**--fullsibpvth** - The p-Value threshold of the sibling filter. For 0,
1 or 2 common "compatible parental genotypes". Such sampled genotypes
are compatible to an offspring genotype according the Mendelian laws and
the typing error rate. Default is 0.001, 0.001, 0.05. This means that if
individual A and B have no common compatible parental genotypes, a
threshold of 0.001 is used. The idea behind these different thresholds
is that if A and B have a common compatible parent pair, this is an
additional hint that A and B are fullsibs. So we should use a less
conservative p-Value threshold (0.05).

**--fullsibH0** - Defines the null hypotheses (PO,HS,U) and their prior
probabilities. Examples: 0,0,1 : FS vs. Unrelated. 1,2,1*: FS vs. PO,
2xHS(HS,Aunt/Uncle), U. 2,4,1#: FS vs. 2xPO (PO and
OP),4xHS(HS,Aunt/Uncle,Grandparent/-child), U.* :default with age
data,#: default without age data

## 5.x FRANz code

Starting code

```{bash}
FRANz --femrepro 2:30 --malerepro 2:30 --Nf 81 --Nm 71 --typingerror 0.01 --fullsibtest --pout "lwed.er01_" --siblingsout "lwed.er01_" --pedigreeout "lwed.er01_" genos_lwedHair_franz.dat

FRANz --Nfmax 2 --Nmmax 5 --femrepro 1:30 --malerepro 1:30 --updatefreqs genoFile_franz.dat
```

# 5 FRANz_run1

## 5.1 Input files

### Format geno files for FRANz

#### Format csv & export

FRANz requires input files to be in a particular format, which is
apparently very similar to one of the Migrate and Phylip programs. they
provide a perl script as well as a GUI on their website
([link](http://legacy.bioinf.uni-leipzig.de/Software/FRANz/CSV_Import.html))
to convert csv files to this format.

To be able to make this csv-to-FRANz conversion, the csv file should be
formatted such that all loci have one column for each allele and
ecological data are stored in columns **before** the alleles.

An example for microsat data is below, where 376 is a genotype ID, 304
is the first allele of the first locus, and so on:

,L109a,L109b,L103a,L103b

376,304,343,436

221,268,308,436

The csv file can include the following columns:

-   genotype IDs - **needs to be at least 10 characters long**
-   sex (optional)
-   birth year (optional)
-   death year (optional)
-   mother ID (optional)
-   sampling location ID (optional)
-   start of alleles - **need to be numbers, not letters, or won't run**

```{r}
# LWED
genos0x_lwed_forFRANz <- names(genos0x_run5_lwedHair_indidOnly) %>%
  map_dfc(~ genos0x_run5_lwedHair_indidOnly %>%
            select(all_of(.x)) %>%
            separate(.x,
                     into = paste0(.x, c("a", "b")),
                     sep = ",")
          ) %>%
  mutate(across(everything(), ~ gsub("0", NA, .))) %>%
  rownames_to_column("sampleID") %>%
  
  # add sampleID_franz, sex, & animalID (animalID is just for merging birthYear)
  merge(., md_run5[, c("sampleID", "sampleID_franz", "animalID", "sex")], by = "sampleID") %>%
  relocate(c(sampleID_franz, sex)) %>%
  
  # add birthYear
  merge(., birthYear_capData_run5[, c("animalID", "birthYear_est")], by = "animalID", all.x = T) %>%
  dplyr::rename("birthYear" = "birthYear_est") %>%
  relocate(birthYear, .before = "sex") %>%
  
  # add deathYear
  merge(., deathYear_capData_run5[, c("animalID", "deathYear_est")], by = "animalID", all.x = T) %>%
  dplyr::rename("deathYear" = "deathYear_est") %>%
  relocate(deathYear, .before = "sex") %>%
  
  # pad sampleID_franz so all are exactly 10 characters (required by FRANz)
  mutate(
    sampleID_franz = str_pad(sampleID_franz, 10, pad = "_", side = "left")
  ) %>%
  select(-sampleID, -animalID) %>%
  
  # replace allele letters with numbers (required by FRANz)
  mutate_at(vars(starts_with(c("INDID", "LWED"))), ~ str_replace(., "A", "4")) %>%
  mutate_at(vars(starts_with(c("INDID", "LWED"))), ~ str_replace(., "T", "7")) %>%
  mutate_at(vars(starts_with(c("INDID", "LWED"))), ~ str_replace(., "C", "3")) %>%
  mutate_at(vars(starts_with(c("INDID", "LWED"))), ~ str_replace(., "G", "6"))
  

# SIMP
genos0x_simp_forFRANz <- names(genos0x_run5_simpHair_indidOnly) %>%
  map_dfc(~ genos0x_run5_simpHair_indidOnly %>%
            select(all_of(.x)) %>%
            separate(.x,
                     into = paste0(.x, c("a", "b")),
                     sep = ",")
          ) %>%
  mutate(across(everything(), ~ gsub("0", NA, .))) %>%
  rownames_to_column("sampleID") %>%
  
  # add sampleID_franz, sex, & animalID (animalID is just for merging birthYear)
  merge(., md_run5[, c("sampleID", "sampleID_franz", "animalID", "sex")], by = "sampleID") %>%
  relocate(c(sampleID_franz, sex)) %>%
  
  # add birthYear
  merge(., birthYear_capData_run5[, c("animalID", "birthYear_est")], by = "animalID", all.x = T) %>%
  dplyr::rename("birthYear" = "birthYear_est") %>%
  relocate(birthYear, .before = "sex") %>%
  
  # add deathYear
  merge(., deathYear_capData_run5[, c("animalID", "deathYear_est")], by = "animalID", all.x = T) %>%
  dplyr::rename("deathYear" = "deathYear_est") %>%
  relocate(deathYear, .before = "sex") %>%
  
  # pad sampleID_franz so all are exactly 10 characters (required by FRANz)
  mutate(
    sampleID_franz = str_pad(sampleID_franz, 10, pad = "_", side = "left")
  ) %>%
  select(-sampleID, -animalID) %>%
  
  # replace allele letters with numbers (required by FRANz)
  mutate_at(vars(starts_with(c("INDID", "SIMP"))), ~ str_replace(., "A", "4")) %>%
  mutate_at(vars(starts_with(c("INDID", "SIMP"))), ~ str_replace(., "T", "7")) %>%
  mutate_at(vars(starts_with(c("INDID", "SIMP"))), ~ str_replace(., "C", "3")) %>%
  mutate_at(vars(starts_with(c("INDID", "SIMP"))), ~ str_replace(., "G", "6"))
```

Export

```{r}
write.csv(genos0x_lwed_forFRANz, "./paper3_demographics/02_relatedness/franz/lwed/franz_run1/genos0x_lwedHair_franz.csv", row.names = F)

write.csv(genos0x_simp_forFRANz, "./paper3_demographics/02_relatedness/franz/simp/franz_run1/genos0x_simpHair_franz.csv", row.names = F)
```

### Convert .csv to .dat

FRANz provides a perl script (csv.pl) to convert .csv files to .dat
files. The manual says that you can also go to a GUI at
<http://legacy.bioinf.uni-leipzig.de/Software/FRANz/CSV_Import.html>,
but I haven't gotten it to work (it just loads forever).

FRANz has examples of how to use csv.pl in their "man" file; I've edited
their example slightly so that the output automatically creates a new
file.

NOTE that column IDs start with 0.

How to put output into file:
<https://askubuntu.com/questions/420981/how-do-i-save-terminal-output-to-a-file>

**NOTE** - use csv_modified.pl; line 164 has an error in the original
script -- original has "\$death_col = $data->[$death_col];", but should
be "\$death = $data->[$death_col];"

```{bash}
# franz example
perl csv.pl --in test.csv --birth_col 1 --death_col 2 --sex_col 3 --data_col 4

# my data
cd /home/rachelvoyt/Documents/UT-Grad/Development/repos/tamGenetics_primatesPeru/paper3_demographics/02_relatedness/franz/

## lwed
perl ./franz_sourceScripts/csv_modified.pl --in ./lwed/franz_run1/genos0x_lwedHair_franz.csv --alleles_per_col 1 --has_header --missing_allele 'NA' --birth_col 1 --death_col 2 --sex_col 3 --data_col 4 > ./lwed/franz_run1/genos0x_lwedHair_franz.dat

## simp
perl ./franz_sourceScripts/csv_modified.pl --in ./simp/franz_run1/genos0x_simpHair_franz.csv --alleles_per_col 1 --has_header --missing_allele 'NA' --birth_col 1 --death_col 2 --sex_col 3 --data_col 4 > ./simp/franz_run1/genos0x_simpHair_franz.dat
```

## 5.2 franz_run1

### LWED

Settings: - femrepro & malerepro 2:30 - Nf = 73 - Nm = 82 -
typingerror - run1a = 0.01 - run1b = 0.03 - run1c = 0.05 - run1d = 0.1 -
fullsibtest - fullsibparental - updatefreqs

```{bash}
cd lwed/franz_run1

# typingerror 0.01
FRANz --femrepro 2:30 --malerepro 2:30 --Nf 73 --Nm 82 --typingerror 0.01 --fullsibtest --fullsibparental --updatefreqs --pout "lwed_er0.01_parentage" --siblingsout "lwed_er0.01_siblings" --pedigreeout "lwed_er0.01_pedigree" genos0x_lwedHair_franz.dat

# typingerror 0.03
FRANz --femrepro 2:30 --malerepro 2:30 --Nf 73 --Nm 82 --typingerror 0.03 --fullsibtest --fullsibparental --updatefreqs --pout "lwed_er0.03_parentage" --siblingsout "lwed_er0.03_siblings" --pedigreeout "lwed_er0.03_pedigree" genos0x_lwedHair_franz.dat

# typingerror 0.05
FRANz --femrepro 2:30 --malerepro 2:30 --Nf 73 --Nm 82 --typingerror 0.05 --fullsibtest --fullsibparental --updatefreqs --pout "lwed_er0.05_parentage" --siblingsout "lwed_er0.05_siblings" --pedigreeout "lwed_er0.05_pedigree" genos0x_lwedHair_franz.dat

# typingerror 0.10
FRANz --femrepro 2:30 --malerepro 2:30 --Nf 73 --Nm 82 --typingerror 0.1 --fullsibtest --fullsibparental --updatefreqs --pout "lwed_er0.10_parentage" --siblingsout "lwed_er0.10_siblings" --pedigreeout "lwed_er0.10_pedigree" genos0x_lwedHair_franz.dat
```

### SIMP

Settings:

-   femrepro & malerepro 2:30
-   Nf = 52
-   Nm = 60
-   typingerror
    -   run1a = 0.01
    -   run1b = 0.03
    -   run1c = 0.05
    -   run1d = 0.1
-   fullsibtest heuristic algorithm

```{bash}
cd ./../../simp/franz_run1

# typingerror 0.01
FRANz --femrepro 2:30 --malerepro 2:30 --Nf 52 --Nm 60 --typingerror 0.01 --fullsibtest --fullsibparental --updatefreqs --pout "simp_er0.01_parentage" --siblingsout "simp_er0.01_siblings" --pedigreeout "simp_er0.01_pedigree" genos0x_simpHair_franz.dat

# typingerror 0.03
FRANz --femrepro 2:30 --malerepro 2:30 --Nf 52 --Nm 60 --typingerror 0.03 --fullsibtest --fullsibparental --updatefreqs --pout "simp_er0.03_parentage" --siblingsout "simp_er0.03_siblings" --pedigreeout "simp_er0.03_pedigree" genos0x_simpHair_franz.dat

# typingerror 0.05
FRANz --femrepro 2:30 --malerepro 2:30 --Nf 52 --Nm 60 --typingerror 0.05 --fullsibtest --fullsibparental --updatefreqs --pout "simp_er0.05_parentage" --siblingsout "simp_er0.05_siblings" --pedigreeout "simp_er0.05_pedigree" genos0x_simpHair_franz.dat

# typingerror 0.10
FRANz --femrepro 2:30 --malerepro 2:30 --Nf 52 --Nm 60 --typingerror 0.1 --fullsibtest --fullsibparental --updatefreqs --pout "simp_er0.10_parentage" --siblingsout "simp_er0.10_siblings" --pedigreeout "simp_er0.10_pedigree" genos0x_simpHair_franz.dat
```

## 5.3 Results

### Full set

#### LWED

```{r}
colnames_parentage <- c("offspring",
                        "lociTyped_offspring",
                        "parent1",
                        "lociTyped_parent1",
                        "parent2",
                        "lociTyped_parent2",
                        "LOD",
                        "posterior",
                        "commonLociTyped",
                        "mismatches",
                        "n_f",
                        "n_m",
                        "pairLOD_parent1",
                        "pairLOD_parent2",
                        "posterior_parent1",
                        "posterior_parent2",
                        "parentage_mlPedigree")

# individual results
po_lwed_run1a <- read.csv("./paper3_demographics/02_relatedness/franz/lwed/franz_run1/lwed_er0.01_parentage.csv", na.strings = c("", "NA")) %>%
  rownames_to_column("temp") %>%
  `colnames<-`(colnames_parentage) %>%
  mutate(franzRun = "run1a")

po_lwed_run1b <- read.csv("./paper3_demographics/02_relatedness/franz/lwed/franz_run1/lwed_er0.03_parentage.csv", na.strings = c("", "NA")) %>%
  rownames_to_column("temp") %>%
  `colnames<-`(colnames_parentage) %>%
  mutate(franzRun = "run1b")

po_lwed_run1c <- read.csv("./paper3_demographics/02_relatedness/franz/lwed/franz_run1/lwed_er0.05_parentage.csv", na.strings = c("", "NA")) %>%
  rownames_to_column("temp") %>%
  `colnames<-`(colnames_parentage) %>%
  mutate(franzRun = "run1c")

po_lwed_run1d <- read.csv("./paper3_demographics/02_relatedness/franz/lwed/franz_run1/lwed_er0.10_parentage.csv", na.strings = c("", "NA")) %>%
  rownames_to_column("temp") %>%
  `colnames<-`(colnames_parentage) %>%
  mutate(franzRun = "run1d")
```

Combo df

```{r}
po_lwed_run1abcd <- rbind(po_lwed_run1a, po_lwed_run1b, po_lwed_run1c, po_lwed_run1d) %>%
  # add birthYear_est to offspring
  merge(., birthYear_run5_hair[, c("sampleID_franz", "birthYear_est")], by.x = "offspring", by.y = "sampleID_franz", all.x = T) %>%
  
  # add sex for parents 1 & 2
  merge(., md_run5_hair[, c("sampleID_franz", "sex")], by.x = "parent1", by.y = "sampleID_franz", all.x = T) %>%
  dplyr::rename("parent1_sex" = "sex") %>%
  merge(., md_run5_hair[, c("sampleID_franz", "sex")], by.x = "parent2", by.y = "sampleID_franz", all.x = T) %>%
  dplyr::rename("parent2_sex" = "sex") %>%
  
  # subset
  relocate(offspring, birthYear_est, parent1, parent1_sex, parent2, parent2_sex) %>%
  filter(!is.na(parent1))
```

#### SIMP

```{r}
# individual dfs
po_simp_run1a <- read.csv("./paper3_demographics/02_relatedness/franz/simp/franz_run1/simp_er0.01_parentage.csv", na.strings = c("", "NA")) %>%
  rownames_to_column("temp") %>%
  `colnames<-`(colnames_parentage) %>%
  mutate(franzRun = "run1a")

po_simp_run1b <- read.csv("./paper3_demographics/02_relatedness/franz/simp/franz_run1/simp_er0.03_parentage.csv", na.strings = c("", "NA")) %>%
  rownames_to_column("temp") %>%
  `colnames<-`(colnames_parentage) %>%
  mutate(franzRun = "run1b")

po_simp_run1c <- read.csv("./paper3_demographics/02_relatedness/franz/simp/franz_run1/simp_er0.05_parentage.csv", na.strings = c("", "NA")) %>%
  rownames_to_column("temp") %>%
  `colnames<-`(colnames_parentage) %>%
  mutate(franzRun = "run1c")

po_simp_run1d <- read.csv("./paper3_demographics/02_relatedness/franz/simp/franz_run1/simp_er0.10_parentage.csv", na.strings = c("", "NA")) %>%
  rownames_to_column("temp") %>%
  `colnames<-`(colnames_parentage) %>%
  mutate(franzRun = "run1d")
```

Combo df

```{r}
po_simp_run1abcd <- rbind(po_simp_run1a, po_simp_run1b, po_simp_run1c, po_simp_run1d) %>%
  # add birthYear_est to offspring
  merge(., birthYear_run5_hair[, c("sampleID_franz", "birthYear_est")], by.x = "offspring", by.y = "sampleID_franz", all.x = T) %>%
  
  # add sex for parents 1 & 2
  merge(., md_run5_hair[, c("sampleID_franz", "sex")], by.x = "parent1", by.y = "sampleID_franz", all.x = T) %>%
  dplyr::rename("parent1_sex" = "sex") %>%
  merge(., md_run5_hair[, c("sampleID_franz", "sex")], by.x = "parent2", by.y = "sampleID_franz", all.x = T) %>%
  dplyr::rename("parent2_sex" = "sex") %>%
  
  # subset
  relocate(offspring, birthYear_est, parent1, parent1_sex, parent2, parent2_sex) %>%
  filter(!is.na(parent1))
```

### Maternity assignments

#### LWED

**Dataframe w/mothers only from run1abcd**

```{r}
po_lwed_run1abcd_mat1 <- po_lwed_run1abcd %>%
  filter(parent1_sex == "F") %>%
  select(!contains("parent2")) %>%
  `colnames<-`(gsub("parent1", "parent", names(.)))

po_lwed_run1abcd_mat2 <- po_lwed_run1abcd %>%
  filter(parent2_sex == "F") %>%
  select(!contains("parent1")) %>%
  `colnames<-`(gsub("parent2", "parent", names(.)))

po_lwed_run1abcd_mothers <- rbind(po_lwed_run1abcd_mat1,
                                  po_lwed_run1abcd_mat2)
```

**Compare assignments b/t error rates**

Retain only assignments with posterior \>= 90%

```{r}
po_lwed_run1_matCompare <- po_lwed_run1abcd_mothers %>%
  # retain only assignments w/posterior >= 0.9
  filter(posterior >= 0.9) %>%
  select(offspring, parent, posterior, franzRun) %>%
  na.omit() %>%
  arrange(franzRun) %>%
  pivot_wider(names_from = franzRun,
              values_from = c(parent, posterior)) %>%
  mutate(
    totalMat = rowSums(select(., contains("parent")) != 0, na.rm = T)
  ) %>%
  rowwise() %>%
  mutate(
    uniqueMat = n_distinct(c_across(parent_run1a:parent_run1d), na.rm = T)
  )
```

**Confirm assignments for pedigree.dat**

Below I'm adding the following variables as checks to determine whether
it's reasonable to confirm maternity assignments and thus use them to
create a pedigree.dat file for franz_run2:

-   offspring_birthYear
-   parent_birthYear
-   offspring_parityStart
-   parent_parityStart
-   offspring_groupName (groupName in offspring_birthYear)
-   parent_groupName (groupName in year offspring_birthYear - 1)

I'm considering the maternity assignment as "confirmed" enough to go
into the pedigree.dat file if the following are true:

-   totalMat \> 1 (more than 1 maternity assigned across run1
    typingerror conditions)
-   uniqueMat == 1
-   offspring_birthYear - parent_parityStart \>= 0
-   offspring_groupName == parent_groupName in off_birthYear - 1

```{r}
po_lwed_run1_matChecks <- po_lwed_run1_matCompare %>%
  # ditch posteriors
  select(!contains("posterior")) %>%
  
  # filter to only 1 unique mat assignment w/2+ totalMat
  filter(totalMat > 1) %>%
  filter(uniqueMat == 1) %>%
  
  # coalesce to single mother_animalID
  mutate(
    parent = coalesce(parent_run1a, parent_run1b, parent_run1c, parent_run1d)
  ) %>%
  select(offspring, parent) %>%
  
  # add birth year est
  merge(., birthYear_run5_hair[, c("sampleID_franz", "birthYear_est")], by.x = "offspring", by.y = "sampleID_franz", all.x = T) %>%
  merge(., birthYear_run5_hair[, c("sampleID_franz", "birthYear_est")], by.x = "parent", by.y = "sampleID_franz", all.x = T) %>%
  dplyr::rename("off_birthYear" = "birthYear_est.x",
                "par_birthYear" = "birthYear_est.y") %>%
  
  # add parity start date
  merge(., capData_parity_firstEntry[, c("sampleID_franz", "parityYear")], by.x = "parent", by.y = "sampleID_franz", all.x = T) %>%
  dplyr::rename("par_parityYear" = "parityYear") %>%
  
  # and offspring_birthYear - parent_parityStart
  mutate(
    diff_birth.parity = as.numeric(off_birthYear) - as.numeric(par_parityYear)
  ) %>%
  relocate(offspring) %>%
  relocate(par_birthYear, .after = diff_birth.parity) %>%
  mutate(
    off_animalID = gsub("_", "", offspring),
    off_animalID = gsub("seq.*$", "", off_animalID),
    par_animalID = gsub("_", "", parent),
    par_animalID = gsub("seq.*$", "", par_animalID)
  ) %>%
  
  # add groups
  merge(., capData_byIndiv[, c("animalID", "captureYear", "groupName")], by.x = c("off_animalID", "off_birthYear"), by.y = c("animalID", "captureYear"), all.x = T) %>%
  dplyr::rename("off_groupName" = "groupName") %>%
  
  mutate(
    temp = as.numeric(off_birthYear) - 1
  ) %>%
  
  merge(., capData_byIndiv[, c("animalID", "captureYear", "groupName")], by.x = c("par_animalID", "temp"), by.y = c("animalID", "captureYear"), all.x = T) %>%
  dplyr::rename("par_groupName" = "groupName") %>%
  select(-temp) %>%
  
  # same groupName?
  mutate(
    groupMatch = off_groupName == par_groupName
  ) %>%
  relocate(offspring, parent, off_birthYear, par_parityYear, diff_birth.parity, par_birthYear, off_groupName, par_groupName, groupMatch, off_animalID)
```

This gives us n = 9 confirmed maternity assignments for LWED

```{r}
po_lwed_run1_matConfirmed <- po_lwed_run1_matChecks %>%
  filter(diff_birth.parity >= 0) %>%
  filter(groupMatch == TRUE) %>%
  mutate(
    franzPed = str_c(parent, offspring)
  )

nrow(po_lwed_run1_matConfirmed)
```

#### SIMP

**Dataframe w/mothers only from run1abcd**

```{r}
po_simp_run1abcd_mat1 <- po_simp_run1abcd %>%
  filter(parent1_sex == "F") %>%
  select(!contains("parent2")) %>%
  `colnames<-`(gsub("parent1", "parent", names(.)))

po_simp_run1abcd_mat2 <- po_simp_run1abcd %>%
  filter(parent2_sex == "F") %>%
  select(!contains("parent1")) %>%
  `colnames<-`(gsub("parent2", "parent", names(.)))

po_simp_run1abcd_mothers <- rbind(po_simp_run1abcd_mat1,
                                  po_simp_run1abcd_mat2)
```

**Compare assignments b/t error rates**

Retain only assignments with posterior \>= 90%

```{r}
po_simp_run1_matCompare <- po_simp_run1abcd_mothers %>%
  # retain only assignments w/posterior >= 0.9
  filter(posterior >= 0.9) %>%
  select(offspring, parent, posterior, franzRun) %>%
  na.omit() %>%
  arrange(franzRun) %>%
  pivot_wider(names_from = franzRun,
              values_from = c(parent, posterior)) %>%
  mutate(
    totalMat = rowSums(select(., contains("parent")) != 0, na.rm = T)
  ) %>%
  rowwise() %>%
  mutate(
    uniqueMat = n_distinct(c_across(parent_run1a:parent_run1d), na.rm = T)
  )
```

**Confirm assignments for pedigree.dat**

Below I'm adding the following variables as checks to determine whether
it's reasonable to confirm maternity assignments and thus use them to
create a pedigree.dat file for franz_run2:

-   offspring_birthYear
-   parent_birthYear
-   offspring_parityStart
-   parent_parityStart
-   offspring_groupName (groupName in offspring_birthYear)
-   parent_groupName (groupName in year offspring_birthYear - 1)

I'm considering the maternity assignment as "confirmed" enough to go
into the pedigree.dat file if the following are true:

-   totalMat \> 1 (more than 1 maternity assigned across run1
    typingerror conditions)
-   uniqueMat == 1
-   offspring_birthYear - parent_parityStart \>= 0
-   offspring_groupName == parent_groupName in off_birthYear

```{r}
po_simp_run1_matChecks <- po_simp_run1_matCompare %>%
  # ditch posteriors
  select(!contains("posterior")) %>%
  
  # filter to only 1 unique mat assignment w/2+ totalMat
  filter(totalMat > 1) %>%
  filter(uniqueMat == 1) %>%
  
  # coalesce to single mother_animalID
  mutate(
    parent = coalesce(parent_run1a, parent_run1b, parent_run1c, parent_run1d)
  ) %>%
  select(offspring, parent) %>%
  
  # add birth year est
  merge(., birthYear_run5_hair[, c("sampleID_franz", "birthYear_est")], by.x = "offspring", by.y = "sampleID_franz", all.x = T) %>%
  merge(., birthYear_run5_hair[, c("sampleID_franz", "birthYear_est")], by.x = "parent", by.y = "sampleID_franz", all.x = T) %>%
  dplyr::rename("off_birthYear" = "birthYear_est.x",
                "par_birthYear" = "birthYear_est.y") %>%
  
  # add parity start date
  merge(., capData_parity_firstEntry[, c("sampleID_franz", "parityYear")], by.x = "parent", by.y = "sampleID_franz", all.x = T) %>%
  dplyr::rename("par_parityYear" = "parityYear") %>%
  
  # and offspring_birthYear - parent_parityStart
  mutate(
    diff_birth.parity = as.numeric(off_birthYear) - as.numeric(par_parityYear)
  ) %>%
  relocate(offspring) %>%
  relocate(par_birthYear, .after = diff_birth.parity) %>%
  mutate(
    off_animalID = gsub("_", "", offspring),
    off_animalID = gsub("seq.*$", "", off_animalID),
    par_animalID = gsub("_", "", parent),
    par_animalID = gsub("seq.*$", "", par_animalID)
  ) %>%
  
  # add groups
  merge(., capData_byIndiv[, c("animalID", "captureYear", "groupName")], by.x = c("off_animalID", "off_birthYear"), by.y = c("animalID", "captureYear"), all.x = T) %>%
  dplyr::rename("off_groupName" = "groupName") %>%
  
  merge(., capData_byIndiv[, c("animalID", "captureYear", "groupName")], by.x = c("par_animalID", "off_birthYear"), by.y = c("animalID", "captureYear"), all.x = T) %>%
  dplyr::rename("par_groupName" = "groupName") %>%
  
  # same groupName?
  mutate(
    groupMatch = off_groupName == par_groupName
  ) %>%
  relocate(offspring, parent, off_birthYear, par_parityYear, diff_birth.parity, par_birthYear, off_groupName, par_groupName, groupMatch, off_animalID)
```

This gives us n = 6 confirmed maternity assignments for SIMP

```{r}
po_simp_run1_matConfirmed <- po_simp_run1_matChecks %>%
  filter(diff_birth.parity >= 0) %>%
  filter(groupMatch == TRUE) %>%
  mutate(
    franzPed = str_c(parent, offspring)
  )

nrow(po_simp_run1_matConfirmed)
```

### Paternity assignments

#### LWED

**Dataframe w/mothers only from run1abcd**

```{r}
po_lwed_run1abcd_pat1 <- po_lwed_run1abcd %>%
  filter(parent1_sex == "M") %>%
  select(!contains("parent2")) %>%
  `colnames<-`(gsub("parent1", "parent", names(.)))

po_lwed_run1abcd_pat2 <- po_lwed_run1abcd %>%
  filter(parent2_sex == "M") %>%
  select(!contains("parent1")) %>%
  `colnames<-`(gsub("parent2", "parent", names(.)))

po_lwed_run1abcd_fathers <- rbind(po_lwed_run1abcd_pat1,
                                  po_lwed_run1abcd_pat2)
```

**Compare assignments b/t error rates**

Retain only assignments with posterior \>= 90%

```{r}
po_lwed_run1_patCompare <- po_lwed_run1abcd_fathers %>%
  # retain only assignments w/posterior >= 0.9
  filter(posterior >= 0.9) %>%
  select(offspring, parent, posterior, franzRun) %>%
  na.omit() %>%
  arrange(franzRun) %>%
  pivot_wider(names_from = franzRun,
              values_from = c(parent, posterior)) %>%
  mutate(
    totalPat = rowSums(select(., contains("parent")) != 0, na.rm = T)
  ) %>%
  rowwise() %>%
  mutate(
    uniquePat = n_distinct(c_across(parent_run1a:parent_run1d), na.rm = T)
  )
```

**Confirm assignments for pedigree.dat**

Below I'm adding the following variables as checks to determine whether
it's reasonable to confirm paternity assignments and thus use them to
create a pedigree.dat file for franz_run2:

-   offspring_birthYear
-   parent_birthYear
-   offspring_groupName (groupName in offspring_birthYear)
-   parent_groupName (groupName in year offspring_birthYear - 1)

I'm considering the paternity assignment as "confirmed" enough to go
into the pedigree.dat file if the following are true:

-   totalPat \> 1 (more than 1 paternity assigned across run1
    typingerror conditions)
-   uniquePat == 1
-   offspring_groupName == parent_groupName in off_birthYear - 1

```{r}
po_lwed_run1_patChecks <- po_lwed_run1_patCompare %>%
  # ditch posteriors
  select(!contains("posterior")) %>%
  
  # filter to only 1 unique mat assignment w/2+ totalMat
  filter(totalPat > 1) %>%
  filter(uniquePat == 1) %>%
  
  # coalesce to single paternity assignment
  mutate(
    parent = coalesce(parent_run1a, parent_run1b, parent_run1c, parent_run1d)
  ) %>%
  select(offspring, parent) %>%
  
  # add birth year est
  merge(., birthYear_run5_hair[, c("sampleID_franz", "birthYear_est")], by.x = "offspring", by.y = "sampleID_franz", all.x = T) %>%
  merge(., birthYear_run5_hair[, c("sampleID_franz", "birthYear_est")], by.x = "parent", by.y = "sampleID_franz", all.x = T) %>%
  dplyr::rename("off_birthYear" = "birthYear_est.x",
                "par_birthYear" = "birthYear_est.y") %>%
  
  # add groups
  mutate(
    off_animalID = gsub("_", "", offspring),
    off_animalID = gsub("seq.*$", "", off_animalID),
    par_animalID = gsub("_", "", parent),
    par_animalID = gsub("seq.*$", "", par_animalID)
  ) %>%
  merge(., capData_byIndiv[, c("animalID", "captureYear", "groupName")], by.x = c("off_animalID", "off_birthYear"), by.y = c("animalID", "captureYear"), all.x = T) %>%
  dplyr::rename("off_groupName" = "groupName") %>%
  
  mutate(
    temp = as.numeric(off_birthYear) - 1
  ) %>%
  
  merge(., capData_byIndiv[, c("animalID", "captureYear", "groupName")], by.x = c("par_animalID", "temp"), by.y = c("animalID", "captureYear"), all.x = T) %>%
  dplyr::rename("par_groupName" = "groupName") %>%
  select(-temp) %>%
  
  # same groupName?
  mutate(
    groupMatch = off_groupName == par_groupName
  ) %>%
  relocate(offspring, parent, off_birthYear, par_birthYear, off_groupName, par_groupName, groupMatch, off_animalID)
```

This gives us n = 9 confirmed paternity assignments for LWED

```{r}
po_lwed_run1_patConfirmed <- po_lwed_run1_patChecks %>%
  filter(groupMatch == TRUE) %>%
  mutate(
    franzPed = str_c(parent, offspring)
  )

nrow(po_lwed_run1_patConfirmed)
```

#### SIMP

**Dataframe w/mothers only from run1abcd**

```{r}
po_simp_run1abcd_pat1 <- po_simp_run1abcd %>%
  filter(parent1_sex == "M") %>%
  select(!contains("parent2")) %>%
  `colnames<-`(gsub("parent1", "parent", names(.)))

po_simp_run1abcd_pat2 <- po_simp_run1abcd %>%
  filter(parent2_sex == "M") %>%
  select(!contains("parent1")) %>%
  `colnames<-`(gsub("parent2", "parent", names(.)))

po_simp_run1abcd_fathers <- rbind(po_simp_run1abcd_pat1,
                                  po_simp_run1abcd_pat2)
```

**Compare assignments b/t error rates**

Retain only assignments with posterior \>= 90%

```{r}
po_simp_run1_patCompare <- po_simp_run1abcd_fathers %>%
  # retain only assignments w/posterior >= 0.9
  filter(posterior >= 0.9) %>%
  select(offspring, parent, posterior, franzRun) %>%
  na.omit() %>%
  arrange(franzRun) %>%
  pivot_wider(names_from = franzRun,
              values_from = c(parent, posterior)) %>%
  mutate(
    totalPat = rowSums(select(., contains("parent")) != 0, na.rm = T)
  ) %>%
  rowwise() %>%
  mutate(
    uniquePat = n_distinct(c_across(parent_run1a:parent_run1d), na.rm = T)
  )
```

**Confirm assignments for pedigree.dat**

Below I'm adding the following variables as checks to determine whether
it's reasonable to confirm paternity assignments and thus use them to
create a pedigree.dat file for franz_run2:

-   offspring_birthYear
-   parent_birthYear
-   offspring_groupName (groupName in offspring_birthYear)
-   parent_groupName (groupName in year offspring_birthYear - 1)

I'm considering the paternity assignment as "confirmed" enough to go
into the pedigree.dat file if the following are true:

-   totalPat \> 1 (more than 1 paternity assigned across run1
    typingerror conditions)
-   uniquePat == 1
-   offspring_groupName == parent_groupName in off_birthYear - 1

```{r}
po_simp_run1_patChecks <- po_simp_run1_patCompare %>%
  # ditch posteriors
  select(!contains("posterior")) %>%
  
  # filter to only 1 unique mat assignment w/2+ totalMat
  filter(totalPat > 1) %>%
  filter(uniquePat == 1) %>%
  
  # coalesce to single paternity assignment
  mutate(
    parent = coalesce(parent_run1a, parent_run1b, parent_run1c, parent_run1d)
  ) %>%
  select(offspring, parent) %>%
  
  # add birth year est
  merge(., birthYear_run5_hair[, c("sampleID_franz", "birthYear_est")], by.x = "offspring", by.y = "sampleID_franz", all.x = T) %>%
  merge(., birthYear_run5_hair[, c("sampleID_franz", "birthYear_est")], by.x = "parent", by.y = "sampleID_franz", all.x = T) %>%
  dplyr::rename("off_birthYear" = "birthYear_est.x",
                "par_birthYear" = "birthYear_est.y") %>%
  
  # add groups
  mutate(
    off_animalID = gsub("_", "", offspring),
    off_animalID = gsub("seq.*$", "", off_animalID),
    par_animalID = gsub("_", "", parent),
    par_animalID = gsub("seq.*$", "", par_animalID)
  ) %>%
  merge(., capData_byIndiv[, c("animalID", "captureYear", "groupName")], by.x = c("off_animalID", "off_birthYear"), by.y = c("animalID", "captureYear"), all.x = T) %>%
  dplyr::rename("off_groupName" = "groupName") %>%
  
  mutate(
    temp = as.numeric(off_birthYear) - 1
  ) %>%
  
  merge(., capData_byIndiv[, c("animalID", "captureYear", "groupName")], by.x = c("par_animalID", "temp"), by.y = c("animalID", "captureYear"), all.x = T) %>%
  dplyr::rename("par_groupName" = "groupName") %>%
  select(-temp) %>%
  
  # same groupName?
  mutate(
    groupMatch = off_groupName == par_groupName
  ) %>%
  relocate(offspring, parent, off_birthYear, par_birthYear, off_groupName, par_groupName, groupMatch, off_animalID)
```

This gives us n = 4 confirmed paternity assignments for SIMP

```{r}
po_simp_run1_patConfirmed <- po_simp_run1_patChecks %>%
  filter(groupMatch == TRUE) %>%
  mutate(
    franzPed = str_c(parent, offspring)
  )

nrow(po_simp_run1_patConfirmed)
```

## 5.4 pedigree.dat

Use confirmed maternity assignments to create pedigree.dat file for
franz_run2

```{r}
# LWED
franz_run1_poList_lwed <- c(sampleList_lwedHair$sampleID_franz,
                            po_lwed_run1_matConfirmed$franzPed,
                            po_lwed_run1_patConfirmed$franzPed) %>%
  as.data.frame() %>%
  dplyr::rename("137" = ".") # 137 total LWED samples

write.table(franz_run1_poList_lwed, file = "./paper3_demographics/02_relatedness/franz/lwed/franz_run2/lwed_run1_pedigree.dat", quote = F, row.names = F)

# SIMP
franz_run1_poList_simp <- c(sampleList_simpHair$sampleID_franz,
                            po_simp_run1_matConfirmed$franzPed,
                            po_simp_run1_patConfirmed$franzPed) %>%
  as.data.frame() %>%
  dplyr::rename("100" = ".") # 100 total SIMP samples

write.table(franz_run1_poList_simp, file = "./paper3_demographics/02_relatedness/franz/simp/franz_run2/simp_run1_pedigree.dat", quote = F, row.names = F)
```

# 6 FRANz_run2

## 6.1 franz_run2

```{bash}
# LWED
cd lwed/franz_run2

FRANz --femrepro 2:30 --malerepro 2:30 --Nf 73 --Nm 82 --fullsibtest --fullsibparental --updatefreqs --pout "lwed_parentage" --siblingsout "lwed_siblings" --pedigreeout "lwed_pedigree" ./../franz_run1/genos0x_lwedHair_franz.dat

# SIMP
cd ./../../simp/franz_run2

FRANz --femrepro 2:30 --malerepro 2:30 --Nf 52 --Nm 60 --fullsibtest --fullsibparental --updatefreqs --pout "simp_parentage" --siblingsout "simp_siblings" --pedigreeout "simp_pedigree" ./../franz_run1/genos0x_simpHair_franz.dat
```

## 6.2 Results

### Parentage

#### Data

```{r}
po_lwed_run2 <- read.csv("./paper3_demographics/02_relatedness/franz/lwed/franz_run2/lwed_parentage.csv", na.strings = c("", "NA")) %>%
  rownames_to_column("temp") %>%
  `colnames<-`(colnames_parentage) %>%
  
  # add birthYear_est to offspring
  merge(., birthYear_run5_hair[, c("sampleID_franz", "birthYear_est")], by.x = "offspring", by.y = "sampleID_franz", all.x = T) %>%
  
  # add sex for parents 1 & 2
  merge(., md_run5_hair[, c("sampleID_franz", "sex")], by.x = "parent1", by.y = "sampleID_franz", all.x = T) %>%
  dplyr::rename("parent1_sex" = "sex") %>%
  merge(., md_run5_hair[, c("sampleID_franz", "sex")], by.x = "parent2", by.y = "sampleID_franz", all.x = T) %>%
  dplyr::rename("parent2_sex" = "sex") %>%
  
  # subset
  relocate(offspring, birthYear_est, parent1, parent1_sex, parent2, parent2_sex) %>%
  filter(!is.na(parent1))

po_simp_run2 <- read.csv("./paper3_demographics/02_relatedness/franz/simp/franz_run2/simp_parentage.csv", na.strings = c("", "NA")) %>%
  rownames_to_column("temp") %>%
  `colnames<-`(colnames_parentage) %>%
  
  # add birthYear_est to offspring
  merge(., birthYear_run5_hair[, c("sampleID_franz", "birthYear_est")], by.x = "offspring", by.y = "sampleID_franz", all.x = T) %>%
  
  # add sex for parents 1 & 2
  merge(., md_run5_hair[, c("sampleID_franz", "sex")], by.x = "parent1", by.y = "sampleID_franz", all.x = T) %>%
  dplyr::rename("parent1_sex" = "sex") %>%
  merge(., md_run5_hair[, c("sampleID_franz", "sex")], by.x = "parent2", by.y = "sampleID_franz", all.x = T) %>%
  dplyr::rename("parent2_sex" = "sex") %>%
  
  # subset
  relocate(offspring, birthYear_est, parent1, parent1_sex, parent2, parent2_sex) %>%
  filter(!is.na(parent1))
```

#### LWED

```{r}
# MATERNITY
po_lwed_run2_mat1 <- po_lwed_run2 %>%
  filter(parent1_sex == "F") %>%
  select(!contains("parent2")) %>%
  `colnames<-`(gsub("parent1", "parent", names(.)))

po_lwed_run2_mat2 <- po_lwed_run2 %>%
  filter(parent2_sex == "F") %>%
  select(!contains("parent1")) %>%
  `colnames<-`(gsub("parent2", "parent", names(.)))

po_lwed_run2_mothers <- rbind(po_lwed_run2_mat1,
                              po_lwed_run2_mat2)

table(po_lwed_run2_mothers$parent)

# PATERNITY
po_lwed_run2_pat1 <- po_lwed_run2 %>%
  filter(parent1_sex == "M") %>%
  select(!contains("parent2")) %>%
  `colnames<-`(gsub("parent1", "parent", names(.)))

po_lwed_run2_pat2 <- po_lwed_run2 %>%
  filter(parent2_sex == "M") %>%
  select(!contains("parent1")) %>%
  `colnames<-`(gsub("parent2", "parent", names(.)))

po_lwed_run2_fathers <- rbind(po_lwed_run2_pat1,
                              po_lwed_run2_pat2)
```

#### SIMP

```{r}
# MATERNITY
po_simp_run2_mat1 <- po_simp_run2 %>%
  filter(parent1_sex == "F") %>%
  select(!contains("parent2")) %>%
  `colnames<-`(gsub("parent1", "parent", names(.)))

po_simp_run2_mat2 <- po_simp_run2 %>%
  filter(parent2_sex == "F") %>%
  select(!contains("parent1")) %>%
  `colnames<-`(gsub("parent2", "parent", names(.)))

po_simp_run2_mothers <- rbind(po_simp_run2_mat1,
                              po_simp_run2_mat2)

# PATERNITY
po_simp_run2_pat1 <- po_simp_run2 %>%
  filter(parent1_sex == "M") %>%
  select(!contains("parent2")) %>%
  `colnames<-`(gsub("parent1", "parent", names(.)))

po_simp_run2_pat2 <- po_simp_run2 %>%
  filter(parent2_sex == "M") %>%
  select(!contains("parent1")) %>%
  `colnames<-`(gsub("parent2", "parent", names(.)))

po_simp_run2_fathers <- rbind(po_simp_run2_pat1,
                              po_simp_run2_pat2)
```

# 7 CKMRsim

## 7.1 Background

Resources:
CKMRsim-example-1 https://eriqande.github.io/CKMRsim/articles/CKMRsim-example-1.html

## 7.2 Packages

```{r}
remotes::install_github("eriqande/CKMRsim")
library(CKMRsim)
library(tidyverse)
# the next step tests to see if mendel was already installed.
# if not, it installs it
if(system.file("bin", package = "CKMRsim") == "") {
  install_mendel(Dir = system.file(package = "CKMRsim"))
}

seqName_shortName <- read.csv("./project_data/seqNames_to_shortNames.csv") %>%
  select(primerName3, seqID3) %>%
  mutate(
    Chrom = sub(".*_(CM[0-9]+\\.[0-9]+).*", "\\1", seqID3)
  )
```

## Genos

Long-format genos should have four columns: Indiv, Locus, gene_copy, and
Allele.

-   Alleles are named with characters (even if they are numbers, they
    must be coerced to characters)
-   The gene_copy column contains either a 1 or a 2 in every row,
    telling us which copy of the gene (in a diploid) is which allele
-   Missing data in the Allele column is given by NA

```{r}
genos0x_lwed_forCKMR <- names(genos0x_run5_lwedHair_indidOnly) %>%
  map_dfc(~ genos0x_run5_lwedHair_indidOnly %>%
            select(all_of(.x)) %>%
            separate(.x,
                     into = paste0(.x, c(".1", ".2")),
                     sep = ",")
          ) %>%
  mutate(across(everything(), ~ gsub("0", NA, .))) %>%
  rownames_to_column("sampleID")

long_genos <- genos0x_lwed_forCKMR %>%
  rename(Indiv = sampleID) %>%
  pivot_longer(
    cols = -Indiv, 
    names_to = c("Locus", "gene_copy"), 
    names_sep = "\\.", 
    values_to = "Allele"
  )
```

## Allele frequency

Allele frequency input format:

-   **Chrom** should be a character or integer denoting which chromosome
    the marker is on. (For example 1 or X or Omy12)
-   **Locus** gives a character vector with the names of the
    markers/loci. Please don't use spaces in the marker names!
-   **Pos** gives the genome coordinates of the marker. This needs to be
    a number (double or integer)
-   **Allele** gives the name of each allele at each locus. This must be
    a character vector. So, in our example, we will coerce the numbers
    we have for each locus into a character vector. Please do not use
    spaces in allele names.
-   **Freq** is the frequency of each allele in the population. These
    should sum to 1.0 over each locus.
-   The remaining columns, **LocIdx** and **AlleIdx** are integer
    indices that get assigned to each locus and to each allele within
    each index. These columns get filled by using the reindex_markers()
    function.

```{r}
locus_names <- unique(long_genos$Locus)

afreqs_ready <- long_genos %>%
  count(Locus, Allele) %>%  
  group_by(Locus) %>%
  mutate(
    Freq = n / sum(n),
    #Chrom = "Unk",
    Pos = as.integer(factor(Locus, levels = locus_names))
  ) %>%
  
  merge(., seqName_shortName[, c("primerName3", "Chrom")], by.x = "Locus", by.y = "primerName3") %>%
  
  ungroup() %>%
  select(Chrom, Pos, Locus, Allele, Freq) %>%
  arrange(Pos, desc(Freq)) %>%
  mutate(AlleIdx = NA,
         LocIdx = NA) %>%
  # NOTE - is very impt to remove NAs from allele frequencies
  filter(!is.na(Allele)) %>%
  # pass to reindex_markers to give allele-freq df for CKRMsim
  reindex_markers()
```

## Create a CKMR object (create_ckmr)

-   **D** : the tibble of allele frequencies that has been run through
    reindex_markers().
-   **kappa_matrix** : A matrix that describes the pairwise
    relationships that you will be wanting to jointly simulate genotypes
    and likelihoods for. Each row is named by what you want to call the
    relationship and there are three columns which give, respectively,
    the probability that a pair with such a relationship share 0, 1, or
    2 genes identical-by-descent. The CKMRsim package comes with a
    matrix called kappas that has this information for 12 relationships:
    -   **MZ** : monozgotic twins (or "self"). This can be used to
        figure out how much power you have for identifying the same
        individual, sampled twice.
    -   **PO** : parent-offspring.
    -   **FS** : full siblings.
    -   **HS** : half siblings.
    -   **GP** : grandparent - grandoffspring.
    -   **AN** : aunt-neice (same as uncle-nephew or any such avuncular
        relationship)
    -   **DFC** : double first cousins.
    -   **FC** : first cousins.
    -   **HC** : half cousins.
    -   **U** : unrelated

```{r}
ckmr <- create_ckmr(
  D = afreqs_ready,
  kappa_matrix = kappas[c("PO", "FS", "HS", "HAN", "U"), ],
  ge_mod_assumed = ge_model_TGIE,
  ge_mod_true = ge_model_TGIE,
  ge_mod_assumed_pars_list = list(epsilon = 0.005),
  ge_mod_true_pars_list = list(epsilon = 0.005)
)
```

## Simulations & log-probabilities

simulate_Qij():

-   **C** : the ckmr object to use for simulation and probabilty calculation.
-   **sim_relats** : the set of true relationships you want to simulate genotypes from.
-   **calc_relats** : the set of assumed relationships you wish to compute genotype probabilities for, from the simulated data.
-   **reps** : for each relationship in sim_relats, the number of genotype pairs to simulate. This is, by default 10,000.

PO = parent-offspring FS = full sib HS = half sib HAN = half aunt-niece
(or uncle/niece) U = unrelated

```{r}
# This simulates a large number of pairwise genotype probabilites
# at each locus
Qs <- simulate_Qij(
  ckmr, 
  calc_relats = c("PO", "FS", "HS", "HAN", "U"),
  sim_relats = c("PO", "FS", "HS", "HAN", "U") 
)
```

extract_logls() function lets us get simulated log-likelihood ratios out of the Qij object we created above and plot a histogram or a density plot of the distributions. This is helpful for developing intuition about things and understanding what is going on

```{r}
# In the following, we extract those genotype probabilities in different
# ways to calculate a variety of different log-likelihood ratios
# for different relationships.  

# in this particular case, we are looking at log likelihood ratios
# for the hypothesis of Parent-Offspring vs Unrelated
PO_U_logls <- extract_logls(
  Qs,
  numer = c(PO = 1),
  denom = c(U = 1)
)

# And we can plot the distribution of those logl ratios for each
# of the different true relationships. 
ggplot(
  PO_U_logls,
  aes(x = logl_ratio, fill = true_relat)
) +
  geom_density(alpha = 0.25)

# can limit focus just to PO vs. U pairs:
ggplot(PO_U_logls %>% filter(true_relat %in% c("PO", "U")),
            aes(x = logl_ratio, fill = true_relat)) +
  geom_density(alpha = 0.25)

# or FS vs. U
FS_U_logls <- extract_logls(Qs,
                            numer = c(FS = 1),
                            denom = c(U = 1))

ggplot(FS_U_logls %>% filter(true_relat %in% c("FS", "U")),
            aes(x = logl_ratio, fill = true_relat)) +
  geom_density(alpha = 0.25)
```

## False neg/pos rates

```{r}
PO_is <- mc_sample_simple(Qs, 
                 nu = "PO",
                 de = "U")
PO_is
```





They use max_mismatch = 500 out of 2500 total genos (i.e., 20%).
0.2\*154 = 30.8

```{r}
matchers <- find_close_matching_genotypes(
  LG = long_genos,
  CK = ckmr,
  max_mismatch = 31
)
matchers
```
