---
title: "tamGenetics_ot2InputFiles"
author: "Rachel Voyt"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1 Overview

This document is for the creation of input files for the tamGenetics OT2 protocols, incuding:

1. PCR1 setup
2. 1:20 dilutions
3. PCR2 setup
4. Bead cleanups

# 2 Packages

```{r}
library(tidyverse)
library(readxl)
```

# 3 Static input values

## 3.1 Labware, pipettes, & tipracks

```{r}
# Labware
xtnPlate <- "biorad_96_wellplate_200ul_pcr"
pcrPlate <- "biorad_96_wellplate_200ul_pcr"
tubeRack <- "opentrons_24_tuberack_eppendorf_1.5ml_safelock_snapcap"
magblock <- "magnetic_module_gen2"

# Pipettes
s20 <- "p20_single_gen2"
m20 <- "p20_multi_gen2"
s300 <- "p300_single_gen2"
m300 <- "p300_multi_gen2"

# Tipracks
tiprack_p20 <- "opentrons_96_tiprack_300ul"
tiprack_p20f <- "opentrons_96_filtertiprack_20ul"
tiprack_p300 <- "opentrons_96_tiprack_300ul"
tiprack_p300f <- "opentrons_96_filtertiprack_200ul"
```

## 3.2 Well values

```{r}
# Row values
rowValues_plate <- rep(LETTERS[1:8], times = 12)
rowValues_tuberack <- rep(LETTERS[1:4], times = 6)

# Column values
colValues_plate <- rep(1:12, times = 8) %>%
  sort() %>%
  as.character()
colValues_tuberack <- rep(1:6, times = 4) %>%
  sort() %>%
  as.character()

# Well values
wellValues_plate <- str_c(rowValues_plate, colValues_plate)
wellValues_tuberack <- str_c(rowValues_tuberack, colValues_tuberack)
```

# 4 OT2 input for PRIMER protocols

## 4.1 Primer rehydration (200 uM)

### 4.1.1 Data

```{r}
plateSpecs_p56 <- read_excel("./primers/02_rehydrationPooling/plateSpecs_tamGenetics_primerPlates5,6.xlsx") %>%
  as.data.frame() %>%
  rename_with(make.names)
```

### 4.1.2 OT2 csv

```{r}
ot2csv_p56Rehydrate <- plateSpecs_p56 %>%
  mutate(source_labware = "nest_12_reservoir_15ml",
         source_slot = 2,
         source_well = c(rep("A2", 70), rep("A3", 70)),
         asp_height = 1,
         dest_labware = "vwr_96_wellplate_1000ul",
         dest_slot = case_when(
           Plate.Name == "tamGenetics_primersPlate5" ~ 3,
           Plate.Name == "tamGenetics_primersPlate6" ~ 6),
         transfer_volume = round(5*nmoles, 1)) %>%
  mutate(dest_well = gsub("0", "", Well.Position)) %>%
  mutate(destRow = str_sub(dest_well, 1, 1),
         destCol = str_sub(dest_well, 2, length(dest_well))) %>%
  arrange(dest_slot, destCol, destRow) %>%
  select(c("source_labware", "source_slot", "source_well", "asp_height", "dest_labware", "dest_slot", "dest_well", "transfer_volume"))

write.csv(ot2csv_p56Rehydrate, "./robotProtocols/inputFiles/tamGenetics_primerRehydration200uM_primerPlates5,6.csv", row.names = F)
```

## 4.2 Primer dilutions & pooling (post-opt2)

### 4.2.1 Data

```{r}
# Primer pool splits
primerPool_splitsOriginal <- read.csv("./primers/03_lociChoices/primerPool_splits.csv")

primerPool2 <- primerPool_splitsOriginal %>%
  filter(primerPool == 2)

primerPool3 <- primerPool_splitsOriginal %>%
  filter(primerPool == 3)

# Primer list post-optimization 2
primers_postOpt2 <- read.csv("./primers/03_lociChoices/primerList_postOpt2.csv")

# Dilution plate post-optimization 2
dilutionPlate <- read.csv("./primers/03_lociChoices/dilutionPlate_postOpt2.csv")
```

### 4.2.2 Primer dilutions

#### Add primers, then water

```{r}
# Add primers to dilution plate
primers_forDilutions <- dilutionPlate %>%
  select(-plate) %>%
  mutate(source_slot = case_when(originalPlate == 1 ~ 2,
                                 originalPlate == 2 ~ 5,
                                 originalPlate == 5 ~ 5,
                                 originalPlate == 6 ~ 8)) %>%
  dplyr::rename("plate" = "originalPlate",
                "source_well" = "originalWell",
                "dest_well" = "dilutionPlate_well",
                "transfer_volume" = "ulPrimer_toAdd") %>%
  mutate(dest_slot = 11) %>%
  mutate(source_labware = "vwr_96_wellplate_1000ul") %>%
  mutate(dest_labware = "biorad_96_wellplate_200ul_pcr") %>%
  select(c("plate", "source_labware", "source_slot", "source_well", "dest_labware", "dest_slot", "dest_well", "transfer_volume", "bp3index"))

# Add water to dilution plate
water_forDilutions <- dilutionPlate %>%
  select(-plate) %>%
  dplyr::rename("plate" = "originalPlate",
                "transfer_volume" = "ulWater_toAdd",
                "dest_well" = "dilutionPlate_well") %>%
  mutate(source_slot = 11) %>%
  mutate(source_well = "A12") %>%
  mutate(dest_slot = 11) %>%
  mutate(source_labware = "biorad_96_wellplate_200ul_pcr") %>%
  mutate(dest_labware = "biorad_96_wellplate_200ul_pcr") %>%
  select(c("plate", "source_labware", "source_slot", "source_well", "dest_labware", "dest_slot", "dest_well", "transfer_volume", "bp3index"))
```

#### Separate by plates

Create separate primer/water adding instructions based on plates present on the OT2 deck - will need to do two rounds, one with plates 1, 2, 3 and one with plates 4, 5, 6.

```{r}
# Plates 1, 2, 3
primersWater_dilutions_plates123 <- rbind(primers_forDilutions, water_forDilutions) %>%
  filter(plate %in% c("1", "2", "3")) %>%
  mutate(plate = ifelse(source_well == "A12", "dilution", plate)) %>%
  dplyr::rename("source_well" = "source_well") %>%
  mutate(plate = gsub("dilution", "dilution123", plate))

# Plates 4, 5, 6 (primers that go into primer pool 3)
primersWater_dilutions_plates456 <- rbind(primers_forDilutions, water_forDilutions) %>%
  filter(plate %in% c("4", "5", "6")) %>%
  mutate(plate = ifelse(source_well == "A12", "dilution", plate)) %>%
  dplyr::rename("source_well" = "source_well") %>%
  mutate(plate = gsub("dilution", "dilution456", plate))
```

#### Add diluted primers to pools

```{r}
vol_primer_toPool <- 4

dilutedPrimers_toPool_plates123 <- primersWater_dilutions_plates123 %>%
  filter(plate == "dilution123") %>%
  select(c("plate", "dest_slot", "dest_well", "bp3index")) %>%
  dplyr::rename("source_slot" = "dest_slot",
                "source_well" = "dest_well") %>%
  mutate(dest_slot = 3) %>%
  mutate(dest_well = case_when(bp3index %in% primerPool2$bp3Index ~ "B1",
                                 bp3index %in% primerPool3$bp3Index ~ "C1")) %>%
  mutate(transfer_volume = vol_primer_toPool) %>%
  mutate(source_labware = "biorad_96_wellplate_200ul_pcr") %>%
  mutate(dest_labware = "opentrons_24_tuberack_eppendorf_1.5ml_safelock_snapcap") %>%
  select(c("plate", "source_labware", "source_slot", "source_well", "dest_labware", "dest_slot", "dest_well", "transfer_volume", "bp3index"))

dilutedPrimers_toPool_plates456 <- primersWater_dilutions_plates456 %>%
  filter(plate == "dilution456") %>%
  select(c("plate", "dest_slot", "dest_well", "bp3index")) %>%
  dplyr::rename("source_slot" = "dest_slot",
                "source_well" = "dest_well") %>%
  mutate(dest_slot = 3) %>%
  mutate(dest_well = case_when(bp3index %in% primerPool2$bp3Index ~ "B1",
                                 bp3index %in% primerPool3$bp3Index ~ "C1")) %>%
  mutate(transfer_volume = vol_primer_toPool) %>%
  mutate(source_labware = "biorad_96_wellplate_200ul_pcr") %>%
  mutate(dest_labware = "opentrons_24_tuberack_eppendorf_1.5ml_safelock_snapcap") %>%
  select(c("plate", "source_labware", "source_slot", "source_well", "dest_labware", "dest_slot", "dest_well", "transfer_volume", "bp3index"))
```

#### Combine steps: Dilute primers & add to pools

```{r}
dilutedPrimers_plates123_forOT2 <- rbind(primersWater_dilutions_plates123, dilutedPrimers_toPool_plates123)

dilutedPrimers_plates456_forOT2 <- rbind(primersWater_dilutions_plates456, dilutedPrimers_toPool_plates456)
```

### 4.2.3 Split pools

#### Transfer info

```{r}
# Plates 1, 2, 3
ot2_plates123d <- primers_postOpt2 %>%
  # Add non-diluted primers
  filter(is.na(action)) %>%
  filter(plate %in% c("1", "2", "3")) %>%
  select(c("plate", "Well.Position", "bp3index")) %>%
  dplyr::rename("source_well" = "Well.Position") %>%
  mutate(col = substr(source_well, 1, 1)) %>%
  mutate(row = substr(source_well, 2, nchar(source_well))) %>%
  arrange(plate, as.numeric(row), col) %>%
  mutate("source_slot" = 
           case_when(plate == "1" ~ 2,
                     plate == "2" ~ 5,
                     plate == "3" ~ 8)) %>%
  mutate("dest_slot" = 3) %>%
  mutate("dest_well" = case_when(bp3index %in% primerPool2$bp3Index ~ "B1",
                                 bp3index %in% primerPool3$bp3Index ~ "C1")) %>%
  mutate("transfer_volume" = vol_primer_toPool) %>%
  mutate("source_labware" = "vwr_96_wellplate_1000ul") %>%
  mutate("dest_labware" = "opentrons_24_tuberack_eppendorf_1.5ml_safelock_snapcap") %>%
  select(c("plate", "source_labware", "source_slot", "source_well", "dest_labware", "dest_slot", "dest_well", "transfer_volume", "bp3index")) %>%
  # Dilute & add diluted primers
  rbind(., dilutedPrimers_plates123_forOT2)

# Plates 4, 5, 6
ot2_plates456d <- primers_postOpt2 %>%
  # Add non-diluted primers
  filter(is.na(action)) %>%
  filter(plate %in% c("4", "5", "6")) %>%
  select(c("plate", "Well.Position", "bp3index")) %>%
   dplyr::rename("source_well" = "Well.Position") %>%
  mutate(col = substr(source_well, 1, 1)) %>%
  mutate(row = substr(source_well, 2, nchar(source_well))) %>%
  arrange(plate, as.numeric(row), col) %>%
  mutate(source_slot = 
           case_when(plate == "4" ~ 2,
                     plate == "5" ~ 5,
                     plate == "6" ~ 8)) %>%
  mutate(dest_slot = 3) %>%
  mutate(dest_well = case_when(bp3index %in% primerPool2$bp3Index ~ "B1",
                                 bp3index %in% primerPool3$bp3Index ~ "C1")) %>%
  mutate(transfer_volume = vol_primer_toPool) %>%
  mutate(source_labware = "vwr_96_wellplate_1000ul") %>%
  mutate(dest_labware = "opentrons_24_tuberack_eppendorf_1.5ml_safelock_snapcap") %>%
  select(c("plate", "source_labware", "source_slot", "source_well", "dest_labware", "dest_slot", "dest_well", "transfer_volume", "bp3index")) %>%
  # Dilute & add diluted primers
  rbind(., dilutedPrimers_plates456_forOT2)

# All plates together
splitPools_ot2 <- rbind(ot2_plates123d, ot2_plates456d)
```

#### Pool volumes: split pools initial

```{r}
# Volumes added to split pool tubes
ot2_plates123d_primersOnly <- ot2_plates123d %>%
  filter(transfer_volume == vol_primer_toPool)

ot2_plates456d_primersOnly <- ot2_plates456d %>%
  filter(transfer_volume == vol_primer_toPool)

# Pool 2
pool2_plates123d <- ot2_plates123d_primersOnly %>%
  filter(dest_well == "B1")
pool2_plates456d <- ot2_plates456d_primersOnly %>%
  filter(dest_well == "B1")

vol_pool2_initial <- sum(as.numeric(pool2_plates123d$transfer_volume)) + sum(as.numeric(pool2_plates456d$transfer_volume))

# Pool 3
pool3_plates123d <- ot2_plates123d_primersOnly %>%
  filter(dest_well == "C1")
pool3_plates456d <- ot2_plates456d_primersOnly %>%
  filter(dest_well == "C1")

vol_pool3_initial <- sum(as.numeric(pool3_plates123d$transfer_volume)) + sum(as.numeric(pool3_plates456d$transfer_volume))
```

#### uM per oligo: split pools initial

c1v1=c2v2
c2=c1v1/v2

```{r}
uM_pool2_initial <- (200*vol_primer_toPool)/vol_pool2_initial
uM_pool3_initial <- (200*vol_primer_toPool)/vol_pool3_initial
```

#### Match uM split pools

Primer pool 2 currently has a slightly higher concentration per oligo; adjust to match pool 3.

##### Water to add

c1v1 = c2v2
v2 = c1v1/c2
water = v2 - v1

```{r}
water_forPool2_initial <- ((uM_pool2_initial*vol_pool2_initial)/uM_pool3_initial) - vol_pool2_initial
```

##### Transfer info

```{r}
uM_splitPoolAdjust_ot2 <- data.frame(
  plate = "water_finalPools",
  source_labware = "opentrons_24_tuberack_eppendorf_1.5ml_safelock_snapcap",
  source_slot = 3,
  source_well = "A6",
  dest_labware = "opentrons_24_tuberack_eppendorf_1.5ml_safelock_snapcap",
  dest_slot = 3,
  dest_well = "B1",
  transfer_volume = water_forPool2_initial,
  bp3index = NA)
```

##### Pool volumes: post-concentration match

```{r}
vol_pool2_postConcMatch <- vol_pool2_initial + water_forPool2_initial
vol_pool3_postConcMatch <- vol_pool3_initial
```

##### uM per oligo: post-concentration match

```{r}
uM_pool2_postConcMatch <- (200*vol_primer_toPool)/vol_pool2_postConcMatch
uM_pool3_postConcMatch <- uM_pool3_initial
```

### 4.2.4 Primer pool 1

#### Transfer info

Transfer 350 ul from each split primer pool.

```{r}
vol_pool2_forPool1 <- 350
vol_pool3_forPool1 <- 350

primerPool1_ot2 <- data.frame(
  plate = rep("water_finalPools", 2),
  source_labware = rep("opentrons_24_tuberack_eppendorf_1.5ml_safelock_snapcap", 2),
  source_slot = rep(3, 2),
  source_well = c("B1", "C1"),
  dest_labware = rep("opentrons_24_tuberack_eppendorf_1.5ml_safelock_snapcap", 2),
  dest_slot = rep(3, 2),
  dest_well = rep("A1", 2),
  transfer_volume = c(vol_pool2_forPool1, vol_pool3_forPool1),
  bp3index = rep(NA, 2))
```

##### Pool volumes: post-pool1

```{r}
vol_pool1_postPool1 <- sum(as.numeric(primerPool1_ot2$transfer_volume))
vol_pool2_postPool1 <- vol_pool2_postConcMatch - vol_pool2_forPool1
vol_pool3_postPool1 <- vol_pool3_postConcMatch - vol_pool3_forPool1
```

##### uM per oligo: pool 1

Since each split pool is now at the same concentration per oligo, combining them results in a pool with the same concentration per oligo. HOWEVER, since the same oligos aren't present in both pools, the concentration per split pool 2 oligos and split pool 3 oligos will actually decrease since the total volume is increasing. 

Keeping this in mind, the concentration per oligo in primer pool 1 can be calculated as follows, using either of the split pools for the starting concentration/volume (since they're both the same):

c1v1 = c2v2
c2 = c1v1/v2

```{r}
uM_pool1_initial <- (uM_pool2_postConcMatch*vol_pool2_forPool1)/vol_pool1_postPool1
```

### 4.2.5 Create SDZ & LA lab aliquots

#### Transfer info

```{r}
sdzLA_subsets_ot2 <- data.frame(
  plate = rep("water_finalPools", 3),
  source_labware = rep("opentrons_24_tuberack_eppendorf_1.5ml_safelock_snapcap", 3),
  source_slot = rep(3, 3),
  source_well = c("A1", "B1", "C1"),
  dest_labware = rep("opentrons_24_tuberack_eppendorf_1.5ml_safelock_snapcap", 3),
  dest_slot = rep(3, 3),
  dest_well = c("A2", "B2", "C2"),
  transfer_volume = c(vol_pool1_postPool1/2, vol_pool2_postPool1/2, vol_pool3_postPool1/2),
  bp3index = rep(NA, 3))
```

#### Pool volumes: post-aliquots

```{r}
vol_pool1_sdzLA <- vol_pool1_postPool1/2
vol_pool2_sdzLA <- vol_pool2_postPool1/2
vol_pool3_sdzLA <- vol_pool3_postPool1/2
```

### 4.2.6 Adjust to 0.25 uM per oligo

#### Water to add

v2 = c1v1/c2
water = v2 - v1 = c1v1/c2 - v1

```{r}
# water to add to each aliquot of pool
water_forPool1 <- ((uM_pool1_initial*vol_pool1_sdzLA)/0.25) - vol_pool1_sdzLA
water_forPool2 <- ((uM_pool2_postConcMatch*vol_pool2_sdzLA)/0.25) - vol_pool2_sdzLA
water_forPool3 <- ((uM_pool3_postConcMatch*vol_pool3_sdzLA)/0.25) - vol_pool3_sdzLA
```

#### Transfer info

```{r}
uM_finalAdjust_ot2 <- data.frame(
  plate = rep("water_finalPools", 6),
  source_labware = rep("opentrons_24_tuberack_eppendorf_1.5ml_safelock_snapcap", 6),
  source_slot = rep(3, 6),
  source_well = c("A6", "A6", "B5", "B6", "C5", "C6"),
  dest_labware = rep("opentrons_24_tuberack_eppendorf_1.5ml_safelock_snapcap", 6),
  dest_slot = rep(3, 6),
  dest_well = c("A1", "A2", "B1", "B2", "C1", "C2"),
  transfer_volume = c(water_forPool1, water_forPool1, water_forPool2, water_forPool2, water_forPool3, water_forPool3),
  bp3index = rep(NA, 6)) %>%
  mutate(transfer_volume = round(transfer_volume, 1))
```

### 4.2.6 Full OT2 input

Combine dilutions, split primer pools, primer pool 1 creation, SDZ/laLab subsets, and uM adjustments:

```{r}
ot2_final_plusPlates <- rbind(splitPools_ot2, uM_splitPoolAdjust_ot2, primerPool1_ot2, sdzLA_subsets_ot2, uM_finalAdjust_ot2) %>%
  mutate(transfer_volume = round(as.numeric(transfer_volume), digits = 1)) %>%
  select(-bp3index) %>%
  mutate(plate = case_when(plate == 1 ~ "primerPlate1",
                           plate == 2 ~ "primerPlate2",
                           plate == 3 ~ "primerPlate3",
                           plate == 4 ~ "primerPlate4",
                           plate == 5 ~ "primerPlate5",
                           plate == 6 ~ "primerPlate6",
                           plate == "dilution123" ~ "dilution123",
                           plate == "dilution456" ~ "dilution456",
                           plate == "water_finalPools" ~ "water_finalPools")) %>%
  mutate(asp_height_mm = 1) %>%
  relocate("asp_height_mm", .after = "source_well") %>%
  mutate(ot2set = case_when(plate %in% c("primerPlate1",
                                         "primerPlate2",
                                         "primerPlate3",
                                         "dilution123")
                            ~ "plates123", 
                            plate %in% c("primerPlate4",
                                         "primerPlate5",
                                         "primerPlate6",
                                         "dilution456",
                                         "water_finalPools")
                            ~ "plates456")) %>%
  relocate(ot2set)
  

ot2_final_noSet_noPlates <- ot2_final_plusPlates %>%
  select(-ot2set, -plate)
```

Export

```{r}
# comma-separated string with end-of-line delimiter
ot2_string <- format_delim(ot2_final_plusPlates,
                           delim = ",",
                           eol = "\\n")

sink("./robotProtocols/inputFiles/tamGenetics_primerDilutionsPooling_postOpt2_string.txt")

ot2_string

sink()

# csv file
write.csv(ot2_final_plusPlates, "./robotProtocols/inputFiles/primerDilutionsPooling_postOpt2_forOT2.csv", row.names = F)
```

# 5 OT2 input for SAMPLES 

## 5.1 Extraction data

### Xtn master lists (27 July 2023)

```{r}
xtnsBlood <- read.csv("./extractions/tamGenetics_xtnsBlood_master_updated27July2023.csv")
xtnsFecal <- read.csv("./extractions/tamGenetics_xtnsFecal_master_updated27July2023.csv")
xtnsHair <- read.csv("./extractions/tamGenetics_xtnsHair_master_updated27July2023.csv")

xtnsAll <- rbind(xtnsBlood, xtnsFecal, xtnsHair)
```

### Xtns to SKIP

#### Xtn records
```{r}
xtns_toSkip_xtnRecords <- xtnsAll %>%
  filter(xtnRemaining != "yes") %>%
  select(c("animalID", "xtnID", "xtnRemaining"))
```

#### tamRun3 records

```{r}
xtns_toSkip_tamRun3Records <- md_tamRun3 %>%
  filter(str_detect(pcr1Notes_p12, "OUT")) %>%
  mutate(
    xtnID = case_when(
      str_detect(xtnUsed_p12, "-") ~ str_c(sampleType, "_p", str_sub(xtnUsed_p12, 1, 1), ".", str_sub(xtnUsed_p12, 3, length(xtnUsed_p12)), "_", tolower(str_sub(pcr1Notes_p12, 1, 2))),
      .default = str_c(sampleType, "_t", xtnUsed_p12, "_", tolower(str_sub(pcr1Notes_p12, 1, 2)))
      ),
    xtnRemaining = "no"
    ) %>%
  select(c("animalID", "xtnID", "xtnRemaining"))
```

#### On-site xtn & bead checks

```{r}
# xtn checks
xtns_toSkip_onsiteChecks <- data.frame(
  xtnID = c("sampleType_p#.w#_e#",
            "sampleType_t#_e#",
            "sampleType_tx#_e#"
            )) %>%
  merge(., xtnsAll[, c("animalID", "xtnID")], by = "xtnID", all.x = T) %>%
  mutate(xtnRemaining = "no") %>%
  select(c("animalID", "xtnID", "xtnRemaining"))

# bead checks
xtnPlate_beadChecks <- data.frame(
  xtnLoc = rep(c("p1", "p2", "p3", "p4"), 2),
  eltn = c(rep("e1", 4), rep("e2", 4))
  ) %>%
  mutate(
    xtnPlate_beadsPresent = case_when(
      xtnLoc == "p1" & eltn == "e1" ~ "tbd",
      xtnLoc == "p1" & eltn == "e2" ~ "tbd",
      xtnLoc == "p2" & eltn == "e1" ~ "tbd",
      xtnLoc == "p2" & eltn == "e2" ~ "tbd",
      xtnLoc == "p3" & eltn == "e1" ~ "tbd",
      xtnLoc == "p3" & eltn == "e2" ~ "tbd",
      xtnLoc == "p4" & eltn == "e1" ~ "tbd",
      xtnLoc == "p4" & eltn == "e2" ~ "tbd")
) %>%
  arrange(xtnLoc)
```

#### All xtns to skip

```{r}
xtns_toSkip <- rbind(xtns_toSkip_xtnRecords, xtns_toSkip_tamRun3Records, xtns_toSkip_onsiteChecks) %>%
  arrange(xtnID) %>%
  unique()
```

#### Update xtn records

Update 'xtnRemaining' and add 'xtnPlate_beadsPresent'

```{r}
xtnsBlood_updated <- xtnsBlood %>%
  mutate(
    xtnRemaining = case_when(
      xtnID %in% xtns_toSkip$xtnID ~ "no",
      .default = xtnRemaining
      )) %>%
  merge(., xtnPlate_beadChecks[, c("xtnLoc", "eltn", "xtnPlate_beadsPresent")], by = c("xtnLoc", "eltn"), all.x = T) %>%
  relocate(xtnPlate_beadsPresent, .before = xtnNotes) %>%
  relocate(c("xtnLoc", "eltn"), .after = xtnID)

xtnsFecal_updated <- xtnsFecal %>%
  mutate(xtnRemaining = case_when(
    xtnID %in% xtns_toSkip$xtnID ~ "no",
    .default = xtnRemaining
  )) %>%
  merge(., xtnPlate_beadChecks[, c("xtnLoc", "eltn", "xtnPlate_beadsPresent")], by = c("xtnLoc", "eltn"), all.x = T) %>%
  relocate(xtnPlate_beadsPresent, .before = xtnNotes) %>%
  relocate(c("xtnLoc", "eltn"), .after = xtnID)

xtnsHair_updated <- xtnsHair %>%
  mutate(xtnRemaining = case_when(
    xtnID %in% xtns_toSkip$xtnID ~ "no",
    .default = xtnRemaining
  )) %>%
  merge(., xtnPlate_beadChecks[, c("xtnLoc", "eltn", "xtnPlate_beadsPresent")], by = c("xtnLoc", "eltn"), all.x = T) %>%
  relocate(xtnPlate_beadsPresent, .before = xtnNotes) %>%
  relocate(c("xtnLoc", "eltn"), .after = xtnID)

xtnsAll_updated <- rbind(xtnsBlood_updated, xtnsFecal_updated, xtnsHair_updated)
```

### Xtns REMAINING by animalID

```{r}
xtnsRemaining_byAnimalID <- xtnsAll_updated %>%
  filter(xtnRemaining == "yes") %>%
  mutate(sampleType_animalID = str_c(sampleType, "_", animalID)) %>%
  group_by(sampleType_animalID) %>%
  summarise_all(~ toString(unique(.))) %>%
  select(c("sampleType_animalID", "xtnID")) %>%
  separate(xtnID, c("xtnID.1", "xtnID.2", "xtnID.3", "xtnID.4"), ",")
```

### Xtns to ADJUST

```{r}
xtns_toAdjust <- xtnsAll_updated %>%
  filter(str_detect(xtnNotes, "dilute")) %>%
  select(c("animalID", "xtnID", "xtnNotes"))
```

## 5.2 PCR1

### 5.2.1 csv headers

```{r}
pcr1Headers <- c("ot2set",
                 "reagent",
                 "source_labware",
                 "source_slot",
                 "source_well",
                 "asp_height",
                 "dest_labware",
                 "dest_slot",
                 "dest_well",
                 "transfer_volume")
```

### 5.2.2 30x3

#### Xtn transfer info

**All samples + dest_wells**

```{r}
# Original 30x3 sample list
bfh_samples <- read.csv("./01_run2_fecalHairBlood/03_run2GTscore/metadata_tamRun2.csv") %>%
  filter(str_detect(pcr1Description, "again|blood")) %>%
  filter(!is.na(sampleType)) %>%
  mutate(xtnTube = gsub("\\s+", "", xtnTube))

# Select negatives
bfh_negsBlood <- xtnsAll_updated %>%
  filter(animalID %in% c("xtnNeg_p1.C4","xtnNeg_p2.C4")) %>%
  filter(eltn == "e1") %>%
  mutate(sampleType = str_c("blood_", sampleType)) %>%
  select(c("sampleType", "animalID", "xtnID"))
bfh_negsFecal <- data.frame(
  sampleType = c("fecal_pcr1Neg1", "fecal_pcr1Neg2"),
  animalID = c("pcr1Neg1", "pcr1Neg2"),
  xtnID = rep(NA, 2))
bfh_negsHair <- xtnsAll_updated %>%
  filter(animalID %in% c("xtnNeg_p3.C4","xtnNeg_p4.C4")) %>%
  filter(eltn == "e1") %>%
  mutate(sampleType = str_c("hair_", sampleType)) %>%
  select(c("sampleType", "animalID", "xtnID"))

bfh_negsAll <- rbind(bfh_negsBlood, bfh_negsFecal, bfh_negsHair) %>%
  rename("xtnTube" = "xtnID") %>%
  mutate(
    sampleType_animalID = case_when(
      str_detect(sampleType, "pcr1Neg") ~ sampleType,
      .default = str_c(sampleType, "_", str_sub(animalID, start = 8, end = -1))
      )
    ) %>%
  select(c("sampleType", "animalID", "xtnTube", "sampleType_animalID"))

# add negs to 30x3 xtn list & reformat
bfh_xtns <- bfh_samples %>%
  select(c("sampleType", "animalID", "xtnTube")) %>%
  mutate(
    sampleType_animalID = str_c(sampleType, "_", animalID),
    xtnTube = case_when(
      sampleType == "fecal" ~ gsub("F", "", xtnTube),
      .default = xtnTube
      )
  ) %>%
  filter(duplicated(sampleType_animalID) == F) %>%
  rbind(., bfh_negsAll) %>%
  mutate(
    xtnID = case_when(
      sampleType == "fecal_pcr1Neg1" ~ "tH2O.1",
      sampleType == "fecal_pcr1Neg2" ~ "tH2O.2",
      str_detect(xtnTube, "Neg") ~ str_c(gsub("_.*", "", sampleType), "_", xtnTube),
      str_detect(xtnTube, "-") ~ str_c(sampleType, "_p", str_sub(xtnTube, 1, 1), ".", str_sub(xtnTube, 3, length(xtnTube)), "_e1"),
      str_detect(xtnTube, "A") ~ str_c(sampleType, "_tx", xtnTube, "_e1"),
      .default = str_c(sampleType, "_t", xtnTube, "_e1")
      ),
    xtnRemaining = case_when(
      xtnID %in% xtns_toSkip$xtnID ~ "no",
      .default = "yes"
      )
    ) %>%
  select(c("sampleType_animalID", "sampleType", "xtnID", "xtnRemaining")) %>%
  # replace "SKIP" xtns
  merge(., xtnsRemaining_byAnimalID, by = "sampleType_animalID", all.x = T) %>%
  # remove columns with all NA
  select(where(function(x) !all(is.na(x)))) %>%
  mutate(
    xtnID = case_when(
      xtnRemaining == "no" ~ xtnID.1,
      .default = xtnID
    )
  ) %>%
  select(c("sampleType_animalID", "sampleType", "xtnID")) %>%
  mutate(
    xtnLoc = case_when(
      str_detect(xtnID, "p") ~ gsub(".*[_]([^.]+)[.].*", "\\1", xtnID),
      str_detect(xtnID, "t") ~ gsub(".*[_]([^_]+)[_].*", "\\1", xtnID)
    ),
    xtnRow = case_when(
      str_detect(xtnID, "p") ~ str_sub(gsub(".*[.]([^_]+)[_].*", "\\1", xtnID), 1, 1),
      str_detect(xtnID, "t") ~ NA
    ),
    xtnCol = case_when(
      str_detect(xtnID, "p") ~ str_sub(gsub(".*[.]([^_]+)[_].*", "\\1", xtnID), 2, length(gsub(".*[.]([^_]+)[_].*", "\\1", xtnID))),
      str_detect(xtnID, "t") ~ NA
    ),
    eltn = case_when(
      str_detect(xtnID, "H2O") ~ NA,
      .default = str_sub(xtnID, -2, length(xtnID))
      )
    ) %>%
  # add whether plates have beads
  merge(., xtnPlate_beadChecks[, c("xtnLoc", "eltn", "xtnPlate_beadsPresent")], by = c("xtnLoc", "eltn"), all.x = T) %>%
  relocate(c("xtnLoc", "eltn"), .after = "xtnID") %>%
  arrange(gsub("_.*", "", sampleType_animalID), str_sub(xtnLoc, 1, 1), as.numeric(str_sub(xtnLoc, 2, length(xtnLoc))), as.numeric(xtnCol), xtnRow) %>%
  # place negatives evenly across plate
  mutate(ID = as.character(row_number())) %>%
  mutate(ID = case_when(
    ID == "5" ~ "11.1",
    ID == "21" ~ "30.1",
    ID == "63" ~ "42.1",
    ID == "64" ~ "52.1",
    ID == "66" ~ "67.1",
    ID == "78" ~ "86.1",
    .default = ID
  )) %>%
  arrange(as.numeric(ID))
```

**Blood samples**

```{r}
bfh_sourceWells_bloodTubes <- bfh_xtns %>%
  filter(!str_detect(xtnID, "tamTiger")) %>%
  filter(str_detect(xtnID, "blood_t")) %>%
  mutate(
    source_well.t = wellValues_tuberack[1:length(xtnID)],
    fecalSet = NA
  ) %>%
  select(c("xtnID", "source_well.t", "fecalSet"))

bfh_sourceWells_fecalTubes1to24 <- bfh_xtns %>%
  filter(str_detect(sampleType, "fecal")) %>%
  .[1:24,] %>%
  mutate(
    source_well.t = wellValues_tuberack[1:length(xtnID)],
    fecalSet = "set4_fecalTubes_1to24"
    ) %>%
  select(c("xtnID", "source_well.t", "fecalSet"))
bfh_sourceWells_fecalTubes25to32 <- bfh_xtns %>%
  filter(str_detect(sampleType, "fecal")) %>%
  .[25:32,] %>%
  mutate(
    source_well.t = wellValues_tuberack[1:length(xtnID)],
    fecalSet = "set5_fecalTubes_25to32"
    ) %>%
  select(c("xtnID", "source_well.t", "fecalSet"))

bfh_sourceWells_hairTubes <- bfh_xtns %>%
  filter(str_detect(xtnID, "hair_t")) %>%
  mutate(
    source_well.t = wellValues_tuberack[1:length(xtnID)],
    fecalSet = NA
  ) %>%
  select(c("xtnID", "source_well.t", "fecalSet"))

bfh_sourceWells_allTubes <- rbind(bfh_sourceWells_bloodTubes,
                                  bfh_sourceWells_fecalTubes1to24,
                                  bfh_sourceWells_fecalTubes25to32,
                                  bfh_sourceWells_hairTubes)

bfh_transferInfo_pp2 <- bfh_xtns %>%
  merge(., bfh_sourceWells_allTubes, all.x = T) %>%
  arrange(as.numeric(ID)) %>%
  mutate(
    ot2set = case_when(
      (str_detect(xtnID, "blood") & str_detect(xtnID, "p1")) ~ "set1_xtnPlate1",
      (str_detect(xtnID, "blood") & str_detect(xtnID, "p2")) ~ "set2_xtnPlate2",
      str_detect(xtnID, "blood_t") ~ "set3_bloodTubes",
      str_detect(sampleType, "fecal") ~ fecalSet,
      str_detect(xtnID, "hair") & str_detect(xtnID, "p3") ~ "set6_xtnPlate3",
      str_detect(xtnID, "hair") & str_detect(xtnID, "p4") ~ "set7_xtnPlate4",
      str_detect(xtnID, "hair_t") ~ "set8_hairTubes",
    ),
    reagent = xtnID,
    source_labware = case_when(
      xtnPlate_beadsPresent == "yes" ~ magblock,
      str_detect(xtnID, "p") ~ pcrPlate,
      str_detect(xtnID, "t") ~ tubeRack
    ),
    source_slot = case_when(
      xtnPlate_beadsPresent == "yes" ~ 7,
      (str_detect(xtnID, "p") & str_detect(xtnID, "e1")) ~ 1,
      (str_detect(xtnID, "p") & str_detect(xtnID, "e2")) ~ 4,
      str_detect(xtnID, "t") ~ 3
    ),
    source_well = case_when(
      str_detect(xtnID, "p") ~ str_c(xtnRow, xtnCol),
      .default = source_well.t),
    asp_height = 1,
    dest_labware = pcrPlate,
    dest_slot = 2,
    dest_well = wellValues_plate,
    transfer_volume = case_when(
      str_detect(sampleType, "blood") ~ 2,
      .default = 4
    )
  ) %>%
  select(pcr1Headers) %>%
  arrange(ot2set) %>%
  mutate(id = row_number())

bfh_transferInfo_pp3 <- bfh_transferInfo_pp2 %>%
  mutate(dest_slot = 5)

bfh_pcr1_xtnTransfer <- rbind(bfh_transferInfo_pp2,
                               bfh_transferInfo_pp3) %>%
  arrange(id, dest_slot) %>%
  select(-id)
```

#### Full OT2 input

```{r}
bfh_pcr1_waterTransfer <- data.frame(
  ot2set = rep("set0.1_water", 64),
  reagent = "mastermix",
  source_labware = pcrPlate,
  source_slot = 6,
  source_well = "B1",
  asp_height = 1,
  dest_labware = pcrPlate,
  dest_slot = c(rep(2, 32), rep(5, 32)),
  dest_well = rep(wellValues_plate[1:32], 2),
  transfer_volume = 2
) %>%
  select(pcr1Headers)

bfh_pcr1_mmTransfer <- data.frame(
  ot2set = rep("set0.2_mm", 24),
  reagent = "mastermix",
  source_labware = pcrPlate,
  source_slot = 6,
  source_well = "A1",
  asp_height = 1,
  dest_labware = pcrPlate,
  dest_slot = c(rep(2, 12), rep(5, 12)),
  dest_well = grep("A", wellValues_plate, value = T),
  transfer_volume = 6
) %>%
  select(pcr1Headers)

bfh_pcr1_ot2Input <- rbind(bfh_pcr1_waterTransfer,
                           bfh_pcr1_mmTransfer,
                           bfh_pcr1_xtnTransfer)
```

Export

```{r}
# comma-separated string with end-of-line delimiter
ot2_string <- format_delim(bfh_pcr1_ot2Input,
                           delim = ",",
                           eol = "\\n")

sink("./robotProtocols/inputFiles/30x3_pcr1_forOT2_string.txt")

ot2_string

sink()

# csv file
write.csv(bfh_pcr1_ot2Input, "./robotProtocols/inputFiles/30x3_pcr1_forOT2.csv", row.names = F)
```

### 5.2.3 tamGenetics

#### Plates 1 & 2 + Strip 1 (blood samples)

Input:
-   extraction plates
-   mastermix w/primer pool 1

Output:
-   2 PCR1 plates
-   1 PCR1 strip

##### Xtns

```{r}
pcr1_xtns_p1e1 <- xtnPlate_map_p1e1 %>%
  select(c("reagent", "well")) %>%
  dplyr::rename("source_well" = "well") %>%
  mutate(source_slot = 1) %>%
  mutate(dest_slot = 3) %>%
  mutate(dest_well = source_well) %>%
  select(c("reagent", "source_slot", "source_well", "dest_slot", "dest_well")) %>%
  filter(!reagent %in% xtnsOut_platesE1$reagent)

pcr1_xtns_p1e2 <- xtnPlate_map_p1e2 %>%
  select(c("reagent", "well")) %>%
  dplyr::rename("source_well" = "well") %>%
  mutate(source_slot = 1) %>%
  mutate(dest_slot = 3) %>%
  mutate(dest_well = source_well) %>%
  select(c("reagent", "source_slot", "source_well", "dest_slot", "dest_well")) %>%
  # Remove if e1 is present
  filter(!source_well %in% pcr1_xtns_p1e1$source_well) %>%
  # Remove if e2 xtn is out
  filter(!reagent %in% xtnsOut_platesE2$reagent)

pcr1_xtns_te1 <- 
```


```{r}
mm_slot2_toPCR96_slot3 <- matrix(c(
  # source slot
  rep(2, 12),
  # source well
  rep("A1", 12),
  # dest slot
  rep(3, 12),
  # dest well
  grep("A", wellValues, value = T)),
  ncol = 4)

xtns_magSlot1_toPCR_slot3 <- matrix(c(
  # source slot
  rep(1, 96),
  # source well
  wellValues,
  # dest slot
  rep(3, 96),
  # dest well
  wellValues),
  ncol = 4)


  
pcr1_p12s1_mm <- 
  
pcr1_p12s1_csv <-
```

#### Plate 3a & 3b (hair samples)

```{r}

```

#### Plate 4a & 4b (hair samples)

```{r}

```

#### Plate 5ab (hair samples)

```{r}

```

## 5.3 1:20 dilutions

## 5.4 PCR2

# SCRAPS

## 5.1 Data

### 5.1.1 Maps & xtns

#### Blood samples

```{r}
# All xtns
xtnMaster_blood <- read.csv("./extractions/tamGenetics_extractionsBlood_master_21July2023.csv")

# Original plate/tube set-up
xtnMap_blood_p1e1 <- xtnMaster_blood %>%
  filter(lysisPlate == 1) %>%
  arrange(lysisCol, lysisRow) %>%
  mutate(sampleType = "blood") %>%
  mutate(xtnLoc = str_c("p", lysisPlate)) %>%
  mutate(well = str_c(lysisRow, lysisCol)) %>%
  mutate(eltn = "e1") %>%
  mutate(reagent = str_c(sampleType, "_", xtnLoc, eltn)) %>%
  select(c("animalID", "sampleType", "xtnLoc", "eltn", "well", "reagent"))
  
xtnMap_blood_p1e2 <- xtnMaster_blood %>%
  filter(lysisPlate == 1) %>%
  arrange(lysisCol, lysisRow) %>%
  mutate(sampleType = "blood") %>%
  mutate(xtnLoc = str_c("p", lysisPlate)) %>%
  mutate(well = str_c(lysisRow, lysisCol)) %>%
  mutate(eltn = "e2") %>%
  mutate(reagent = str_c(sampleType, "_", xtnLoc, eltn)) %>%
  select(c("animalID", "sampleType", "xtnLoc", "eltn", "well", "reagent"))

xtnMap_blood_p2e1 <- xtnMaster_blood %>%
  filter(lysisPlate == 2) %>%
  arrange(lysisCol, lysisRow) %>%
  mutate(sampleType = "blood") %>%
  mutate(xtnLoc = str_c("p", lysisPlate)) %>%
  mutate(well = str_c(lysisRow, lysisCol)) %>%
  mutate(eltn = "e1") %>%
  mutate(reagent = str_c(sampleType, "_", xtnLoc, eltn)) %>%
  select(c("animalID", "sampleType", "xtnLoc", "eltn", "well", "reagent"))
  
xtnMap_blood_p2e2 <- xtnMaster_blood %>%
  filter(lysisPlate == 2) %>%
  arrange(lysisCol, lysisRow) %>%
  mutate(sampleType = "blood") %>%
  mutate(xtnLoc = str_c("p", lysisPlate)) %>%
  mutate(well = str_c(lysisRow, lysisCol)) %>%
  mutate(eltn = "e2") %>%
  mutate(reagent = str_c(sampleType, "_", xtnLoc, eltn)) %>%
  select(c("animalID", "sampleType", "xtnLoc", "eltn", "well", "reagent"))

xtnMap_blood_tubesE1 <- xtnMaster_blood %>%
  filter(str_detect(lysisPlate, "tube")) %>%
  mutate(sampleType = "blood") %>%
  mutate(xtnLoc = str_c("t", sampleTube)) %>%
  mutate(well = NA) %>%
  mutate(eltn = "e1") %>%
  mutate(reagent = str_c(sampleType, "_", xtnLoc, eltn)) %>%
  select(c("animalID", "sampleType", "xtnLoc", "eltn", "well", "reagent"))

xtnMap_blood_tubesE2 <- xtnMaster_blood %>%
  filter(str_detect(lysisPlate, "tube")) %>%
  mutate(sampleType = "blood") %>%
  mutate(xtnLoc = str_c("t", sampleTube)) %>%
  mutate(well = NA) %>%
  mutate(eltn = "e2") %>%
  mutate(reagent = str_c(sampleType, "_", xtnLoc, eltn)) %>%
  select(c("animalID", "sampleType", "xtnLoc", "eltn", "well", "reagent"))

# Completed PLATE xtns
xtnPlates_blood_e1 <- xtnMaster_blood %>%
  filter(xtnPlate %in% c(1, 2)) %>%
  arrange(xtnCol, xtnRow) %>%
  mutate(sampleType = "blood") %>%
  mutate(xtnLoc = str_c("p", xtnPlate)) %>%
  mutate(well = str_c(xtnRow, xtnCol)) %>%
  mutate(eltn = "e1") %>%
  mutate(reagent = str_c(sampleType, "_", xtnLoc, eltn)) %>%
  select(c("animalID", "sampleType", "xtnLoc", "eltn", "well", "reagent"))

xtnPlates_blood_e2 <- xtnMaster_blood %>%
  filter(xtnPlate %in% c(1, 2)) %>%
  arrange(xtnCol, xtnRow) %>%
  mutate(sampleType = "blood") %>%
  mutate(xtnLoc = str_c("p", xtnPlate)) %>%
  mutate(well = str_c(xtnRow, xtnCol)) %>%
  mutate(eltn = "e2") %>%
  mutate(reagent = str_c(sampleType, "_", xtnLoc, eltn)) %>%
  select(c("animalID", "sampleType", "xtnLoc", "eltn", "well", "reagent"))

# Completed TUBE xtns
xtnTubes_blood_e1 <- xtnMaster_blood %>%
  filter(!is.na(xtnTube)) %>%
  mutate(sampleType = "blood") %>%
  mutate(xtnLoc = str_c("t", xtnTube)) %>%
  mutate(well = NA) %>%
  mutate(eltn = "e1") %>%
  mutate(reagent = str_c(sampleType, "_", xtnLoc, eltn)) %>%
  mutate(mapID = case_when(
    str_detect(lysisPlate, "tube") ~ reagent,
    .default = str_c("blood_p", lysisPlate, "e1_", lysisRow, lysisCol))) %>%
  select(c("animalID", "sampleType", "mapID", "xtnLoc", "eltn", "well", "reagent"))

xtnTubes_blood_e2 <- xtnMaster_blood %>%
  filter(!is.na(xtnTube)) %>%
  mutate(sampleType = "blood") %>%
  mutate(xtnLoc = str_c("t", xtnTube)) %>%
  mutate(well = NA) %>%
  mutate(eltn = "e2") %>%
  mutate(reagent = str_c(sampleType, "_", xtnLoc, eltn)) %>%
  mutate(mapID = case_when(
    str_detect(lysisPlate, "tube") ~ reagent,
    .default = str_c("blood_p", lysisPlate, "e2_", lysisRow, lysisCol))) %>%
  select(c("animalID", "sampleType", "mapID", "xtnLoc", "eltn", "well", "reagent"))
```

#### Fecal samples

```{r}
xtnMaster_fecal <- read.csv("./extractions/tamGenetics_extractionsFecal_master_21July2023.csv")

xtnMap_fecal_tubesE1 <- xtnMaster_fecal %>%
  filter(elution == "E1") %>%
  mutate(sampleType = "fecal") %>%
  mutate(xtnLoc = str_c("t", sampleTube)) %>%
  mutate(well = NA) %>%
  mutate(eltn = tolower(elution)) %>%
  mutate(reagent = str_c(sampleType, "_", xtnLoc, eltn)) %>%
  select(c("animalID", "sampleType", "xtnLoc", "eltn", "well", "reagent"))

xtnMap_fecal_tubesE2 <- xtnMaster_fecal %>%
  filter(elution == "E2") %>%
  mutate(sampleType = "fecal") %>%
  mutate(xtnLoc = str_c("t", sampleTube)) %>%
  mutate(well = NA) %>%
  mutate(eltn = tolower(elution)) %>%
  mutate(reagent = str_c(sampleType, "_", xtnLoc, eltn)) %>%
  select(c("animalID", "sampleType", "xtnLoc", "eltn", "well", "reagent"))

xtnTubes_fecal_e1 <- xtnMap_fecal_tubesE1 %>%
  mutate(mapID = reagent) %>%
  select(c("animalID", "sampleType", "mapID", "xtnLoc", "eltn", "well", "reagent"))

xtnTubes_fecal_e2 <- xtnMap_fecal_tubesE2 %>%
  mutate(mapID = reagent) %>%
  select(c("animalID", "sampleType", "mapID", "xtnLoc", "eltn", "well", "reagent"))
```

#### Hair samples

```{r}
xtnMaster_hair <- read.csv("./extractions/tamGenetics_extractionsHair_master_21July2023.csv") %>%
  mutate(xtnTube = gsub("\\s+", "", xtnTube))

# Original plate/tube set-up
xtnMap_hair_p3e1 <- xtnMaster_hair %>%
  filter(xtnPlate == 3) %>%
  mutate(sampleType = "hair") %>%
  mutate(xtnLoc = str_c("p", xtnPlate)) %>%
  mutate(eltn = "e1") %>%
  arrange(xtnCol, xtnRow) %>%
  mutate(well = str_c(xtnRow, xtnCol)) %>%
  mutate(reagent = str_c("hair_", xtnLoc, eltn)) %>%
  select(c("animalID", "sampleType", "xtnLoc", "eltn", "well", "reagent"))

xtnMap_hair_p3e2 <- xtnMaster_hair %>%
  filter(xtnPlate == 3) %>%
  mutate(sampleType = "hair") %>%
  mutate(xtnLoc = str_c("p", xtnPlate)) %>%
  mutate(eltn = "e2") %>%
  arrange(xtnCol, xtnRow) %>%
  mutate(well = str_c(xtnRow, xtnCol)) %>%
  mutate(reagent = str_c("hair_", xtnLoc, eltn)) %>%
  select(c("animalID", "sampleType", "xtnLoc", "eltn", "well", "reagent"))

xtnMap_hair_p4e1 <- xtnMaster_hair %>%
  filter(xtnPlate == 4) %>%
  mutate(sampleType = "hair") %>%
  mutate(xtnLoc = str_c("p", xtnPlate)) %>%
  mutate(eltn = "e1") %>%
  arrange(xtnCol, xtnRow) %>%
  mutate(well = str_c(xtnRow, xtnCol)) %>%
  mutate(reagent = str_c("hair_", xtnLoc, eltn)) %>%
  select(c("animalID", "sampleType", "xtnLoc", "eltn", "well", "reagent"))

xtnMap_hair_p4e2 <- xtnMaster_hair %>%
  filter(xtnPlate == 4) %>%
  mutate(sampleType = "hair") %>%
  mutate(xtnLoc = str_c("p", xtnPlate)) %>%
  mutate(eltn = "e2") %>%
  arrange(xtnCol, xtnRow) %>%
  mutate(well = str_c(xtnRow, xtnCol)) %>%
  mutate(reagent = str_c("hair_", xtnLoc, eltn)) %>%
  select(c("animalID", "sampleType", "xtnLoc", "eltn", "well", "reagent"))

xtnMap_hair_tubesE1 <- xtnMaster_hair %>%
  filter(is.na(xtnPlate)) %>%
  filter(elution != "E2") %>%
  filter(!grepl("-", xtnStatus)) %>%
  filter(!is.na(xtnTube)) %>%
  mutate(sampleType = "hair") %>%
  mutate(xtnLoc = 
           case_when(
             str_detect(xtnTube, "A,B") ~ str_c("tx", gsub("A,B", "A", xtnTube)),
             .default = str_c("t", xtnTube))) %>%
  mutate(eltn = "e1") %>%
  mutate(well = NA) %>%
  mutate(reagent = str_c(sampleType, "_", xtnLoc, eltn)) %>%
  select(c("animalID", "sampleType", "xtnLoc", "eltn", "well", "reagent"))

xtnMap_hair_tubesE2 <- xtnMaster_hair %>%
  filter(is.na(xtnPlate)) %>%
  filter(elution != "E2") %>%
  filter(!grepl("-", xtnStatus)) %>%
  filter(!is.na(xtnTube)) %>%
  mutate(sampleType = "hair") %>%
  mutate(xtnLoc = 
           case_when(
             str_detect(xtnTube, "A,B") ~ str_c("tx", gsub("A,B", "B", xtnTube)),
             .default = str_c("t", xtnTube))) %>%
  mutate(eltn = "e2") %>%
  mutate(well = NA) %>%
  mutate(reagent = str_c(sampleType, "_", xtnLoc, eltn)) %>%
  select(c("animalID", "sampleType", "xtnLoc", "eltn", "well", "reagent"))

# Completed PLATE xtns
xtnPlates_hair_p3e1 <- xtnMaster_hair %>%
  filter(xtnPlate == 3) %>%
  mutate(sampleType = "hair") %>%
  mutate(xtnLoc = str_c("p", xtnPlate)) %>%
  mutate(eltn = "e1") %>%
  arrange(xtnCol, xtnRow) %>%
  mutate(well = str_c(xtnRow, xtnCol)) %>%
  mutate(reagent = str_c("hair_", xtnLoc, eltn)) %>%
  select(c("animalID", "sampleType", "xtnLoc", "eltn", "well", "reagent"))

xtnPlates_hair_p3e2 <- xtnMaster_hair %>%
  filter(xtnPlate == 3) %>%
  mutate(sampleType = "hair") %>%
  mutate(xtnLoc = str_c("p", xtnPlate)) %>%
  mutate(eltn = "e2") %>%
  arrange(xtnCol, xtnRow) %>%
  mutate(well = str_c(xtnRow, xtnCol)) %>%
  mutate(reagent = str_c("hair_", xtnLoc, eltn)) %>%
  select(c("animalID", "sampleType", "xtnLoc", "eltn", "well", "reagent"))

xtnPlates_hair_p4e1 <- xtnMaster_hair %>%
  filter(xtnPlate == 4) %>%
  mutate(sampleType = "hair") %>%
  mutate(xtnLoc = str_c("p", xtnPlate)) %>%
  mutate(eltn = "e1") %>%
  arrange(xtnCol, xtnRow) %>%
  mutate(well = str_c(xtnRow, xtnCol)) %>%
  mutate(reagent = str_c("hair_", xtnLoc, eltn)) %>%
  select(c("animalID", "sampleType", "xtnLoc", "eltn", "well", "reagent"))

xtnPlates_hair_p4e2 <- xtnMaster_hair %>%
  filter(xtnPlate == 4) %>%
  mutate(sampleType = "hair") %>%
  mutate(xtnLoc = str_c("p", xtnPlate)) %>%
  mutate(eltn = "e2") %>%
  arrange(xtnCol, xtnRow) %>%
  mutate(well = str_c(xtnRow, xtnCol)) %>%
  mutate(reagent = str_c("hair_", xtnLoc, eltn)) %>%
  select(c("animalID", "sampleType", "xtnLoc", "eltn", "well", "reagent"))

# Completed TUBE xtns
xtnTubes_hair_e1 <- xtnMaster_hair %>%
  filter(!is.na(xtnTube)) %>%
  filter(elution != "E2") %>%
  filter(!is.na(xtnTube)) %>%
  mutate(sampleType = "hair") %>%
  mutate(xtnLoc = 
           case_when(
             str_detect(xtnTube, "A,B") ~ str_c("tx", gsub("A,B", "A", xtnTube)),
             .default = str_c("t", xtnTube))) %>%
  mutate(well = NA) %>%
  mutate(eltn = "e1") %>%
  mutate(reagent = str_c(sampleType, "_", xtnLoc, eltn)) %>%
  mutate(mapID =
           case_when(
             grepl("-", xtnStatus) ~ gsub("replacement for plate xtn ", "", xtnStatus),
             str_detect(xtnTube, "A,B") ~ str_c("hair_tx", gsub("A,B", "A", xtnTube, "_e1")),
             .default = str_c("hair_t", xtnTube, eltn))) %>%
  mutate(mapID = 
           case_when(
             grepl("-", mapID) ~ str_c("hair_p", gsub("(.*)-(.*)", "\\1", mapID), "e1_", gsub("(.*)-(.*)", "\\2", mapID)),
             .default = mapID)) %>%
  select(c("animalID", "sampleType", "mapID", "xtnLoc", "eltn", "well", "reagent"))

xtnTubes_hair_e2 <- xtnMaster_hair %>%
  filter(!is.na(xtnTube)) %>%
  filter(elution != "E2") %>%
  filter(!is.na(xtnTube)) %>%
  mutate(sampleType = "hair") %>%
  mutate(xtnLoc = 
           case_when(
             str_detect(xtnTube, "A,B") ~ str_c("tx", gsub("A,B", "B", xtnTube)),
             .default = str_c("t", xtnTube))) %>%
  mutate(well = NA) %>%
  mutate(eltn = "e2") %>%
  mutate(reagent = str_c(sampleType, "_", xtnLoc, eltn)) %>%
  mutate(mapID =
           case_when(
             grepl("-", xtnStatus) ~ gsub("replacement for plate xtn ", "", xtnStatus),
             str_detect(xtnTube, "A,B") ~ str_c("hair_tx", gsub("A,B", "B", xtnTube, "_e2")),
             .default = str_c("hair_t", xtnTube, "_e2"))) %>%
  mutate(mapID = 
           case_when(
             grepl("-", mapID) ~ str_c("hair_p", gsub("(.*)-(.*)", "\\1", mapID), "e2_", gsub("(.*)-(.*)", "\\2", mapID)),
             .default = mapID)) %>%
  select(c("animalID", "sampleType", "mapID", "xtnLoc", "eltn", "well", "reagent"))
```

### 5.1.2 Xtns to SKIP

Extractions known to be OUT or non-functional

#### Xtn records

(blood & fecal master xtn files don't have xtn status)

```{r}
tx9B <- data.frame(sampleType = "hair",
                   xtnLoc = "tx9B",
                   eltn = "e2",
                   well = NA,
                   reagent = "hair_tx9B")

xtnsOut_hairMaster <- xtnMaster_hair %>%
  filter(str_detect(xtnStatus, c("OUT|nonfxnl"))) %>%
  mutate(sampleType = "hair") %>%
  mutate(xtnLoc = 
           case_when(
             is.na(xtnPlate) ~ str_c("tx", gsub("A,B", "A", xtnTube)),
             .default = str_c("p", xtnPlate))) %>%
  mutate(eltn = tolower(elution)) %>%
  mutate(well = str_c(xtnRow, xtnCol)) %>%
  mutate(reagent = str_c(sampleType, "_", xtnLoc, eltn)) %>%
  select(c("sampleType", "xtnLoc", "eltn", "well", "reagent")) %>%
  rbind(., tx9B)
```

#### tamRun3 records

```{r}
# tamRun3 metadata
md_tamRun3 <- read.csv("./03_tamRun3/03_run3GTscore/tamRun3_metadata_5June2023.csv")

t16e2 <- data.frame(sampleType = "hair",
                   xtnLoc = "t16e2",
                   eltn = "e2",
                   well = NA,
                   reagent = "hair_t16e2")

xtnsOut_tamun3 <- md_tamRun3 %>%
  filter(str_detect(pcr1Notes_p12, "OUT")) %>%
  select(c("pcr1Notes_p12", "xtnUsed_p12", "sampleType")) %>%
  mutate(xtnsOut_plate =
           case_when(
             grepl("-", xtnUsed_p12) ~ str_c(sampleType, "_p", str_sub(xtnUsed_p12, 1, 1)),
             .default = str_c(sampleType, "_t", xtnUsed_p12))) %>%
  mutate(xtnsOut_loc = 
           case_when(
             str_detect(pcr1Notes_p12, fixed("e1 is out", ignore_case = T)) ~ str_c(xtnsOut_plate, "_e1"),
             .default = str_c(xtnsOut_plate, "_e2"))) %>%
  mutate(well = 
           case_when(
             grepl("-", xtnUsed_p12) ~ str_sub(xtnUsed_p12, 3, length(xtnUsed_p12)))) %>%
  select(c("xtnsOut_loc", "well")) %>%
  separate(., xtnsOut_loc, into = c("sampleType", "xtnLoc", "eltn")) %>%
  mutate(reagent = str_c(sampleType, "_", xtnLoc, eltn)) %>%
  rbind(., t16e2)
```

#### On-site xtn checks

```{r}
xtnsOut_onsiteChecks <- matrix(c("sampleType_p#_eltn_well",
                                 "sampleType_t#_eltn_NA"),
                               ncol = 1) %>%
  `colnames<-`(., "xtnsOut") %>%
  as.data.frame() %>%
  separate(., xtnsOut, into = c("sampleType", "xtnLoc", "eltn", "well"), sep = "_") %>%
  mutate(reagent = str_c(sampleType, "_", xtnLoc, eltn))
```

#### Full list

```{r}
xtnsOut <- rbind(xtnsOut_hairMaster, xtnsOut_tamun3, xtnsOut_onsiteChecks)

xtnsOut_platesE1 <- xtnsOut %>%
  filter(grepl("p", xtnLoc)) %>%
  filter(eltn == "e1")

xtnsOut_platesE2 <- xtnsOut %>%
  filter(grepl("p", xtnLoc)) %>%
  filter(eltn == "e2")

xtnsOut_tubesE1 <- xtnsOut %>%
  filter(grepl("t", xtnLoc)) %>%
  filter(!grepl("tamTiger", xtnLoc)) %>%
  filter(eltn == "e1")

xtnsOut_tubesE2 <- xtnsOut %>%
  filter(grepl("t", xtnLoc)) %>%
  filter(!grepl("tamTiger", xtnLoc)) %>%
  filter(eltn == "e2")
```

```{r}
xtnList_blood_e1 <- xtnMaster_blood %>%
  mutate(sampleType =
           "blood") %>%
  mutate(animalID =
           case_when(animalID == "BLANK" ~ str_c("neg_", xtnPlate, xtnRow, xtnCol, "_e1"),
                              .default = animalID)) %>%
  mutate(origLocID =
           case_when(
             !str_detect(lysisPlate, "tube") ~ str_c("p", lysisPlate, "_", lysisRow, lysisCol),
             str_detect(lysisPlate, "tube") ~ str_c("t", xtnTube, "_blood_e1"))) %>%
  mutate(origLoc =
           case_when(
             grepl("t", origLocID) ~ "tube",
             .default = "plate")) %>%
  mutate(xtn =
           case_when(
             is.na(xtnTube) ~ str_c("p", xtnPlate, "_", xtnRow, xtnCol, "_e1"),
             .default = str_c("t", xtnTube, "_blood_e1"))) %>%
  mutate(xtnType =
           case_when(grepl("t", xtn) ~ "tube",
                     .default = "plate")) %>%
  select(c("animalID", "sampleType", "origLoc", "origLocID", "xtnType", "xtn"))

xtnList_blood_e2 <- xtnList_blood_e1 %>%
  mutate(origLocID = gsub("e1", "e2", origLocID)) %>%
  mutate(xtn = gsub("e1", "e2", xtn))

xtnList_blood <- rbind(xtnList_blood_e1, xtnList_blood_e2)

# Fecal xtns


# Hair xtns
xtnMaster_hair <- read.csv("./extractions/tamGenetics_extractionsHair_master_21July2023.csv")

xtnList_hair_e1 <- xtnMaster_hair %>%
  filter(!if_all(c(xtnTube, xtnPlate), is.na)) %>%
  mutate(animalID =
           case_when(
             animalID == "BLANK" ~ str_c("neg_", xtnPlate, xtnRow, xtnCol, "_e1"),
             .default = animalID)) %>%
  mutate(sampleType =
           "hair") %>%
  mutate(origLocID =
           case_when(
             grepl("-", xtnStatus) ~ gsub("replacement for plate xtn ", "", xtnStatus),
             is.na(xtnPlate) ~ str_c("t", xtnTube, "_hair_", tolower(elution)),
             .default = str_c("p", xtnPlate, "_", xtnRow, xtnCol, "_", tolower(elution)))) %>%
  mutate(origLocID = 
           case_when(
             grepl("-", origLocID) ~ str_c("p", gsub("-", "_", origLocID), "_", tolower(elution)),
             .default = origLocID)) %>%
  mutate(origLoc = 
           case_when(
             grepl("tamTiger", origLocID) ~ "plate",
             grepl("t", origLocID) ~ "tube",
             .default = "plate")) %>%
  mutate(xtn =
           case_when(
             is.na(xtnTube) ~ str_c("p", xtnPlate, "_", xtnRow, xtnCol, "_", tolower(elution)),
             .default = str_c("t", xtnTube, "_hair_", tolower(elution)))) %>%
  mutate(xtnType =
           case_when(
             grepl("tamTiger", xtn) ~ "plate",
             grepl("t", xtn) ~ "tube",
             .default = "plate")) %>%
  select(c("animalID", "sampleType", "origLoc", "origLocID", "xtnType", "xtn"))

xtnList_hair_e2 <- xtnList_hair_e1 %>%
  mutate(origLocID = gsub("e1", "e2", origLocID)) %>%
  mutate(xtn = gsub("e1", "e2", xtn))

xtnList_hair <- rbind(xtnList_hair_e1, xtnList_hair_e2)

# Full xtn list
xtnList_all <- rbind(xtnList_blood, xtnList_fecal, xtnList_hair)
```

